### 2023-05-15 ms.Rmd
Add first (incomplete) MS draft

---
title: 'Bayesian assessment of heterogeneity in psychological phenomena'
shorttitle: Heterogeneity in psychology
leftheader: Heterogeneity in psychology
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: Department of Social Psychology, Tilburg University
    email: m.j.vuorre@tilburguniversity.edu
  - name: Niall Bolger
    affiliation: 2
    address: Department of Psychology, Columbia University
    email: nb2229@columbia.edu
affiliation:
  - id: 1
    institution: Department of Social Psychology, Tilburg University
  - id: 1
    institution: Department of Psychology, Columbia University
abstract: |
  Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development. Yet, established methods and practices for calculating and reporting heterogeneity are lacking in their treatment of uncertainty regarding estimates that describe the magnitude of heterogeneity in the phenomenon under study. There is then a need for more complete treatments of heterogeneity in psychological phenomena, which we aim to deliver in the current work.
  We provide a walkthrough of probabilistic methods for i. estimating, ii. summarising, and iii. reporting heterogeneity using bayesian hierarchical models. These methods are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
  We hope that the proposed methods help researchers to move from reporting point estimates to drawing inferences about magnitudes and differences in heterogeneity in psychological phenomena across different populations.
keywords: variation, heterogeneity, hierarchical models, mixed models, methodology
authornote: |
  \noindent \textbf{Working paper. Proceed with caution.}
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: references.bib
floatsintext: yes
numbersections: false
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: jou
header-includes:
  - |
    \usepackage{float}
    \setlength{\parskip}{4pt}
output: 
  papaja::apa6_pdf:
    number_sections: true
    keep_tex: false
    highlight: kate
  papaja::apa6_docx: 
    number_sections: true
    keep_tex: false
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(janitor)
library(scales)
library(ellipse)
library(knitr)
library(lme4)
library(patchwork)
library(latex2exp)
library(ggpp)
library(distributional)
library(ggstance)
library(tidybayes)
library(parameters)
library(brms)
library(tidyverse)

# Output options
opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  warning = FALSE,
  error = TRUE,
  message = FALSE
)

# Plotting options
theme_set(
  theme_linedraw(base_size = 12) +
    theme(
      strip.background = element_rect(fill = "gray90", color = "gray90"),
      strip.text = element_text(color = "black", hjust = 0),
      panel.grid = element_blank()
    )
)

# Model estimation options
dir.create("models", FALSE)
options(mc.cores = 4)
```

```{r data, include = FALSE}
# Load, save, & clean example data from Bolger et al. (2019)
dir.create("data", FALSE)
path <- "data/bolger-etal.zip"
if (!file.exists(path)) {
  download.file(
    "https://github.com/kzee/heterogeneityproject/archive/refs/heads/master.zip",
    destfile = path
  )
  unzip(path, exdir = dirname(path), junkpaths = TRUE)
}

# Save variables as numerical & factor (might need for more complex contrast/eff/etc codes)
dat <- read_csv("data/heterogeneity_dataset1_traitvalence.csv") %>% 
  # Cleaning as in Bolger et al 2019
  filter(
    response.keys == "up",
    scale(rt) < 3,
    !(id %in% c(250, 257, 272))
  ) %>% 
  # Some more cleaning
  mutate(
    person = fct_anon(factor(id)),
    valence = factor(valenceE, levels = c(-0.5, 0.5), labels = c("Negative", "Positive")),
    rt = logrt,
    .keep = "none"
  ) %>% 
  select(person, valence, rt) %>% 
  arrange(person)
```

```{r contrast-coding, include = FALSE}
contrasts(dat$valence) <- c(-0.5, 0.5)
contrasts(dat$valence)
```

# Introduction

When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not. Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon at hand if the treatment effect is highly variable from one individual to another.

## Current work

Our aim is to outline methods for studying and communicating heterogeneity in a manner that fully takes uncertainty into account. In the first section, we introduce the example data used throughout the manuscript and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.

We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods: We briefly review key principles of probabilistic inference and how they affect inferences about heterogeneity. We then return to the data example and examine effect heterogeneity from different angles using probabilistic inference.

We further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.

We conclude by discussing additional benefits of probabilistic inference in psychology in general, and for multilevel models specifically. We emphasize that these benefits come with little or no cost to applied practitioners. We also discuss our findings regarding causal effect heterogeneity in light of this new perspective.

# Review of expected heterogeneity in hierarchical models

To begin our exposition, we reproduce the analyses presented in @bolgerCausalProcessesPsychology2019. In their study, which replicated findings first presented in @scholerInflatingDeflatingSelf2014, 62 participants saw twenty positively and twenty negatively valenced words, and judged whether each word was self-descriptive or not. Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that response times to positively valenced words would be shorter than to negatively valenced words [@scholerInflatingDeflatingSelf2014].

```{r}
#| label: tab-dat1
dat %>% 
  head() %>% 
  apa_table(
    span_text_columns = FALSE, 
    caption = "First six rows of example dataset 1.", 
    note = "rt is log-transformed."
  )
```

In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data (https://github.com/kzee/heterogeneityproject). We first cleaned the data as in @bolgerCausalProcessesPsychology2019, leaving `r nrow(dat)` trials from 59 participants that were endorsed as self-relevant (see Table \@ref(tab:tab-dat1)). Then, we estimated the same statistical model with these data:

\begin{equation}
\text{RT}_{ij} \sim \operatorname{Normal}(\eta_{ij}, \sigma^2).
\end{equation}

In words, we model the (log) reaction time of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*, which can differ between trials $i$ and individuals $j$) and standard deviation $\sigma$ (*sigma*). The lack of subscripts on $\sigma$ indicates that we assume it to be invariant across trials and subjects, but note that this assumption is straightforward to relax. Then, we specify a model of the mean $\eta_{ij}$ such that the regression coefficients capture our substantive questions:

\begin{equation}
\eta_{ij} = \beta_0 + \gamma_{0j} + (\beta_1 + \gamma_{1j})\text{V}_{ij}.
\end{equation}

This equation includes two kinds of parameters: $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V), do not have subscripts. That means that they are not further modelled, and therefore sometimes called "fixed". We contrast coded valence such that negative stimuli are coded as -0.5, and positive stimuli as 0.5. This coding results in an intercept that corresponds to the average reaction time across negative and positive stimuli, and a slope coefficient that reflects the magnitude of the difference in log(rt) between negative and positive words (see Appendix). The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$ are subject-specific deviations from the average intercept and slope, respectively. That is, $\beta_0 + \gamma_{01}$ is the average reaction time for subject $j=1$. The $\gamma$ coefficients have subscripts, which in this context indicates that they themselves are further modelled, and therefore sometimes called "random" [@gelmanDataAnalysisUsing2007]. As is standard in multilevel modelling, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:

\begin{equation}
\begin{bmatrix} 
\gamma_{0} \\ \gamma_{1}
\end{bmatrix} \sim 
\operatorname{MVN}(
\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
\begin{bmatrix} 
\tau_{\gamma_{0}} &\rho_{\beta_0\beta_1} \\ 
\rho_{\beta_0\beta_1} &\tau_{\gamma_{1}} 
\end{bmatrix}
).
\end{equation}

In these equations, we assume that the subject-specific deviations are have means of zero, standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*). Note that the $\gamma$s are zero-centered because the means are added to them in the above linear mixed-effects equation for $\eta_{ij}$. What these equations mean substantively is that the extent to which the effect of valence varies around the distribution's mean $\beta_1$ is estimated by the standard deviation $\tau_{\gamma_1}$, and $\rho$ indicates the extent to which individuals who have greater slopes also have greater or smaller average reaction times. With data described in Table \@ref(tab:tab-dat1) loaded to R, we can estimate this model using standard maximum likelihood methods as implemented in the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2022] (see the Appendix for code details.)

```{r fit1-lmer}
fit1_lmer <- lmer(
  rt ~ 1 + valence + (1 + valence | person),
  data = dat
)
```

```{r}
# Create a table of coefficients
tab1_lmer <- model_parameters(fit1_lmer, effects = "all") %>% 
  tibble() %>% 
  select(Parameter, Coefficient, starts_with("CI_"))

# Bootstrap CIs for variance parameters
tab1_lmer[3:6,3:4] <- confint(fit1_lmer, method = "boot", nsim = 100)[c(1,3,2,4),1:2]
```

```{r fit1-brms}
fit1 <- brm(
  rt ~ 1 + valence + (1 + valence | person),
  data = dat,
  file = "models/brm-fit-1"
)
```

```{r}
#| label: tbl-fit-1-brms
#| cache: false

tab1_brms <- parameters(fit1, centrality = "mean", effects = "all") %>% 
  tibble() %>% 
  select(Parameter, Mean, starts_with("CI_"))
```

```{r}
#| label: tbl-fit-1
#| cache: false

m1_parnames <- c(
  "$\\beta_0$", "$\\beta_1$", 
  "$\\tau_0$", "$\\tau_1$", 
  "$\\rho$", "$\\sigma$"
)

bind_cols(
  tab1_lmer %>% 
    mutate(
      across(where(is.numeric), ~number(., .01)),
      MLE = str_glue("{Coefficient} [{CI_low}, {CI_high}]"),
    ) %>% 
    select(Parameter, MLE),
  tab1_brms %>% 
    mutate(
      across(where(is.numeric), ~number(., .01)),
      Bayes = str_glue("{Mean} [{CI_low}, {CI_high}]"),
    ) %>% 
    select(Bayes)
) %>% 
  mutate(
    Parameter = m1_parnames
  ) %>% 
  apa_table(
    span_text_columns = FALSE,
    caption = "Parameter estimates from Model 1.",
    note = "MLE: Maximum Likelihood Estimate. Numbers in brackets are 95\\% Confidence (MLE) or Credibility (Bayes) Intervals (CIs). CIs for MLE (co)variance parameters are calculated with 100 bootstrap iterations.",
    escape = FALSE
  )
```

We show a traditional summary of the estimated parameters from this model in the MLE column of Table \@ref(tab:tbl-fit-1). The expected effect of positive valence on RT for the average person is approximately `r number(tab1_lmer[2,2,T], .01)` log seconds. Moreover, the expected standard deviation describing this effect's degree of variability in the population is approximately `r number(tab1_lmer[4,2,T], .01)`. These point estimates together define the expected normal distribution of valence effects in the population (see Figure \@ref(fig:hdist-1)).

```{r}
#| label: hdist-1
#| fig.height: 5
#| fig.cap: \textbf{Top}. Expected heterogeneity distribution of valence effects from Example 1. The interval displays the 90\% heterogeneity interval ($HI_{90}$), which captures the central 90\% of values in the distribution. \textbf{Bottom}. Model-estimated person-specific valence effects ($\gamma_1$; dark points) and corresponding coefficients from linear regression models estimated separately for each person (light points).

fit1_coef_lm <- tibble(fit1$data) %>% 
  mutate(person = factor(as.numeric(person))) %>% 
  group_by(person) %>% 
  group_modify(~as_tibble(coef(lm(rt ~ valence, data = .))[2])) %>% 
  ungroup() %>% 
  select(person, Coefficient = value) %>% 
  mutate(model = "lm")
fit1_coef_lmer <- tibble(coef(fit1_lmer)$person)[,2] %>% 
  rownames_to_column("person") %>% 
  select(person, Coefficient = valence1) %>% 
  mutate(model = "lmer")
fit1_coef_brms <- as_tibble(coef(fit1)$person[,,"valence1"]) %>% 
  rownames_to_column("person") %>% 
  select(person, Coefficient = Estimate) %>% 
  mutate(model = "brms")

coefs <- bind_rows(
  fit1_coef_lm,
  fit1_coef_lmer,
  fit1_coef_brms
) %>% 
  pivot_wider(names_from = model, values_from = Coefficient) %>% 
  arrange(lm) %>% 
  mutate(i = 1:n())

(ggplot() +
    coord_cartesian(xlim = c(-0.7, 0.4)) +
    stat_halfeye(
      aes(
        xdist = dist_normal(tab1_brms$Mean[2], tab1_brms$Mean[4]),
      ), 
      .width = c(.90), 
      p_limits = c(.0001, .9999), 
      slab_fill = "grey70",
      slab_color = "grey10",
      slab_linewidth = .4,
      size = 2, justification = .01
    ) +
    stat_halfeye(
      aes(
        xdist = dist_normal(
          mean(fit1_coef_lm$Coefficient), 
          sd(fit1_coef_lm$Coefficient)
        )
      ),
      p_limits = c(.0001, .9999), 
      point_interval = NULL,
      slab_fill = NA,
      slab_color = "black",
      slab_linewidth = .4
    ) +
    scale_y_continuous(
      expand = expansion(c(0.0025, 0.05))
    ) +
    labs(
      y = "Probability",
      x = "Valence effect"
    ) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      plot.margin = unit(c(8,8,0,8), "pt")
    )) /
  (coefs %>% 
     ggplot(aes(brms, i)) +
     coord_cartesian(xlim = c(-0.7, 0.4)) +
     scale_y_continuous(
       "Person",
       expand = expansion(0.025)
     ) +
     scale_x_continuous(
       "Valence effect"
     ) +
     geom_segment(
       aes(x = lm, xend = brms, y = i, yend = i),
       linewidth = .1
     ) +
     geom_point(aes(lm), shape = 21, fill = "white") +
     geom_point(shape = 21, fill = "grey70") +
     theme(
       plot.margin = unit(c(0,8,8,8), "pt"),
       axis.text.y = element_blank(),
       axis.ticks.y = element_blank(),
     )
  ) +
  plot_layout(heights = c(3.5, 6.5))
```

```{r hi90, include = FALSE}
hi90_num <- tab1_lmer[2,2,T] + qnorm(c(0.05, .95))*tab1_lmer[4,2,T]
hi90 <- paste(
  number(hi90_num, .01), 
  collapse = ", "
)
hi90_low <- paste(
  number(tab1_lmer[2,3,T] + qnorm(c(0.05, .95))*tab1_lmer[4,3,T], .01), 
  collapse = ", "
)
```

We can use the numbers in Table \@ref(tab:tbl-fit-1) to communicate the degree of heterogeneity in the effect of valence in the population: The standard deviation $\tau_1=$ `r number(tab1_lmer[4,2,T], .01)`. However, that quantity alone is unable to properly characterize the heterogeneity in actionable terms. We were therefore motivated to find alternative concise summaries of the degree of heterogeneity of slopes in the population. To begin with, following @bolgerCausalProcessesPsychology2019, we used the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes a range within which a certain percentage of the population's slopes are expected to fall (interval in bottom of Figure \@ref(fig:hdist-1)). Later on, we describe alternative summaries of heterogeneity that address different questions about variability, and afford different ways of communicating and visualizing heterogeneity.

By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$), but because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties in estimated quantities. This heterogeneity interval can be calculated by first specifying the desired probability limits. For a 90% interval, we use the .05 and .95 percentiles. Then, we pass those percentiles to the normal quantile function ($q$) using the estimated mean and standard deviation to obtain the interval: $HI_{90} = q([.05, .95], \hat\beta_1, \hat\tau_1)$. Therefore (see Table \@ref(tab:tbl-fit-1)), $HI_{90} =$ $q([.05, .95],$ `r number(tab1_lmer[2,2,T], .01)`, `r number(tab1_lmer[4,2,T], .01)`) = [`r hi90`]. In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates, indicated with hats in the equation: We expect 90% of the effects in the population to fall within this interval.

Critically, while valid, this calculation ignores the uncertainty inherent in the estimated parameters. That is, the uncertainty with which we have estimated the population parameters, indicated by e.g. the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward in calculating $HI_{90}$: We have only presented the expected heterogeneity interval, but not the range of credible values that it might take. For instance, if we re-calculate $HI_{90}$ based on the lowest CI bounds in Table \@ref(tab:tbl-fit-1), the result is a heterogeneity interval that indicates 90% of the population's slopes falling in the [`r hi90_low`] interval. In contrast to the first interval, this interval implies much less heterogeneity and that 90% of the population's slopes are expected to fall below zero.

We have now arrived at the crux of the current work: How should we estimate and communicate causal effect heterogeneity such that the fundamental uncertainty in the estimated quantities (the mean and standard deviation of the distribution of slopes in the population) is retained?

# Interlude on probabilistic estimation in practice {-}

<!-- todo introduce citations and topics as we go; delete this section -->

We propose that probabilistic methods are uniquely applicable to the problem presented above. In what follows, we describe practical calculations and visualizations that make use of bayesian methods in estimating and communicating causal effect heterogeneity. Frequentist methods, such as maximum likelihood above, aim at estimating optimum values of estimated quantities, typically called point estimates. In contrast, probabilistic methods aim at describing distributions of plausible values that could underlie hypothetical data generating processes. In this section, we briefly describe bayesian estimation such that we can meaningfully use it in the remainder of our proposition. For more complete treatments, we refer readers to excellent existing literature on the topic [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].

Lorem ipsum bayesian propagandum.

# Probabilistic assessment of heterogeneity

We now return to examining the heterogeneity in valence effects but with using bayesian methods. First, we use the R package brms to estimate the same model as above [@burknerAdvancedBayesianMultilevel2018; @burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2022]. The goal of bayesian inference is the multivariate posterior probability distribution of the model's parameters. However, because closed-form solutions are not available for posterior distributions of many important types of statistical models, in practice bayesian methods rely on drawing many random samples from the posterior distribution [@gelmanBayesianDataAnalysis2013]. Therefore, "estimate" here means drawing a large number of---here, 4,000---random samples from the posterior distribution. These samples can then be used to calculate and summarise the posterior distribution of any desired quantity. In practice, one obtains (e.g.) 4,000 samples from the posterior distribution, which can then be summarized using familiar data processing techniques. Below, we show the posterior mean and 2.5 and 97.5 percentiles of the parameters' distributions in the rightmost column of Table \@ref(tab:tbl-fit-1). Note that because we used the default noninformative prior distributions, summaries of the resulting posterior distributions are numerically almost identical to the maximum likelihood estimates. To highlight the issue at hand, Table \@ref(tab:tbl-samples-1) shows the first six rows of samples from their joint distribution.

```{r}
#| label: tbl-samples-1
#| cache: false

post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_"), regex = TRUE) %>% 
  tibble() %>% 
  mutate(b = b_valence1, sd = sd_person__valence1, .keep = "none") %>% 
  mutate(hi_low = b + qnorm(.05)*sd, hi_high = b + qnorm(.95)*sd) %>% 
  mutate(Sample = 1:n(), .before = 1)
post1_sum <- post1 %>% 
  mean_qi(hi_low, hi_high)
post1 %>% 
  setNames(c("Sample", m1_parnames)) %>% 
  head() %>% 
  apa_table(
    span_text_columns = FALSE,
    caption = "First six rows of samples from the posterior distribution of Model 1.",
    escape = FALSE,
    placement = "h"
  )
```

To return to the $HI_{90}$, we could reproduce the above calculations for the expected interval using the posterior modes. Instead, to carry uncertainty forward from parameters to (e.g.) $HI_{90}$, we calculate samples of $HI_{90}$ from samples of the model's parameters (Table \@ref(tab:tbl-fit-1); Figure \@ref(fig:hdist-2)A). Specifically, because an interval is defined by its endpoints, the resulting posterior distribution is a two-dimensional distribution of lower and upper limits of the heterogeneity interval (Figure \@ref(fig:hdist-2)B).

<!-- todo panel b: explain / give insight into why panel b is a negative correlation -->
<!-- todo panel d requires a lot more introduction--move into its own figure & section -->

```{r}
#| label: hdist-2
#| fig.height: 6.4
#| fig.cap: \textbf{A}. Scatterplot of 4,000 random samples from the joint posterior distribution of the mean $\beta_1$ and standard deviation $\tau_1$ of the gaussian population distribution of valence effects. \textbf{B}. Scatterplot of 4,000 samples from the joint posterior distribution of the $HI_{90}$ lower ($\beta_1 + \phi_{0.05}\tau_1$; x-axis) and upper ($\beta_1 + \phi_{0.95}\tau_1$; y-axis) limits, calculated from samples in panel A. \textbf{C}. A sample of 200 $HI_{90}$s arranged in increasing magnitude of the standard deviation. Red symbols in each panel indicate the posterior mean quantities; e.g. the expected heterogeneity interval in C. \textbf{D}. Histogram of 4,000 samples from the posterior distribution of the proportion of negative effects in the population distribution of effects defined by $\beta_1$ and $\tau_1$. The red point and interval in the bottom indicate the proportion's posterior mean and 95\%CI.

post1 <- post1 %>%
  arrange(hi_low) %>% 
  mutate(i = 1:n())

hcol <- "#cc4778"
p_scatter <- post1 %>% 
  ggplot(aes(b, sd)) +
  labs(x = bquote(beta[1]), y = bquote(tau[1])) +
  labs(
    x = "Mean of slope distribution",
    y = "SD of slope distribution"
  ) +
  geom_point(shape = 1, alpha = .25, size = .75) +
  stat_summary_xy(col = hcol, .fun.x = mean, .fun.y = mean, size = 2) +
  theme(
    aspect.ratio = 1
  )

p_scatter_hi <- post1 %>% 
  ggplot(aes(hi_low, hi_high)) +
  labs(
    x = TeX("$HI_{90}$ lower limit"), 
    y = TeX("$HI_{90}$ upper limit")
  ) +
  geom_point(shape = 1, alpha = .25, size = .75) +
  stat_summary_xy(col = hcol, .fun.x = mean, .fun.y = mean, size = 2) +
  theme(
    aspect.ratio = 1
  )

post1_snip <- post1 %>% 
  filter(i %in% as.integer(seq(1, 4000, length.out = 200)))
p_intervals <- post1_snip %>% 
  ggplot(aes(y = as.numeric(i), yend = as.numeric(i))) +
  scale_fill_brewer(
    "HI limit",
    palette = "Set1",
    aesthetics = c("fill", "color"),
    breaks = c("hi_low", "hi_high"),
    labels = c("Low", "High")
  ) +
  scale_y_continuous(
    expand = expansion(0.01)
  ) +
  labs(
    x = "Slope magnitude",
    y = "Posterior sample"
  ) +
  geom_segment(
    aes(x = hi_low, xend = hi_high),
    linewidth = .1, col = "gray50"
  ) +
  geom_point(
    data = post1_snip %>% 
      pivot_longer(c(hi_low, hi_high)),
    aes(x = value), col = "gray50",
    size = .25, alpha = .5
  ) +
  geom_segment(
    data = post1 %>% 
      summarise(hi_low = mean(hi_low), hi_high = mean(hi_high)) %>% 
      mutate(i = 2000),
    aes(x = hi_low, xend = hi_high),
    col = hcol
  ) +
  geom_point(
    data = post1 %>% 
      summarise(hi_low = mean(hi_low), hi_high = mean(hi_high)) %>% 
      pivot_longer(c(hi_low, hi_high)) %>% 
      mutate(i = 2000),
    aes(x = value),
    col = hcol
  ) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p_slopezero <- post1 %>% 
  mutate(
    p_slopes_under_zero = pnorm(0, b, sd)
  ) %>% 
  ggplot(aes(p_slopes_under_zero)) +
  scale_y_continuous(
    "Count",
    expand = expansion(c(0.0025, .1))
  ) +
  scale_x_continuous(
    "Proportion of negative slopes"
  ) +
  stat_histinterval(
    breaks = 30,
    outline_bars = TRUE,
    slab_color = "gray20",
    slab_fill = "gray80",
    slab_linewidth = .2,
    justification = .01,
    point_color = hcol, interval_color = hcol,
    .width = .90
  ) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

((p_scatter |
    p_scatter_hi) /
    (p_intervals |
       p_slopezero)) +
  plot_annotation(tag_levels = "A")
```

```{r}
p1s <- post1 %>% 
  mean_qi(hi_low, hi_high) %>% 
  mutate(
    across(where(is.numeric), ~number(., .01)),
    l = str_glue("[{hi_low.lower}, {hi_low.upper}]"),
    u = str_glue("[{hi_high.lower}, {hi_high.upper}]")
  )
```

Summarising a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds. The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u`. Communicating these two credibility intervals is not only cumbersome, but also ignores the correlation between the heterogeneity interval endpoints' posterior distributions (Figure \@ref(fig:hdist-2)B)). We therefore turn to alternative summaries of the population distribution of slopes.

## Alternatives to the heterogeneity interval

To reiterate, the model's estimated distribution of slopes is a normal distribution with mean $\beta_1$ and standard deviation $\tau_1$. The 90% heterogeneity interval reports the $5^{th}$ and $95^{th}$ quantiles of the distribution, of whose posterior distributions are correlated. But we are not limited to describing the plausible ranges of slopes in the population. In fact, for many applications, quantities such as proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.

```{r}
tmp <- p_slopezero$data %>% 
  mean_qi(p_slopes_under_zero, .width = .95) %>% 
  mutate(
    across(where(is.numeric), ~number(., .01)),
    p = str_glue("{p_slopes_under_zero}, [{.lower}, {.upper}]")
  )

tmp2 <- p_slopezero$data %>% 
  mean_qi(1-p_slopes_under_zero, .width = .95) %>% 
  mutate(
    across(where(is.numeric), ~percent(., .1)),
    p = str_glue("{`1 - p_slopes_under_zero`}, [{.lower}, {.upper}]")
  )
```

To make this concrete, we ask "What proportion of individuals in the population have a negative effect of valence?" To answer, we calculate the probability density below zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$, and obtain samples of this density. Figure \@ref(fig:hdist-2)D then shows these samples' histogram, and the posterior mean and 95% Credibility Interval (`r tmp$p`): We expect about `r percent(as.numeric(tmp$p_slopes_under_zero))` to be negative, but with 95% credibility, this value could be as low as `r percent(as.numeric(tmp$.lower))` or as high as `r percent(as.numeric(tmp$.upper))`. Alternatively [@bolgerCausalProcessesPsychology2019, p.605-606], the model predicts that `r tmp2$p` of individuals in the population would show reversals of the valence effect.

```{r}
tmp <- post1 %>% 
  summarise(
    post_prob_90p_negative = percent(
      sum(hi_high < 0) / n(), 
      .01
    )
  ) %>% 
  pull(1)
```

Additionally, we might ask, for example, "What is the posterior probability that (e.g.) 90% of slopes in the population are negative?" To answer, we calculate, for each posterior sample of $\beta_1$ and $\tau_1$, whether the resulting $HI_{90}$'s upper limit is below zero. Visually, this calculation represents drawing a vertical line at 0 on the y-axis of Figure \@ref(fig:hdist-2)B and calculating the proportion of samples below zero over all the samples. The result, `r tmp`, indicates that we expect `r tmp` of slopes in the population to be negative. Note that this quantity cannot be calculated from the point estimates alone.

## Does heterogeneity matter?

```{r}
tmp <- post1 %>% 
  mutate(ratio = abs(sd / b)) %>% 
  mean_qi(ratio) %>% 
  mutate(
    across(where(is.numeric), ~number(., .1)),
    r = str_glue("{ratio} [{.lower}, {.upper}]")
  )
```

The above two quantities summarized the heterogeneity of valence effects in the population in absolute terms. To ask whether heterogeneity is substantial enough to matter, we might directly calculate the posterior mean and (e.g.) 95%CI of $\tau_1$ to decide whether the spread of effects in the population is large enough to be of practical significance. Alternatively, we might focus on the relative size of heterogeneity when compared to the average effect. [@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the standard deviation is 0.25 or greater in magnitude than the average effect, heterogeneity is noteworthy. With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r tmp$r` (posterior mean and 95% CI), suggesting that the degree of heterogeneity in the effect of valence is noteworthy. (We reiterate the point made by @bolgerCausalProcessesPsychology2019 that the cutoff value of 0.25 is arbitrary and researchers should choose the cutoff based on their substantive goals.)

### Smallest effect size of interest

todo: Examine proportions of effects in context of hypothesized SESOIs.

# Example 3: Comparing heterogeneity between groups

todo: Model separate random effects distributions for e.g. labs in a large replication project, then compare heterogeneity between labs. (In the current running example men vs women would be obvious & easy--maybe create a fake gender variable?)

# Example 4: Multivariate heterogeneity

todo: Extend Study 4 from Bolger et al., studying correlations of random effects between timepoints among the same individuals.

# Discussion

# Competing interests

The author(s) declare no competing interests.

# Author contributions

Conceptualization: MV & NB    
Methodology: MV   
Software: MV    
Validation: MV    
Formal Analysis: MV   
Writing -- Original Draft: MV & NB    
Writing -- Review & Editing: MV & NB    
Visualization: MV & NB   
Project Administration: MV

# References {-}

::: {#refs custom-style="Bibliography"}
:::

\clearpage
\onecolumn

# Appendix: Code listings

\setcounter{page}{1}
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{A-\arabic{table}}
\renewcommand{\thefigure}{A-\arabic{figure}}

## Contrast coding

By default, R uses dummy coding (called "treatment contrasts" in R) for categorical predictors, meaning that one category is used as a baseline, and then contrasts between the baseline category and the other categories are estimated. It is often useful to use contrast coding (called "sum to zero contrasts" in R) instead, where the levels of a categorical factor are assigned codes that sum to zero: In this manner, the intercept of the model is at the midpoint between the category levels, and consequently the other parameters are easier to interpret. To use contrast codes for Valence, in the example data, the R code is `contrasts(dat$valence) <- "contr.sum"`. However, this has the slight drawback that the levels are assigned numerical codes of -1 and 1, and the resulting parameter estimate indicates a contrast to the "average" valence, and must be multiplied by two to give the difference between the conditions. We therefore used the more readily interpretable values of -0.5 and 0.5:

```{r contrast-coding, echo = TRUE, eval = TRUE}
```

In addition to making the resulting parameter estimate directly indicate the difference between conditions, this keeps the `valence` variable in R as a factor, which ensures that downstream functions using that variable will appropriately treat it as a categorical variable with interpretable labels (e.g. for figures.) For useful references to predictor coding in R, see [@rabeHyprPackageHypothesisdriven2020; @schadHowCapitalizePriori2020].

## Estimating a multilevel model using maximum likelihood

```{r fit1-lmer, echo=TRUE, eval=FALSE}
```

## Calculating a heterogeneity interval

```{r hi90-appendix, echo = -1}
options(digits = 2)
mean <- -0.16 # Mean of slope distribution point estimate from Table 1
sd <- 0.12 # Standard deviation of slope distribution point estimate from Table 1
quantiles <- c(0.05, 0.95) # Desired quantiles (these define 90%)
mean + qnorm(quantiles)*sd
```

## Estimating a multilevel model using probabilistic inference

```{r fit1-brms, echo=TRUE, eval=FALSE}
```

todo: think about if this appendix is needed at all

### 2023-06-27 ms.Rmd
Clarification
Update draft

@@
     corresponding: yes
-    address: Department of Social Psychology, Tilburg University
+    address: Tilburg School of Social and Behavioral Sciences, Tilburg University
     email: m.j.vuorre@tilburguniversity.edu
+  - name: Matthew Kay
+    affiliation: 2
   - name: Niall Bolger
-    affiliation: 2
-    address: Department of Psychology, Columbia University
-    email: nb2229@columbia.edu
+    affiliation: 3
 affiliation:
@@
     institution: Department of Social Psychology, Tilburg University
-  - id: 1
+  - id: 2
+    institution: Computer Science and Communication Studies, Northwestern University
+  - id: 3
     institution: Department of Psychology, Columbia University
@@
   Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development.
-Yet, established methods and practices for calculating and reporting heterogeneity are lacking in their treatment of uncertainty regarding estimates that describe the magnitude of heterogeneity in the phenomenon under study.
-There is then a need for more complete treatments of heterogeneity in psychological phenomena, which we aim to deliver in the current work.
+Yet, established methods and practices for calculating and describing heterogeneity are lacking in how they treat uncertainty regarding heterogeneity in the phenomenon under study.
+In the current work, we aim to deliver methods and guidelines for more complete treatments of heterogeneity in psychological phenomena.
   We provide a walkthrough of probabilistic methods for i.
@@
 summarising, and iii.
-reporting heterogeneity using bayesian hierarchical models.
-These methods are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
+describing heterogeneity.
+The methods we propose---bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
   We hope that the proposed methods help researchers to move from reporting point estimates to drawing inferences about magnitudes and differences in heterogeneity in psychological phenomena across different populations.
@@
 library(knitr)
+library(latex2exp)
 library(lme4)
+library(ggpp)
 library(patchwork)
 library(latex2exp)
-library(ggpp)
-library(distributional)
+library(ggdist)
 library(ggstance)
 library(tidybayes)
+library(distributional)
 library(parameters)
@@
 theme_set(
-  theme_linedraw(base_size = 12) +
+  theme_classic(base_size = 9) +
     theme(
-      strip.background = element_rect(fill = "gray90", color = "gray90"),
       strip.text = element_text(color = "black", hjust = 0),
-      panel.grid = element_blank()
+      panel.grid = element_blank(),
+      line = element_line(linewidth = .25)
     )
@@
 dir.create("models", FALSE)
-options(mc.cores = 4)
+options(
+  mc.cores = 4,
+  brms.backend = ifelse(require(cmdstanr), "cmdstanr", "rstan")
+)
+
 ```
@@
     person = fct_anon(factor(id)),
+    trial,
     valence = factor(valenceE, levels = c(-0.5, 0.5), labels = c("Negative", "Positive")),
-    rt = logrt,
-    .keep = "none"
+    rt = logrt
   ) %>% 
-  select(person, valence, rt) %>% 
-  arrange(person)
+  select(person, trial, valence, rt) %>% 
+  arrange(person, trial)
 ```
@@
 When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not.
-Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon at hand if the treatment effect is highly variable from one individual to another.
+Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon if the treatment effect is highly variable from one individual to another.
+The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].
+
+One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still tend to analyze data with models that cannot address between-person variability, such as ANOVA [@bolgerCausalProcessesPsychology2019].
+Although more appropriate mixed-effects (or multilevel, hierarchical) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to estimate and describe heterogeneity to a sufficiently informative degree.
+In the current work, we propose several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
 
@@
 Our aim is to outline methods for studying and communicating heterogeneity in a manner that fully takes uncertainty into account.
-In the first section, we introduce the example data used throughout the manuscript and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.
+We first introduce an example dataset from social psychology and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.
+We then outline useful summaries of heterogeneity that focus on informative numerical and graphical descriptions, and note how they are limited by established methods' focus on point estimates of model parameters.
 
-We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods: We briefly review key principles of probabilistic inference and how they affect inferences about heterogeneity.
-We then return to the data example and examine effect heterogeneity from different angles using probabilistic inference.
+We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods.
+By estimating probability distributions over model parameters, these bayesian methods are uniquely suitable for examining heterogeneity and researchers' uncertainty about it.
+Moreover, advances in computational procedures and statistical software make these methods available to researchers with a minimal learning curve beyond established maximum likelihood based methods.
 
-We further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.
+In the following sections, we detail calculations and graphics that make use of bayesian methods in communicating descriptions of heterogeneity such that uncertainty in the model parameters and, subsequently, heterogeneity, is taken into account.
+We then further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.
 
@@
 
+## Model 1
+
 ```{r}
@@
     span_text_columns = FALSE, 
+    digits = c(0,0,0,2),
     caption = "First six rows of example dataset 1.", 
@@
 In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data (https://github.com/kzee/heterogeneityproject).
-We first cleaned the data as in @bolgerCausalProcessesPsychology2019, leaving `r nrow(dat)` trials from 59 participants that were endorsed as self-relevant (see Table \@ref(tab:tab-dat1)).
-Then, we estimated the same statistical model with these data:
+We first cleaned the data as in @bolgerCausalProcessesPsychology2019, leaving `r number(nrow(dat), big.mark = ",")` trials from 59 participants that were endorsed as self-relevant.
+(We present all our code in the online analysis supplement.) We show a sample of these data in Table \@ref(tab:tab-dat1).
+Then, we estimated the same statistical model with these data (Model 1; equations 1.1-1.3).
+First, we model the (log) reaction time of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*, which can differ between trials $i$ and individuals $j$) and standard deviation $\sigma$ (*sigma*):
 
 \begin{equation}
+\tag{1.1}
 \text{RT}_{ij} \sim \operatorname{Normal}(\eta_{ij}, \sigma^2).
@@
 
-In words, we model the (log) reaction time of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*, which can differ between trials $i$ and individuals $j$) and standard deviation $\sigma$ (*sigma*).
-The lack of subscripts on $\sigma$ indicates that we assume it to be invariant across trials and subjects, but note that this assumption is straightforward to relax.
+The lack of subscripts on $\sigma$ indicates that we assume it to be invariant across trials and subjects.
 Then, we specify a model of the mean $\eta_{ij}$ such that the regression coefficients capture our substantive questions:
@@
 \begin{equation}
+\tag{1.2}
 \eta_{ij} = \beta_0 + \gamma_{0j} + (\beta_1 + \gamma_{1j})\text{V}_{ij}.
@@
 That means that they are not further modelled, and therefore sometimes called "fixed".
-We contrast coded valence such that negative stimuli are coded as -0.5, and positive stimuli as 0.5.
-This coding results in an intercept that corresponds to the average reaction time across negative and positive stimuli, and a slope coefficient that reflects the magnitude of the difference in log(rt) between negative and positive words (see Appendix).
+We contrast coded valence such that negative words are coded as -0.5, and positive words as 0.5.
+This coding results in an intercept that corresponds to the average reaction time across negative and positive words, and a slope coefficient that reflects the difference in log(rt) between negative and positive words.
 The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$ are subject-specific deviations from the average intercept and slope, respectively.
 That is, $\beta_0 + \gamma_{01}$ is the average reaction time for subject $j=1$.
-The $\gamma$ coefficients have subscripts, which in this context indicates that they themselves are further modelled, and therefore sometimes called "random" [@gelmanDataAnalysisUsing2007].
+The $\gamma$ coefficients have subscripts, which in this context indicates that they are further modelled, and therefore sometimes called "random" [@gelmanDataAnalysisUsing2007].
 As is standard in multilevel modelling, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:
@@
 \begin{equation}
+\tag{1.3}
 \begin{bmatrix} 
@@
 \begin{bmatrix} 
-\tau_{\gamma_{0}} &\rho_{\beta_0\beta_1} \\ 
-\rho_{\beta_0\beta_1} &\tau_{\gamma_{1}} 
+\tau_{\gamma_{0}} &\rho_{\gamma_0\gamma_1} \\ 
+\rho_{\gamma_0\gamma_1} &\tau_{\gamma_{1}} 
 \end{bmatrix}
@@
 
-In these equations, we assume that the subject-specific deviations are have means of zero, standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
+In these equations, we assume that the subject-specific deviations have means of zero, standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
 Note that the $\gamma$s are zero-centered because the means are added to them in the above linear mixed-effects equation for $\eta_{ij}$.
-What these equations mean substantively is that the extent to which the effect of valence varies around the distribution's mean $\beta_1$ is estimated by the standard deviation $\tau_{\gamma_1}$, and $\rho$ indicates the extent to which individuals who have greater slopes also have greater or smaller average reaction times.
-With data described in Table \@ref(tab:tab-dat1) loaded to R, we can estimate this model using standard maximum likelihood methods as implemented in the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2022] (see the Appendix for code details.)
+What these equations mean substantively is that the extent to which the effect of valence varies around the average effect ($\beta_1$) is estimated by the standard deviation $\tau_{\gamma_1}$.
+Moreover, $\rho$ indicates the extent to which individuals' average reaction times correlate with how much their reaction time is affected by valence.
+
+With data described in Table \@ref(tab:tab-dat1) loaded to R, we can estimate this model using standard maximum likelihood methods as implemented in the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].
 
@@
   data = dat,
-  file = "models/brm-fit-1"
+  file = "models/brm-fit-1",
+  control = list(adapt_delta = .95)
 )
@@
 
-bind_cols(
 tab1_lmer %>% 
@@
   ) %>% 
-    select(Parameter, MLE),
-  tab1_brms %>% 
-    mutate(
-      across(where(is.numeric), ~number(., .01)),
-      Bayes = str_glue("{Mean} [{CI_low}, {CI_high}]"),
-    ) %>% 
-    select(Bayes)
-) %>% 
+  select(Parameter, `Estimate [95\\%CI]` = MLE) %>% 
   mutate(
@@
     caption = "Parameter estimates from Model 1.",
-    note = "MLE: Maximum Likelihood Estimate.
-Numbers in brackets are 95\\% Confidence (MLE) or Credibility (Bayes) Intervals (CIs).
-CIs for MLE (co)variance parameters are calculated with 100 bootstrap iterations.",
     escape = FALSE
   )
+mu1n <- tab1_lmer[2,2,T]
+mu1 <- number(mu1n, .01)
+sd1n <- tab1_lmer[4,2,T]
+sd1 <- number(sd1n, .01)
 ```
 
-We show a traditional summary of the estimated parameters from this model in the MLE column of Table \@ref(tab:tbl-fit-1).
-The expected effect of positive valence on RT for the average person is approximately `r number(tab1_lmer[2,2,T], .01)` log seconds.
-Moreover, the expected standard deviation describing this effect's degree of variability in the population is approximately `r number(tab1_lmer[4,2,T], .01)`.
-These point estimates together define the expected normal distribution of valence effects in the population (see Figure \@ref(fig:hdist-1)).
+We show a traditional summary of the estimated parameters from this model in Table \@ref(tab:tbl-fit-1).
+The expected effect of positive valence on RT for the average person is approximately `r mu1` log seconds, and the expected standard deviation describing this effect's degree of variability in the population is approximately `r sd1`.
 
-```{r}
-#| label: hdist-1
-#| fig.height: 5
-#| fig.cap: \textbf{Top}.
-Expected heterogeneity distribution of valence effects from Example 1.
-The interval displays the 90\% heterogeneity interval ($HI_{90}$), which captures the central 90\% of values in the distribution.
-\textbf{Bottom}.
-Model-estimated person-specific valence effects ($\gamma_1$; dark points) and corresponding coefficients from linear regression models estimated separately for each person (light points).
+## Heterogeneity distribution
 
-fit1_coef_lm <- tibble(fit1$data) %>% 
-  mutate(person = factor(as.numeric(person))) %>% 
-  group_by(person) %>% 
-  group_modify(~as_tibble(coef(lm(rt ~ valence, data = .))[2])) %>% 
-  ungroup() %>% 
-  select(person, Coefficient = value) %>% 
-  mutate(model = "lm")
-fit1_coef_lmer <- tibble(coef(fit1_lmer)$person)[,2] %>% 
-  rownames_to_column("person") %>% 
-  select(person, Coefficient = valence1) %>% 
-  mutate(model = "lmer")
-fit1_coef_brms <- as_tibble(coef(fit1)$person[,,"valence1"]) %>% 
-  rownames_to_column("person") %>% 
-  select(person, Coefficient = Estimate) %>% 
-  mutate(model = "brms")
-
-coefs <- bind_rows(
-  fit1_coef_lm,
-  fit1_coef_lmer,
-  fit1_coef_brms
-) %>% 
-  pivot_wider(names_from = model, values_from = Coefficient) %>% 
-  arrange(lm) %>% 
-  mutate(i = 1:n())
-
-(ggplot() +
-    coord_cartesian(xlim = c(-0.7, 0.4)) +
-    stat_halfeye(
-      aes(
-        xdist = dist_normal(tab1_brms$Mean[2], tab1_brms$Mean[4]),
-      ), 
-      .width = c(.90), 
-      p_limits = c(.0001, .9999), 
-      slab_fill = "grey70",
-      slab_color = "grey10",
-      slab_linewidth = .4,
-      size = 2, justification = .01
-    ) +
-    stat_halfeye(
-      aes(
-        xdist = dist_normal(
-          mean(fit1_coef_lm$Coefficient), 
-          sd(fit1_coef_lm$Coefficient)
-        )
-      ),
-      p_limits = c(.0001, .9999), 
-      point_interval = NULL,
-      slab_fill = NA,
-      slab_color = "black",
-      slab_linewidth = .4
-    ) +
-    scale_y_continuous(
-      expand = expansion(c(0.0025, 0.05))
-    ) +
-    labs(
-      y = "Probability",
-      x = "Valence effect"
-    ) +
-    theme(
-      axis.text = element_blank(),
-      axis.ticks = element_blank(),
-      axis.title.x = element_blank(),
-      plot.margin = unit(c(8,8,0,8), "pt")
-    )) /
-  (coefs %>% 
-     ggplot(aes(brms, i)) +
-     coord_cartesian(xlim = c(-0.7, 0.4)) +
-     scale_y_continuous(
-       "Person",
-       expand = expansion(0.025)
-     ) +
-     scale_x_continuous(
-       "Valence effect"
-     ) +
-     geom_segment(
-       aes(x = lm, xend = brms, y = i, yend = i),
-       linewidth = .1
-     ) +
-     geom_point(aes(lm), shape = 21, fill = "white") +
-     geom_point(shape = 21, fill = "grey70") +
-     theme(
-       plot.margin = unit(c(0,8,8,8), "pt"),
-       axis.text.y = element_blank(),
-       axis.ticks.y = element_blank(),
-     )
-  ) +
-  plot_layout(heights = c(3.5, 6.5))
-```
-
-```{r hi90, include = FALSE}
-hi90_num <- tab1_lmer[2,2,T] + qnorm(c(0.05, .95))*tab1_lmer[4,2,T]
+```{r descriptors, include = FALSE}
+# Heterogeneity interval
+hi90_num <- mu1n + qnorm(c(0.05, .95)) * sd1n
 hi90 <- paste(
@@
 )
+
+# Proportions
+p_less_zero <- percent(pnorm(0, mu1n, sd1n), .1)
+p_rope <- percent(pnorm(0.1, mu1n, sd1n) - pnorm(-0.1, mu1n, sd1n), .1)
+
+# Ratio
+ratio <- number(abs(sd1n / mu1n), .01)
 ```
 
-We can use the numbers in Table \@ref(tab:tbl-fit-1) to communicate the degree of heterogeneity in the effect of valence in the population: The standard deviation $\tau_1=$ `r number(tab1_lmer[4,2,T], .01)`.
-However, that quantity alone is unable to properly characterize the heterogeneity in actionable terms.
-We were therefore motivated to find alternative concise summaries of the degree of heterogeneity of slopes in the population.
-To begin with, following @bolgerCausalProcessesPsychology2019, we used the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes a range within which a certain percentage of the population's slopes are expected to fall (interval in bottom of Figure \@ref(fig:hdist-1)).
-Later on, we describe alternative summaries of heterogeneity that address different questions about variability, and afford different ways of communicating and visualizing heterogeneity.
+```{r}
+#| label: hdist-1
+#| fig.height: 2.2
+#| fig.width: 8
+#| fig.env: "figure*"
+#| fig.cap: The heterogeneity distribution and various descriptions of the expected heterogeneity in valence effects from Model 1.
+\textbf{A}.
+The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
+Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean.
+\textbf{B}.
+The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area.
+\textbf{C}.
+Proportion of negative valence effects in the population (dark).
+\textbf{D}.
+Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).
 
-By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$), but because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties in estimated quantities.
-This heterogeneity interval can be calculated by first specifying the desired probability limits.
-For a 90% interval, we use the .05 and .95 percentiles.
-Then, we pass those percentiles to the normal quantile function ($q$) using the estimated mean and standard deviation to obtain the interval: $HI_{90} = q([.05, .95], \hat\beta_1, \hat\tau_1)$.
-Therefore (see Table \@ref(tab:tbl-fit-1)), $HI_{90} =$ $q([.05, .95],$ `r number(tab1_lmer[2,2,T], .01)`, `r number(tab1_lmer[4,2,T], .01)`) = [`r hi90`].
-In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates, indicated with hats in the equation: We expect 90% of the effects in the population to fall within this interval.
+p0 <- ggplot() +
+  aes(xdist = dist_normal(mu1n, sd1n)) +
+  scale_x_continuous(breaks = extended_breaks()) +
+  coord_cartesian(xlim = c(-0.55, 0.25)) +
+  theme(
+    axis.title.y = element_blank(),
+    axis.title.x = element_text(lineheight = rel(0.5))
+    ) +
+  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white")
+d1 <- dist_normal(mu1n, sd1n) %>% 
+  mean_qi(.width = c(.66, .9))
 
-Critically, while valid, this calculation ignores the uncertainty inherent in the estimated parameters.
+p1 <- p0 +
+  stat_slab(
+    aes(fill = after_stat(level)), 
+    .width = c(.68, .95), 
+    p_limits = c(.000001, .999999), 
+    slab_color = "grey10",
+    slab_linewidth = .4
+  ) +
+  geom_text(
+    label = TeX("$\\tau_1$"),
+    x = mu1n + .05, y = .175,
+    size = 3,
+    parse = TRUE
+  ) +
+  annotate(
+    "segment",
+    x = mu1n, xend = mu1n + sd1n, y = .1, yend = .1,
+    linewidth = .35,
+    arrow = arrow(length = unit(4, "pt"), type = "closed")
+  ) +
+  geom_text(
+    label = TeX("$\\beta_1$"),
+    x = mu1n-.05, y = .075,
+    size = 3,
+    parse = TRUE
+  ) +
+  annotate(
+    "segment",
+    x = mu1n, xend = mu1n, y = .1, yend = 0,
+    linewidth = .35,
+    arrow = arrow(length = unit(4, "pt"), type = "closed")
+  ) +
+  theme(axis.title.y = element_text(angle = 90))
+
+p2 <- p0 + 
+  stat_slab(
+    aes(fill = after_stat(level)), 
+    .width = c(.90, 1), 
+    p_limits = c(.000001, .999999), 
+    slab_color = "grey10",
+    slab_linewidth = .4
+  ) +
+  geom_text(
+    label = TeX("$HI_{90}$"),
+    x = mu1n, y = .175,
+    size = 3,
+    parse = TRUE
+  ) +
+  annotate(
+    "segment",
+    x = d1$.lower[2], xend = d1$.upper[2], y = .1, yend = .1,
+    linewidth = .35,
+    arrow = arrow(length = unit(4, "pt"), type = "closed", ends = "both")
+  )
+
+p3 <- p0 + 
+  stat_slab(
+    aes(fill = after_stat(x < 0)), 
+    p_limits = c(.000001, .999999), 
+    slab_color = "grey10",
+    slab_linewidth = .4
+  ) +
+  geom_text(
+    label = TeX('$p^d$'),
+    x = mu1n, y = .175,
+    size = 3,
+    parse = TRUE
+  )
+
+p4 <- p0 + 
+  stat_slab(
+    aes(fill = after_stat(between(x, -.1, .1))), 
+    p_limits = c(.000001, .999999), 
+    slab_color = "grey10",
+    slab_linewidth = .4
+  ) +
+  geom_text(
+    label = TeX('$p^{ROPE}$'),
+    x = 0, y = .07,
+    size = 3,
+    parse = TRUE
+  )
+
+(p1 | p2 | p3 | p4) &
+  scale_y_continuous(
+    expand = expansion(c(0.0, 0.05))
+  ) &
+  labs(
+    y = "Density",
+    x = "Valence effect"
+  ) &
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "none"
+  ) &
+  plot_annotation(tag_levels = "A")
+```
+
+The point estimates in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:hdist-1)A.
+As such, Normal(`r mu1`, `r sd1`^2^) is a valid description of heterogeneity in valence effects.
+However, it is an incomplete description of heterogeneity for two reasons.
+First, it does not necessarily communicate the degree of heterogeneity in clear and actionable terms.
+Second, it ignores the uncertainty with which the heterogeneity parameters are estimated.
+
+### Interval descriptors
+
+Having estimated the parameters of the heterogeneity distribution, we now turn to using them to define alternative, more informative descriptions of the degree of heterogeneity in valence effects in the population.
+
+To begin with, and following @bolgerCausalProcessesPsychology2019, we used the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes a range within which a certain percentage of the population's slopes are expected to fall.
+By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$), but because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties.
+
+To calculate a heterogeneity interval, we first specify the desired probability limits.
+For a 90% interval, we use the .05 and .95 percentiles (which define the central 90% of the distribution).
+Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1`, `r sd1`) = [`r hi90`].
+In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
+We illustrate this interval in Figure \@ref(fig:hdist-1)B.
+
+### Proportion descriptors
+
+For many applications, quantities such as proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.
+For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" Thus, we first define a critical value, here zero.
+We then pass that value and the valence effect's estimated mean and standard deviation to the cumulative normal distribution function ($p$; `pnorm()` in R) to obtain the probability density in this distribution that falls below zero: $p^d = p(0, \beta_1, \tau_1) = p(0,$ `r mu1`, `r sd1`) = `r p_less_zero`.
+Specifically, this metric indicates the proportion of individuals in the population with negative effects.
+In words, we have estimated that `r p_less_zero` of the population respond faster to positively valenced words.
+We show this area under the curve in Figure \@ref(fig:hdist-1)C.
+
+However, using zero as a critical value might not be sufficiently informative especially when theory allows defining a smallest effect sizes of interest, or what is known as a region of practical equivalence (to zero) [@anvariUsingAnchorbasedMethods2021; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014].
+For example, we might know from theory that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
+To calculate, we can again use the cumulative normal distribution function ($p$) to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = p(0.1, \beta_1, \tau_1) - p(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
+In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero (within the [-0.1, 0.1] interval).
+We visualize this area under the curve in Figure \@ref(fig:hdist-1)D.
+
+### Ratio descriptors
+
+While the interval and ratio descriptors allow determining where the population's slope values are likely to fall, they do so in absolute terms, such as log(rt) in the running example.
+One way to contextualize the magnitude of heterogeneity is to put it in relative terms.
+To do so, we can compare the standard deviation of the valence effect distribution to its mean.
+[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the ratio of the standard deviation to the average effect is 0.25 or greater, then heterogeneity can be deemed noteworthy.
+With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r ratio`, suggesting that the degree of heterogeneity in valence effects is noteworthy.
+(We reiterate the point made by @bolgerCausalProcessesPsychology2019 that the cutoff value of 0.25 is arbitrary and researchers should choose the cutoff based on their substantive goals.)
+
+## Uncertainty
+
+Although informative, the expected normal distribution of valence effects, or any transformation of those point estimates such as the ones illustrated in Figure \@ref(fig:hdist-1), ignores the uncertainty inherent in the estimated parameters, and therefore cannot communicate the analyst's uncertainty about the degree of heterogeneity.
 That is, the uncertainty with which we have estimated the population parameters, indicated by e.g.
-the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward in calculating $HI_{90}$: We have only presented the expected heterogeneity interval, but not the range of credible values that it might take.
-For instance, if we re-calculate $HI_{90}$ based on the lowest CI bounds in Table \@ref(tab:tbl-fit-1), the result is a heterogeneity interval that indicates 90% of the population's slopes falling in the [`r hi90_low`] interval.
+the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward to the calculations described above.
+For example, we only presented the expected heterogeneity interval, but not the range of credible values that it might take.
+For instance, if we re-calculate $HI_{90}$ based on the lowest 95%CI bounds in Table \@ref(tab:tbl-fit-1), the result is a heterogeneity interval that indicates 90% of the population's slopes falling in the [`r hi90_low`] interval.
 In contrast to the first interval, this interval implies much less heterogeneity and that 90% of the population's slopes are expected to fall below zero.
 
-We have now arrived at the crux of the current work: How should we estimate and communicate causal effect heterogeneity such that the fundamental uncertainty in the estimated quantities (the mean and standard deviation of the distribution of slopes in the population) is retained?
-
-# Interlude on probabilistic estimation in practice {-}
-
-<!-- todo introduce citations and topics as we go; delete this section -->
-
-We propose that probabilistic methods are uniquely applicable to the problem presented above.
-In what follows, we describe practical calculations and visualizations that make use of bayesian methods in estimating and communicating causal effect heterogeneity.
-Frequentist methods, such as maximum likelihood above, aim at estimating optimum values of estimated quantities, typically called point estimates.
-In contrast, probabilistic methods aim at describing distributions of plausible values that could underlie hypothetical data generating processes.
-In this section, we briefly describe bayesian estimation such that we can meaningfully use it in the remainder of our proposition.
-For more complete treatments, we refer readers to excellent existing literature on the topic [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
-
-Lorem ipsum bayesian propagandum.
+We have now arrived at the crux of the current work: How should we estimate and describe heterogeneity in psychological phenomena such that the fundamental uncertainty in the estimated parameters is retained?
 
@@
 
-We now return to examining the heterogeneity in valence effects but with using bayesian methods.
-First, we use the R package brms to estimate the same model as above [@burknerAdvancedBayesianMultilevel2018; @burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2022].
-The goal of bayesian inference is the multivariate posterior probability distribution of the model's parameters.
-However, because closed-form solutions are not available for posterior distributions of many important types of statistical models, in practice bayesian methods rely on drawing many random samples from the posterior distribution [@gelmanBayesianDataAnalysis2013].
-Therefore, "estimate" here means drawing a large number of---here, 4,000---random samples from the posterior distribution.
-These samples can then be used to calculate and summarise the posterior distribution of any desired quantity.
-In practice, one obtains (e.g.) 4,000 samples from the posterior distribution, which can then be summarized using familiar data processing techniques.
-Below, we show the posterior mean and 2.5 and 97.5 percentiles of the parameters' distributions in the rightmost column of Table \@ref(tab:tbl-fit-1).
-Note that because we used the default noninformative prior distributions, summaries of the resulting posterior distributions are numerically almost identical to the maximum likelihood estimates.
-To highlight the issue at hand, Table \@ref(tab:tbl-samples-1) shows the first six rows of samples from their joint distribution.
+We think that probabilistic (i.e.
+"bayesian") methods are uniquely applicable to estimating and describing heterogeneity in psychological phenomena in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
+Moreover, these methods are widely available in modern software packages; here we used the R package brms, which uses familiar R formulas for specifying bayesian models [@burknerAdvancedBayesianMultilevel2018; @burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2023].
 
@@
 
-post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_"), regex = TRUE) %>% 
+post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) %>% 
   tibble() %>% 
-  mutate(b = b_valence1, sd = sd_person__valence1, .keep = "none") %>% 
-  mutate(hi_low = b + qnorm(.05)*sd, hi_high = b + qnorm(.95)*sd) %>% 
-  mutate(Sample = 1:n(), .before = 1)
-post1_sum <- post1 %>% 
-  mean_qi(hi_low, hi_high)
+  rownames_to_column("Sample")
+post1_sum <- fit1 %>% 
+  parameters(effects = "all", centrality = "mean", dispersion = TRUE) %>% 
+  tibble() %>% 
+  mutate(
+    Mean = number(Mean, .01), 
+    SD = number(SD, .001)
+  ) %>% 
+  select(Parameter, Mean, SD) %>% 
+  pivot_longer(c(Mean, SD), names_to = "Sample") %>%
+  pivot_wider(names_from = Parameter, values_from = value)
+
+set.seed(191)
 post1 %>% 
+  slice_sample(n = 6) %>% 
+  arrange(as.integer(Sample)) %>% 
+  mutate(
+    across(-Sample, ~number(., .01))
+  ) %>%
+  add_row(post1_sum) %>% 
   setNames(c("Sample", m1_parnames)) %>% 
-  head() %>% 
   apa_table(
     span_text_columns = FALSE,
-    caption = "First six rows of samples from the posterior distribution of Model 1.",
+    font_size = "small",
+    midrules = 6,
+    caption = "Six random samples from Model 1's parameters' posterior distributions, and their summaries.", 
     escape = FALSE,
@@
 
-To return to the $HI_{90}$, we could reproduce the above calculations for the expected interval using the posterior modes.
-Instead, to carry uncertainty forward from parameters to (e.g.) $HI_{90}$, we calculate samples of $HI_{90}$ from samples of the model's parameters (Table \@ref(tab:tbl-fit-1); Figure \@ref(fig:hdist-2)A).
-Specifically, because an interval is defined by its endpoints, the resulting posterior distribution is a two-dimensional distribution of lower and upper limits of the heterogeneity interval (Figure \@ref(fig:hdist-2)B).
+The goal of bayesian inference is the multivariate posterior probability distribution of the model's parameters.
+However, because closed-form solutions are not available for posterior distributions of many important types of statistical models, in practice bayesian methods rely on drawing many random samples from the posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016].
+Therefore, "estimate" means a large number of random samples from the posterior distribution.
+These samples can then be used to calculate and summarize the posterior distribution of any desired quantity.
 
-<!-- todo panel b: explain / give insight into why panel b is a negative correlation -->
-<!-- todo panel d requires a lot more introduction--move into its own figure & section -->
+In practice, one obtains (e.g.) 4,000 samples from the posterior distribution, and then summarizes them using familiar data processing techniques.
+To illustrate, Table \@ref(tab:tbl-samples-1) shows six random samples from the parameters' posterior distributions; the two bottom rows show their means and standard deviations.
+Note that because we used the brms package's default noninformative prior distributions, summaries of the resulting posterior distributions are numerically almost identical to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
+
+## Heterogeneity distribution
+
+Armed with the bayesian estimates (Table \@ref(tab:tbl-samples-1)), we now return to the distribution of valence effects in the population.
+We first redraw the expected heterogeneity distribution from Figure \@ref(fig:hdist-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:hdist-2)A.
+Superimposed to that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 random samples from the posterior distribution.
+From these curves we glean that the distribution of valence effects might well be narrower (i.e.
+less heterogeneity) or wider (more heterogeneity) than is suggested by the point estimates.
+
+Moreover, the location of these curves differ: Some curves are further to the left (average effect is more negative), and some further to the right (average effect is more positive).
+The distribution of these curves represents our uncertainty over the heterogeneity distribution of valence effects in the population.
+Thus, a good summary of heterogeneity that incorporates uncertainty must include information about uncertainty in both the scale (standard deviation) and location (mean) of the heterogeneity distribution.
 
@@
 #| label: hdist-2
-#| fig.height: 6.4
-#| fig.cap: \textbf{A}.
-Scatterplot of 4,000 random samples from the joint posterior distribution of the mean $\beta_1$ and standard deviation $\tau_1$ of the gaussian population distribution of valence effects.
+#| fig.height: 3.4
+#| fig.width: 4
+#| fig-cap: Posterior distribution of the heterogeneity distribution of valence effects (\textbf{A}) and its 90\% heterogeneity interval (\textbf{B}).
+\textbf{A}.
+Posterior mean heterogeneity distribution in dark blue, and 100 random draws from the distribution's posterior distribution in light blue.
 \textbf{B}.
-Scatterplot of 4,000 samples from the joint posterior distribution of the $HI_{90}$ lower ($\beta_1 + \phi_{0.05}\tau_1$; x-axis) and upper ($\beta_1 + \phi_{0.95}\tau_1$; y-axis) limits, calculated from samples in panel A.
-\textbf{C}.
-A sample of 200 $HI_{90}$s arranged in increasing magnitude of the standard deviation.
-Red symbols in each panel indicate the posterior mean quantities; e.g.
-the expected heterogeneity interval in C.
-\textbf{D}.
-Histogram of 4,000 samples from the posterior distribution of the proportion of negative effects in the population distribution of effects defined by $\beta_1$ and $\tau_1$.
-The red point and interval in the bottom indicate the proportion's posterior mean and 95\%CI.
+100 random draws from the 90\% heterogeneity interval's posterior distribution.
 
-post1 <- post1 %>%
-  arrange(hi_low) %>% 
-  mutate(i = 1:n())
+set.seed(99)
+x <- seq(-0.55, 0.25, length = 301)
 
-hcol <- "#cc4778"
-p_scatter <- post1 %>% 
-  ggplot(aes(b, sd)) +
-  labs(x = bquote(beta[1]), y = bquote(tau[1])) +
-  labs(
-    x = "Mean of slope distribution",
-    y = "SD of slope distribution"
-  ) +
-  geom_point(shape = 1, alpha = .25, size = .75) +
-  stat_summary_xy(col = hcol, .fun.x = mean, .fun.y = mean, size = 2) +
-  theme(
-    aspect.ratio = 1
+calc_pdf <- function(posterior) {
+  posterior %>% 
+    crossing(x) %>% 
+    mutate(
+      pdf = dnorm(x, b_valence1, sd_person__valence1),
+      cdf = pnorm(x, b_valence1, sd_person__valence1)
     )
+}
 
-p_scatter_hi <- post1 %>% 
-  ggplot(aes(hi_low, hi_high)) +
-  labs(
-    x = TeX("$HI_{90}$ lower limit"), 
-    y = TeX("$HI_{90}$ upper limit")
-  ) +
-  geom_point(shape = 1, alpha = .25, size = .75) +
-  stat_summary_xy(col = hcol, .fun.x = mean, .fun.y = mean, size = 2) +
-  theme(
-    aspect.ratio = 1
-  )
+dist_mean <- post1 %>% 
+  mean_qi(b_valence1, sd_person__valence1) %>% 
+  calc_pdf()
 
-post1_snip <- post1 %>% 
-  filter(i %in% as.integer(seq(1, 4000, length.out = 200)))
-p_intervals <- post1_snip %>% 
-  ggplot(aes(y = as.numeric(i), yend = as.numeric(i))) +
-  scale_fill_brewer(
-    "HI limit",
-    palette = "Set1",
-    aesthetics = c("fill", "color"),
-    breaks = c("hi_low", "hi_high"),
-    labels = c("Low", "High")
+dist_samples <- post1 %>% 
+  slice_sample(n = 100) %>% 
+  calc_pdf()
+
+p1 <- dist_mean %>% 
+  ggplot(aes(x, pdf)) +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0),
+    breaks = extended_breaks()
   ) +
   scale_y_continuous(
-    expand = expansion(0.01)
+    "Density",
+    expand = expansion(c(0.001, 0.05))
   ) +
-  labs(
-    x = "Slope magnitude",
-    y = "Posterior sample"
+  geom_line(
+    linewidth = .6,
+    col = "dodgerblue4"
   ) +
-  geom_segment(
-    aes(x = hi_low, xend = hi_high),
-    linewidth = .1, col = "gray50"
+  geom_line(
+    data = dist_samples,
+    aes(group = Sample),
+    col = "dodgerblue1",
+    alpha = .33, 
+    linewidth = .15
   ) +
-  geom_point(
-    data = post1_snip %>% 
-      pivot_longer(c(hi_low, hi_high)),
-    aes(x = value), col = "gray50",
-    size = .25, alpha = .5
+  theme(
+    axis.title.x = element_blank(), 
+    plot.margin = unit(c(4,4,16,4), "pt")
+  )
+p2 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>% 
+  arrange(b_valence1) %>% 
+  mutate(i = 1:n()) %>% 
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1)
+  ) %>% 
+  ggplot() +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0),
+    breaks = extended_breaks()
   ) +
-  geom_segment(
-    data = post1 %>% 
-      summarise(hi_low = mean(hi_low), hi_high = mean(hi_high)) %>% 
-      mutate(i = 2000),
-    aes(x = hi_low, xend = hi_high),
-    col = hcol
-  ) +
-  geom_point(
-    data = post1 %>% 
-      summarise(hi_low = mean(hi_low), hi_high = mean(hi_high)) %>% 
-      pivot_longer(c(hi_low, hi_high)) %>% 
-      mutate(i = 2000),
-    aes(x = value),
-    col = hcol
-  ) +
-  guides(color = guide_legend(override.aes = list(size = 2))) +
+  aes(xdist = dist, y = i) +
+  labs(y = "Sample") +
+  stat_pointinterval(
+    .width = .9,
+    interval_size_range = c(.05, .1),
+    alpha = .5,
+    color = "dodgerblue1"
+  )
+(p1 / p2) &
+  coord_cartesian(xlim = c(-0.55, 0.25)) &
   theme(
@@
     axis.ticks.y = element_blank()
-  )
-
-p_slopezero <- post1 %>% 
-  mutate(
-    p_slopes_under_zero = pnorm(0, b, sd)
-  ) %>% 
-  ggplot(aes(p_slopes_under_zero)) +
-  scale_y_continuous(
-    "Count",
-    expand = expansion(c(0.0025, .1))
-  ) +
-  scale_x_continuous(
-    "Proportion of negative slopes"
-  ) +
-  stat_histinterval(
-    breaks = 30,
-    outline_bars = TRUE,
-    slab_color = "gray20",
-    slab_fill = "gray80",
-    slab_linewidth = .2,
-    justification = .01,
-    point_color = hcol, interval_color = hcol,
-    .width = .90
-  ) +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-((p_scatter |
-    p_scatter_hi) /
-    (p_intervals |
-       p_slopezero)) +
+  ) &
+  plot_layout(heights = c(75, 25)) &
   plot_annotation(tag_levels = "A")
@@
 
+### Interval descriptors
+
+To carry uncertainty forward from model parameters to their transformations, such as $HI_{90}$, we reproduce the calculations described above, but instead of using the point estimates we repeat the calculation over each posterior sample.
+In this manner, we obtain (samples from) posterior distributions of the parameter transformations, such as $HI_{90}$ (Figure \@ref(fig:hdist-2)B).
+
 ```{r}
 p1s <- post1 %>% 
+  mutate(
+    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+  ) %>% 
   mean_qi(hi_low, hi_high) %>% 
@@
 The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u`.
-Communicating these two credibility intervals is not only cumbersome, but also ignores the correlation between the heterogeneity interval endpoints' posterior distributions (Figure \@ref(fig:hdist-2)B)).
-We therefore turn to alternative summaries of the population distribution of slopes.
+Thus, to adequately describe an effects heterogeneity interval, researchers must communicate two separate credibility intervals.
+However, communicating these two credibility intervals is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
+Because of these limitations, the heterogeneity interval may not always be the most desirable description of heterogeneity.
 
-## Alternatives to the heterogeneity interval
-
-To reiterate, the model's estimated distribution of slopes is a normal distribution with mean $\beta_1$ and standard deviation $\tau_1$.
-The 90% heterogeneity interval reports the $5^{th}$ and $95^{th}$ quantiles of the distribution, of whose posterior distributions are correlated.
-But we are not limited to describing the plausible ranges of slopes in the population.
-In fact, for many applications, quantities such as proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.
+### Proportion descriptors
 
 ```{r}
-tmp <- p_slopezero$data %>% 
-  mean_qi(p_slopes_under_zero, .width = .95) %>% 
+#| label: hdist-3
+#| fig.height: 2.2
+#| fig.width: 4
+#| fig-cap: \textbf{A}.
+100 random draws from the valence effect distribution's posterior distribution.
+Darker area under the curve represents the proportion of negative effects in the population for that posterior draw ($p^d$).
+\textbf{B}.
+Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI.
+
+p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>%
+  arrange(b_valence1) %>%
+  # slice(seq(1, nrow(.), by = 2)) %>%
+  mutate(i = 1:n()) %>% 
   mutate(
-    across(where(is.numeric), ~number(., .01)),
-    p = str_glue("{p_slopes_under_zero}, [{.lower}, {.upper}]")
+    dist = dist_normal(b_valence1, sd_person__valence1)
+  ) %>% 
+  ggplot() +
+  scale_fill_manual(values = c("dodgerblue1", "dodgerblue3"), na.value = "white") +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0),
+    breaks = extended_breaks()
+  ) +
+  scale_y_continuous(
+    "Sample",
+    expand = expansion(0)
+  ) +
+  aes(xdist = dist, y = i) +
+  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
+  stat_slab(
+    aes(fill = after_stat(x < 0)), 
+    height = 15,
+    slab_color = "grey10",
+    slab_linewidth = .1,
+    alpha = .5,
+    p_limits = c(.01, .99),
+    show.legend = FALSE
   )
+p2 <- post1 %>% 
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1),
+    pd = cdf(dist, 0),
+    prope = percent(cdf(dist, 0.1) - cdf(dist, -0.1), .1)
+  ) %>% 
+  ggplot(aes(pd)) +
+  scale_y_continuous(
+    "Count",
+    expand = expansion(0)
+  ) +
+  scale_x_continuous(
+    TeX('$p^d$'),
+    breaks = extended_breaks(),
+    labels = ~percent(., 1)
+  ) +
+  stat_histinterval(
+    breaks = 50,
+    outline_bars = TRUE,
+    slab_color = "gray10",
+    slab_fill = alpha("dodgerblue1", .33),
+    slab_linewidth = .2,
+    justification = .01,
+    point_color = "dodgerblue4", 
+    interval_color = "dodgerblue4",
+    .width = .90
+  )
+(p1 | p2) &
+  theme(
+    axis.ticks.y = element_blank(),
+    axis.text.y = element_blank()
+  ) &
+  plot_annotation(tag_levels = "A")
+```
 
-tmp2 <- p_slopezero$data %>% 
-  mean_qi(1-p_slopes_under_zero, .width = .95) %>% 
+```{r}
+tmp <- post1 %>% 
   mutate(
-    across(where(is.numeric), ~percent(., .1)),
-    p = str_glue("{`1 - p_slopes_under_zero`}, [{.lower}, {.upper}]")
+    dist = dist_normal(b_valence1, sd_person__valence1),
+    pd = cdf(dist, 0),
+    rope = cdf(dist, 0.1) - cdf(dist, -0.1),
+    ratio = abs(sd_person__valence1 / b_valence1)
+  ) %>% 
+  mean_qi(pd, rope, ratio) %>% 
+  mutate(
+    pd_r = str_glue(
+      "{percent(pd, .1)} [{percent(pd.lower, .1)}, {percent(pd.upper, .1)}]"
+    ),
+    pd_r_n = str_glue(
+      "{percent(1-pd, .1)} [{percent(1-pd.upper, .1)}, {percent(1-pd.lower, .1)}]"
+    ),
+    rope_r = str_glue(
+      "{percent(rope, .1)} [{percent(rope.lower, .1)}, {percent(rope.upper, .1)}]"
+    ),
+    rope_r_n = str_glue(
+      "{percent(1-rope, .1)} [{percent(1-rope.upper, .1)}, {percent(1-rope.lower, .1)}]"
+    ),
+    ratio_r = str_glue(
+      "{number(ratio, .01)} [{number(ratio.lower, .01)}, {number(ratio.upper, .01)}]"
+    )
   )
@@
 
-To make this concrete, we ask "What proportion of individuals in the population have a negative effect of valence?" To answer, we calculate the probability density below zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$, and obtain samples of this density.
-Figure \@ref(fig:hdist-2)D then shows these samples' histogram, and the posterior mean and 95% Credibility Interval (`r tmp$p`): We expect about `r percent(as.numeric(tmp$p_slopes_under_zero))` to be negative, but with 95% credibility, this value could be as low as `r percent(as.numeric(tmp$.lower))` or as high as `r percent(as.numeric(tmp$.upper))`.
-Alternatively [@bolgerCausalProcessesPsychology2019, p.605-606], the model predicts that `r tmp2$p` of individuals in the population would show reversals of the valence effect.
+An alternative and often more useful description of heterogeneity is the proportion of the population whose effects fall above or below some critical value.
+In many applications, zero is a useful critical value because it allows us to calculate proportions of individuals with negative and positive effects.
+Figure \@ref(fig:hdist-3)A visualizes 100 random draws of the heterogeneity distribution, such that the area under each curve below zero is filled in darker grey.
+That area represents the proportion of individuals who respond faster to positive words.
+
+Therefore, we can ask "What proportion of individuals in the population have a negative effect of valence?" To answer, we calculate the probability density below zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$, and obtain samples of this density (Figure \@ref(fig:hdist-3)A).
+Figure \@ref(fig:hdist-3)B then shows these samples' histogram, and the posterior mean and 95% Credibility Interval (`r tmp$pd_r`).
+In words, we expect about `r percent(as.numeric(tmp$pd), .1)` to be negative, but with 95% credibility, this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
+Alternatively [@bolgerCausalProcessesPsychology2019, p.605-606], the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
+
+However, a sign test against zero is not always the most informative quantity because it discards all information about the magnitude of effects.
+A more theoretically motivated approach is to argue for and define a smallest effect size of interest (SESOI) which then defines a region of practical equivalence (ROPE) to zero, and then quantify uncertainty for the proposition that the effect is "practically equivalent to zero".
+Figure \@ref(fig:hdist-4)A shows 100 random draws of the heterogeneity distribution of valence effects, such that the area under the curve within a ROPE of [-0.1, 0.1] ($p^{ROPE}$) is filled with a darker color.
+That shaded area represents the posterior probability, for each draw, that the effect is practically equivalent to zero.
+Figure \@ref(fig:hdist-4)B then shows a histogram of all 4,000 posterior draws of $p^{ROPE}$ with their mean and 95%CI.
+We learn that `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
 
 ```{r}
-tmp <- post1 %>% 
-  summarise(
-    post_prob_90p_negative = percent(
-      sum(hi_high < 0) / n(), 
-      .01
+#| label: hdist-4
+#| fig.height: 2.2
+#| fig.width: 4
+#| fig-cap: \textbf{A}.
+100 random draws from the valence effect distribution's posterior distribution.
+Dark area under the curve represents the proportion of effects within a region of practical equivalence to zero (ROPE) in the population for that posterior draw ($p^{ROPE}$).
+\textbf{B}.
+Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI.
+p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>% 
+  arrange(b_valence1) %>% 
+  # slice(seq(1, nrow(.), by = 2)) %>%
+  mutate(i = 1:n()) %>% 
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1)
+  ) %>% 
+  ggplot() +
+  scale_fill_manual(
+    values = c("dodgerblue1", "dodgerblue3"), 
+    na.value = "white"
+    ) +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0),
+    breaks = extended_breaks()
+  ) +
+  scale_y_continuous(
+    "Sample",
+    expand = expansion(0)
+  ) +
+  aes(xdist = dist, y = i) +
+  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
+  stat_slab(
+    aes(fill = after_stat(near(x, 0, .1))), 
+    height = 15,
+    slab_color = "grey10",
+    slab_linewidth = .1,
+    alpha = .5,
+    p_limits = c(.01, .99),
+    show.legend = FALSE
   )
+p2 <- post1 %>% 
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1),
+    pd = cdf(dist, 0),
+    prope = cdf(dist, 0.1) - cdf(dist, -0.1)
   ) %>% 
-  pull(1)
+  ggplot(aes(prope)) +
+  scale_y_continuous(
+    "Count",
+    expand = expansion(0)
+  ) +
+  scale_x_continuous(
+    TeX('$p^{ROPE}$'),
+    breaks = extended_breaks(),
+    labels = ~percent(., 1)
+  ) +
+  stat_histinterval(
+    breaks = 50,
+    outline_bars = TRUE,
+    slab_color = "gray10",
+    slab_fill = alpha("dodgerblue1", .33),
+    slab_linewidth = .2,
+    justification = .01,
+    point_color = "dodgerblue4", 
+    interval_color = "dodgerblue4",
+    .width = .90
+  )
+(p1 | p2) &
+  theme(
+    axis.ticks.y = element_blank(),
+    axis.text.y = element_blank()
+  ) &
+  plot_annotation(tag_levels = "A")
 ```
 
-Additionally, we might ask, for example, "What is the posterior probability that (e.g.) 90% of slopes in the population are negative?" To answer, we calculate, for each posterior sample of $\beta_1$ and $\tau_1$, whether the resulting $HI_{90}$'s upper limit is below zero.
-Visually, this calculation represents drawing a vertical line at 0 on the y-axis of Figure \@ref(fig:hdist-2)B and calculating the proportion of samples below zero over all the samples.
-The result, `r tmp`, indicates that we expect `r tmp` of slopes in the population to be negative.
-Note that this quantity cannot be calculated from the point estimates alone.
+So far, these examples have highlighted the importance of quantifying uncertainty in measures of heterogeneity.
+Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^d$=`r percent(tmp$pd, .1)` and $p^{ROPE}$=`r percent(tmp$rope, .1)`.
+However, with 95% confidence, these values might be as great as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
 
-## Does heterogeneity matter?
+### Ratio descriptors
+
+Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$.
+Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, which we then transformed into 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ in Figure \@ref(fig:hdist-5)B.
+The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)`.
 
 ```{r}
-tmp <- post1 %>% 
-  mutate(ratio = abs(sd / b)) %>% 
-  mean_qi(ratio) %>% 
+#| label: hdist-5
+#| fig.height: 2.2
+#| fig.width: 4
+#| fig-cap: \textbf{A}.
+4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
+\textbf{B}.
+Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95%CI.
+p1 <- post1 %>% 
+  ggplot(aes(b_valence1, sd_person__valence1)) +
+  scale_y_continuous(
+    TeX("$\\tau_1$")
+  ) +
+  scale_x_continuous(
+    TeX("$\\beta_1$")
+  ) +
+  geom_point(
+    shape = 1,
+    col = "dodgerblue3",
+    alpha = .25,
+    size = .5
+  )
+p2 <- post1 %>% 
+  mutate(ratio = abs(sd_person__valence1 / b_valence1)) %>% 
+  ggplot(aes(ratio)) +
+    scale_y_continuous(
+    "Count",
+    expand = expansion(0)
+  ) +
+  scale_x_continuous(
+    TeX('$\\frac{\\tau_1}{\\beta_1}$'),
+    breaks = extended_breaks()
+  ) +
+  stat_histinterval(
+    breaks = 50,
+    outline_bars = TRUE,
+    slab_color = "gray10",
+    slab_fill = alpha("dodgerblue1", .33),
+    slab_linewidth = .2,
+    justification = .01,
+    point_color = "dodgerblue4", 
+    interval_color = "dodgerblue4",
+    .width = .90
+  ) +
+  theme(
+    axis.ticks.y = element_blank(),
+    axis.text.y = element_blank(),
+  )
+(p1 | p2) &
+  theme(
+    aspect.ratio = 1
+  ) &
+  plot_annotation(tag_levels = "A")
+```
+
+# Comparing heterogeneity between groups
+
+We have now established methods for assessing and describing between-person heterogeneity in an experimental effect in a way that keeps track and describes uncertainty in this variability.
+In the following example, we take this idea further by addressing the following question: Does heterogeneity differ between groups?
+These groups could be e.g.
+different demographic groups that participated in an experiment ("Is the valence effect more variable among younger vs.
+older adults?") or different research groups in a multi-site project.
+
+Accurately representing the uncertainty in model parameters is critical for addressing questions about differences: It is not sufficient to report that the standard deviation of valence effects was greater among one group compared to another.
+Instead, we need to quantify uncertainty in the differences, and the methods described above extend naturally to such questions.
+
+As an example, we randomly divided the participants in the example data to two Groups A and B.
+These could represent, for example, two demographic groups such as younger and older participants.
+We then modified Model 1 to estimate separate population-level parameters ($\beta$, $\tau$, $\rho$) for the groups, allowing us to examine not only whether the average persons differ between the groups, but also the extent to which the two groups differ in heterogeneity.
+
+```{r}
+set.seed(999)
+dat2 <- distinct(dat, person) %>% 
   mutate(
-    across(where(is.numeric), ~number(., .1)),
-    r = str_glue("{ratio} [{.lower}, {.upper}]")
+    g = sample(
+      c("a", "b"), 
+      length(unique(dat$person)), 
+      replace = TRUE
+    ),
+    g = factor(g)
+  ) %>% 
+  right_join(dat)
+contrasts(dat2$g) <- c(-0.5, 0.5)
+```
+
+```{r fit2-brms}
+fit2 <- brm(
+  bf(
+    rt ~ 0 + g / valence + (1 + valence | gr(person, by = g)),
+    sigma ~ 0 + g
+  ),
+  family = brmsfamily("gaussian", link = "identity", link_sigma = "identity"),
+  data = dat2,
+  file = "models/brm-fit-2",
+  control = list(adapt_delta = .95)
 )
@@
 
-The above two quantities summarized the heterogeneity of valence effects in the population in absolute terms.
-To ask whether heterogeneity is substantial enough to matter, we might directly calculate the posterior mean and (e.g.) 95%CI of $\tau_1$ to decide whether the spread of effects in the population is large enough to be of practical significance.
-Alternatively, we might focus on the relative size of heterogeneity when compared to the average effect.
-[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the standard deviation is 0.25 or greater in magnitude than the average effect, heterogeneity is noteworthy.
-With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r tmp$r` (posterior mean and 95% CI), suggesting that the degree of heterogeneity in the effect of valence is noteworthy.
-(We reiterate the point made by @bolgerCausalProcessesPsychology2019 that the cutoff value of 0.25 is arbitrary and researchers should choose the cutoff based on their substantive goals.)
+Table \@ref(tab:tbl-fit-2) shows the model's estimated parameters and the heterogeneity metrics $p^d$, $p^{ROPE}$ (based on a ROPE of [-0.1, 0.1]), and $\frac{\tau_1}{\beta_1}$.
+Using these estimates (samples from each parameter's posterior distribution), we can extend the calculations from above to calculate differences between groups.
+Because we randomly divided participants into two groups, there are no broad differences between groups.
 
-### Smallest effect size of interest
+```{r}
+#| label: tbl-fit-2
+#| cache: false
 
-todo: Examine proportions of effects in context of hypothesized SESOIs.
+m2_parnames <- c(
+  "$\\beta_0$", 
+  "$\\beta_1$", 
+  "$\\tau_0$", 
+  "$\\tau_1$", 
+  "$\\rho$",
+  "$\\sigma$"
+)
+post2 <- gather_draws(fit2, `b_g.*`, `sd_.*`, `cor_.*`, `b_sigma.*`, regex = TRUE) %>% 
+  ungroup() %>% 
+  mutate(
+    g = str_extract(.variable, "g[a|b]") %>% str_remove("g"),
+    .variable = .variable %>% str_replace_all(":", "_") %>% str_remove_all("([_|:]g[a|b])+")
+  )
 
-# Example 3: Comparing heterogeneity between groups
+post2 <- post2 %>% 
+  pivot_wider(names_from = .variable, values_from = .value) %>% 
+  mutate(
+    dist = dist_normal(`b_valence1`, sd_person__valence1),
+    pd = cdf(dist, 0),
+    prope = cdf(dist, 0.1) - cdf(dist, -0.1),
+    ratio = abs(sd_person__valence1 / `b_valence1`)
+  ) %>% 
+  select(-dist)
 
-todo: Model separate random effects distributions for e.g.
-labs in a large replication project, then compare heterogeneity between labs.
-(In the current running example men vs women would be obvious & easy--maybe create a fake gender variable?)
-
-# Example 4: Multivariate heterogeneity
-
-todo: Extend Study 4 from Bolger et al., studying correlations of random effects between timepoints among the same individuals.
+post2 %>% 
+  pivot_longer(b:ratio) %>% 
+  pivot_wider(names_from = g, values_from = value) %>% 
+  rename(A = a, B = b) %>% 
+  mutate(`B - A` = B - A) %>% 
+  pivot_longer(c(A, B, `B - A`), names_to = "g") %>% 
+  mutate(name = fct_inorder(name)) %>% 
+  group_by(g, name) %>% 
+  mean_qi(value) %>% 
+  mutate(
+    r = if_else(
+      name %in% c("pd", "prope"),
+      str_glue("{percent(value, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]"),
+      str_glue("{number(value, .01)} [{number(.lower, .01)}, {number(.upper, .01)}]")
+      )
+  ) %>% 
+  mutate(
+    name2 = rep(
+      c(
+        m2_parnames, 
+        "$p^d$", 
+        "$p^{ROPE}$", 
+        "$\\frac{\\tau_1}{\\beta_1}$"
+        ), 
+      times = 3
+    )
+  ) %>% 
+  select(Parameter = name2, g, r) %>% 
+  mutate(g = str_glue("Group {toupper(g)}")) %>% 
+  pivot_wider(names_from = g, values_from = r) %>% 
+  apa_table(
+    span_text_columns = FALSE,
+    font_size = "scriptsize",
+    caption = "Parameter estimates from Model 2.",
+    escape = FALSE
+  )
+```
 
 # Discussion
+
+In the current work, we illustrated the use of practical descriptors of between-person heterogeneity in psychological phenomena.
+Moreover, we emphasized throughout that probabilistic modelling allows carrying uncertainty forward from model parameters to heterogeneity descriptors.
+The resulting metrics are useful because they not only convey the analyst's expectation of the magnitude of heterogeneity, but also her uncertainty in the description.
+In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, allows easily comparing these heterogeneity descriptors across groups.
+
+We believe that psychology should take a meta-theoretic step forward and try to incorporate an understanding of heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019].
+In order to do so, descriptions of heterogeneity must include measures of uncertainty.
+
+## Limitations
+
+## Conclusion
 
@@
 
-todo: think about if this appendix is needed at all
+todo: think about if this appendix is needed at all because code will be open.

### 2023-06-27 ms.Rmd
Delete appendix

@@
 :::
-
-\clearpage
-\onecolumn
-
-# Appendix: Code listings
-
-\setcounter{page}{1}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{A-\arabic{table}}
-\renewcommand{\thefigure}{A-\arabic{figure}}
-
-## Contrast coding
-
-By default, R uses dummy coding (called "treatment contrasts" in R) for categorical predictors, meaning that one category is used as a baseline, and then contrasts between the baseline category and the other categories are estimated.
-It is often useful to use contrast coding (called "sum to zero contrasts" in R) instead, where the levels of a categorical factor are assigned codes that sum to zero: In this manner, the intercept of the model is at the midpoint between the category levels, and consequently the other parameters are easier to interpret.
-To use contrast codes for Valence, in the example data, the R code is `contrasts(dat$valence) <- "contr.sum"`.
-However, this has the slight drawback that the levels are assigned numerical codes of -1 and 1, and the resulting parameter estimate indicates a contrast to the "average" valence, and must be multiplied by two to give the difference between the conditions.
-We therefore used the more readily interpretable values of -0.5 and 0.5:
-
-```{r contrast-coding, echo = TRUE, eval = TRUE}
-```
-
-In addition to making the resulting parameter estimate directly indicate the difference between conditions, this keeps the `valence` variable in R as a factor, which ensures that downstream functions using that variable will appropriately treat it as a categorical variable with interpretable labels (e.g.
-for figures.) For useful references to predictor coding in R, see [@rabeHyprPackageHypothesisdriven2020; @schadHowCapitalizePriori2020].
-
-## Estimating a multilevel model using maximum likelihood
-
-```{r fit1-lmer, echo=TRUE, eval=FALSE}
-```
-
-## Calculating a heterogeneity interval
-
-```{r hi90-appendix, echo = -1}
-options(digits = 2)
-mean <- -0.16 # Mean of slope distribution point estimate from Table 1
-sd <- 0.12 # Standard deviation of slope distribution point estimate from Table 1
-quantiles <- c(0.05, 0.95) # Desired quantiles (these define 90%)
-mean + qnorm(quantiles)*sd
-```
-
-## Estimating a multilevel model using probabilistic inference
-
-```{r fit1-brms, echo=TRUE, eval=FALSE}
-```
-
-todo: think about if this appendix is needed at all because code will be open.

### 2023-07-04 ms.Rmd
Use R native pipe

@@
 # Save variables as numerical & factor (might need for more complex contrast/eff/etc codes)
-dat <- read_csv("data/heterogeneity_dataset1_traitvalence.csv") %>% 
+dat <- read_csv("data/heterogeneity_dataset1_traitvalence.csv") |> 
   # Cleaning as in Bolger et al 2019
@@
     !(id %in% c(250, 257, 272))
-  ) %>% 
+  ) |> 
   # Some more cleaning
@@
     rt = logrt
-  ) %>% 
-  select(person, trial, valence, rt) %>% 
+  ) |> 
+  select(person, trial, valence, rt) |> 
   arrange(person, trial)
@@
 #| label: tab-dat1
-dat %>% 
-  head() %>% 
+dat |> 
+  head() |> 
   apa_table(
@@
 # Create a table of coefficients
-tab1_lmer <- model_parameters(fit1_lmer, effects = "all") %>% 
-  tibble() %>% 
+tab1_lmer <- model_parameters(fit1_lmer, effects = "all") |> 
+  tibble() |> 
   select(Parameter, Coefficient, starts_with("CI_"))
@@
 
-tab1_brms <- parameters(fit1, centrality = "mean", effects = "all") %>% 
-  tibble() %>% 
+tab1_brms <- parameters(fit1, centrality = "mean", effects = "all") |> 
+  tibble() |> 
   select(Parameter, Mean, starts_with("CI_"))
@@
 
-tab1_lmer %>% 
+tab1_lmer |> 
   mutate(
@@
     MLE = str_glue("{Coefficient} [{CI_low}, {CI_high}]"),
-  ) %>% 
-  select(Parameter, `Estimate [95\\%CI]` = MLE) %>% 
+  ) |> 
+  select(Parameter, `Estimate [95\\%CI]` = MLE) |> 
   mutate(
     Parameter = m1_parnames
-  ) %>% 
+  ) |> 
   apa_table(
@@
   scale_fill_manual(values = c("grey90", "grey80"), na.value = "white")
-d1 <- dist_normal(mu1n, sd1n) %>% 
+d1 <- dist_normal(mu1n, sd1n) |> 
   mean_qi(.width = c(.66, .9))
@@
 
-post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) %>% 
-  tibble() %>% 
+post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) |> 
+  tibble() |> 
   rownames_to_column("Sample")
-post1_sum <- fit1 %>% 
-  parameters(effects = "all", centrality = "mean", dispersion = TRUE) %>% 
-  tibble() %>% 
+post1_sum <- fit1 |> 
+  parameters(effects = "all", centrality = "mean", dispersion = TRUE) |> 
+  tibble() |> 
   mutate(
@@
     SD = number(SD, .001)
-  ) %>% 
-  select(Parameter, Mean, SD) %>% 
-  pivot_longer(c(Mean, SD), names_to = "Sample") %>%
+  ) |> 
+  select(Parameter, Mean, SD) |> 
+  pivot_longer(c(Mean, SD), names_to = "Sample") |>
   pivot_wider(names_from = Parameter, values_from = value)
@@
 set.seed(191)
-post1 %>% 
-  slice_sample(n = 6) %>% 
-  arrange(as.integer(Sample)) %>% 
+post1 |> 
+  slice_sample(n = 6) |> 
+  arrange(as.integer(Sample)) |> 
   mutate(
     across(-Sample, ~number(., .01))
-  ) %>%
-  add_row(post1_sum) %>% 
-  setNames(c("Sample", m1_parnames)) %>% 
+  ) |>
+  add_row(post1_sum) |> 
+  setNames(c("Sample", m1_parnames)) |> 
   apa_table(
@@
 calc_pdf <- function(posterior) {
-  posterior %>% 
-    crossing(x) %>% 
+  posterior |> 
+    crossing(x) |> 
     mutate(
@@
 
-dist_mean <- post1 %>% 
-  mean_qi(b_valence1, sd_person__valence1) %>% 
+dist_mean <- post1 |> 
+  mean_qi(b_valence1, sd_person__valence1) |> 
   calc_pdf()
 
-dist_samples <- post1 %>% 
-  slice_sample(n = 100) %>% 
+dist_samples <- post1 |> 
+  slice_sample(n = 100) |> 
   calc_pdf()
 
-p1 <- dist_mean %>% 
+p1 <- dist_mean |> 
   ggplot(aes(x, pdf)) +
@@
   )
-p2 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>% 
-  arrange(b_valence1) %>% 
-  mutate(i = 1:n()) %>% 
+p2 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
+  arrange(b_valence1) |> 
+  mutate(i = 1:n()) |> 
   mutate(
     dist = dist_normal(b_valence1, sd_person__valence1)
-  ) %>% 
+  ) |> 
   ggplot() +
@@
 ```{r}
-p1s <- post1 %>% 
+p1s <- post1 |> 
   mutate(
@@
     hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
-  ) %>% 
-  mean_qi(hi_low, hi_high) %>% 
+  ) |> 
+  mean_qi(hi_low, hi_high) |> 
   mutate(
@@
 
-p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>%
-  arrange(b_valence1) %>%
-  # slice(seq(1, nrow(.), by = 2)) %>%
-  mutate(i = 1:n()) %>% 
+p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |>
+  arrange(b_valence1) |>
+  # slice(seq(1, nrow(.), by = 2)) |>
+  mutate(i = 1:n()) |> 
   mutate(
     dist = dist_normal(b_valence1, sd_person__valence1)
-  ) %>% 
+  ) |> 
   ggplot() +
@@
   )
-p2 <- post1 %>% 
+p2 <- post1 |> 
   mutate(
@@
     prope = percent(cdf(dist, 0.1) - cdf(dist, -0.1), .1)
-  ) %>% 
+  ) |> 
   ggplot(aes(pd)) +
@@
 ```{r}
-tmp <- post1 %>% 
+tmp <- post1 |> 
   mutate(
@@
     ratio = abs(sd_person__valence1 / b_valence1)
-  ) %>% 
-  mean_qi(pd, rope, ratio) %>% 
+  ) |> 
+  mean_qi(pd, rope, ratio) |> 
   mutate(
@@
 Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI.
-p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) %>% 
-  arrange(b_valence1) %>% 
-  # slice(seq(1, nrow(.), by = 2)) %>%
-  mutate(i = 1:n()) %>% 
+p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
+  arrange(b_valence1) |> 
+  # slice(seq(1, nrow(.), by = 2)) |>
+  mutate(i = 1:n()) |> 
   mutate(
     dist = dist_normal(b_valence1, sd_person__valence1)
-  ) %>% 
+  ) |> 
   ggplot() +
@@
     prope = cdf(dist, 0.1) - cdf(dist, -0.1)
-  ) %>% 
+  ) |> 
   ggplot(aes(prope)) +
@@
 Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95%CI.
-p1 <- post1 %>% 
+p1 <- post1 |> 
   ggplot(aes(b_valence1, sd_person__valence1)) +
@@
   )
-p2 <- post1 %>% 
-  mutate(ratio = abs(sd_person__valence1 / b_valence1)) %>% 
+p2 <- post1 |> 
+  mutate(ratio = abs(sd_person__valence1 / b_valence1)) |> 
   ggplot(aes(ratio)) +
@@
 set.seed(999)
-dat2 <- distinct(dat, person) %>% 
+dat2 <- distinct(dat, person) |> 
   mutate(
@@
     g = factor(g)
-  ) %>% 
+  ) |> 
   right_join(dat)
@@
 )
-post2 <- gather_draws(fit2, `b_g.*`, `sd_.*`, `cor_.*`, `b_sigma.*`, regex = TRUE) %>% 
-  ungroup() %>% 
-  mutate(
-    g = str_extract(.variable, "g[a|b]") %>% str_remove("g"),
-    .variable = .variable %>% str_replace_all(":", "_") %>% str_remove_all("([_|:]g[a|b])+")
-  )
-
-post2 <- post2 %>% 
-  pivot_wider(names_from = .variable, values_from = .value) %>% 
+post2 <- gather_draws(fit2, `b_g.*`, `sd_.*`, `cor_.*`, `b_sigma.*`, regex = TRUE) |> 
+  ungroup() |> 
+  mutate(
+    g = str_extract(.variable, "g[a|b]") |> str_remove("g"),
+    .variable = .variable |> str_replace_all(":", "_") |> str_remove_all("([_|:]g[a|b])+")
+  )
+
+post2 <- post2 |> 
+  pivot_wider(names_from = .variable, values_from = .value) |> 
   mutate(
@@
     ratio = abs(sd_person__valence1 / `b_valence1`)
-  ) %>% 
+  ) |> 
   select(-dist)
 
-post2 %>% 
-  pivot_longer(b:ratio) %>% 
-  pivot_wider(names_from = g, values_from = value) %>% 
-  rename(A = a, B = b) %>% 
-  mutate(`B - A` = B - A) %>% 
-  pivot_longer(c(A, B, `B - A`), names_to = "g") %>% 
-  mutate(name = fct_inorder(name)) %>% 
-  group_by(g, name) %>% 
-  mean_qi(value) %>% 
+post2 |> 
+  pivot_longer(b:ratio) |> 
+  pivot_wider(names_from = g, values_from = value) |> 
+  rename(A = a, B = b) |> 
+  mutate(`B - A` = B - A) |> 
+  pivot_longer(c(A, B, `B - A`), names_to = "g") |> 
+  mutate(name = fct_inorder(name)) |> 
+  group_by(g, name) |> 
+  mean_qi(value) |> 
   mutate(
@@
       )
-  ) %>% 
+  ) |> 
   mutate(
@@
     )
-  ) %>% 
-  select(Parameter = name2, g, r) %>% 
-  mutate(g = str_glue("Group {toupper(g)}")) %>% 
-  pivot_wider(names_from = g, values_from = r) %>% 
+  ) |> 
+  select(Parameter = name2, g, r) |> 
+  mutate(g = str_glue("Group {toupper(g)}")) |> 
+  pivot_wider(names_from = g, values_from = r) |> 
   apa_table(

### 2023-09-07 ms.Rmd
Incoming revisions

@@
   Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development.
-Yet, established methods and practices for calculating and describing heterogeneity are lacking in how they treat uncertainty regarding heterogeneity in the phenomenon under study.
-In the current work, we aim to deliver methods and guidelines for more complete treatments of heterogeneity in psychological phenomena.
-  We provide a walkthrough of probabilistic methods for i.
+Yet, established methods and practices for describing heterogeneity underestimate the true uncertainty in those estimates.
+In the current work, we aim to deliver methods and guidelines for more rigorous treatments of heterogeneity in psychological phenomena.
+  We provide a walkthrough of methods for i.
 estimating, ii.
@@
 describing heterogeneity.
-The methods we propose---bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
-  We hope that the proposed methods help researchers to move from reporting point estimates to drawing inferences about magnitudes and differences in heterogeneity in psychological phenomena across different populations.
+Although other methods for incorporating estimation uncertainty are available, the methods we propose---involving bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
+  We hope that the proposed methods help researchers to widen their inferential focus from statistical averages to include statistical heterogeneity in psychological phenomena.
 keywords: variation, heterogeneity, hierarchical models, mixed models, methodology
 authornote: |
-  \noindent \textbf{Working paper. Proceed with caution.}
+  \noindent \textbf{This working paper is not yet peer reviewed.}
 wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
@@
 
-```
-
-```{r data, include = FALSE}
+dir.create("data", FALSE)
+```
+
+# Introduction
+
+When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not.
+Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon if the treatment effect is highly variable from one individual to another.
+The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].
+
+One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still tend to analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019].
+Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to estimate and describe heterogeneity.
+In the current work, we propose several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
+
+## Current work
+
+Our aim is to outline methods for studying and communicating heterogeneity in a manner that fully takes uncertainty into account.
+We first introduce an example dataset from social psychology and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.
+We then outline useful summaries of heterogeneity that focus on informative numerical and graphical descriptions, and note how they are limited by established methods' focus on point estimates of model parameters.
+
+We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods.
+By estimating probability distributions over model parameters, these bayesian methods are uniquely suitable for examining heterogeneity and researchers' uncertainty about it.
+Moreover, advances in computational procedures and statistical software make these methods available to researchers with a minimal learning curve beyond established maximum likelihood based methods.
+
+In the following sections, we detail calculations and graphics that make use of bayesian methods in communicating descriptions of heterogeneity such that uncertainty in the model parameters and, subsequently, heterogeneity, is taken into account.
+We then further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.
+
+We conclude by discussing additional benefits of probabilistic inference in psychology in general, and for multilevel models specifically.
+We emphasize that these benefits come with little or no cost to applied practitioners.
+We also discuss our findings regarding causal effect heterogeneity in light of this new perspective.
+
+# Review of expected heterogeneity in hierarchical models
+
+To begin our exposition, we reproduce the analyses presented in @bolgerCausalProcessesPsychology2019.
+In their study, which replicated findings first presented in @scholerInflatingDeflatingSelf2014, 62 participants saw twenty positively and twenty negatively valenced words, and judged whether each word was self-descriptive or not.
+Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that response times to positively valenced words would be shorter than to negatively valenced words [@scholerInflatingDeflatingSelf2014].
+
+```{r data-bolgeretal, include = FALSE}
 # Load, save, & clean example data from Bolger et al. (2019)
-dir.create("data", FALSE)
 path <- "data/bolger-etal.zip"
@@
   )
-  unzip(path, exdir = dirname(path), junkpaths = TRUE)
+  unzip(path, exdir = str_remove(path, ".zip"), junkpaths = TRUE)
 }
@@
 # Save variables as numerical & factor (might need for more complex contrast/eff/etc codes)
-dat <- read_csv("data/heterogeneity_dataset1_traitvalence.csv") |> 
+dat <- read_csv("data/bolger-etal/heterogeneity_dataset1_traitvalence.csv") |> 
   # Cleaning as in Bolger et al 2019
@@
 ```
-
-# Introduction
-
-When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not.
-Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon if the treatment effect is highly variable from one individual to another.
-The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].
-
-One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still tend to analyze data with models that cannot address between-person variability, such as ANOVA [@bolgerCausalProcessesPsychology2019].
-Although more appropriate mixed-effects (or multilevel, hierarchical) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to estimate and describe heterogeneity to a sufficiently informative degree.
-In the current work, we propose several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
-
-## Current work
-
-Our aim is to outline methods for studying and communicating heterogeneity in a manner that fully takes uncertainty into account.
-We first introduce an example dataset from social psychology and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.
-We then outline useful summaries of heterogeneity that focus on informative numerical and graphical descriptions, and note how they are limited by established methods' focus on point estimates of model parameters.
-
-We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods.
-By estimating probability distributions over model parameters, these bayesian methods are uniquely suitable for examining heterogeneity and researchers' uncertainty about it.
-Moreover, advances in computational procedures and statistical software make these methods available to researchers with a minimal learning curve beyond established maximum likelihood based methods.
-
-In the following sections, we detail calculations and graphics that make use of bayesian methods in communicating descriptions of heterogeneity such that uncertainty in the model parameters and, subsequently, heterogeneity, is taken into account.
-We then further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.
-
-We conclude by discussing additional benefits of probabilistic inference in psychology in general, and for multilevel models specifically.
-We emphasize that these benefits come with little or no cost to applied practitioners.
-We also discuss our findings regarding causal effect heterogeneity in light of this new perspective.
-
-# Review of expected heterogeneity in hierarchical models
-
-To begin our exposition, we reproduce the analyses presented in @bolgerCausalProcessesPsychology2019.
-In their study, which replicated findings first presented in @scholerInflatingDeflatingSelf2014, 62 participants saw twenty positively and twenty negatively valenced words, and judged whether each word was self-descriptive or not.
-Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that response times to positively valenced words would be shorter than to negatively valenced words [@scholerInflatingDeflatingSelf2014].
 
@@
     digits = c(0,0,0,2),
-    caption = "First six rows of example dataset 1.", 
+    caption = "First six rows of example dataset 1 (Bolger et al., 2019; Exp 1).", 
     note = "rt is log-transformed."
@@
 This equation includes two kinds of parameters: $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V), do not have subscripts.
-That means that they are not further modelled, and therefore sometimes called "fixed".
+That means that they are not further modelled except by way of (optional) prior distributions with fixed hyperparameters.
+In line with existing conventions in bayesian inference and software, we refer to these as *population level* parameters.
+From the frequentist perspective, these terms are constants (i.e.
+do not potentially vary as functions of covariates) and are therefore sometimes "fixed".
+For all practical purposes, because we do not discuss prior distributions with constant hyperparameters, readers can equate the terms "fixed" and "population level" parameters in the current discussion.
+
 We contrast coded valence such that negative words are coded as -0.5, and positive words as 0.5.
@@
 That is, $\beta_0 + \gamma_{01}$ is the average reaction time for subject $j=1$.
-The $\gamma$ coefficients have subscripts, which in this context indicates that they are further modelled, and therefore sometimes called "random" [@gelmanDataAnalysisUsing2007].
+The $\gamma$ coefficients have subscripts, which in this context indicates that they are further modelled on covariates---here, index variables for "person" in the data.
+We refer to these as "group level" parameters, in line with conventions.
+In the frequentist discourse, these parameters are called "random" parameters [@gelmanDataAnalysisUsing2007].
+Again, readers can equate the term "group level" parameters with "random" parameters in the ongoing discussion.
+
 As is standard in multilevel modelling, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:
@@
   )
-mu1n <- tab1_lmer[2,2,T]
-mu1 <- number(mu1n, .01)
-sd1n <- tab1_lmer[4,2,T]
-sd1 <- number(sd1n, .01)
+mu1 <- number(as.numeric(tab1_lmer[2,2:4,T]))
+sd1 <- number(as.numeric(tab1_lmer[4,2:4,T]))
+mu1n <- tab1_lmer[2,2,TRUE]
+sd1n <- tab1_lmer[4,2,TRUE]
 ```
@@
 We show a traditional summary of the estimated parameters from this model in Table \@ref(tab:tbl-fit-1).
-The expected effect of positive valence on RT for the average person is approximately `r mu1` log seconds, and the expected standard deviation describing this effect's degree of variability in the population is approximately `r sd1`.
+For the average person, the estimated effect of valence on RT is `r mu1[1]` log seconds, with a 95% bootstrap confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`.
+The estimated standard deviation of between-person differences in the valence effects is `r sd1[1]` log seconds, with a 95% CI of [`r sd1[2]`, `r sd1[3]`].
 
@@
 #| fig.cap: The heterogeneity distribution and various descriptions of the expected heterogeneity in valence effects from Model 1.
-\textbf{A}.
+**A**.
 The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
 Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean.
-\textbf{B}.
+**B**.
 The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area.
-\textbf{C}.
+**C**.
 Proportion of negative valence effects in the population (dark).
-\textbf{D}.
+**D**.
 Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).
@@
 
-While the interval and ratio descriptors allow determining where the population's slope values are likely to fall, they do so in absolute terms, such as log(rt) in the running example.
-One way to contextualize the magnitude of heterogeneity is to put it in relative terms.
-To do so, we can compare the standard deviation of the valence effect distribution to its mean.
+Although the interval and proportion descriptors show where the population's slope values are likely to fall, they do so in absolute terms, such as log(rt) in the running example.
+A contrasting, relative way to show heterogeneity is to express it as a ratio of the standard deviation to the mean.
+In our example this is the ratio of the standard deviation to the mean of the valence effect.
+
 [@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the ratio of the standard deviation to the average effect is 0.25 or greater, then heterogeneity can be deemed noteworthy.
@@
 We first redraw the expected heterogeneity distribution from Figure \@ref(fig:hdist-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:hdist-2)A.
-Superimposed to that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 random samples from the posterior distribution.
+Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 randomly sampled values from the joint posterior distribution of $\beta_1$ and $\tau_1$.
 From these curves we glean that the distribution of valence effects might well be narrower (i.e.
@@
 The distribution of these curves represents our uncertainty over the heterogeneity distribution of valence effects in the population.
-Thus, a good summary of heterogeneity that incorporates uncertainty must include information about uncertainty in both the scale (standard deviation) and location (mean) of the heterogeneity distribution.
+Thus, a good summary of heterogeneity that incorporates uncertainty must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
+
+```{r}
+set.seed(99)
+tmp <- post1 |> 
+  slice_sample(n = 4) |> 
+  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |> 
+  mutate(i = 1:n())
+
+tmp <- tmp |> 
+  mutate(
+    plot = map2(
+      b_valence1, sd_person__valence1, 
+      ~tibble(dist = dist_normal(.x, .y)) |> 
+        ggplot(aes(xdist = dist)) +
+        scale_x_continuous(
+          breaks = extended_breaks(3)
+        ) +
+        coord_cartesian(xlim = c(-0.4, 0.1), clip = "off") +
+        scale_y_continuous(
+          expand = expansion(c(0.0, 0.1))
+        ) +
+        stat_slab(slab_fill = "grey30") +
+        theme(
+          axis.title = element_blank(),
+          axis.ticks.x = element_line(size = rel(0.33)),
+          axis.ticks.y = element_blank(),
+          axis.text.y = element_blank(),
+          axis.text.x = element_text(size = rel(0.6)),
+          axis.line.y = element_blank()
+          )
+      )
+  )
+
+p1 <- post1 |> 
+  ggplot(aes(b_valence1, sd_person__valence1)) +
+  scale_x_continuous(
+    expand = expansion(.33)
+  ) +
+  scale_y_continuous(
+    expand = expansion(.33)
+  ) +
+  geom_point(shape = 1, col = "dodgerblue") +
+  geom_curve(
+    data = tmp |> 
+      mutate(y = sd_person__valence1),
+    aes(x = b_valence1, xend = 0,
+        y = y, yend = y),
+    alpha = .33, linewidth = .75, curvature = .2
+  ) +
+  theme(aspect.ratio = 1)
+
+p1 + 
+  inset_element(tmp$plot[[1]], 0.7, 0.1, 0.3, 0.28)
+```
 
@@
 #| fig.width: 4
-#| fig-cap: Posterior distribution of the heterogeneity distribution of valence effects (\textbf{A}) and its 90\% heterogeneity interval (\textbf{B}).
-\textbf{A}.
+#| fig-cap: Posterior distribution of the heterogeneity distribution of valence effects (**A**) and its 90\% heterogeneity interval (**B**).
+**A**.
 Posterior mean heterogeneity distribution in dark blue, and 100 random draws from the distribution's posterior distribution in light blue.
-\textbf{B}.
+**B**.
 100 random draws from the 90\% heterogeneity interval's posterior distribution.
@@
 
-To carry uncertainty forward from model parameters to their transformations, such as $HI_{90}$, we reproduce the calculations described above, but instead of using the point estimates we repeat the calculation over each posterior sample.
-In this manner, we obtain (samples from) posterior distributions of the parameter transformations, such as $HI_{90}$ (Figure \@ref(fig:hdist-2)B).
+To carry uncertainty forward from model parameters to their transformations, such as $HI_{90}$, we reproduce the calculations described above, but instead of using the typical (expected) values of the mean and standard deviation, we repeat the calculation for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
+In this manner, we get (samples from) posterior distributions of the parameter transformations, such as $HI_{90}$ (Figure \@ref(fig:hdist-2)B).
 
@@
 #| fig.width: 4
-#| fig-cap: \textbf{A}.
+#| fig-cap: **A**.
 100 random draws from the valence effect distribution's posterior distribution.
 Darker area under the curve represents the proportion of negative effects in the population for that posterior draw ($p^d$).
-\textbf{B}.
+**B**.
 Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI.
@@
 However, a sign test against zero is not always the most informative quantity because it discards all information about the magnitude of effects.
-A more theoretically motivated approach is to argue for and define a smallest effect size of interest (SESOI) which then defines a region of practical equivalence (ROPE) to zero, and then quantify uncertainty for the proposition that the effect is "practically equivalent to zero".
+A more theoretically motivated approach is to define a smallest effect size of interest (SESOI) which then defines a region of practical equivalence (ROPE) where the effect is deemed "practically equivalent to zero".
 Figure \@ref(fig:hdist-4)A shows 100 random draws of the heterogeneity distribution of valence effects, such that the area under the curve within a ROPE of [-0.1, 0.1] ($p^{ROPE}$) is filled with a darker color.
-That shaded area represents the posterior probability, for each draw, that the effect is practically equivalent to zero.
+That shaded area under the curve represents the posterior probability, for each random sample from the join posterior distribution of $\beta_1$ and $\tau_1$, that the effect is practically equivalent to zero.
 Figure \@ref(fig:hdist-4)B then shows a histogram of all 4,000 posterior draws of $p^{ROPE}$ with their mean and 95%CI.
@@
 #| fig.width: 4
-#| fig-cap: \textbf{A}.
+#| fig-cap: **A**.
 100 random draws from the valence effect distribution's posterior distribution.
 Dark area under the curve represents the proportion of effects within a region of practical equivalence to zero (ROPE) in the population for that posterior draw ($p^{ROPE}$).
-\textbf{B}.
+**B**.
 Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI.
@@
   scale_fill_manual(
-    values = c("dodgerblue1", "dodgerblue3"), 
+    values = c("dodgerblue1", "dodgerblue4"), 
     na.value = "white"
@@
 So far, these examples have highlighted the importance of quantifying uncertainty in measures of heterogeneity.
-Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^d$=`r percent(tmp$pd, .1)` and $p^{ROPE}$=`r percent(tmp$rope, .1)`.
-However, with 95% confidence, these values might be as great as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
+Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^d$ = `r percent(tmp$pd, .1)` and $p^{ROPE}$ = `r percent(tmp$rope, .1)`.
+However, with 95% confidence, these values might be as small as `r percent(tmp$pd.lower, .1)` and `r percent(tmp$rope.lower, .1)`, and as large as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
 
@@
 #| fig.width: 4
-#| fig-cap: \textbf{A}.
+#| fig-cap: **A**.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
-\textbf{B}.
+**B**.
 Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95%CI.

### 2023-09-07 ms.Rmd
Small fixes

@@
 For a 90% interval, we use the .05 and .95 percentiles (which define the central 90% of the distribution).
-Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1`, `r sd1`) = [`r hi90`].
+Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
 In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
@@
 #| fig.width: 4
-#| fig-cap: **A**.
+#| fig-cap: "**A**.
 100 random draws from the valence effect distribution's posterior distribution.
@@
 **B**.
-Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI.
+Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI."
 
@@
 **B**.
-Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI.
+Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI."
+
 p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
@@
 #| fig.width: 4
-#| fig-cap: **A**.
+#| fig-cap: Panel **A**.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
 **B**.
-Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95%CI.
+Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\\%CI.
+
 p1 <- post1 |> 

### 2023-09-07 ms.Rmd
Remove weird figure

@@
 ```{r}
-set.seed(99)
-tmp <- post1 |> 
-  slice_sample(n = 4) |> 
-  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |> 
-  mutate(i = 1:n())
-
-tmp <- tmp |> 
-  mutate(
-    plot = map2(
-      b_valence1, sd_person__valence1, 
-      ~tibble(dist = dist_normal(.x, .y)) |> 
-        ggplot(aes(xdist = dist)) +
-        scale_x_continuous(
-          breaks = extended_breaks(3)
-        ) +
-        coord_cartesian(xlim = c(-0.4, 0.1), clip = "off") +
-        scale_y_continuous(
-          expand = expansion(c(0.0, 0.1))
-        ) +
-        stat_slab(slab_fill = "grey30") +
-        theme(
-          axis.title = element_blank(),
-          axis.ticks.x = element_line(size = rel(0.33)),
-          axis.ticks.y = element_blank(),
-          axis.text.y = element_blank(),
-          axis.text.x = element_text(size = rel(0.6)),
-          axis.line.y = element_blank()
-          )
-      )
-  )
-
-p1 <- post1 |> 
-  ggplot(aes(b_valence1, sd_person__valence1)) +
-  scale_x_continuous(
-    expand = expansion(.33)
-  ) +
-  scale_y_continuous(
-    expand = expansion(.33)
-  ) +
-  geom_point(shape = 1, col = "dodgerblue") +
-  geom_curve(
-    data = tmp |> 
-      mutate(y = sd_person__valence1),
-    aes(x = b_valence1, xend = 0,
-        y = y, yend = y),
-    alpha = .33, linewidth = .75, curvature = .2
-  ) +
-  theme(aspect.ratio = 1)
-
-p1 + 
-  inset_element(tmp$plot[[1]], 0.7, 0.1, 0.3, 0.28)
-```
-
-```{r}
 #| label: hdist-2

### 2023-10-30 ms.Rmd
WIP
First draft of complete analyses.

@@
 ---
-title: 'Bayesian assessment of heterogeneity in psychological phenomena'
+title: 'On investigating heterogeneity in psychological phenomena'
 shorttitle: Heterogeneity in psychology
@@
     address: Tilburg School of Social and Behavioral Sciences, Tilburg University
-    email: m.j.vuorre@tilburguniversity.edu
+    email: mjvuorre@uvt.nl
   - name: Matthew Kay
@@
   - id: 1
-    institution: Department of Social Psychology, Tilburg University
+    institution: Department of Social Psychology, Tilburg School of Social and Behavioral Sciences, Tilburg University
   - id: 2
@@
   Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development.
-Yet, established methods and practices for describing heterogeneity underestimate the true uncertainty in those estimates.
+Yet, established methods and practices for studying heterogeneity underestimate the true uncertainty in those estimates.
 In the current work, we aim to deliver methods and guidelines for more rigorous treatments of heterogeneity in psychological phenomena.
@@
 Although other methods for incorporating estimation uncertainty are available, the methods we propose---involving bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
-  We hope that the proposed methods help researchers to widen their inferential focus from statistical averages to include statistical heterogeneity in psychological phenomena.
+We hope that the methods we discuss would help researchers to widen their inferential focus from statistical averages to include statistical heterogeneity in psychological phenomena.
 keywords: variation, heterogeneity, hierarchical models, mixed models, methodology
@@
   \noindent \textbf{This working paper is not yet peer reviewed.}
-wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
+wordcount: 5196
 bibliography: references.bib
@@
 library(knitr)
+library(readxl)
 library(latex2exp)
@@
 library(distributional)
+library(posterior)
 library(parameters)
@@
       strip.text = element_text(color = "black", hjust = 0),
-      panel.grid = element_blank(),
-      line = element_line(linewidth = .25)
+      strip.background = element_rect(color = NA, fill = NA),
+      strip.text.x = element_text(size = rel(0.9)),
+      line = element_line(linewidth = .25),
+      plot.tag = element_text(size = rel(1))
     )
@@
 options(
-  mc.cores = 4,
-  brms.backend = ifelse(require(cmdstanr), "cmdstanr", "rstan")
-)
-
+  mc.cores = 4
+)
+if (require(cmdstanr)) {
+  options(
+    brms.backend = "cmdstanr",
+    brms.threads = 2
+  )
+}  
 dir.create("data", FALSE)
@@
 One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still tend to analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019].
-Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to estimate and describe heterogeneity.
-In the current work, we propose several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
+Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to sufficiently estimate and describe heterogeneity.
+In the current work, we highlight several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
 
@@
 
-# Save variables as numerical & factor (might need for more complex contrast/eff/etc codes)
+# Save variables as numerical & factor
 dat <- read_csv("data/bolger-etal/heterogeneity_dataset1_traitvalence.csv") |>
@@
   mutate(
-    person = fct_anon(factor(id)),
+    person = factor(id),
     trial,
-    valence = factor(valenceE, levels = c(-0.5, 0.5), labels = c("Negative", "Positive")),
+    valence = factor(
+      valenceE,
+      levels = c(-0.5, 0.5),
+      labels = c("Negative", "Positive")
+    ),
     rt = logrt
@@
   arrange(person, trial)
-```
-
-```{r contrast-coding, include = FALSE}
+
 contrasts(dat$valence) <- c(-0.5, 0.5)
@@
     span_text_columns = FALSE,
-    digits = c(0,0,0,2),
+    font_size = "small",
+    digits = c(0, 0, 0, 2),
     caption = "First six rows of example dataset 1 (Bolger et al., 2019; Exp 1).",
@@
 The lack of subscripts on $\sigma$ indicates that we assume it to be invariant across trials and subjects.
-Then, we specify a model of the mean $\eta_{ij}$ such that the regression coefficients capture our substantive questions:
+(Note that there are better alternatives for modelling reaction times (RT) than modelling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specify a model of the mean $\eta_{ij}$ such that the regression coefficients capture our substantive questions:
 
@@
 This coding results in an intercept that corresponds to the average reaction time across negative and positive words, and a slope coefficient that reflects the difference in log(rt) between negative and positive words.
-The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$ are subject-specific deviations from the average intercept and slope, respectively.
-That is, $\beta_0 + \gamma_{01}$ is the average reaction time for subject $j=1$.
+The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$ are person-specific deviations from the average intercept and slope, respectively.
+That is, $\beta_0 + \gamma_{01}$ is the average reaction time for person $j=1$.
 The $\gamma$ coefficients have subscripts, which in this context indicates that they are further modelled on covariates---here, index variables for "person" in the data.
@@
 ```{r fit1-lmer}
-fit1_lmer <- lmer(
+fit_lmer <- lmer(
   rt ~ 1 + valence + (1 + valence | person),
@@
 # Create a table of coefficients
-tab1_lmer <- model_parameters(fit1_lmer, effects = "all") |> 
+tab_1 <- model_parameters(fit_lmer, effects = "all") |>
   tibble() |>
@@
 # Bootstrap CIs for variance parameters
-tab1_lmer[3:6,3:4] <- confint(fit1_lmer, method = "boot", nsim = 100)[c(1,3,2,4),1:2]
-```
-
-```{r fit1-brms}
-fit1 <- brm(
-  rt ~ 1 + valence + (1 + valence | person),
-  data = dat,
-  file = "models/brm-fit-1",
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r}
-#| label: tbl-fit-1-brms
-#| cache: false
-
-tab1_brms <- parameters(fit1, centrality = "mean", effects = "all") |> 
-  tibble() |> 
-  select(Parameter, Mean, starts_with("CI_"))
+tab_1[3:6, 3:4] <- confint(
+  fit_lmer,
+  method = "boot",
+  nsim = 100,
+  oldNames = FALSE
+)[c(1, 3, 2, 4), 1:2]
+
+# Get approx SEs
+tab_1 <- tab_1 |> 
+  mutate(SE = (CI_high - Coefficient) / 1.96) |> 
+  select(1, 2, 5, 3:4)
 ```
@@
 
-tab1_lmer |> 
-  mutate(
-    across(where(is.numeric), ~number(., .01)),
-    MLE = str_glue("{Coefficient} [{CI_low}, {CI_high}]"),
-  ) |> 
-  select(Parameter, `Estimate [95\\%CI]` = MLE) |> 
-  mutate(
+tab_1 |>
+  mutate(
+    mutate(across(where(is.numeric), ~number(.x, .01))),
+    CI = str_glue("[{CI_low}, {CI_high}]"),
+    Coefficient = ifelse(
+      str_starts(Coefficient, "-"), 
+      Coefficient, 
+      paste0("\\phantom{-}", Coefficient)
+    ),
     Parameter = m1_parnames
+  ) |> 
+  select(
+    Parameter, Coefficient, SE, CI
   ) |>
@@
     span_text_columns = FALSE,
+    font_size = "small",
     caption = "Parameter estimates from Model 1.",
@@
   )
-mu1 <- number(as.numeric(tab1_lmer[2,2:4,T]))
-sd1 <- number(as.numeric(tab1_lmer[4,2:4,T]))
-mu1n <- tab1_lmer[2,2,TRUE]
-sd1n <- tab1_lmer[4,2,TRUE]
+mu1 <- number(as.numeric(tab_1[2, 2:4, TRUE]))
+sd1 <- number(as.numeric(tab_1[4, 2:4, TRUE]))
+mu1n <- tab_1[2, 2, TRUE]
+sd1n <- tab_1[4, 2, TRUE]
 ```
@@
 hi90_low <- paste(
-  number(tab1_lmer[2,3,T] + qnorm(c(0.05, .95))*tab1_lmer[4,3,T], .01), 
+  number(
+    tab_1[2, 3, TRUE] + qnorm(c(0.05, .95)) * 
+      tab_1[4, 3, TRUE], .01
+  ),
   collapse = ", "
@@
 ```{r}
-#| label: hdist-1
-#| fig.height: 2.2
+#| label: fig-1
+#| fig.height: 2
 #| fig.width: 8
 #| fig.env: "figure*"
-#| fig.cap: The heterogeneity distribution and various descriptions of the expected heterogeneity in valence effects from Model 1.
-**A**.
+#| fig.align: "center"
+#| fig.cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
+A.
 The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
 Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean.
-**B**.
+B.
 The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area.
-**C**.
+C.
 Proportion of negative valence effects in the population (dark).
-**D**.
+D.
 Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).
 
-p0 <- ggplot() +
+fig_0 <- ggplot() +
   aes(xdist = dist_normal(mu1n, sd1n)) +
@@
 
-p1 <- p0 +
+fig_1 <- fig_0 +
   stat_slab(
@@
     label = TeX("$\\beta_1$"),
-    x = mu1n-.05, y = .075,
+    x = mu1n - .05, y = .075,
     size = 3,
@@
 
-p2 <- p0 + 
+fig_2 <- fig_0 +
   stat_slab(
@@
 
-p3 <- p0 + 
+fig_3 <- fig_0 +
   stat_slab(
@@
   geom_text(
-    label = TeX('$p^d$'),
+    label = TeX("$p^d$"),
     x = mu1n, y = .175,
@@
 
-p4 <- p0 + 
+fig_4 <- fig_0 +
   stat_slab(
@@
   geom_text(
-    label = TeX('$p^{ROPE}$'),
+    label = TeX("$p^{ROPE}$"),
     x = 0, y = .07,
@@
 
-(p1 | p2 | p3 | p4) &
+(fig_1 | fig_2 | fig_3 | fig_4) &
   scale_y_continuous(
@@
 
-The point estimates in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:hdist-1)A.
+The point estimates in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:fig-1)A.
 As such, Normal(`r mu1`, `r sd1`^2^) is a valid description of heterogeneity in valence effects.
@@
 In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
-We illustrate this interval in Figure \@ref(fig:hdist-1)B.
+We illustrate this interval in Figure \@ref(fig:fig-1)B.
 
@@
 In words, we have estimated that `r p_less_zero` of the population respond faster to positively valenced words.
-We show this area under the curve in Figure \@ref(fig:hdist-1)C.
+We show this area under the curve in Figure \@ref(fig:fig-1)C.
 
@@
 In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero (within the [-0.1, 0.1] interval).
-We visualize this area under the curve in Figure \@ref(fig:hdist-1)D.
+We visualize this area under the curve in Figure \@ref(fig:fig-1)D.
 
@@
 
-Although informative, the expected normal distribution of valence effects, or any transformation of those point estimates such as the ones illustrated in Figure \@ref(fig:hdist-1), ignores the uncertainty inherent in the estimated parameters, and therefore cannot communicate the analyst's uncertainty about the degree of heterogeneity.
+Although informative, the expected normal distribution of valence effects, or any transformation of those point estimates such as the ones illustrated in Figure \@ref(fig:fig-1), ignores the uncertainty inherent in the estimated parameters, and therefore cannot communicate the analyst's uncertainty about the degree of heterogeneity.
 That is, the uncertainty with which we have estimated the population parameters, indicated by e.g.
@@
 "bayesian") methods are uniquely applicable to estimating and describing heterogeneity in psychological phenomena in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
-Moreover, these methods are widely available in modern software packages; here we used the R package brms, which uses familiar R formulas for specifying bayesian models [@burknerAdvancedBayesianMultilevel2018; @burknerBrmsPackageBayesian2017; @rcoreteamLanguageEnvironmentStatistical2023].
+Moreover, these methods are widely available in modern software packages: For typical scenarios, probabilistic models are as easily estimated as their maximum likelihood counterparts.
+
+```{r fit1-brms}
+fit_1 <- brm(
+  rt ~ 1 + valence + (1 + valence | person),
+  data = dat,
+  file = "models/brm-fit-1",
+  control = list(adapt_delta = .95)
+)
+```
+
+```{r}
+#| label: ppcheck-1
+#| eval: false
+
+# Quick model checking
+pairs(fit_1$fit, pars = variables(fit_1)[1:6])
+mcmc_plot(fit_1, type = "rhat_hist")
+pp_check(fit_1)
+```
+
+```{r}
+#| label: tbl-fit-1-brms
+#| cache: false
+
+tab1 <- parameters(fit_1, centrality = "mean", effects = "all") |>
+  tibble() |>
+  select(Parameter, Mean, starts_with("CI_"))
+```
 
@@
 
-post1 <- as.data.frame(fit1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) |> 
+post1 <- as.data.frame(fit_1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) |>
   tibble() |>
   rownames_to_column("Sample")
-post1_sum <- fit1 |> 
+post1_sum <- fit_1 |>
   parameters(effects = "all", centrality = "mean", dispersion = TRUE) |>
@@
   mutate(
-    across(-Sample, ~number(., .01))
+    across(-Sample, ~ number(., .01))
   ) |>
@@
 Therefore, "estimate" means a large number of random samples from the posterior distribution.
-These samples can then be used to calculate and summarize the posterior distribution of any desired quantity.
-
-In practice, one obtains (e.g.) 4,000 samples from the posterior distribution, and then summarizes them using familiar data processing techniques.
-To illustrate, Table \@ref(tab:tbl-samples-1) shows six random samples from the parameters' posterior distributions; the two bottom rows show their means and standard deviations.
-Note that because we used the brms package's default noninformative prior distributions, summaries of the resulting posterior distributions are numerically almost identical to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
+These samples can then be used to calculate, summarize, and visualize the posterior distribution of any desired quantity.
+
+In practice, one obtains (e.g.) 4,000 samples from the posterior distribution using computational algorithms [e.g.
+@standevelopmentteamStanModelingLanguage2023] through accessible software [e.g., @burknerBrmsPackageBayesian2017], and then summarizes them using familiar data processing techniques [e.g.
+@rcoreteamLanguageEnvironmentStatistical2023].
+Here, we used the R package brms [@burknerBrmsPackageBayesian2017] to specify the model and then draw random samples from its posterior distribution [@standevelopmentteamStanModelingLanguage2023].
+The MCMC estimation algorithm completed in about 2 minutes on a modern laptop.
+We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
+
+Table \@ref(tab:tbl-samples-1) shows six random samples from Model 1's posterior distribution; the two bottom rows show their means and standard deviations.
+Note that because we used the brms package's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
 
@@
 Armed with the bayesian estimates (Table \@ref(tab:tbl-samples-1)), we now return to the distribution of valence effects in the population.
-We first redraw the expected heterogeneity distribution from Figure \@ref(fig:hdist-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:hdist-2)A.
+We first redraw the expected heterogeneity distribution from Figure \@ref(fig:fig-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:fig-2)A (thick dark blue curve).
 Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 randomly sampled values from the joint posterior distribution of $\beta_1$ and $\tau_1$.
@@
 
-Moreover, the location of these curves differ: Some curves are further to the left (average effect is more negative), and some further to the right (average effect is more positive).
-The distribution of these curves represents our uncertainty over the heterogeneity distribution of valence effects in the population.
-Thus, a good summary of heterogeneity that incorporates uncertainty must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
-
-```{r}
-#| label: hdist-2
-#| fig.height: 3.4
-#| fig.width: 4
-#| fig-cap: Posterior distribution of the heterogeneity distribution of valence effects (**A**) and its 90\% heterogeneity interval (**B**).
-**A**.
-Posterior mean heterogeneity distribution in dark blue, and 100 random draws from the distribution's posterior distribution in light blue.
-**B**.
-100 random draws from the 90\% heterogeneity interval's posterior distribution.
+<!-- TODO: Conceptually link this to Figure 1 Panel A -->
+
+```{r}
+#| label: calc-dfs
 
 set.seed(99)
-x <- seq(-0.55, 0.25, length = 301)
-
-calc_pdf <- function(posterior) {
+x <- seq(-0.6, 0.3, length = 301)
+
+calculate_density <- function(posterior, b, sd) {
   posterior |>
@@
     mutate(
-      pdf = dnorm(x, b_valence1, sd_person__valence1),
-      cdf = pnorm(x, b_valence1, sd_person__valence1)
+      pdf = dnorm(x, {{b}}, {{sd}}),
+      cdf = pnorm(x, {{b}}, {{sd}})
     )
@@
 
-dist_mean <- post1 |> 
+post1_density_mean <- post1 |>
   mean_qi(b_valence1, sd_person__valence1) |>
-  calc_pdf()
-
-dist_samples <- post1 |> 
+  calculate_density(b_valence1, sd_person__valence1)
+
+post1_density_samples <- post1 |>
   slice_sample(n = 100) |>
-  calc_pdf()
-
-p1 <- dist_mean |> 
-  ggplot(aes(x, pdf)) +
+  calculate_density(b_valence1, sd_person__valence1)
+
+post1_density_ribbon <- post1 |>
+  select(b_valence1, sd_person__valence1) |>
+  calculate_density(b_valence1, sd_person__valence1) |> 
+  group_by(x) |>
+  curve_interval(.width = 0.9)
+```
+
+```{r}
+#| label: person-specific-valence-effects
+#| eval: false
+
+# Just sketching a figure of person-specific valence slopes
+coef(fit_lmer)$person |> 
+  as.data.frame() |> 
+  rownames_to_column("person") |> 
+  left_join(
+    coef(fit_1)$person |> 
+      as.data.frame() |> 
+      rownames_to_column("person"),
+    by = join_by(person)
+) |> 
+  select(person, MLE = valence1, Bayes = Estimate.valence1) |> 
+  pivot_longer(-person) |> 
+  ggplot(aes(value, name)) +
+  geom_line(aes(group = person)) +
+  geom_point()
+summary(fit_1)
+```
+
+```{r}
+#| label: fig-2
+#| fig.height: 3
+#| fig.width: 8
+#| fig.env: "figure*"
+#| fig.align: "center"
+#| fig.cap: "Bayesian estimates of the heterogeneity distribution of valence effects.
+A.
+Probability density function (PDF) curves.
+The thick line illustrates the expected (point estimate; posterior mean) PDF of the valence effect heterogeneity distribution.
+Thin lines show 100 random samples of the PDF's posterior distribution, and facilitate visual communication of uncertainty in the PDF and, therefore, the valence effects' heterogeneity distribution.
+Dark line segments are inside the [-0.1, 0.1] ROPE, light segments are outside.
+The y-axis is in arbitrary probability density units at each point of the x-axis.
+B.
+The heterogeneity distribution's posterior mean PDF curve (line), and 90\\%CI (ribbon).
+C.
+Cumulative density function (CDF) curves, annotated as A.
+D.
+Histogram of the posterior distribution of $p^d$ (CDF segments below zero): This value describes the proportion of individuals with negative valence effects on logRT.
+Point and interval representes the posterior mean and 95\\%CI.
+E.
+As C, but represents uncertainty with a 95\\% confidence ribbon.
+F.
+Approximate posterior density of $p^d$."
+
+# Shared Y axis
+scale_y_01 <- scale_y_continuous(
+  breaks = extended_breaks(7),
+  expand = expansion(c(0.005, 0.005)),
+  limits = c(0, 1)
+)
+
+# Posterior samples of pdf and cdf curves
+p_cdf_samples <- post1_density_mean |>
+  ggplot() +
+  aes(x, cdf) +
+  ylab("Cumulative density") +
+  scale_color_manual(
+    values = c("dodgerblue1", "dodgerblue4")
+  ) +
   scale_x_continuous(
@@
   ) +
-  scale_y_continuous(
-    "Density",
-    expand = expansion(c(0.001, 0.05))
+  scale_y_01 +
+  geom_vline(
+    xintercept = 0, 
+    lty = "dotted", 
+    linewidth = .25
   ) +
@@
   ) +
+  theme(legend.position = "none")
+
+p_pdf_samples <- p_cdf_samples +
+  aes(y = pdf) +
   geom_line(
-    data = dist_samples,
+    data = post1_density_samples,
     aes(group = Sample),
@@
   ) +
+  scale_y_continuous(
+    breaks = extended_breaks(7),
+    expand = expansion(c(0.005, 0.005)),
+  ) +
+  ylab("Probability density") +
   theme(
-    axis.title.x = element_blank(), 
-    plot.margin = unit(c(4,4,16,4), "pt")
-  )
-p2 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
-  arrange(b_valence1) |> 
-  mutate(i = 1:n()) |> 
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1)
-  ) |> 
-  ggplot() +
+    axis.title.x = element_blank()
+  )
+
+p_cdf_samples <- p_cdf_samples +
+  geom_line(
+    data = post1_density_samples,
+    aes(group = Sample, color = after_stat(between(x, -0.1, 0.1))),
+    alpha = .33,
+    linewidth = .15
+  )
+
+# Posterior mean and 95%CI pdf and cdf curves
+p_cdf_ribbon <- post1_density_ribbon |>
+  ggplot(aes(x = x)) +
+  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper) +
   scale_x_continuous(
@@
   ) +
-  aes(xdist = dist, y = i) +
-  labs(y = "Sample") +
-  stat_pointinterval(
-    .width = .9,
-    interval_size_range = c(.05, .1),
-    alpha = .5,
-    color = "dodgerblue1"
-  )
-(p1 / p2) &
-  coord_cartesian(xlim = c(-0.55, 0.25)) &
+  scale_y_continuous(
+    breaks = extended_breaks(7),
+    expand = expansion(c(0.005, 0.005))
+  ) +
+  geom_vline(
+    xintercept = 0, 
+    lty = "dotted", 
+    linewidth = .25
+  ) +
+  geom_lineribbon(
+    linewidth = .5,
+    color = "dodgerblue4",
+    fill = alpha("dodgerblue1", .33)
+  ) +
   theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  ) &
-  plot_layout(heights = c(75, 25)) &
+    legend.position = "none",
+    axis.title.y = element_blank()
+  )
+
+p_pdf_ribbon <- p_cdf_ribbon +
+  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
+  theme(
+    axis.title.x = element_blank(), 
+  )
+
+# Add summaries of CDF < 0 to CDF curves as inset plots
+p_p0 <- post1 |>
+  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |>
+  ggplot() +
+  scale_x_continuous(
+    expand = expansion(c(0.0, 0.1))
+  ) +
+  scale_y_01 +
+  theme_void(base_size = 9) + 
+  theme(
+    plot.tag.position = c(.55, .75), 
+    plot.tag = element_text(size = rel(1))
+  )
+p_p0_samples <- p_p0 + 
+  stat_histinterval(
+    aes(y = cdf(dist, 0)),
+    breaks = 50,
+    fill = alpha("dodgerblue1", 0.5),
+    color = "dodgerblue4",
+    side = "right",
+    .width = .95,
+    size = 1,
+    scale = 0.5,
+    justification = 0.025
+  )
+p_p0_density <- p_p0 + 
+  stat_slabinterval(
+    aes(y = cdf(dist, 0)),
+    fill = alpha("dodgerblue1", 0.5),
+    color = "dodgerblue4",
+    side = "right",
+    .width = .95,
+    size = 1,
+    scale = 0.5,
+    justification = 0.025
+  )
+p_cdf_samples <- p_cdf_samples +
+  inset_element(
+    p_p0_samples,
+    0, 0, .2, 1, 
+    align_to = "panel", 
+    on_top = TRUE
+  )
+p_cdf_ribbon <- p_cdf_ribbon +
+  inset_element(
+    p_p0_density,
+    0, 0, .2, 1, 
+    align_to = "panel", 
+    on_top = TRUE
+  )
+
+p_cdf <- p_cdf_samples | p_cdf_ribbon
+
+p_pdf <- p_pdf_samples | p_pdf_ribbon
+
+(p_pdf / p_cdf) +
   plot_annotation(tag_levels = "A")
@@
 
+Moreover, the location of these curves differ: Some curves are further to the left (valence effect for the average person is more negative), and some further to the right (effect for the average person is more positive).
+The distribution of these curves represents our uncertainty over the heterogeneity distribution of valence effects in the population, and can be more completely summarized graphically as a ribbon: Figure \@ref(fig:fig-2)B shows the 95%CI of the valence effect's probability density for each value of the effect on the x-axis.
+A sufficient description of heterogeneity must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
+
+However, depicting the heterogeneity distribution visually as a probability density function (PDF) curve has its drawbacks.
+First, it appears to us visually more challenging to read the degree of uncertainty from a PDF, even when the underlying parameters' uncertainty is summarized as an interval (Figure \@ref(fig:fig-2)B).
+Second, for many applications, the y-axis is not informative: We do not care that the probability density of the curve is (e.g.) 3.0 at some specific value of the valence effect.
+
+Therefore, in Figure \@ref(fig:fig-2)C we depict the heterogeneity distribution as a *cumulative density* function (CDF).
+We again superimposed 100 random samples from the curve's posterior distribution on the posterior mean curve, and summarize the entire posterior distribution with a credibility ribbon in Figure \@ref(fig:fig-2)D.
+We believe the CDF is particularly useful as a descriptor of heterogeneity, because the y-axis specifically describes the proportion of population with valence effects below (or above) some specific value of the effect, as we highlight below.
+
 ### Interval descriptors
 
-To carry uncertainty forward from model parameters to their transformations, such as $HI_{90}$, we reproduce the calculations described above, but instead of using the typical (expected) values of the mean and standard deviation, we repeat the calculation for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
-In this manner, we get (samples from) posterior distributions of the parameter transformations, such as $HI_{90}$ (Figure \@ref(fig:hdist-2)B).
-
-```{r}
-p1s <- post1 |> 
+```{r}
+p1n <- post1 |>
   mutate(
@@
   ) |>
-  mean_qi(hi_low, hi_high) |> 
-  mutate(
-    across(where(is.numeric), ~number(., .01)),
+  mean_qi(hi_low, hi_high)
+p1s <- p1n |>
+  mutate(
+    across(where(is.numeric), ~ number(., .01)),
     l = str_glue("[{hi_low.lower}, {hi_low.upper}]"),
@@
 
-Summarising a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
-The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u`.
-Thus, to adequately describe an effects heterogeneity interval, researchers must communicate two separate credibility intervals.
-However, communicating these two credibility intervals is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
-Because of these limitations, the heterogeneity interval may not always be the most desirable description of heterogeneity.
+Above, we described the heterogeneity interval as a range of values where a specific percentage of the population's slopes are expected to fall (e.g.
+$HI_{90}$ for a 90% heterogeneity interval).
+However, a single interval cannot accommodate the uncertainty with which the underlying parameters are estimated.
+To carry uncertainty forward from model parameters to their transformations, such as the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
+Consequently, we get 4,000 samples from the $HI_{90}$s posterior distribution (Figure \@ref(fig:intervals)).
+
+Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
+The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (Figure \@ref(fig:intervals)B).
+Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.
+
+Figure \@ref(fig:intervals)A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
 
@@
 
-```{r}
-#| label: hdist-3
-#| fig.height: 2.2
-#| fig.width: 4
-#| fig-cap: "**A**.
-100 random draws from the valence effect distribution's posterior distribution.
-Darker area under the curve represents the proportion of negative effects in the population for that posterior draw ($p^d$).
-**B**.
-Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI."
-
-p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |>
-  arrange(b_valence1) |>
-  # slice(seq(1, nrow(.), by = 2)) |>
-  mutate(i = 1:n()) |> 
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1)
-  ) |> 
-  ggplot() +
-  scale_fill_manual(values = c("dodgerblue1", "dodgerblue3"), na.value = "white") +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0),
-    breaks = extended_breaks()
-  ) +
-  scale_y_continuous(
-    "Sample",
-    expand = expansion(0)
-  ) +
-  aes(xdist = dist, y = i) +
-  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
-  stat_slab(
-    aes(fill = after_stat(x < 0)), 
-    height = 15,
-    slab_color = "grey10",
-    slab_linewidth = .1,
-    alpha = .5,
-    p_limits = c(.01, .99),
-    show.legend = FALSE
-  )
-p2 <- post1 |> 
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1),
-    pd = cdf(dist, 0),
-    prope = percent(cdf(dist, 0.1) - cdf(dist, -0.1), .1)
-  ) |> 
-  ggplot(aes(pd)) +
-  scale_y_continuous(
-    "Count",
-    expand = expansion(0)
-  ) +
-  scale_x_continuous(
-    TeX('$p^d$'),
-    breaks = extended_breaks(),
-    labels = ~percent(., 1)
-  ) +
-  stat_histinterval(
-    breaks = 50,
-    outline_bars = TRUE,
-    slab_color = "gray10",
-    slab_fill = alpha("dodgerblue1", .33),
-    slab_linewidth = .2,
-    justification = .01,
-    point_color = "dodgerblue4", 
-    interval_color = "dodgerblue4",
-    .width = .90
-  )
-(p1 | p2) &
-  theme(
-    axis.ticks.y = element_blank(),
-    axis.text.y = element_blank()
-  ) &
-  plot_annotation(tag_levels = "A")
-```
+A complementary description of heterogeneity is the proportion of the population whose effects fall above or below some critical value.
+For example, we can calculate proportions with negative and positive effects by using zero as the critical value.
+In this example, we ask "What proportion of individuals in the population have negative effects of valence?"
 
@@
 
-An alternative and often more useful description of heterogeneity is the proportion of the population whose effects fall above or below some critical value.
-In many applications, zero is a useful critical value because it allows us to calculate proportions of individuals with negative and positive effects.
-Figure \@ref(fig:hdist-3)A visualizes 100 random draws of the heterogeneity distribution, such that the area under each curve below zero is filled in darker grey.
-That area represents the proportion of individuals who respond faster to positive words.
-
-Therefore, we can ask "What proportion of individuals in the population have a negative effect of valence?" To answer, we calculate the probability density below zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$, and obtain samples of this density (Figure \@ref(fig:hdist-3)A).
-Figure \@ref(fig:hdist-3)B then shows these samples' histogram, and the posterior mean and 95% Credibility Interval (`r tmp$pd_r`).
-In words, we expect about `r percent(as.numeric(tmp$pd), .1)` to be negative, but with 95% credibility, this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
+To answer, we calculate the cumulative probability density at zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$.
+We show 100 random samples of the CDF in Figure \@ref(fig:fig-2)C with a vertical line superimposed at zero.
+The y-axis value where each CDF crosses zero on the x-axis indicates the population proportion of negative valence effects.
+
+Critically, we also show a histogram of all 4,000 posterior samples of that proportion in the top left margin of Figure \@ref(fig:fig-2)C, with the associated 95%CI: According to the model, the proportion of individuals in the population with negative valence effects is approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)` (with 95% confidence).
 Alternatively [@bolgerCausalProcessesPsychology2019, p.605-606], the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
 
-However, a sign test against zero is not always the most informative quantity because it discards all information about the magnitude of effects.
-A more theoretically motivated approach is to define a smallest effect size of interest (SESOI) which then defines a region of practical equivalence (ROPE) where the effect is deemed "practically equivalent to zero".
-Figure \@ref(fig:hdist-4)A shows 100 random draws of the heterogeneity distribution of valence effects, such that the area under the curve within a ROPE of [-0.1, 0.1] ($p^{ROPE}$) is filled with a darker color.
-That shaded area under the curve represents the posterior probability, for each random sample from the join posterior distribution of $\beta_1$ and $\tau_1$, that the effect is practically equivalent to zero.
-Figure \@ref(fig:hdist-4)B then shows a histogram of all 4,000 posterior draws of $p^{ROPE}$ with their mean and 95%CI.
-We learn that `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
-
-```{r}
-#| label: hdist-4
-#| fig.height: 2.2
-#| fig.width: 4
-#| fig-cap: "**A**.
-100 random draws from the valence effect distribution's posterior distribution.
-Dark area under the curve represents the proportion of effects within a region of practical equivalence to zero (ROPE) in the population for that posterior draw ($p^{ROPE}$).
-**B**.
-Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI."
-
-p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
-  arrange(b_valence1) |> 
-  # slice(seq(1, nrow(.), by = 2)) |>
-  mutate(i = 1:n()) |> 
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1)
-  ) |> 
-  ggplot() +
-  scale_fill_manual(
-    values = c("dodgerblue1", "dodgerblue4"), 
-    na.value = "white"
-    ) +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0),
-    breaks = extended_breaks()
-  ) +
-  scale_y_continuous(
-    "Sample",
-    expand = expansion(0)
-  ) +
-  aes(xdist = dist, y = i) +
-  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
-  stat_slab(
-    aes(fill = after_stat(near(x, 0, .1))), 
-    height = 15,
-    slab_color = "grey10",
-    slab_linewidth = .1,
-    alpha = .5,
-    p_limits = c(.01, .99),
-    show.legend = FALSE
-  )
-p2 <- post1 |> 
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1),
-    pd = cdf(dist, 0),
-    prope = cdf(dist, 0.1) - cdf(dist, -0.1)
-  ) |> 
-  ggplot(aes(prope)) +
-  scale_y_continuous(
-    "Count",
-    expand = expansion(0)
-  ) +
-  scale_x_continuous(
-    TeX('$p^{ROPE}$'),
-    breaks = extended_breaks(),
-    labels = ~percent(., 1)
-  ) +
-  stat_histinterval(
-    breaks = 50,
-    outline_bars = TRUE,
-    slab_color = "gray10",
-    slab_fill = alpha("dodgerblue1", .33),
-    slab_linewidth = .2,
-    justification = .01,
-    point_color = "dodgerblue4", 
-    interval_color = "dodgerblue4",
-    .width = .90
-  )
-(p1 | p2) &
-  theme(
-    axis.ticks.y = element_blank(),
-    axis.text.y = element_blank()
-  ) &
-  plot_annotation(tag_levels = "A")
-```
-
-So far, these examples have highlighted the importance of quantifying uncertainty in measures of heterogeneity.
+We can use the posterior distribution of the valence effect's heterogeneity in other useful manners too.
+For example, if theory has helped us define a smallest effect size of interest (SESOI) to determine a range of magnitudes that are "practically equivalent to zero" (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically equivalent to zero valence effects.
+We colored the curves in Figure \@ref(fig:fig-2)C such that curve segments within [-0.1, 0.1] are highlighted with a darker color.
+Those darker line segments represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$).
+To quantify uncertainty in $p^{ROPE}$ we then aggregate the segments' to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
+We note that the ROPE of [-0.1, 0.1] here was arbitrary and picked just to illustrate the example.
+
+So far, these examples have highlighted the importance of quantifying uncertainty in descriptions of heterogeneity.
 Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^d$ = `r percent(tmp$pd, .1)` and $p^{ROPE}$ = `r percent(tmp$rope, .1)`.
-However, with 95% confidence, these values might be as small as `r percent(tmp$pd.lower, .1)` and `r percent(tmp$rope.lower, .1)`, and as large as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
+However, with 95% confidence, these values might be as small as `r percent(tmp$pd.lower, .1)` and `r percent(tmp$rope.lower, .1)`, or as large as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
 
@@
 Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, which we then transformed into 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ in Figure \@ref(fig:hdist-5)B.
-The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)`.
+The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
 
@@
 #| fig.width: 4
-#| fig-cap: Panel **A**.
+#| fig.cap: Panel A.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
-**B**.
-Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\\%CI.
+B.
+Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\%CI.
 
@@
     shape = 1,
-    col = "dodgerblue3",
+    col = "dodgerblue1",
     alpha = .25,
-    size = .5
+    size = .75
   )
@@
   scale_x_continuous(
-    TeX('$\\frac{\\tau_1}{\\beta_1}$'),
+    TeX("$\\frac{\\tau_1}{\\beta_1}$"),
     breaks = extended_breaks()
@@
   stat_histinterval(
-    breaks = 50,
-    outline_bars = TRUE,
-    slab_color = "gray10",
-    slab_fill = alpha("dodgerblue1", .33),
-    slab_linewidth = .2,
-    justification = .01,
-    point_color = "dodgerblue4", 
-    interval_color = "dodgerblue4",
-    .width = .90
+    breaks = 100,
+    fill = alpha("dodgerblue1", 0.5),
+    color = "dodgerblue4",
+    .width = .95,
+    size = 1,
+    justification = 0.01
   ) +
@@
 
-# Comparing heterogeneity between groups
-
-We have now established methods for assessing and describing between-person heterogeneity in an experimental effect in a way that keeps track and describes uncertainty in this variability.
-In the following example, we take this idea further by addressing the following question: Does heterogeneity differ between groups?
-These groups could be e.g.
-different demographic groups that participated in an experiment ("Is the valence effect more variable among younger vs.
-older adults?") or different research groups in a multi-site project.
-
-Accurately representing the uncertainty in model parameters is critical for addressing questions about differences: It is not sufficient to report that the standard deviation of valence effects was greater among one group compared to another.
-Instead, we need to quantify uncertainty in the differences, and the methods described above extend naturally to such questions.
-
-As an example, we randomly divided the participants in the example data to two Groups A and B.
-These could represent, for example, two demographic groups such as younger and older participants.
-We then modified Model 1 to estimate separate population-level parameters ($\beta$, $\tau$, $\rho$) for the groups, allowing us to examine not only whether the average persons differ between the groups, but also the extent to which the two groups differ in heterogeneity.
-
-```{r}
-set.seed(999)
-dat2 <- distinct(dat, person) |> 
-  mutate(
-    g = sample(
-      c("a", "b"), 
-      length(unique(dat$person)), 
-      replace = TRUE
-    ),
-    g = factor(g)
-  ) |> 
-  right_join(dat)
-contrasts(dat2$g) <- c(-0.5, 0.5)
-```
-
-```{r fit2-brms}
+Throughout these examples, we have seen that---in these example data---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects.
+While useful in its own right, next we see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing it across multiple populations.
+
+# Comparing heterogeneity
+
+Next, we move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of study units.
+To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2023 addressing differences in between person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment.
+In @mahVariabilitySubjectsFree2023's Experiment 3, 260 individuals studied a list of twenty target words.
+After a short break, participants then either freely recalled as many of the target words as they could (Free recall group, N = 123) or recalled as many target words as they could when prompted with related cue words (Cued recall group, N = 137).
+Thus, the Free and Cued recall tasks had different groups of participants but the same target words.
+We show a sample of their data in Table \@ref(tab:tab-dat2).
+
+```{r data-mah-lindsay-3, include = FALSE}
+# Experiment 3: task type is between-person
+path <- "data/mah-lindsay-experiment-3.zip"
+if (!file.exists(path)) {
+  download.file(
+    "https://files.ca-1.osf.io/v1/resources/pfhu9/providers/osfstorage/?zip=",
+    destfile = path
+  )
+  unzip(path, exdir = str_remove(path, ".zip"), junkpaths = TRUE)
+}
+
+dat <- read_excel(
+  "data/mah-lindsay-experiment-3/CuedVsFree_DRM_2022_Item_Data.xlsx"
+) |>
+  filter(!Exclude) |>
+  mutate(
+    person = factor(idnumber),
+    task = factor(test_type, levels = c("FR", "CR"), labels = c("Free", "Cued")),
+    accuracy
+  ) |>
+  select(person, task, target, accuracy) |>
+  # Two observations were miscoded
+  mutate(accuracy = if_else(accuracy == 2, 1, accuracy)) |> 
+  arrange(task, person, target)
+```
+
+With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2023 found that participants who completed the cued recall task showed greater heterogeneity in memory performance than did the free recall group: The Cued:Free recall variance ratio was 1.33 (with a [1.14, 1.54] 95% bootstrap interval).
+@mahVariabilitySubjectsFree2023 also confirmed this result by comparing multilevel models that did and didn't allow for different between-person variabilities in memory ability for the two groups, and found similar results across three different experiments.
+
+```{r tab-dat2}
+dat |>
+  slice(4:6, .by = task) |> 
+  apa_table(
+    span_text_columns = FALSE,
+    font_size = "small",
+    escape = FALSE,
+    digits = 0,
+    caption = "Six rows of example dataset 2 (Mah \\& Lindsay, 2023; Exp 3)."
+  )
+```
+
+Here, we reproduce and extend those analyses to illustrate how our graphical and numerical descriptions of heterogeneity generalize to describing differences in heterogeneity.
+We ask three questions about between-task differences in heterogeneity: (1) To what extent is memory performance more variable between people in the cued recall task compared to the free recall task?
+(2) To what extent is memory performance variability between target words different across cued recall and free recall tasks?
+And (3) How consistent is target word heterogeneity across the two tasks: Are target words associated with good memory performance in cued recall experiments the same words that are associated with good memory performance in free recall experiments?
+
+To answer these questions, we model the $i$th total recall accuracy in 1 to `r nrow(dat)`, of person $j$ in 1 to `r length(unique(dat$person))`, word $k$ in 1 to `r length(unique(dat$target))`, and task $m$ in {F (free recall), C (cued recall)} as Bernoulli distributed, where the probability of an accurate answer is determined by the rate parameter $\pi_{ijkm}$.
+As is common with generalized linear models, we model the rate parameter through a nonlinear link function.
+In this example, we use the cumulative normal density function ($\Phi$; "probit"), but other link functions are also available, such as the common logit.
+Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.
+
+\begin{align*}
+\tag{2}
+\text{Accuracy}_{ijkm} &\sim \operatorname{Bernoulli}\left(\pi_{ijkm}\right) \\
+\pi_{ijkm} &= \Phi\left(\eta_{ijkm}\right) \\
+\eta_{ijkm} &= \beta_{m} + \gamma_{jm} + \delta_{km} \\
+\gamma_{[m:F]} &\sim 
+  \operatorname{Normal}\left(0, \tau_{\gamma_{[m:F]}} \right) \\
+\gamma_{[m:C]} &\sim 
+  \operatorname{Normal}\left(0, \tau_{\gamma_{[m:C]}} \right) \\
+\begin{bmatrix} 
+  \delta_{[m:F]} \\ \delta_{[m:C]}
+\end{bmatrix} &\sim 
+  \operatorname{MVN}\left(
+  \begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
+  \begin{bmatrix} 
+  \tau_{\delta_{[m:F]}} &\rho_\delta \\ 
+  \rho_\delta &\tau_{\delta_{[m:C]}} 
+\end{bmatrix}\right).
+\end{align*}
+
+This Model 2 of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$).
+In addition, instead of coding the task type (free recall vs.
+cued recall) using predictor coding schemes such as contrast or dummy coding, we have index-coded task using subscripts $_{m:F}$ to stand for Free recall parameters, and $_{m:C}$ for parameters pertaining to the Cued recall task.
+This reparameterization is useful, because it allows us to quantify heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference.
+Notice also that we model $\gamma$ using to independent normal distributions, and $\delta$ with a multivariate normal distribution.
+Because different persons participated in the two tasks, we cannot assess whether participant-specific abilities are correlated across the tasks.
+But we can assess this for target items, which were common across the tasks.
+
+```{r fit-brms-ml3}
+#| message: false
+
+model <- bf(
+  accuracy ~ 0 + task + 
+    (1 | gr(person, by = task)) + 
+    (0 + task | target) 
+)
 fit2 <- brm(
-  bf(
-    rt ~ 0 + g / valence + (1 + valence | gr(person, by = g)),
-    sigma ~ 0 + g
-  ),
-  family = brmsfamily("gaussian", link = "identity", link_sigma = "identity"),
-  data = dat2,
-  file = "models/brm-fit-2",
+  model,
+  family = bernoulli("probit"),
+  data = dat,
+  silent = 0,
+  file = "models/brm-fit-ml3-target",
   control = list(adapt_delta = .95)
@@
 
-Table \@ref(tab:tbl-fit-2) shows the model's estimated parameters and the heterogeneity metrics $p^d$, $p^{ROPE}$ (based on a ROPE of [-0.1, 0.1]), and $\frac{\tau_1}{\beta_1}$.
-Using these estimates (samples from each parameter's posterior distribution), we can extend the calculations from above to calculate differences between groups.
-Because we randomly divided participants into two groups, there are no broad differences between groups.
-
-```{r}
-#| label: tbl-fit-2
-#| cache: false
+```{r}
+#| label: ppcheck-2
+#| eval: false
+
+# Quick model checking
+pairs(fit2$fit, pars = variables(fit2)[1:7])
+mcmc_plot(fit2, type = "rhat_hist")
+pp_check(fit2, type = "bars_grouped", group = "task", ndraws = 100)
+```
+
+We estimated the model as above by drawing 4,000 random samples from the posterior distribution [@burknerBrmsPackageBayesian2017].
+We then confirmed graphically and numerically that the estimation algorithm had converged, and that the model performed adequately using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
+We summarise the model's posterior distribution in Table \@ref(tab:tbl-samples-2).
+
+```{r}
+#| label: tbl-samples-2
+#| cache: true
 
 m2_parnames <- c(
-  "$\\beta_0$", 
-  "$\\beta_1$", 
-  "$\\tau_0$", 
-  "$\\tau_1$", 
-  "$\\rho$",
-  "$\\sigma$"
-)
-post2 <- gather_draws(fit2, `b_g.*`, `sd_.*`, `cor_.*`, `b_sigma.*`, regex = TRUE) |> 
-  ungroup() |> 
-  mutate(
-    g = str_extract(.variable, "g[a|b]") |> str_remove("g"),
-    .variable = .variable |> str_replace_all(":", "_") |> str_remove_all("([_|:]g[a|b])+")
-  )
-
-post2 <- post2 |> 
-  pivot_wider(names_from = .variable, values_from = .value) |> 
-  mutate(
-    dist = dist_normal(`b_valence1`, sd_person__valence1),
-    pd = cdf(dist, 0),
-    prope = cdf(dist, 0.1) - cdf(dist, -0.1),
-    ratio = abs(sd_person__valence1 / `b_valence1`)
-  ) |> 
-  select(-dist)
-
+  "$\\beta_{m:F}$", "$\\beta_{m:C}$",
+  "$\\tau_{\\gamma_{m:F}}$", "$\\tau_{\\gamma_{m:C}}$",
+  "$\\tau_{\\delta_{m:F}}$", "$\\tau_{\\delta_{m:C}}$",
+  "$\\rho_\\delta$"
+)
+
+post2 <- as_draws_df(
+  fit2, 
+  variable = c("b_", "sd_", "cor_"), 
+  regex = TRUE
+) |> 
+  select(-.iteration, -.chain) |> 
+  relocate(.draw, .before = b_taskFree)
+
+set.seed(191)
 post2 |>
-  pivot_longer(b:ratio) |> 
-  pivot_wider(names_from = g, values_from = value) |> 
-  rename(A = a, B = b) |> 
-  mutate(`B - A` = B - A) |> 
-  pivot_longer(c(A, B, `B - A`), names_to = "g") |> 
-  mutate(name = fct_inorder(name)) |> 
-  group_by(g, name) |> 
-  mean_qi(value) |> 
-  mutate(
-    r = if_else(
-      name %in% c("pd", "prope"),
-      str_glue("{percent(value, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]"),
-      str_glue("{number(value, .01)} [{number(.lower, .01)}, {number(.upper, .01)}]")
-      )
-  ) |> 
-  mutate(
-    name2 = rep(
-      c(
-        m2_parnames, 
-        "$p^d$", 
-        "$p^{ROPE}$", 
-        "$\\frac{\\tau_1}{\\beta_1}$"
-        ), 
-      times = 3
-    )
-  ) |> 
-  select(Parameter = name2, g, r) |> 
-  mutate(g = str_glue("Group {toupper(g)}")) |> 
-  pivot_wider(names_from = g, values_from = r) |> 
+  slice_sample(n = 6) |>
+  mutate(.draw = as.character(.draw)) |> 
+  add_row(
+    post2 |> 
+      summarise_draws(
+        Mean = ~mean(.),
+        SD = ~sd(.)
+      ) |> 
+      pivot_longer(c(Mean, SD), names_to = ".draw") |>
+      pivot_wider(names_from = variable, values_from = value)
+  ) |> 
+  setNames(c("Sample", m2_parnames)) |> 
   apa_table(
@@
     font_size = "scriptsize",
-    caption = "Parameter estimates from Model 2.",
-    escape = FALSE
-  )
-```
+    midrules = 6,
+    digits = 2,
+    caption = "Six random samples from Model 2's parameters' posterior distributions, and their summaries.",
+    escape = FALSE,
+    placement = "h"
+  )
+```
+
+```{r}
+#| label: calc-2-heterogeneity
+
+# Calculate all heterogeneity metrics and differences
+post2 <- post2 |> 
+  mutate(
+    # Persons
+    dist_person_free = dist_normal(b_taskFree, `sd_person__Intercept:taskFree`),
+    dist_person_cued = dist_normal(b_taskCued, `sd_person__Intercept:taskCued`),
+    pd_person_free = 1 - cdf(dist_person_free, 0),
+    pd_person_cued = 1 - cdf(dist_person_cued, 0),
+    ratio_person_free = abs(`sd_person__Intercept:taskFree` / b_taskFree),
+    ratio_person_cued = abs(`sd_person__Intercept:taskCued` / b_taskCued),
+    hi_low_free = b_taskFree + qnorm(0.05) * `sd_person__Intercept:taskFree`,
+    hi_high_free = b_taskFree + qnorm(0.95) * `sd_person__Intercept:taskFree`,
+    hi_low_cued = b_taskCued + qnorm(0.05) * `sd_person__Intercept:taskCued`,
+    hi_high_cued = b_taskCued + qnorm(0.95) * `sd_person__Intercept:taskCued`,
+    # Differences
+    b_difference = b_taskCued - b_taskFree,
+    sd_person_difference = 
+      `sd_person__Intercept:taskCued` - `sd_person__Intercept:taskFree`,
+    pd_person_difference = pd_person_cued - pd_person_free,
+    ratio_person_difference = ratio_person_cued - ratio_person_free,
+    hi_low_difference = hi_low_cued - hi_low_free,
+    hi_high_difference = hi_high_cued - hi_high_free,
+    # Targets
+    dist_target_free = dist_normal(b_taskFree, `sd_target__taskFree`),
+    dist_target_cued = dist_normal(b_taskCued, `sd_target__taskCued`),
+    pd_target_free = 1 - cdf(dist_target_free, 0),
+    pd_target_cued = 1 - cdf(dist_target_cued, 0),
+    ratio_target_free = abs(`sd_target__taskFree` / b_taskFree),
+    ratio_target_cued = abs(`sd_target__taskCued` / b_taskCued),
+    # Differences
+    sd_target_difference = 
+      `sd_target__taskCued` - `sd_target__taskFree`,
+    pd_target_difference = pd_target_cued - pd_target_free,
+    ratio_target_difference = ratio_target_cued - ratio_target_free,
+    .before = starts_with("cor_")
+  ) |> 
+  select(-starts_with("dist_"))
+
+post2_sum <- post2 |> 
+  mutate(
+    across(b_taskFree:b_taskCued, list(p = ~pnorm(.x)))
+  ) |> 
+  summarize_draws(
+    mean, ~quantile2(.x, probs = c(0.025, 0.975))
+  ) |> 
+  mutate(
+    x = str_glue("{number(mean, .01)} [{number(q2.5, .01)}, {number(q97.5, .01)}]")
+  )
+```
+
+## Comparing between-person heterogeneity across tasks
+
+Descriptively, we reproduced @mahVariabilitySubjectsFree2023's finding that participants' memory performance was more heterogeneous in the cued recall than in the free recall experiment (columns 4 and 5 in Table \@ref(tab:tbl-samples-2)).
+We show the relevant estimated quantities, and the implied heterogeneity distributions in Figure \@ref(fig:fig-2-person).
+
+The top panel of Figure \@ref(fig:fig-2-person)A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall).
+Notice that all quantities in this model are in the probit space ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` in the free and cued recall conditions, respectively, in the probit space, they can be converted to probabilities, in which the quantities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]`.
+However, this conversion does not work in general and we therefore restrict descriptions to probits in what follows.
+
+Second, and more importantly, the second panel in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks, and their difference.
+The model estimates the proportion of individuals who perform above chance to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
+
+Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of Figure \@ref(fig:fig-2-person)A shows that the degree of relative heterogeneity is identical across the two tasks.
+(We truncated the Ratio axis at [-10, 10] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
+The difference in ratios is very uncertain partly for the same reason.)
+
+Moreover, we depict the heterogeneity distributions---both as PDFs and CDFs---in Figure \@ref(fig:fig-2-person)B.
+These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks: First, we see that the majority of the free recall CDF is to the left of the zero (chance) line, indicating that the majority of this population is predicted to perform worse than chance.
+This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-chance performance only for `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
+Second, we see that the slope of the cued recall CDF is less steep than that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.
+
+```{r}
+#| label: fig-2-person-calc
+
+x <- seq(-2.2, 2.2, length = 201)
+
+tmp_person <- bind_rows(
+  "Free" = post2 |>
+    select(b_taskFree, `sd_person__Intercept:taskFree`) |>
+    calculate_density(b_taskFree, `sd_person__Intercept:taskFree`) |> 
+    select(x, pdf, cdf),
+  "Cued" = post2 |>
+    select(b_taskCued, `sd_person__Intercept:taskCued`) |>
+    calculate_density(b_taskCued, `sd_person__Intercept:taskCued`) |> 
+    select(x, pdf, cdf),
+  .id = "Task"
+)  |> 
+  group_by(x, Task) |>
+  curve_interval(.width = 0.9)
+```
+
+```{r}
+#| label: model-2-ratio
+#| include: false
+
+# Some ratios are weird
+post2 |> 
+  select(
+    starts_with("b_"), 
+    starts_with("sd_person"), 
+    starts_with("ratio_person"), 
+    -ends_with("_difference")
+  ) |> 
+  select(
+    contains("free")
+  ) |> 
+  filter(ratio_person_free > 10)
+```
+
+```{r}
+#| label: fig-2-person
+#| fig-width: 8
+#| fig-height: 3.6
+#| fig.env: "figure*"
+#| fig.cap: "Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
+A.
+Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
+Differences calculated as Cued - Free recall.
+B.
+Probability density (top) and cumulative density functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
+The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance."
+
+p2_pars_person <- post2 |> 
+  select(contains("b_") | contains("person")) |> 
+  pivot_longer(everything()) |> 
+  mutate(
+    name = str_to_lower(name) |> 
+      str_remove("_person") |> 
+      str_remove("_intercept:task") |> 
+      str_remove("task")
+  ) |> 
+  separate(name, c("type", "group")) |> 
+  mutate(
+    type = factor(
+      type, 
+      levels = c("b", "sd", "pd", "ratio"),
+      labels = c("Mean", "SD", "Pr(Above chance)", "Ratio")
+    ),
+    group = factor(
+      group,
+      levels = c("free", "cued", "difference"),
+      labels = c("Free", "Cued", "Difference")
+    ) |> fct_rev(),
+    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
+  ) |> 
+  ggplot(aes(value, group, col = group, fill = group)) +
+  scale_color_brewer(
+    "Recall task",
+    palette = "Dark2",
+    aesthetics = c("color", "fill"),
+    direction = -1
+  ) +
+  scale_y_discrete(
+    expand = expansion(c(0.1, 0.25))
+  ) +
+  scale_x_continuous(
+    "Parameter value",
+    expand = expansion(0.025)
+  ) +
+  geom_vline(
+    xintercept = 0, linewidth = .2, alpha = .25
+  ) +
+  stat_histinterval(
+    slab_alpha = .25,
+    breaks = 50,
+    normalize = "panels",
+    .width = .95,
+    size = 1,
+    justification = 0.01
+  ) +
+  facet_wrap("type", ncol = 1, scales = "free") +
+  theme(
+    legend.position = "none",
+    axis.title.y = element_blank()
+  )
+
+p2_cdf_person <- tmp_person |> 
+  ggplot(aes(x = x)) +
+  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
+  geom_vline(
+    xintercept = 0, linewidth = .2, alpha = .25
+  ) +
+  scale_color_brewer(
+    "Recall task",
+    palette = "Dark2",
+    aesthetics = c("color", "fill"),
+    direction = -1
+  ) +
+  scale_x_continuous(
+    TeX("$\\Phi(Pr(Correct))$"),
+    expand = expansion(0.01),
+    breaks = extended_breaks(5),
+    sec.axis = sec_axis(
+      name = "Proportion correct",
+      trans = ~pnorm(.), 
+      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
+      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
+    )
+  ) +
+  scale_y_continuous(
+    breaks = extended_breaks(7),
+    expand = expansion(c(0.005, 0.005))
+  ) +
+  ylab("CDF") +
+  geom_ribbon(
+    alpha = .2, col = NA
+  ) +
+  geom_line(
+    linewidth = .5
+  ) +
+  theme(
+    legend.position = "none",
+    panel.border = element_rect(fill = NA)
+  )
+
+p2_pdf_person <- p2_cdf_person %+% 
+  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
+  ylab("PDF")
+
+p2_cdf_person_inset <- post2 |> 
+  select(pd_person_free, pd_person_cued) |> 
+  pivot_longer(everything()) |> 
+  ggplot() +
+  scale_y_01 +
+  scale_x_continuous(
+    expand = expansion(0)
+  ) +
+  scale_color_brewer(
+    "Recall task",
+    palette = "Dark2",
+    aesthetics = c("color", "fill"),
+    direction = -1
+  ) +
+  stat_slab(
+    aes(y = value, fill = name),
+    side = "right",
+    .width = .95,
+    size = 1,
+    scale = 0.5,
+    alpha = .2,
+    justification = 0.025
+  ) +
+  stat_pointinterval(
+    aes(y = value, col = name),
+    side = "right",
+    .width = .95,
+    size = 1,
+    justification = 0.025
+  ) +
+  theme_void() +
+  theme(legend.position = "none")
+
+p2_cdf_person <- p2_cdf_person +
+  inset_element(
+    p2_cdf_person_inset,
+    0, 0, .2, 1, 
+    align_to = "panel", 
+    on_top = TRUE
+  )
+
+p2_pcdf_person <- p2_pdf_person / p2_cdf_person
+
+(p2_pars_person | wrap_elements(full = p2_pcdf_person)) +
+  plot_annotation(tag_levels = "A")
+```
+
+Finally, we return to the heterogeneity interval (HI).
+The $HI_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task.
+While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (`r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (`r post2_sum$x[post2_sum$variable=="hi_high_difference"]`).
+Moreover, studying Figure \@ref(fig:fig-2-person)B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
+
+## Comparing between-target heterogeneity across tasks
+
+Naturally, heterogeneity across people is the more theoretically important phenomenon for psychologists.
+However, examining and understanding heterogeneity in other sampled units can be both theoretically and methodologically important.
+Without weighing in on this difference too strongly, for the sake of our example analyses we next turn to our second and third questions regarding potential differences and consistency in between-target word heterogeneity.
+
+```{r}
+#| label: fig-2-target-calc
+
+x <- seq(-2.2, 2.2, length = 201)
+
+tmp_target <- bind_rows(
+  "Free" = post2 |>
+    select(b_taskFree, sd_target__taskFree) |>
+    calculate_density(b_taskFree, `sd_target__taskFree`) |> 
+    select(x, pdf, cdf),
+  "Cued" = post2 |>
+    select(b_taskCued, `sd_target__taskCued`) |>
+    calculate_density(b_taskCued, `sd_target__taskCued`) |> 
+    select(x, pdf, cdf),
+  .id = "Task"
+)  |> 
+  group_by(x, Task) |>
+  curve_interval(.width = 0.9)
+```
+
+```{r}
+#| label: fig-2-target
+#| fig.width: 8
+#| fig.height: 3.6
+#| fig.env: "figure*"
+#| fig.cap: "Estimated between-target word heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
+A.
+Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
+Differences calculated as Cued - Free recall.
+B.
+Probability density (top) and cumulative density functions (bottom) of the two tasks' heterogeneity distributions (green: free recall, red: cued recall).
+The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance.
+C.
+Posterior mean (dark), and 100 posterior draws (light) of the correlation between target words' proportions correct in the free (x-axis) and cued recall (y-axis) tasks.
+Ellipses indicate the 90th percentile of the bivariate normal distribution."
+
+p2_pars_target <- post2 |> 
+  select(contains("target")) |> 
+  rename_with(
+    ~str_to_lower(.) |> 
+      str_remove("target_") |> 
+      str_remove_all("_task")
+  ) |> 
+  pivot_longer(everything()) |> 
+  separate(name, c("type", "group"), extra = "merge") |> 
+  mutate(
+    type = factor(
+      type, 
+      levels = c("sd", "pd", "ratio", "cor"),
+      labels = c("SD", "Pr(Above chance)", "Ratio", "Cor(Free, Cued)")
+    ),
+    group = factor(
+      group,
+      levels = c("free", "cued", "difference", "free_cued"),
+      labels = c("Free", "Cued", "Difference", "Correlation")
+    ) |> fct_rev(),
+    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
+  ) |> 
+  ggplot(aes(value, group, color = group, fill = group)) +
+  scale_color_manual(
+    "Recall task",
+    values = c(rep("#7570B3", 2), "#D95F02", "#1B9E77"),
+    aesthetics = c("color", "fill")
+  ) +
+  scale_y_discrete(
+    expand = expansion(c(0.1, 0.25))
+  ) +
+  scale_x_continuous(
+    "Parameter value",
+    expand = expansion(0.1)
+  ) +
+  geom_vline(
+    xintercept = 0, linewidth = .2, alpha = .25
+  ) +
+  stat_histinterval(
+    breaks = 50,
+    normalize = "panels",
+    .width = .95,
+    size = 1,
+    slab_alpha = 0.2,
+    justification = 0.01
+  ) +
+  facet_wrap("type", ncol = 1, scales = "free") +
+  theme(
+    legend.position = "none",
+    axis.title.y = element_blank()
+  )
+
+p2_cdf_target <- tmp_target |> 
+  ggplot(aes(x = x)) +
+  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
+  geom_vline(
+    xintercept = 0, linewidth = .2, alpha = .25
+  ) +
+  scale_color_brewer(
+    "Recall task",
+    palette = "Dark2",
+    aesthetics = c("color", "fill"),
+    direction = -1
+  ) +
+  scale_x_continuous(
+    TeX("$\\Phi(Pr(Correct))$"),
+    expand = expansion(0.01),
+    breaks = extended_breaks(5),
+    sec.axis = sec_axis(
+      name = "Proportion correct",
+      trans = ~pnorm(.), 
+      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
+      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
+    )
+  ) +
+  scale_y_continuous(
+    breaks = extended_breaks(7),
+    expand = expansion(c(0.005, 0.005))
+  ) +
+  ylab("CDF") +
+  geom_ribbon(
+    alpha = .2, col = NA
+  ) +
+  geom_line(
+    linewidth = .5
+  ) +
+  theme(
+    legend.position = "none",
+    panel.border = element_rect(fill = NA)
+  )
+
+p2_pdf_target <- p2_cdf_target %+% 
+  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
+  ylab("PDF")
+
+p2_cdf_target_inset <- post2 |> 
+  select(pd_target_free, pd_target_cued) |> 
+  pivot_longer(everything()) |> 
+  ggplot() +
+  scale_y_01 +
+  scale_x_continuous(
+    expand = expansion(0)
+  ) +
+  scale_color_brewer(
+    "Recall task",
+    palette = "Dark2",
+    aesthetics = c("color", "fill"),
+    direction = -1
+  ) +
+  stat_slab(
+    aes(y = value, fill = name),
+    side = "right",
+    .width = .95,
+    size = 1,
+    scale = 0.5,
+    alpha = .2,
+    justification = 0.025
+  ) +
+  stat_pointinterval(
+    aes(y = value, col = name),
+    side = "right",
+    .width = .95,
+    size = 1,
+    justification = 0.025
+  ) +
+  theme_void() +
+  theme(legend.position = "none")
+
+p2_cdf_target <- p2_cdf_target +
+  inset_element(
+    p2_cdf_target_inset,
+    0, 0, .2, 1, 
+    align_to = "panel", 
+    on_top = TRUE
+  )
+
+p2_pcdf_target <-  p2_pdf_target / p2_cdf_target
+
+# Ellipses
+.ellipse <- function(data) {
+  data |>
+    mutate(
+      e = pmap(
+        list(
+          cor_target__taskFree__taskCued,
+          sd_target__taskFree,
+          sd_target__taskCued,
+          b_taskFree,
+          b_taskCued
+        ),
+        ~ ellipse(
+          x = ..1,
+          scale = c(..2, ..3),
+          centre = c(..4, ..5),
+          level = .90
+        ) |>
+          as_tibble()
+      )
+    ) |>
+    unnest(e)
+}
+samples <- as_draws_df(
+  fit2,
+  variable = c(
+    "b_",
+    "sd_",
+    "cor_"
+  ),
+  regex = TRUE
+)
+
+set.seed(99)
+ellipse_samples <- samples |>
+  slice_sample(n = 100) |>
+  .ellipse()
+
+ellipse_median <- samples |>
+  summarise_draws(median) |>
+  pivot_wider(names_from = variable, values_from = median) |>
+  .ellipse()
+
+# Target word coefficients, data and model
+dat2_target_coefs <- left_join(
+  dat |> 
+    summarise(
+      p = qnorm(mean(accuracy)),
+      .by = c(task, target)
+    ) |> 
+    pivot_wider(
+      names_from = task, values_from = p
+    ),
+  coef(fit2)$target |> 
+    as.data.frame() |> 
+    rownames_to_column("target") |> 
+    tibble() |> 
+    select(
+      target, 
+      .Free = Estimate.taskFree,
+      .Cued = Estimate.taskCued
+    )
+)
+
+lims <- c(-1.8, 1.8)
+
+p2_ellipse <- ellipse_median |>
+  ggplot(aes(x, y)) +
+  coord_cartesian(
+    ylim = lims,
+    xlim = lims
+  ) +
+  geom_vline(
+    xintercept = 0, linewidth = .1, linetype = "dashed"
+  ) +
+  geom_hline(
+    yintercept = 0, linewidth = .1, linetype = "dashed"
+  ) +
+  geom_abline(
+    linewidth = .1, linetype = "dashed"
+  ) +
+  scale_x_continuous(
+    TeX("$\\Phi(Pr(Correct))~Free~recall$"),
+    breaks = seq(-1.5, 1.5, by = .5),
+    sec.axis = sec_axis(
+      trans = ~ pnorm(.),
+      name = "Pr(Correct) Free recall",
+      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
+      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
+    )
+  ) +
+  scale_y_continuous(
+    TeX("$\\Phi(Pr(Correct))~Cued~recall$"),
+    breaks = seq(-1.5, 1.5, by = .5),
+    sec.axis = sec_axis(
+      trans = ~ pnorm(.),
+      name = "Pr(Correct) Cued recall",
+      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
+      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
+    )
+  ) +
+  geom_path(
+    data = ellipse_samples,
+    aes(group = .draw),
+    alpha = .075, color = "#7570B3"
+  ) +
+  geom_path(
+    color = "#7570B3"
+  ) +
+  # geom_segment(
+  #   data = dat2_target_coefs,
+  #   aes(x = Free, xend = .Free, y = Cued, yend = .Cued)
+  # ) +
+  # geom_point(
+  #   data = dat2_target_coefs,
+  #   aes(x = Free, y = Cued),
+  #   shape = 21, fill = "white"
+  # ) +
+  # geom_point(
+  #   data = dat2_target_coefs,
+  #   aes(x = .Free, y = .Cued),
+  #   color = "dodgerblue4"
+  # ) +
+  theme(
+    aspect.ratio = 1
+  )
+
+(wrap_elements(full = p2_pars_target) | wrap_elements(full = p2_pcdf_target) | p2_ellipse) +
+  plot_annotation(tag_levels = "A")
+```
+
+The results regarding differences in memory performance heterogeneity across target words were similar to those as observed regarding heterogeneity in people's memory performances.
+Figure \@ref(fig:fig-2-target)A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
+The interpretation of this difference is quantitatively similar to that observed about people above: Units of study exhibit greater memory task performance variability in the cued recall than in the free recall task.
+However, because this experiment used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
+
+Moreover, we observe similar differences in between-target word heterogeneity between the two tasks' as we did in the above analysis of between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than chance performance in the cued recall than in the free recall task.
+Yet, the ratio of the standard deviation to the mean again appeared identical across the two tasks.
+
+However, this analysis of target word heterogeneity afforded an additional piece of information: There was a clear positive correlation between free and cued recall memory performance (bottom panel of Figure \@ref(fig:fig-2-target)B, and Figure \@ref(fig:fig-2-target)C).
+The posterior mean and 95%CI of this correlation was `r post2_sum$x[post2_sum$variable=="cor_target__taskFree__taskCued"]`.
+What this correlation means substantively is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task.
+@bolgerCausalProcessesPsychology2019 found the conceptually similar result within a correlation of random effects that individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.
+
+Although outside our current scope, we note that these analyses might indicate exciting new avenues for this line of inquiry.
+One explanation for the between-person heterogeneity difference is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2023].
+Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (columns 6 and 7 in Table \@ref(tab:tbl-samples-2); top panel in Figure \@ref(fig:fig-2-target)A).
+Further, we observed across two kinds of study units (people, target words) that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
 
@@
 
-In the current work, we illustrated the use of practical descriptors of between-person heterogeneity in psychological phenomena.
+In the current work, we illustrated the use of practical descriptors of heterogeneity in psychological phenomena with examples drawn from the published literature in both social and cognitive psychology.
+Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019, who described the importance and available methods for examining heterogeneity in causal effects in psychology, and describe how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
+We believe our currently proposed methods to be particularly useful to researchers because they allow doing just that: Incorporating uncertainty in the analyses at every stage, from modelling to communicating resulting inferences.
+
+- Researchers already focusing on heterogeneity
+- awesome and my comments
+- coin flipping (standard deviations reported but not inferential quantities (e.g. pr positive))
+- in light of what we've presented they are not communicated in sufficient detail
+
 Moreover, we emphasized throughout that probabilistic modelling allows carrying uncertainty forward from model parameters to heterogeneity descriptors.
@@
 
+## Practical implementation
+
+We provide all the code supporting this manuscript at [todo].
+There, we implement the graphical and computational methods for investigating heterogeneity presented here in the R programming language [@rcoreteamLanguageEnvironmentStatistical2023].
+Moreover, we make use of the ecosystem of R packages for bayesian computation, including brms [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] and ggdist [@ggdist2023].
+
 ## Limitations
 
+- We assume gaussian distributions but see rouder
+
 ## Conclusion
 
+# Data and code availability
+
+All the code supporting this manuscript is available at <https://github.com/mvuorre/heterogeneity-uncertainty> and archived on Zenodo at [todo].
+We reused openly available data from @bolgerCausalProcessesPsychology2019 and @mahVariabilitySubjectsFree2023.
+
+# Author contributions
+
+<!-- https://casrai.org/credit/ -->
+
+Conceptualization: MV, NB\
+Formal Analysis: MV\
+Methodology: MV, NB, MK\
+Project Administration: MV\
+Software: MV, MK\
+Visualization: MV, MK, NB\
+Writing – Original Draft: MV\
+Writing – Review & Editing: MV, NB, MK
+
 # Competing interests
@@
 The author(s) declare no competing interests.
-
-# Author contributions
-
-Conceptualization: MV & NB    
-Methodology: MV   
-Software: MV    
-Validation: MV    
-Formal Analysis: MV   
-Writing -- Original Draft: MV & NB    
-Writing -- Review & Editing: MV & NB    
-Visualization: MV & NB   
-Project Administration: MV
 
@@
 :::
+
+<!-- Prepare appendix section -->
+\clearpage
+\setcounter{page}{1}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\renewcommand{\thefigure}{A\arabic{figure}}
+\onecolumn
+
+# Appendix
+
+```{r}
+#| label: intervals
+#| fig.height: 2.4
+#| fig.width: 8
+#| fig.cap: Bayesian estimates of the 90% heterogeneity interval of valence effects.
+A.
+Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$.
+B.
+Histograms of 4,000 samples of the HI90 lower (left) and upper (right) limits with their posterior means and 95%CIs as points and intervals.
+C.
+100 random samples from the $HI_{90}$ posterior distribution, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
+
+p1 <- post1 |>
+  mutate(
+    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+  ) |>
+  ggplot(aes(hi_low, hi_high)) +
+  labs(x = "HI90 lower limit", y = "HI90 upper limit") +
+  geom_point(
+    alpha = .25,
+    shape = 1,
+    size = 0.75,
+    col = "dodgerblue1"
+  ) +
+  theme(aspect.ratio = 1)
+
+p2 <- post1 |>
+  mutate(
+    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+  ) |>
+  select(Sample, hi_low, hi_high) |>
+  pivot_longer(-Sample) |>
+  ggplot(aes(value, group = name)) +
+  scale_y_continuous(
+    expand = expansion(c(0, 0.1))
+  ) +
+  scale_x_continuous(
+    "Valence effects"
+  ) +
+  stat_histinterval(
+    breaks = 100,
+    fill = alpha("dodgerblue1", .33),
+    color = "dodgerblue4",
+    .width = c(.95),
+    justification = .01
+  ) +
+  theme(
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank()
+  )
+
+p3 <- distinct(post1_density_samples, Sample, b_valence1, sd_person__valence1) |>
+  arrange(b_valence1) |>
+  mutate(i = 1:n()) |>
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1)
+  ) |>
+  ggplot() +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0.025),
+    breaks = extended_breaks()
+  ) +
+  scale_y_continuous(
+    expand = expansion(0.025)
+  ) +
+  labs(y = "Posterior sample") +
+  stat_pointinterval(
+    aes(xdist = dist, y = i),
+    .width = .9,
+    interval_size_range = c(.15, .3),
+    alpha = .5,
+    color = "dodgerblue1"
+  ) +
+  geom_pointinterval(
+    data = post1 |>
+      summarise(
+        across(c(b_valence1, sd_person__valence1), mean)
+      ) |>
+      mutate(
+        hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+        hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+      ),
+    aes(y = 50, x = b_valence1, xmin = hi_low, xmax = hi_high),
+    color = "dodgerblue4"
+  )
+
+(p1 | p2 | p3) +
+  # coord_cartesian(xlim = c(-0.6, 0.3)) &
+  # scale_x_continuous(
+  #   "Valence effect",
+  #   expand = expansion(0),
+  #   breaks = extended_breaks(7)
+  # )) +
+  # plot_layout(widths = c(6, 4)) +
+  plot_annotation(tag_levels = "A")
+```

### 2023-10-31 ms.Rmd
Add SD ratio to ex2

@@
       `sd_person__Intercept:taskCued` - `sd_person__Intercept:taskFree`,
+    sd_person_ratio = 
+      `sd_person__Intercept:taskCued` / `sd_person__Intercept:taskFree`,
     pd_person_difference = pd_person_cued - pd_person_free,
@@
 
-Second, and more importantly, the second panel in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+Second, and more importantly, the second row in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
 The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks, and their difference.

### 2023-11-01 ms.Rmd
Small additions

@@
 The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
+If we used rule of thumb of 0.25 in @bolgerCausalProcessesPsychology2019, with these results we could say *with confidence* that heterogeneity in valence effects is notable.
 
@@
 Throughout these examples, we have seen that---in these example data---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects.
-While useful in its own right, next we see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing it across multiple populations.
+We will next see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing it across multiple populations.
 
@@
 
-- We assume gaussian distributions but see rouder
+In our example analyses, we have brushed many important modelling decisions under the rug in order to focus on the main topic of heterogeneity.
+First, in example 1 we analyzed reaction times by simply taking log-transforming the RTs.
+Proper analyses of RTs benefit from models that make more realistic assumptions about the data generating process underlying a reaction time response, but here we necessarily excluded this complication for reasons of brevity.
+
+Our exposition and interpretation of heterogeneity relied on a critical assumption in line with standard practices in multilevel and generalized linear mixed modelling; that of (multivariate) normality of the person-level parameters.
+Assuming that random effects are normally distributed is a computationally and conceptually useful fiction, and we recognize that it is unlikely to hold exactly in real psychological phenomena.
+Excellent work by Haaf, Rouder, and colleagues have explored alternatives to continuous normal distributions of random effects [e.g., @haafDevelopingConstraintBayesian2017; @haafDonAccountingVariability2019].
 
 ## Conclusion
+
+We hope that the computational, graphical, and conceptual tools that we have discussed here prove useful to researchers interested in better understanding heterogeneity in psychological phenomena.
 

### 2023-11-10 ms.Rmd
An early complete draft.

--- 2023-10-31 ms.Rmd
@@
 authornote: |
-  \noindent \textbf{This working paper is not yet peer reviewed.}
-wordcount: 5196
+  \textbf{This working paper is not yet peer reviewed.}
+wordcount: "5,837"
 bibliography: references.bib
@@
     \usepackage{float}
-    \setlength{\parskip}{4pt}
+    \setlength{\parskip}{0.5em}
 output: 
@@
 library(readxl)
-library(latex2exp)
 library(lme4)
@@
 opts_chunk$set(
+  fig.width = 6,
+  fig.align = "center",
   eval = TRUE,
@@
 )
-if (require(cmdstanr)) {
+if ("cmdstanr" %in% installed.packages()[,"Package"]) {
   options(
@@
 
-# Introduction
-
 When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not.
-Yet, establishing that an independent variable affects a dependent variable on average may not be a sufficient description of the phenomenon if the treatment effect is highly variable from one individual to another.
+Yet, establishing that an independent variable affects a dependent variable may not be a sufficient description of the phenomenon if the treatment effect is variable from one individual to another.
 The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].
 
-One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still tend to analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019].
+One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still commonly analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019].
 Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to sufficiently estimate and describe heterogeneity.
-In the current work, we highlight several practical methods, descriptions, and visualizations that allow researchers to better estimate, interpret, and report heterogeneity in a manner that maximizes empirical results' impact on theory building.
-
-## Current work
-
-Our aim is to outline methods for studying and communicating heterogeneity in a manner that fully takes uncertainty into account.
-We first introduce an example dataset from social psychology and review established procedures for estimating and communicating the expected heterogeneity in causal effects in the population.
-We then outline useful summaries of heterogeneity that focus on informative numerical and graphical descriptions, and note how they are limited by established methods' focus on point estimates of model parameters.
-
-We then turn to our proposed method of interrogating and communicating causal effect heterogeneity using probabilistic methods.
-By estimating probability distributions over model parameters, these bayesian methods are uniquely suitable for examining heterogeneity and researchers' uncertainty about it.
-Moreover, advances in computational procedures and statistical software make these methods available to researchers with a minimal learning curve beyond established maximum likelihood based methods.
-
-In the following sections, we detail calculations and graphics that make use of bayesian methods in communicating descriptions of heterogeneity such that uncertainty in the model parameters and, subsequently, heterogeneity, is taken into account.
-We then further extend the practice and consequent benefits of probabilistic inference for causal effect heterogeneity to comparing heterogeneity between populations and across timepoints in further example analyses.
-
-We conclude by discussing additional benefits of probabilistic inference in psychology in general, and for multilevel models specifically.
-We emphasize that these benefits come with little or no cost to applied practitioners.
-We also discuss our findings regarding causal effect heterogeneity in light of this new perspective.
+
+When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023] which, in isolation, might be difficult to interpret.
+That is, although we are excited about the increased focus on heterogeneity, we offer that commonly used descriptions of it leave room for improvement.
+
+Because this focus on heterogeneity in psychological phenomena is relatively new, work remains to be done in developing models, descriptions, and reporting practices that maximize investigations' impact on theory development.
+Our goal is to contribute to that work.
+We describe and illustrate the use of numerical and graphical descriptions of heterogeneity that (1) Go beyond model parameters to describe heterogeneity in clear and actionable terms, and (2) Take uncertainty in model parameters into account.
+
+Our plan is as follows.
+We first review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology.
+We then describe additional ways in which model parameters can be interrogated and transformed to describe distributions of causal effects in the population.
+We then move beyond summarizing expected degrees of heterogeneity to describing distributions of plausible degrees of heterogeneity.
+Finally, we extend these methods to compare heterogeneity across different populations using an example dataset from cognitive psychology.
 
@@
 In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data (https://github.com/kzee/heterogeneityproject).
-We first cleaned the data as in @bolgerCausalProcessesPsychology2019, leaving `r number(nrow(dat), big.mark = ",")` trials from 59 participants that were endorsed as self-relevant.
+We first wrangled the data as in @bolgerCausalProcessesPsychology2019, which led to a sample of `r number(nrow(dat), big.mark = ",")` trials from 59 participants that were endorsed as self-relevant.
 (We present all our code in the online analysis supplement.) We show a sample of these data in Table \@ref(tab:tab-dat1).
-Then, we estimated the same statistical model with these data (Model 1; equations 1.1-1.3).
-First, we model the (log) reaction time of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*, which can differ between trials $i$ and individuals $j$) and standard deviation $\sigma$ (*sigma*):
+Then, we estimated the same statistical model (Model 1; equations 1.1-1.3).
+We modeled the log-transformed reaction time (logRT) of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials, as indicated by the lack of subscripts:
 
@@
 \tag{1.1}
-\text{RT}_{ij} \sim \operatorname{Normal}(\eta_{ij}, \sigma^2).
+\text{logRT}_{ij} \sim \operatorname{Normal}\left(\eta_{ij}, \sigma^2\right).
 \end{equation}
 
-The lack of subscripts on $\sigma$ indicates that we assume it to be invariant across trials and subjects.
-(Note that there are better alternatives for modelling reaction times (RT) than modelling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specify a model of the mean $\eta_{ij}$ such that the regression coefficients capture our substantive questions:
+(We note that there are better and more informative alternatives than modeling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specified a model of the mean of the logRT distribution ($\eta_{ij}$) such that the regression coefficients captured our substantive questions:
 
@@
 \tag{1.2}
-\eta_{ij} = \beta_0 + \gamma_{0j} + (\beta_1 + \gamma_{1j})\text{V}_{ij}.
+\eta_{ij} = \beta_0 + \gamma_{0j} + \left(\beta_1 + \gamma_{1j}\right)\text{V}_{ij}.
 \end{equation}
@@
 This equation includes two kinds of parameters: $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V), do not have subscripts.
-That means that they are not further modelled except by way of (optional) prior distributions with fixed hyperparameters.
-In line with existing conventions in bayesian inference and software, we refer to these as *population level* parameters.
-From the frequentist perspective, these terms are constants (i.e.
-do not potentially vary as functions of covariates) and are therefore sometimes "fixed".
-For all practical purposes, because we do not discuss prior distributions with constant hyperparameters, readers can equate the terms "fixed" and "population level" parameters in the current discussion.
-
-We contrast coded valence such that negative words are coded as -0.5, and positive words as 0.5.
-This coding results in an intercept that corresponds to the average reaction time across negative and positive words, and a slope coefficient that reflects the difference in log(rt) between negative and positive words.
-The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$ are person-specific deviations from the average intercept and slope, respectively.
+In the frequentist tradition, these parameters are considered constants---not modelled on covariates---and typically referred to as "fixed" parameters [e.g.
+@raudenbushHierarchicalLinearModels2002].
+The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$, have the subscript $j$ to indicate that they are person-specific deviations from the average intercept and slope, respectively.
 That is, $\beta_0 + \gamma_{01}$ is the average reaction time for person $j=1$.
-The $\gamma$ coefficients have subscripts, which in this context indicates that they are further modelled on covariates---here, index variables for "person" in the data.
-We refer to these as "group level" parameters, in line with conventions.
-In the frequentist discourse, these parameters are called "random" parameters [@gelmanDataAnalysisUsing2007].
-Again, readers can equate the term "group level" parameters with "random" parameters in the ongoing discussion.
-
-As is standard in multilevel modelling, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:
+In frequentist texts, these are typically called "random" parameters, because they are modeled as varying randomly according to a specified distribution.
+Following standard multilevel modeling assumptions, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:
 
@@
 \begin{bmatrix} 
-\gamma_{0} \\ \gamma_{1}
+  \gamma_0 \\ \gamma_1
 \end{bmatrix} \sim 
-\operatorname{MVN}(
+\operatorname{MVN}\left(
   \begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
-\begin{bmatrix} 
-\tau_{\gamma_{0}} &\rho_{\gamma_0\gamma_1} \\ 
-\rho_{\gamma_0\gamma_1} &\tau_{\gamma_{1}} 
-\end{bmatrix}
-).
+  \begin{pmatrix} 
+    \tau_0 &\rho \\ 
+    \rho &\tau_1 
+  \end{pmatrix}
+\right).
 \end{equation}
 
-In these equations, we assume that the subject-specific deviations have means of zero, standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
-Note that the $\gamma$s are zero-centered because the means are added to them in the above linear mixed-effects equation for $\eta_{ij}$.
-What these equations mean substantively is that the extent to which the effect of valence varies around the average effect ($\beta_1$) is estimated by the standard deviation $\tau_{\gamma_1}$.
-Moreover, $\rho$ indicates the extent to which individuals' average reaction times correlate with how much their reaction time is affected by valence.
-
-With data described in Table \@ref(tab:tab-dat1) loaded to R, we can estimate this model using standard maximum likelihood methods as implemented in the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].
+In this equation, we assume that the subject-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
+Perhaps confusingly, $\tau$s and $\rho$ are also sometimes called random effects because they describe random (co)variations of the person-specific effects.
+To be clear, despite this naming convention they are "fixed" features of the population, not of any one group or individual.
+
+What these equations mean substantively is that the extent to which the effect of valence on logRT varies around the average effect ($\beta_1$) is estimated by the standard deviation $\tau_1$.
+$\tau_0$, on the other hand, describes the standard deviation of the population of individuals' average logRTs across negatively and positively valenced words (intercepts).
+Moreover, $\rho$ indicates the extent to which individuals' average logRTs correlate with how much their logRTs are affected by valence.
+
+Finally, we contrast coded valence such that negative words were assigned -0.5, and positive words 0.5.
+This coding results in an intercept that corresponds to the average reaction time across negative and positive words, and a slope that reflects the difference in logRT between negative and positive words.
+
+With data described in Table \@ref(tab:tab-dat1), we can estimate this model using standard maximum likelihood methods as implemented in, for example, the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].
 
@@
 # Bootstrap CIs for variance parameters
-tab_1[3:6, 3:4] <- confint(
+path <- "models/lmer-fit1-bootstrap.rds"
+if (!file.exists(path)) {
+  boots <- confint(
     fit_lmer,
     method = "boot",
-  nsim = 100,
+    nsim = 1000,
     oldNames = FALSE
-)[c(1, 3, 2, 4), 1:2]
+  )
+  boots <- boots[c(5, 6, 1, 3, 4, 2), ]
+  tab_1[, 3:4] <- boots
+  saveRDS(tab_1, path)
+} else {tab_1 <- readRDS(path)}
 
@@
   mutate(SE = (CI_high - Coefficient) / 1.96) |> 
-  select(1, 2, 5, 3:4)
+  select(Parameter, Coefficient, SE, CI_low, CI_high)
 ```
@@
   )
-mu1 <- number(as.numeric(tab_1[2, 2:4, TRUE]))
-sd1 <- number(as.numeric(tab_1[4, 2:4, TRUE]))
 mu1n <- tab_1[2, 2, TRUE]
+mu1n_low <- tab_1[2, "CI_low", TRUE]
 sd1n <- tab_1[4, 2, TRUE]
+sd1n_low <- tab_1[4, "CI_low", TRUE]
+mu1 <- number(
+  as.numeric(tab_1[2, c("Coefficient", "CI_low", "CI_high"), TRUE]), 
+  .01
+)
+sd1 <- number(
+  as.numeric(tab_1[4, c("Coefficient", "CI_low", "CI_high"), TRUE]), 
+  .01
+)
 ```
@@
 We show a traditional summary of the estimated parameters from this model in Table \@ref(tab:tbl-fit-1).
-For the average person, the estimated effect of valence on RT is `r mu1[1]` log seconds, with a 95% bootstrap confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`.
-The estimated standard deviation of between-person differences in the valence effects is `r sd1[1]` log seconds, with a 95% CI of [`r sd1[2]`, `r sd1[3]`].
+For the average person, the estimated effect of valence on logRT is `r mu1[1]` log seconds, with a 95% bootstrap confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`.
+The estimated standard deviation of valence effects in the population is `r sd1[1]` log seconds, with a 95% CI of [`r sd1[2]`, `r sd1[3]`].
 
@@
 # Heterogeneity interval
-hi90_num <- mu1n + qnorm(c(0.05, .95)) * sd1n
+hi90_num <- qnorm(c(0.05, .95), mu1n, sd1n)
 hi90 <- paste(
@@
 hi90_low <- paste(
-  number(
-    tab_1[2, 3, TRUE] + qnorm(c(0.05, .95)) * 
-      tab_1[4, 3, TRUE], .01
-  ),
+  number(qnorm(c(0.05, .95), mu1n_low, sd1n_low), .01),
   collapse = ", "
@@
 p_rope <- percent(pnorm(0.1, mu1n, sd1n) - pnorm(-0.1, mu1n, sd1n), .1)
-
 # Ratio
 ratio <- number(abs(sd1n / mu1n), .01)
+ratio_low <- number(abs(sd1n_low / mu1n_low), .01)
 ```
@@
 #| label: fig-1
-#| fig.height: 2
-#| fig.width: 8
+#| fig.height: 1.8
 #| fig.env: "figure*"
@@
     label = TeX("$\\beta_1$"),
-    x = mu1n - .05, y = .075,
+    x = mu1n - .075, y = .075,
     size = 3,
@@
   geom_text(
-    label = TeX("$p^d$"),
+    label = TeX("$p^-$"),
     x = mu1n, y = .175,
@@
     label = TeX("$p^{ROPE}$"),
-    x = 0, y = .07,
+    x = 0.0, y = 0.5, hjust = 0,
     size = 3,
@@
 
-The point estimates in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:fig-1)A.
-As such, Normal(`r mu1`, `r sd1`^2^) is a valid description of heterogeneity in valence effects.
+Rows 2 and 4 in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:fig-1)A.
+In other words, our point estimate of the distribution of valence effects is Normal(`r mu1[1]`, `r sd1[1]`^2^).
 However, it is an incomplete description of heterogeneity for two reasons.
-First, it does not necessarily communicate the degree of heterogeneity in clear and actionable terms.
+First, it does not necessarily communicate the degree of heterogeneity in clear and actionable terms: In addition to the SD, we want quantities that describe heterogeneity in clearer terms.
+Below, we introduce several metrics that e.g.
+directly describe where a given proportion of the slopes are expected to fall.
 Second, it ignores the uncertainty with which the heterogeneity parameters are estimated.
+That is, we need ways to describe (uncertainty) distributions of distributions.
+
+We now turn to this first issue and outline several additional metrics of heterogeneity afforded by the model's parameter estimates.
 
@@
 
-Having estimated the parameters of the heterogeneity distribution, we now turn to using them to define alternative, more informative descriptions of the degree of heterogeneity in valence effects in the population.
-
-To begin with, and following @bolgerCausalProcessesPsychology2019, we used the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes a range within which a certain percentage of the population's slopes are expected to fall.
-By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$), but because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties.
+First, we can use the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes the range within which a certain percentage of the population's slopes are expected to fall [@bolgerCausalProcessesPsychology2019].
+To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$).
+However, because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties.
+The appropriate percentage to describe with a heterogeneity interval is determined by the substantive and communicative aims at hand; for our illustration 90% seemed reasonable.
 
 To calculate a heterogeneity interval, we first specify the desired probability limits.
-For a 90% interval, we use the .05 and .95 percentiles (which define the central 90% of the distribution).
-Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution.
+Then, we pass those probabilities and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to obtain the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
 In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
@@
 
-For many applications, quantities such as proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.
-For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" Thus, we first define a critical value, here zero.
-We then pass that value and the valence effect's estimated mean and standard deviation to the cumulative normal distribution function ($p$; `pnorm()` in R) to obtain the probability density in this distribution that falls below zero: $p^d = p(0, \beta_1, \tau_1) = p(0,$ `r mu1`, `r sd1`) = `r p_less_zero`.
-Specifically, this metric indicates the proportion of individuals in the population with negative effects.
-In words, we have estimated that `r p_less_zero` of the population respond faster to positively valenced words.
-We show this area under the curve in Figure \@ref(fig:fig-1)C.
-
-However, using zero as a critical value might not be sufficiently informative especially when theory allows defining a smallest effect sizes of interest, or what is known as a region of practical equivalence (to zero) [@anvariUsingAnchorbasedMethods2021; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014].
+In contrast, proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.
+For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" In other words, we ask what proportion of the heterogeneity distribution is below zero.
+To answer, we pass zero and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
+This number---$p^-$ for proportion of population with negative effects---is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
+We illustrate this probability in Figure \@ref(fig:fig-1)C.
+
+However, using zero as a critical value might not be sufficiently informative, especially when theory allows specifying a smallest effect size of interest, or what is known as a region of practical equivalence [ROPE, @anvariUsingAnchorbasedMethods2021; @lakensEquivalenceTestingPsychological2018; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014].
+In common applications, ROPE is used to statistically infer whether an estimated parameter, such as the effect of valence on logRT for the average person, is practically significant.
+But we can equally well use a theory-informed region of parameter values to describe and make inferences about the heterogeneity distribution of this effect in the population.
+
 For example, we might know from theory that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
-To calculate, we can again use the cumulative normal distribution function ($p$) to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = p(0.1, \beta_1, \tau_1) - p(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
-In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero (within the [-0.1, 0.1] interval).
-We visualize this area under the curve in Figure \@ref(fig:fig-1)D.
+To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
+In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero.
+Note that this statement's validity critically depends on the chosen interval, which we arbitrarily set as [-0.1, 0.1] in this example.
+We visualize this probability in Figure \@ref(fig:fig-1)D.
 
@@
 
-Although the interval and proportion descriptors show where the population's slope values are likely to fall, they do so in absolute terms, such as log(rt) in the running example.
-A contrasting, relative way to show heterogeneity is to express it as a ratio of the standard deviation to the mean.
-In our example this is the ratio of the standard deviation to the mean of the valence effect.
+Although the interval and proportion descriptors show where the population's slopes are likely to fall, they do so in absolute terms such as log(rt) in the running example.
+A contrasting or *relative* way to describe heterogeneity is to express it as a ratio of the standard deviation to the mean.
+The benefit of this descriptor is that in isolation a standard deviation says very little, but when directly compared to the mean effect, it allows for a sufficient description of the relative degree of heterogeneity of the phenomenon under study.
 
@@
 
-Although informative, the expected normal distribution of valence effects, or any transformation of those point estimates such as the ones illustrated in Figure \@ref(fig:fig-1), ignores the uncertainty inherent in the estimated parameters, and therefore cannot communicate the analyst's uncertainty about the degree of heterogeneity.
+Although informative, the expected normal distribution of valence effects and its transformations, such as the ones illustrated in Figure \@ref(fig:fig-1), ignore the uncertainty inherent in the estimated parameters.
+The point estimates alone cannot fully communicate the model's knowledge about the degree of heterogeneity.
 That is, the uncertainty with which we have estimated the population parameters, indicated by e.g.
-the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward to the calculations described above.
+the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward to the calculations described above or any description based on them.
+
 For example, we only presented the expected heterogeneity interval, but not the range of credible values that it might take.
-For instance, if we re-calculate $HI_{90}$ based on the lowest 95%CI bounds in Table \@ref(tab:tbl-fit-1), the result is a heterogeneity interval that indicates 90% of the population's slopes falling in the [`r hi90_low`] interval.
-In contrast to the first interval, this interval implies much less heterogeneity and that 90% of the population's slopes are expected to fall below zero.
+If we re-calculate $HI_{90}$ based on the lowest 95%CI bounds in Table \@ref(tab:tbl-fit-1), the resulting heterogeneity interval indicates that 90% of the population's slopes would fall in the [`r hi90_low`] interval.
+In contrast to the expected interval, this interval implies that 90% of the population's slopes are expected to fall below zero.
 
@@
 We think that probabilistic (i.e.
-"bayesian") methods are uniquely applicable to estimating and describing heterogeneity in psychological phenomena in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
-Moreover, these methods are widely available in modern software packages: For typical scenarios, probabilistic models are as easily estimated as their maximum likelihood counterparts.
+"bayesian") methods are uniquely able to address this challenge.
+Bayesian methods estimate and describe heterogeneity in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
+For typical scenarios, probabilistic models are as easy to use as their maximum likelihood counterparts [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] Moreover, many advanced models *require* bayesian methods [@muthenBayesianStructuralEquation2012; @asparouhovDynamicStructuralEquation2018].
 
@@
 
-In practice, one obtains (e.g.) 4,000 samples from the posterior distribution using computational algorithms [e.g.
+In practice, one obtains (for example) 4,000 samples from the posterior distribution using computational algorithms [e.g.
 @standevelopmentteamStanModelingLanguage2023] through accessible software [e.g., @burknerBrmsPackageBayesian2017], and then summarizes them using familiar data processing techniques [e.g.
 @rcoreteamLanguageEnvironmentStatistical2023].
-Here, we used the R package brms [@burknerBrmsPackageBayesian2017] to specify the model and then draw random samples from its posterior distribution [@standevelopmentteamStanModelingLanguage2023].
-The MCMC estimation algorithm completed in about 2 minutes on a modern laptop.
+Here, we used the R package brms [@burknerBrmsPackageBayesian2017] to specify the model and then draw random samples from its posterior distribution.
+The MCMC estimation algorithm completed in about 5 seconds on a modern laptop.
 We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
 
-Table \@ref(tab:tbl-samples-1) shows six random samples from Model 1's posterior distribution; the two bottom rows show their means and standard deviations.
-Note that because we used the brms package's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
+Table \@ref(tab:tbl-samples-1) shows six random samples from the posterior distribution of Model 1's population-level parameters ("fixed" parameters in the frequentist nomenclature).
+The two bottom rows show their means and standard deviations (which correspond to frequentist standard errors).
+Note that because we used brms's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
 
@@
 Armed with the bayesian estimates (Table \@ref(tab:tbl-samples-1)), we now return to the distribution of valence effects in the population.
+We now have 4,000 samples from this heterogeneity distribution's posterior distribution.
 We first redraw the expected heterogeneity distribution from Figure \@ref(fig:fig-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:fig-2)A (thick dark blue curve).
-Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 randomly sampled values from the joint posterior distribution of $\beta_1$ and $\tau_1$.
+Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 random posterior samples of $\beta_1$ and $\tau_1$.
 From these curves we glean that the distribution of valence effects might well be narrower (i.e.
 less heterogeneity) or wider (more heterogeneity) than is suggested by the point estimates.
-
-<!-- TODO: Conceptually link this to Figure 1 Panel A -->
 
@@
 #| fig.height: 3
-#| fig.width: 8
+#| fig.width: 5.4
 #| fig.env: "figure*"
-#| fig.align: "center"
 #| fig.cap: "Bayesian estimates of the heterogeneity distribution of valence effects.
@@
 D.
-Histogram of the posterior distribution of $p^d$ (CDF segments below zero): This value describes the proportion of individuals with negative valence effects on logRT.
+Histogram of the posterior distribution of $p^-$ (CDF segments below zero): This value describes the proportion of individuals with negative valence effects on logRT.
 Point and interval representes the posterior mean and 95\\%CI.
@@
 F.
-Approximate posterior density of $p^d$."
+Approximate posterior density of $p^-$."
 
@@
 Moreover, the location of these curves differ: Some curves are further to the left (valence effect for the average person is more negative), and some further to the right (effect for the average person is more positive).
-The distribution of these curves represents our uncertainty over the heterogeneity distribution of valence effects in the population, and can be more completely summarized graphically as a ribbon: Figure \@ref(fig:fig-2)B shows the 95%CI of the valence effect's probability density for each value of the effect on the x-axis.
+The distribution of these curves represents our current knowledge about the heterogeneity distribution of valence effects in the population---given these data and this model---and can be more completely summarized graphically as a ribbon: Figure \@ref(fig:fig-2)B shows the 95%CI of the valence effect's probability density for each value of the effect on the x-axis.
 A sufficient description of heterogeneity must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
@@
 However, depicting the heterogeneity distribution visually as a probability density function (PDF) curve has its drawbacks.
-First, it appears to us visually more challenging to read the degree of uncertainty from a PDF, even when the underlying parameters' uncertainty is summarized as an interval (Figure \@ref(fig:fig-2)B).
-Second, for many applications, the y-axis is not informative: We do not care that the probability density of the curve is (e.g.) 3.0 at some specific value of the valence effect.
+First, it appears to us visually more challenging to read the degree of uncertainty from a PDF, even when the underlying parameters' uncertainty is summarized as a ribbon around the posterior mean (Figure \@ref(fig:fig-2)B).
+Second, for many applications, the y-axis is not informative: We do not care that the probability density of the curve is (for example) 3.0 at some specific value of the valence effect.
 
 Therefore, in Figure \@ref(fig:fig-2)C we depict the heterogeneity distribution as a *cumulative density* function (CDF).
-We again superimposed 100 random samples from the curve's posterior distribution on the posterior mean curve, and summarize the entire posterior distribution with a credibility ribbon in Figure \@ref(fig:fig-2)D.
-We believe the CDF is particularly useful as a descriptor of heterogeneity, because the y-axis specifically describes the proportion of population with valence effects below (or above) some specific value of the effect, as we highlight below.
+We again superimposed 100 random samples from the curve's posterior distribution on the posterior mean, and summarize the entire posterior distribution with a 95% credibility ribbon in Figure \@ref(fig:fig-2)D.
+We believe the CDF is particularly useful as a descriptor of heterogeneity, because the y-axis describes a directly interpretable quantity: Namely, the proportion of population with valence effects below some specific value.
 
@@
 Figure \@ref(fig:intervals)A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
+For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2023] the scalar descriptors discussed below.
 
@@
 
-To answer, we calculate the cumulative probability density at zero of $\operatorname{Normal}(\beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$.
-We show 100 random samples of the CDF in Figure \@ref(fig:fig-2)C with a vertical line superimposed at zero.
-The y-axis value where each CDF crosses zero on the x-axis indicates the population proportion of negative valence effects.
-
-Critically, we also show a histogram of all 4,000 posterior samples of that proportion in the top left margin of Figure \@ref(fig:fig-2)C, with the associated 95%CI: According to the model, the proportion of individuals in the population with negative valence effects is approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)` (with 95% confidence).
-Alternatively [@bolgerCausalProcessesPsychology2019, p.605-606], the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
-
-We can use the posterior distribution of the valence effect's heterogeneity in other useful manners too.
-For example, if theory has helped us define a smallest effect size of interest (SESOI) to determine a range of magnitudes that are "practically equivalent to zero" (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically equivalent to zero valence effects.
+To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$.
+We show 100 posterior samples of the CDF in Figure \@ref(fig:fig-2)C with a vertical line superimposed at zero.
+The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$).
+We also show a histogram of all 4,000 posterior samples of that proportion in the top left margin of Figure \@ref(fig:fig-2)C, with the associated 95%CI.
+The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
+Alternatively, the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
+
+Moreover, if theory allows defining a range of parameter values that are practically equivalent to zero (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically negligible effects.
 We colored the curves in Figure \@ref(fig:fig-2)C such that curve segments within [-0.1, 0.1] are highlighted with a darker color.
@@
 So far, these examples have highlighted the importance of quantifying uncertainty in descriptions of heterogeneity.
-Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^d$ = `r percent(tmp$pd, .1)` and $p^{ROPE}$ = `r percent(tmp$rope, .1)`.
+Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^-$ = `r percent(tmp$pd, .1)` and $p^{ROPE}$ = `r percent(tmp$rope, .1)`.
 However, with 95% confidence, these values might be as small as `r percent(tmp$pd.lower, .1)` and `r percent(tmp$rope.lower, .1)`, or as large as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.
@@
 Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$.
-Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, which we then transformed into 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ in Figure \@ref(fig:hdist-5)B.
+Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (Figure \@ref(fig:hdist-5)B).
 The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
+If we used the 1/4 rule of thumb suggested in @bolgerCausalProcessesPsychology2019, with these results we could say *with confidence* that heterogeneity in valence effects is notable.
 
@@
 
-Throughout these examples, we have seen that---in these example data---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects.
-While useful in its own right, next we see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing it across multiple populations.
+We have seen that---with these example data and this model---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects.
+We will next see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing its magnitude across multiple populations.
 
@@
 
-Next, we move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of study units.
+Next, we move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across potentially multiple populations of different kinds of study units.
 To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2023 addressing differences in between person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment.
 In @mahVariabilitySubjectsFree2023's Experiment 3, 260 individuals studied a list of twenty target words.
-After a short break, participants then either freely recalled as many of the target words as they could (Free recall group, N = 123) or recalled as many target words as they could when prompted with related cue words (Cued recall group, N = 137).
+After a short break, they then either freely recalled as many of the target words as they could (Free recall group, N = 123) or recalled as many target words as they could when prompted with related cue words (Cued recall group, N = 137).
 Thus, the Free and Cued recall tasks had different groups of participants but the same target words.
-We show a sample of their data in Table \@ref(tab:tab-dat2).
+We show a sample of these data in Table \@ref(tab:tab-dat2).
 
@@
 
-Here, we reproduce and extend those analyses to illustrate how our graphical and numerical descriptions of heterogeneity generalize to describing differences in heterogeneity.
+We extend those analyses to illustrate how our descriptions of heterogeneity generalize to differences in heterogeneity in ways that inform audiences beyond variance ratios and model comparisons.
 We ask three questions about between-task differences in heterogeneity: (1) To what extent is memory performance more variable between people in the cued recall task compared to the free recall task?
@@
 As is common with generalized linear models, we model the rate parameter through a nonlinear link function.
-In this example, we use the cumulative normal density function ($\Phi$; "probit"), but other link functions are also available, such as the common logit.
+In this example, we use the cumulative normal density function ($\Phi$, or "probit"), but other link functions are also available, such as the common logit.
 Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.
@@
   \begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
-  \begin{bmatrix} 
+  \begin{pmatrix} 
     \tau_{\delta_{[m:F]}} &\rho_\delta \\ 
     \rho_\delta &\tau_{\delta_{[m:C]}} 
-\end{bmatrix}\right).
+  \end{pmatrix}
+\right).
 \end{align*}
@@
 
-We estimated the model as above by drawing 4,000 random samples from the posterior distribution [@burknerBrmsPackageBayesian2017].
+We estimated Model 2 exactly as Model 1, by drawing 4,000 random samples from its posterior distribution [@burknerBrmsPackageBayesian2017].
 We then confirmed graphically and numerically that the estimation algorithm had converged, and that the model performed adequately using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
@@
 The top panel of Figure \@ref(fig:fig-2-person)A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall).
-Notice that all quantities in this model are in the probit space ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` in the free and cued recall conditions, respectively, in the probit space, they can be converted to probabilities, in which the quantities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]`.
+Notice that the model's parameters refer to probits ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` probits in the free and cued recall conditions, respectively, the corresponding probabilities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued_p"]`.
 However, this conversion does not work in general and we therefore restrict descriptions to probits in what follows.
-
-Second, and more importantly, the second row in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
-The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
-The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks, and their difference.
-The model estimates the proportion of individuals who perform above chance to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
-
-Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of Figure \@ref(fig:fig-2-person)A shows that the degree of relative heterogeneity is identical across the two tasks.
-(We truncated the Ratio axis at [-10, 10] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
-The difference in ratios is very uncertain partly for the same reason.)
-
-Moreover, we depict the heterogeneity distributions---both as PDFs and CDFs---in Figure \@ref(fig:fig-2-person)B.
-These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks: First, we see that the majority of the free recall CDF is to the left of the zero (chance) line, indicating that the majority of this population is predicted to perform worse than chance.
-This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-chance performance only for `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
-Second, we see that the slope of the cued recall CDF is less steep than that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.
 
@@
 #| label: fig-2-person
-#| fig-width: 8
-#| fig-height: 3.6
+#| fig.height: 3.2
 #| fig.env: "figure*"
@@
 p2_pars_person <- post2 |> 
-  select(contains("b_") | contains("person")) |> 
+  select(contains("b_") | contains("person"), -sd_person_ratio) |> 
   pivot_longer(everything()) |> 
@@
 
-Finally, we return to the heterogeneity interval (HI).
+Second, and more importantly, the second row in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
+
+The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks ($p^+$), and their difference.
+The model estimates the proportion of individuals who perform above chance to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
+Notice that the this quantity refers to population proportions and is not a z-score.
+
+Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of Figure \@ref(fig:fig-2-person)A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
+(We truncated the Ratio axis at [-10, 10] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
+The difference in ratios is very uncertain partly for the same reason.)
+
+We also depict the heterogeneity distributions as PDFs and CDFs in Figure \@ref(fig:fig-2-person)B.
+These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks: First, we see that the majority of the free recall CDF (green) is to the left of zero (chance), indicating that the majority of this population is predicted to perform worse than chance.
+This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-chance performance only for a proportion of `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
+Second, we see that the slope of the cued recall CDF (red) is less steep than that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.
+
+Finally, we turn to the heterogeneity interval (HI).
 The $HI_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task.
-While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (`r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (`r post2_sum$x[post2_sum$variable=="hi_high_difference"]`).
-Moreover, studying Figure \@ref(fig:fig-2-person)B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
+While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $HI_90$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
+Studying Figure \@ref(fig:fig-2-person)B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
 
@@
 
-Naturally, heterogeneity across people is the more theoretically important phenomenon for psychologists.
-However, examining and understanding heterogeneity in other sampled units can be both theoretically and methodologically important.
-Without weighing in on this difference too strongly, for the sake of our example analyses we next turn to our second and third questions regarding potential differences and consistency in between-target word heterogeneity.
+Between-person heterogeneity is typically the more theoretically important phenomenon for psychologists than differences in model parameters between other randomly sampled study units, such as stimuli.
+However, examining heterogeneity in other sampled units can be both theoretically and methodologically important.
+Recognizing this difference, and for the sake of our example analyses we next turn to our second and third questions regarding potential differences and consistencies in between-target word heterogeneity.
 
@@
 #| label: fig-2-target
-#| fig.width: 8
-#| fig.height: 3.6
+#| fig.height: 4
+#| fig.width: 7.4
 #| fig.env: "figure*"
@@
 
-lims <- c(-1.8, 1.8)
-
 p2_ellipse <- ellipse_median |>
@@
   coord_cartesian(
-    ylim = lims,
-    xlim = lims
+    xlim = c(-1.8, 1.5),
+    ylim = c(-1.5, 1.8)
   ) +
@@
 Figure \@ref(fig:fig-2-target)A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
-The interpretation of this difference is quantitatively similar to that observed about people above: Units of study exhibit greater memory task performance variability in the cued recall than in the free recall task.
-However, because this experiment used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
-
-Moreover, we observe similar differences in between-target word heterogeneity between the two tasks' as we did in the above analysis of between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than chance performance in the cued recall than in the free recall task.
-Yet, the ratio of the standard deviation to the mean again appeared identical across the two tasks.
-
-However, this analysis of target word heterogeneity afforded an additional piece of information: There was a clear positive correlation between free and cued recall memory performance (bottom panel of Figure \@ref(fig:fig-2-target)B, and Figure \@ref(fig:fig-2-target)C).
+The interpretation of this difference is quantitatively similar to that observed about people above: Both people and target words exhibit greater memory task performance variability in the cued recall than in the free recall task.
+However, because @mahVariabilitySubjectsFree2023 used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
+
+Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than chance performance in the cued recall than in the free recall task.
+Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared nearly identical across the two tasks.
+
+This analysis of target word heterogeneity afforded an additional piece of information, because the same target words were used across the two tasks, and thus allowed answering our third question.
+There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of Figure \@ref(fig:fig-2-target)A, and Figure \@ref(fig:fig-2-target)C).
 The posterior mean and 95%CI of this correlation was `r post2_sum$x[post2_sum$variable=="cor_target__taskFree__taskCued"]`.
 What this correlation means substantively is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task.
-@bolgerCausalProcessesPsychology2019 found the conceptually similar result within a correlation of random effects that individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.
-
-Although outside our current scope, we note that these analyses might indicate exciting new avenues for this line of inquiry.
-One explanation for the between-person heterogeneity difference is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2023].
-Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (columns 6 and 7 in Table \@ref(tab:tbl-samples-2); top panel in Figure \@ref(fig:fig-2-target)A).
-Further, we observed across two kinds of study units (people, target words) that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
+@bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.
+
+Although outside our current scope, we note that our study of Model 2's results might indicate exciting new avenues for this line of inquiry.
+One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2023].
+Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (Figure \@ref(fig:fig-2-target)A).
+Moreover, we observed across two kinds of study units (people, target words) that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
 
@@
 
-In the current work, we illustrated the use of practical descriptors of heterogeneity in psychological phenomena with examples drawn from the published literature in both social and cognitive psychology.
-Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019, who described the importance and available methods for examining heterogeneity in causal effects in psychology, and describe how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
+In the current work, we illustrated the use of practical descriptors of heterogeneity in psychological phenomena with examples drawn from social and cognitive psychology.
+Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects in psychology---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
 We believe our currently proposed methods to be particularly useful to researchers because they allow doing just that: Incorporating uncertainty in the analyses at every stage, from modelling to communicating resulting inferences.
 
-- Researchers already focusing on heterogeneity
-- awesome and my comments
-- coin flipping (standard deviations reported but not inferential quantities (e.g. pr positive))
-- in light of what we've presented they are not communicated in sufficient detail
-
-Moreover, we emphasized throughout that probabilistic modelling allows carrying uncertainty forward from model parameters to heterogeneity descriptors.
-The resulting metrics are useful because they not only convey the analyst's expectation of the magnitude of heterogeneity, but also her uncertainty in the description.
-In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, allows easily comparing these heterogeneity descriptors across groups.
-
-We believe that psychology should take a meta-theoretic step forward and try to incorporate an understanding of heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019].
-In order to do so, descriptions of heterogeneity must include measures of uncertainty.
-
-## Practical implementation
-
-We provide all the code supporting this manuscript at [todo].
-There, we implement the graphical and computational methods for investigating heterogeneity presented here in the R programming language [@rcoreteamLanguageEnvironmentStatistical2023].
-Moreover, we make use of the ecosystem of R packages for bayesian computation, including brms [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] and ggdist [@ggdist2023].
+We emphasized throughout that a probabilistic ("bayesian") approach is uniquely positioned to answer the needs of researchers interested in heterogeneity.
+Bayesian methods allow carrying uncertainty forward from model parameters to descriptors of heterogeneity and beyond.
+The resulting metrics are useful because they not only convey analysts' expectations regarding heterogeneity, but more fully convey their states of knowledge regarding heterogeneity, including degrees of certainty.
+In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
+We further elucidate these points in our Online Supplementary Analysis [todo].
+
+We believe that psychology, broadly speaking, is methodologically and theoretically at a ripe position to taking a meta-theoretic step forward and try to incorporate an understanding of heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019].
+In order to do so, descriptions of heterogeneity must include measures of uncertainty, and we hope the techniques illustrated here help researchers do so.
 
@@
 
-- We assume gaussian distributions but see rouder
+In our example analyses, we have brushed many important modelling decisions under the rug in order to focus on the main topic of heterogeneity.
+First, in example 1 we analyzed reaction times by simply log-transforming reaction times.
+More informative analyses of RTs would make use of models that make more realistic assumptions about the data generating process underlying reaction time responses, but here we necessarily excluded this complication for reasons of brevity.
+
+Our exposition and interpretation of heterogeneity relied on a critical assumption in line with standard practices in multilevel and generalized linear mixed modelling; that of (multivariate) normality of the group-level parameters.
+Assuming that random effects are normally distributed is a computationally and conceptually useful fiction, and we recognize that it is unlikely to hold exactly in real psychological phenomena.
+Haaf, Rouder, and colleagues have explored alternatives to continuous normal distributions of random effects [e.g., @haafDevelopingConstraintBayesian2017; @haafDonAccountingVariability2019].
 
 ## Conclusion
+
+We hope that the conceptual, computational, and graphical tools that we have discussed here prove useful to researchers interested in better understanding heterogeneity in psychological phenomena.
 
@@
 #| fig.height: 2.4
-#| fig.width: 8
 #| fig.cap: Bayesian estimates of the 90% heterogeneity interval of valence effects.

### 2024-02-05 ms.Rmd
Use .Renviron for mcmc config

@@
 library(knitr)
+library(cmdstanr)
 library(readxl)
@@
 dir.create("models", FALSE)
+MAX_CORES <- as.numeric(Sys.getenv("MAX_CORES"), 4)
 options(
-  mc.cores = 4
-)
-if ("cmdstanr" %in% installed.packages()[,"Package"]) {
-  options(
-    brms.backend = "cmdstanr",
-    brms.threads = 2
-  )
-}  
-dir.create("data", FALSE)
+  brms.backend = Sys.getenv("BRMS_BACKEND", "rstan"), 
+  brms.threads = MAX_CORES %/% 4,
+  mc.cores = MAX_CORES
+)
 ```
@@
 ```{r data-bolgeretal, include = FALSE}
+dir.create("data", FALSE)
 # Load, save, & clean example data from Bolger et al. (2019)
@@
   \begin{pmatrix} 
-    \tau_0 &\rho \\ 
+    \tau_0 & \\ 
     \rho &\tau_1 
@@
   \begin{pmatrix} 
-    \tau_{\delta_{[m:F]}} &\rho_\delta \\ 
+    \tau_{\delta_{[m:F]}} & \\ 
     \rho_\delta &\tau_{\delta_{[m:C]}} 

### 2024-03-21 ms.Rmd
Comment Grice et al. (2020)

@@
 Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects in psychology---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
-We believe our currently proposed methods to be particularly useful to researchers because they allow doing just that: Incorporating uncertainty in the analyses at every stage, from modelling to communicating resulting inferences.
+
+Although prior work on developing metrics of heterogeneity, and placing experimental effect sizes in context of person-specific effects exists, it mostly has ignored estimation uncertainty and thus remained purely descriptive.
+For example, @gricePersonsEffectSizes2020 describe a method whereby analysts simply count the number of individuals whose point estimate of an effect is concordant with a hypothesis.
+But such counting ignores estimation uncertainty in both the person-specific effects and variability among them.
+By accounting for these uncertainties, the methods described here go beyond description and allow inference to be drawn regarding populations and individuals with confidence.
 

### 2024-04-09 ms.Rmd
Convert chunk options

@@
 
-```{r setup, include = FALSE}
+```{r}
+#| label = "setup",
+#| include = FALSE
 # Packages
@@
 
-```{r data-bolgeretal, include = FALSE}
+```{r}
+#| label = "data-bolgeretal",
+#| include = FALSE
 dir.create("data", FALSE)
@@
 
-```{r fit1-lmer}
+```{r}
+#| label = "fit1-lmer"
 fit_lmer <- lmer(
@@
 
-```{r descriptors, include = FALSE}
+```{r}
+#| label = "descriptors",
+#| include = FALSE
 # Heterogeneity interval
@@
 
-```{r fit1-brms}
+```{r}
+#| label = "fit1-brms"
 fit_1 <- brm(
@@
 
-```{r data-mah-lindsay-3, include = FALSE}
+```{r}
+#| label = "data-mah-lindsay-3",
+#| include = FALSE
 # Experiment 3: task type is between-person
@@
 
-```{r tab-dat2}
+```{r}
+#| label = "tab-dat2"
 dat |>
@@
 
-```{r fit-brms-ml3}
+```{r}
+#| label = "fit-brms-ml3"
 #| message: false

### 2024-06-07 ms.qmd
Manuscript revision

--- 2024-03-21 ms.Rmd
@@
 ---
-title: 'On investigating heterogeneity in psychological phenomena'
-shorttitle: Heterogeneity in psychology
-leftheader: Heterogeneity in psychology
+title: Communicating causal effect heterogeneity
+shorttitle: Communicating heterogeneity
 author: 
   - name: Matti Vuorre
-    affiliation: 1
-    corresponding: yes
-    address: Tilburg School of Social and Behavioral Sciences, Tilburg University
+    corresponding: true
+    orcid: 0000-0001-5052-066X
     email: mjvuorre@uvt.nl
+    affiliation: 
+      - ref: 1
   - name: Matthew Kay
-    affiliation: 2
+    affiliation:
+      - ref: 2
   - name: Niall Bolger
-    affiliation: 3
     affiliation:
+      - ref: 3
+affiliations:
   - id: 1
-    institution: Department of Social Psychology, Tilburg School of Social and Behavioral Sciences, Tilburg University
+    name: Tilburg University
+    department: Department of Social Psychology
   - id: 2
-    institution: Computer Science and Communication Studies, Northwestern University
+    name: Northwestern University
+    department: Computer Science and Communication Studies
   - id: 3
-    institution: Department of Psychology, Columbia University
-abstract: |
-  Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development.
-Yet, established methods and practices for studying heterogeneity underestimate the true uncertainty in those estimates.
-In the current work, we aim to deliver methods and guidelines for more rigorous treatments of heterogeneity in psychological phenomena.
-We provide a walkthrough of methods for i.
-estimating, ii.
-summarising, and iii.
-describing heterogeneity.
-Although other methods for incorporating estimation uncertainty are available, the methods we propose---involving bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters.
-We hope that the methods we discuss would help researchers to widen their inferential focus from statistical averages to include statistical heterogeneity in psychological phenomena.
-keywords: variation, heterogeneity, hierarchical models, mixed models, methodology
-authornote: |
-  \textbf{This working paper is not yet peer reviewed.}
-wordcount: "5,837"
-bibliography: references.bib
-floatsintext: yes
-numbersections: false
-linenumbers: no
-draft: no
-mask: no
-figurelist: no
-tablelist: no
-footnotelist: no
-documentclass: apa7
-classoption: jou
-header-includes:
-  - |
-    \usepackage{float}
-    \setlength{\parskip}{0.5em}
-output: 
-  papaja::apa6_pdf:
-    number_sections: true
-    keep_tex: false
-    highlight: kate
-  papaja::apa6_docx: 
-    number_sections: true
-    keep_tex: false
+    name: Columbia University
+    department: Department of Psychology
+abstract: "Advances in experimental, data collection, and analysis methods have brought population variability in psychological phenomena to the fore.
+Yet, current practices for interpreting such heterogeneity underestimate the uncertainty inevitable in any statistical summary.
+Causal effect heterogeneity is best thought of as distribution of effects with a mean (average person’s effect) and variance (between-person differences).
+This average or expected heterogeneity distribution can be further summarized as a specific heterogeneity interval [@bolgerCausalProcessesPsychology2019].
+However, because the underlying mean and variance parameters will be estimated with uncertainty in empirical studies, these expected distributions and intervals will underestimate the true range of effect values in the population.
+Using Bayesian hierarchical models, and with the aid of empirical datasets, we provide a walk-through of tools for effective heterogeneity reporting and display.
+We cover interval, proportion, and ratio measures of heterogeneity and their interpretation.
+These tools can be a spur to theory building, allowing researchers to widen their focus from population averages to population heterogeneity in psychological phenomena."
+keywords: [heterogeneity, uncertainty, variation, hierarchical model, statistics]
+bibliography: bibliography.bib
+mainfont: Libertinus Serif
+numbered-lines: true
+floatsintext: true
+format:
+  apaquarto-pdf: 
+    documentmode: doc
+    keep-tex: true
+    link-citations: true
+  apaquarto-docx: default
 ---
 
-```{r setup, include = FALSE}
+```{r}
+#| label: setup
+#| include: false
+
 # Packages
-library(papaja)
 library(janitor)
@@
 library(ellipse)
-library(knitr)
 library(cmdstanr)
@@
 library(tidyverse)
+library(knitr)
 
@@
   warning = FALSE,
-  error = TRUE,
+  error = FALSE,
   message = FALSE
 )
+options(digits = 2)
 
@@
 
-# Model estimation options
+# Model estimation options, customize in .Renviron
 dir.create("models", FALSE)
-MAX_CORES <- as.numeric(Sys.getenv("MAX_CORES"), 4)
 options(
   brms.backend = Sys.getenv("BRMS_BACKEND", "rstan"), 
-  brms.threads = MAX_CORES %/% 4,
-  mc.cores = MAX_CORES
-)
-```
-
-When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not.
-Yet, establishing that an independent variable affects a dependent variable may not be a sufficient description of the phenomenon if the treatment effect is variable from one individual to another.
-The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].
-
-One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still commonly analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019].
-Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to sufficiently estimate and describe heterogeneity.
-
-When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023] which, in isolation, might be difficult to interpret.
-That is, although we are excited about the increased focus on heterogeneity, we offer that commonly used descriptions of it leave room for improvement.
-
-Because this focus on heterogeneity in psychological phenomena is relatively new, work remains to be done in developing models, descriptions, and reporting practices that maximize investigations' impact on theory development.
-Our goal is to contribute to that work.
-We describe and illustrate the use of numerical and graphical descriptions of heterogeneity that (1) Go beyond model parameters to describe heterogeneity in clear and actionable terms, and (2) Take uncertainty in model parameters into account.
+  brms.threads = as.numeric(Sys.getenv("MAX_CORES"), 1),
+  mc.cores = as.numeric(Sys.getenv("MAX_CORES"), 4)
+)
+```
+
+When building and testing theories, psychologists have long focused on asking whether an effect exists or not, and what its magnitude might be.
+Yet, establishing that an independent variable affects a dependent variable, possibly to some specific extent, may not be a sufficient description of the phenomenon if the effect varies appreciably from one treatment unit (e.g.
+person) to another.
+The relevance of such variation in the effect, or *heterogeneity*, for theory development is recognized yet typically insufficiently described in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013; @gricePersonsEffectSizes2020; @richters2021].
+
+One reason for the scarcity of reporting and sufficiently interpreting heterogeneity is that psychologists still commonly analyze data with models that obscure its assessment, such as traditional ANOVA [@bolgerCausalProcessesPsychology2019].
+However, more informative modeling is not the only challenge: Although more informative hierarchical (or multilevel, mixed-effects) models are becoming widespread, most users do not yet have the conceptual and practical tools to benefit from the greater explanatory power such models afford.
+
+When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022; @gricePersonsEffectSizes2020], graphical displays that don't yield numerical estimates of hypothetical data-generating mechanisms[@beck2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023].
+These, as we will show, provide an incomplete picture of variation: If (e.g.) a treatment effect is found for 60% or participants in a sample but the uncertainty inherent in that percentage is not communicated, we cannot make inferential conclusions about the effect's prevalence in the population.
+Therefore, to communicate heterogeneity effectively we need not only meaningful measures of it, but also effective methods for describing uncertainty in them.
+Our goal in this paper is to address this challenge by illustrating measures of heterogeneity and how to communicate it, both numerically and graphically, in ways that take uncertainty into account.
+
+<!-- Fix quarto rendering problem in documentmode: jou. -->
+
+<!-- \clearpage -->
+
+<!-- Fix -->
 
 Our plan is as follows.
-We first review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology.
-We then describe additional ways in which model parameters can be interrogated and transformed to describe distributions of causal effects in the population.
+First, we review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology and describe additional ways in which model parameters can be transformed to describe distributions of causal effects in the population.
+We review the concepts and computations underlying three heterogeneity metrics: Individual effect’s mean and standard deviation; their heterogeneity interval; and the prevalence proportion.
 We then move beyond summarizing expected degrees of heterogeneity to describing distributions of plausible degrees of heterogeneity.
@@
 
-# Review of expected heterogeneity in hierarchical models
+# Review of heterogeneity
 
@@
 In their study, which replicated findings first presented in @scholerInflatingDeflatingSelf2014, 62 participants saw twenty positively and twenty negatively valenced words, and judged whether each word was self-descriptive or not.
-Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that response times to positively valenced words would be shorter than to negatively valenced words [@scholerInflatingDeflatingSelf2014].
-
-```{r data-bolgeretal, include = FALSE}
+Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that responses to positively valenced words would be faster than to negatively valenced words [@scholerInflatingDeflatingSelf2014].
+
+```{r}
+#| label: data-bolgeretal
+#| include: false
+
 dir.create("data", FALSE)
@@
   mutate(
-    person = factor(id),
+    person = fct_anon(factor(id)),
     trial,
@@
 
-```{r}
-#| label: tab-dat1
+In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data.
+We first wrangled the data as in @bolgerCausalProcessesPsychology2019, which led to a sample of `r number(nrow(dat), big.mark = ",")` trials from `r length(unique(dat$person))` participants that were endorsed as self-descriptive.
+(We present our code in the online analysis supplement.) We show a sample of these data in @tbl-dat1.
+
+```{r}
+#| label: tbl-dat1
+#| tbl-cap: First six rows of example dataset 1.
+#| apa-note: Reaction time is log-transformed.
+
 dat |>
   head() |>
-  apa_table(
-    span_text_columns = FALSE,
-    font_size = "small",
-    digits = c(0, 0, 0, 2),
-    caption = "First six rows of example dataset 1 (Bolger et al., 2019; Exp 1).",
-    note = "rt is log-transformed."
-  )
-```
-
-In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data (https://github.com/kzee/heterogeneityproject).
-We first wrangled the data as in @bolgerCausalProcessesPsychology2019, which led to a sample of `r number(nrow(dat), big.mark = ",")` trials from 59 participants that were endorsed as self-relevant.
-(We present all our code in the online analysis supplement.) We show a sample of these data in Table \@ref(tab:tab-dat1).
+  rename(`Reaction time` = rt) |> 
+  rename_with(~str_to_sentence(.x)) |> 
+  kable()
+```
+
 Then, we estimated the same statistical model (Model 1; equations 1.1-1.3).
-We modeled the log-transformed reaction time (logRT) of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials, as indicated by the lack of subscripts:
-
+We modeled the log-transformed reaction time (logRT) of person $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials as indicated by the lack of subscripts:
+
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-(We note that there are better and more informative alternatives than modeling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specified a model of the mean of the logRT distribution ($\eta_{ij}$) such that the regression coefficients captured our substantive questions:
-
+```
+(We recognize that there are better alternatives to modeling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specified a model of the mean of the logRT distribution ($\eta_{ij}$) such that the regression coefficients captured our substantive questions:
+
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-This equation includes two kinds of parameters: $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V), do not have subscripts.
-In the frequentist tradition, these parameters are considered constants---not modelled on covariates---and typically referred to as "fixed" parameters [e.g.
+```
+This equation includes two sets of parameters: The first set contains $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V).
+Parameters in this set do not have subscripts: In the frequentist tradition, they are considered constants---not modelled on covariates---and typically referred to as "fixed" parameters [e.g.
 @raudenbushHierarchicalLinearModels2002].
 The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$, have the subscript $j$ to indicate that they are person-specific deviations from the average intercept and slope, respectively.
-That is, $\beta_0 + \gamma_{01}$ is the average reaction time for person $j=1$.
-In frequentist texts, these are typically called "random" parameters, because they are modeled as varying randomly according to a specified distribution.
+That is, $\beta_0 + \gamma_{01}$ is the intercept (average reaction time) for person $j=1$.
+In frequentist texts, these are typically called "random" parameters because they are modeled as varying randomly according to a specified distribution.
 Following standard multilevel modeling assumptions, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-In this equation, we assume that the subject-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
+```
+In this equation, we assume that the person-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
 Perhaps confusingly, $\tau$s and $\rho$ are also sometimes called random effects because they describe random (co)variations of the person-specific effects.
-To be clear, despite this naming convention they are "fixed" features of the population, not of any one group or individual.
+To be clear, despite this naming convention they are features of the population, not of any one group or individual.
 
@@
 
-With data described in Table \@ref(tab:tab-dat1), we can estimate this model using standard maximum likelihood methods as implemented in, for example, the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].
-
-```{r fit1-lmer}
+With data shown in @tbl-dat1, we can estimate this model using standard (restricted) maximum likelihood methods as implemented in, for example, the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].
+
+```{r}
+#| label: fit1-lmer
+
 fit_lmer <- lmer(
@@
   )
-  boots <- boots[c(5, 6, 1, 3, 4, 2), ]
-  tab_1[, 3:4] <- boots
-  saveRDS(tab_1, path)
-} else {tab_1 <- readRDS(path)}
+  saveRDS(boots, path)
+} else {
+  boots <- readRDS(path)
+}
+tab_1[3:6, 3:4] <- boots[c(1, 3, 2, 4), ]
 
@@
   select(Parameter, Coefficient, SE, CI_low, CI_high)
-```
-
-```{r}
-#| label: tbl-fit-1
-#| cache: false
 
@@
   "$\\rho$", "$\\sigma$"
-)
-
-tab_1 |>
-  mutate(
-    mutate(across(where(is.numeric), ~number(.x, .01))),
-    CI = str_glue("[{CI_low}, {CI_high}]"),
-    Coefficient = ifelse(
-      str_starts(Coefficient, "-"), 
-      Coefficient, 
-      paste0("\\phantom{-}", Coefficient)
-    ),
-    Parameter = m1_parnames
-  ) |> 
-  select(
-    Parameter, Coefficient, SE, CI
-  ) |>
-  apa_table(
-    span_text_columns = FALSE,
-    font_size = "small",
-    caption = "Parameter estimates from Model 1.",
-    escape = FALSE
 )
@@
 
-We show a traditional summary of the estimated parameters from this model in Table \@ref(tab:tbl-fit-1).
-For the average person, the estimated effect of valence on logRT is `r mu1[1]` log seconds, with a 95% bootstrap confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`.
-The estimated standard deviation of valence effects in the population is `r sd1[1]` log seconds, with a 95% CI of [`r sd1[2]`, `r sd1[3]`].
-
-## Heterogeneity distribution
-
-```{r descriptors, include = FALSE}
+```{r}
+#| label: tbl-lmer
+#| tbl-cap: Parameter estimates from Model 1 (ML).
+
+tab_1 |>
+  mutate(
+    mutate(across(where(is.numeric), ~number(.x, .01))),
+    `95% CI` = str_glue("[{CI_low}, {CI_high}]"),
+    Parameter = m1_parnames
+  ) |> 
+  select(
+    Parameter, Coefficient, SE, `95% CI`
+  ) |>
+  kable()
+```
+
+We show a conventional summary of this model's estimated parameters in @tbl-lmer.
+For the average person, the estimated effect of positive valence on logRT is `r mu1[1]` log seconds, with a 95% confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`.
+The estimated standard deviation of valence effects in the population is `r sd1[1]` log seconds.
+The lme4 software package does not report a standard error or CI for (co)variance parameters by default, and we therefore calculated it by bootstrapping.
+The resulting 95% bootstrap CI of the valence effect's standard deviation was \[`r sd1[2]`, `r sd1[3]`\].
+
+## Heterogeneity distribution at maximum likelihood estimate of $\beta_1$ and $\tau_1$
+
+```{r}
+#| label: descriptors
+#| include: false
+
 # Heterogeneity interval
@@
 #| fig.height: 1.8
-#| fig.env: "figure*"
-#| fig.align: "center"
-#| fig.cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
+#| apa-twocolumn: true
+#| fig-cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
 A.
@@
 
-Rows 2 and 4 in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:fig-1)A.
+Rows 2 and 4 in @tbl-lmer define the expected normal distribution of valence effects in the population, visualized in @fig-1 A.
 In other words, our point estimate of the distribution of valence effects is Normal(`r mu1[1]`, `r sd1[1]`^2^).
-However, it is an incomplete description of heterogeneity for two reasons.
-First, it does not necessarily communicate the degree of heterogeneity in clear and actionable terms: In addition to the SD, we want quantities that describe heterogeneity in clearer terms.
+However, this distribution is an incomplete description of heterogeneity for two reasons.
+First, it does not incorporate uncertainty in the two determinants of heterogeneity, that is $\beta_1$ and $tau_1$: If they are precisely estimated, i.e.
+when uncertainty regarding them is negligible, the distribution and any quantities calculated from it characterize the population well.
+On the other hand, if they are estimated with considerable uncertainty, the distribution or its transformations would not characterize the population well.
+We return to this key issue below.
+Second, the distribution or its parameters do not, for many purposes, communicate heterogeneity in clear and actionable terms.
 Below, we introduce several metrics that e.g.
 directly describe where a given proportion of the slopes are expected to fall.
-Second, it ignores the uncertainty with which the heterogeneity parameters are estimated.
-That is, we need ways to describe (uncertainty) distributions of distributions.
-
-We now turn to this first issue and outline several additional metrics of heterogeneity afforded by the model's parameter estimates.
 
@@
 
-First, we can use the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes the range within which a certain percentage of the population's slopes are expected to fall [@bolgerCausalProcessesPsychology2019].
+First, we can use the point estimates in @tbl-lmer to construct an expected *heterogeneity interval* that describes the range within which a certain percentage of the population's slopes are expected to fall [@bolgerCausalProcessesPsychology2019].
 To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$).
@@
 For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution.
-Then, we pass those probabilities and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to obtain the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
-In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
-We illustrate this interval in Figure \@ref(fig:fig-1)B.
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = \[`r hi90`\].
+In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
+We illustrate this interval in @fig-1 B.
 
@@
 
-In contrast, proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries.
-For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" In other words, we ask what proportion of the heterogeneity distribution is below zero.
-To answer, we pass zero and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
-This number---$p^-$ for proportion of population with negative effects---is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
-We illustrate this probability in Figure \@ref(fig:fig-1)C.
+The above HI summarizes where a given proportion of individuals' effects lie.
+In contrast, some applications might find it more informative to summarize proportions of effects above or below some critical value, or within some critical range.
+For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" In other words, we ask a question of *prevalence*: What proportion of the heterogeneity distribution is below zero?
+We label this quantity $p^-$ for proportion of population with negative effects.
+
+To answer, we pass zero (the critical value) and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
+This number is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
+We illustrate this probability in @fig-1 C.
 
@@
 
-For example, we might know from theory that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
+For example, let us imagine that a theory states that valence effects in the interval \[-0.1, 0.1\] are practically equivalent to zero.
 To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
 In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero.
-Note that this statement's validity critically depends on the chosen interval, which we arbitrarily set as [-0.1, 0.1] in this example.
-We visualize this probability in Figure \@ref(fig:fig-1)D.
+Note that this statement's validity critically depends on the chosen interval.
+We visualize this probability in @fig-1 D.
 
@@
 
-Although the interval and proportion descriptors show where the population's slopes are likely to fall, they do so in absolute terms such as log(rt) in the running example.
-A contrasting or *relative* way to describe heterogeneity is to express it as a ratio of the standard deviation to the mean.
-The benefit of this descriptor is that in isolation a standard deviation says very little, but when directly compared to the mean effect, it allows for a sufficient description of the relative degree of heterogeneity of the phenomenon under study.
-
-[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the ratio of the standard deviation to the average effect is 0.25 or greater, then heterogeneity can be deemed noteworthy.
+Although the interval and proportion descriptors describe where the population's slopes are likely to fall, they do so in absolute terms such as logRT in the running example.
+A contrasting or *relative* way to describe heterogeneity is to express it as a ratio of the effect's standard deviation to its mean.
+Such relative metrics are concise and can be useful especially when the absolute units are difficult to interpret, or when comparing heterogeneity across different populations or experimental conditions (see below).
+This ratio, expressed simply as the fraction $\frac{\tau_1}{\beta_1}$ is `r ratio` in the current example.
+
+[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that heterogeneity can be deemed noteworthy when the ratio of the standard deviation to the average effect is 0.25 or greater: A ratio greater than 1/4 indicates a $HI_{95}$ whose limits extend to effects one-half and one-and-a-half times that of the average effect.
 With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r ratio`, suggesting that the degree of heterogeneity in valence effects is noteworthy.
-(We reiterate the point made by @bolgerCausalProcessesPsychology2019 that the cutoff value of 0.25 is arbitrary and researchers should choose the cutoff based on their substantive goals.)
-
-## Uncertainty
-
-Although informative, the expected normal distribution of valence effects and its transformations, such as the ones illustrated in Figure \@ref(fig:fig-1), ignore the uncertainty inherent in the estimated parameters.
-The point estimates alone cannot fully communicate the model's knowledge about the degree of heterogeneity.
-That is, the uncertainty with which we have estimated the population parameters, indicated by e.g.
-the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward to the calculations described above or any description based on them.
-
-For example, we only presented the expected heterogeneity interval, but not the range of credible values that it might take.
-If we re-calculate $HI_{90}$ based on the lowest 95%CI bounds in Table \@ref(tab:tbl-fit-1), the resulting heterogeneity interval indicates that 90% of the population's slopes would fall in the [`r hi90_low`] interval.
-In contrast to the expected interval, this interval implies that 90% of the population's slopes are expected to fall below zero.
-
+While this heuristic can sometimes be useful, we urge users to apply domain-specific knowledge when considering critical values or thresholds whenever possible.
+
+## Missing uncertainty
+
+The expected normal distribution of valence effects and its transformations (@fig-1) ignore uncertainty inherent in the estimated parameters.
+That is, we calculated $HI_{90}$, $p^-$, and the other heterogeneity measures from the point estimates $\beta_1$ = `r mu1[1]` and $\tau_1$ = `r sd1[1]`.
+We did not use any information about the precision, or uncertainty, with which these parameters were estimated.
 We have now arrived at the crux of the current work: How should we estimate and describe heterogeneity in psychological phenomena such that the fundamental uncertainty in the estimated parameters is retained?
 
-# Probabilistic assessment of heterogeneity
-
-We think that probabilistic (i.e.
-"bayesian") methods are uniquely able to address this challenge.
-Bayesian methods estimate and describe heterogeneity in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
-For typical scenarios, probabilistic models are as easy to use as their maximum likelihood counterparts [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] Moreover, many advanced models *require* bayesian methods [@muthenBayesianStructuralEquation2012; @asparouhovDynamicStructuralEquation2018].
-
-```{r fit1-brms}
+# Incorporating inferential uncertainty to assessments of heterogeneity
+
+Assessments of heterogeneity involve combining information about fixed and random effects; to fully incorporate inferential uncertainty then requires accounting for their uncertainties in combination.
+Probabilistic, that is, Bayesian methods are uniquely able to address this challenge.
+Modern Bayesian methods, by obtaining draws from the joint posterior distribution of all model parameters presumed to underlie the observed data, allow incorporating posterior uncertainty in combinations of parameters such as the ones highlighted above [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
+For typical scenarios, Bayesian models are as easy to use as their maximum likelihood counterparts [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018].
+
+```{r}
+#| label: fit1-brms
+
 fit_1 <- brm(
@@
 )
+
+post1 <- as_draws_df(
+  fit_1, 
+  variable = c("b_", "sd_", "cor_", "sigma"), 
+  regex = TRUE
+)
 ```
@@
 
-```{r}
-#| label: tbl-fit-1-brms
-#| cache: false
-
-tab1 <- parameters(fit_1, centrality = "mean", effects = "all") |>
-  tibble() |>
-  select(Parameter, Mean, starts_with("CI_"))
-```
+The output of Bayesian computations is the multivariate posterior probability distribution of the model's parameters.
+However, closed-form solutions are not available for multivariate posterior distributions of many important types of statistical models.
+Therefore in practice modern Bayesian methods rely on algorithms that yield many random draws from the multivariate posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016].
+These draws can then be used to calculate, summarize, and visualize any desired quantity of the multivariate posterior such as means, variances, correlations, proportions above or below zero, and so on.
+@tbl-samples-1 illustrates this, showing six random draws of the posteriors of $\beta_1$ and $\tau_1$ (rows).
+We then computed their ratio in the third column, which then represents (draws from) the ratio's posterior distribution and can be summarized, visualized, etc.
 
@@
 #| label: tbl-samples-1
-#| cache: false
-
-post1 <- as.data.frame(fit_1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) |>
-  tibble() |>
-  rownames_to_column("Sample")
-post1_sum <- fit_1 |>
+#| tbl-cap: Random draws from $\beta_1$, $\tau_1$, and their ratio's posterior.
+
+post1 |> 
+  head() |> 
+  select(c(2, 4)) |> 
+  mutate(ratio = sd_person__valence1 / b_valence1) |> 
+  set_names(m1_parnames[c(2, 4)], "$\\frac{\\tau_1}{\\beta_1}$") |> 
+  kable()
+```
+
+In practice, one obtains (for example) 4,000 samples from the posterior distribution using Markov Chain Monte Carlo algorithms [e.g.
+@standevelopmentteamStanModelingLanguage2023] through accessible software [e.g., @burknerBrmsPackageBayesian2017], and then summarizes them using familiar data processing techniques [e.g.
+@dplyr2023; @rcoreteamLanguageEnvironmentStatistical2023].
+Here, we used the R package brms [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] to specify the model and then draw random samples from its posterior distribution.
+The MCMC estimation algorithm completed in about 5 seconds on a modern laptop.
+We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
+
+@tbl-fit-1 shows summaries of Model 1's population-level parameters' ("fixed" parameters in the frequentist nomenclature) posterior distributions.
+The second and third columns show their means and standard deviations (which correspond to frequentist standard errors).
+Note that because we used brms's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in @tbl-lmer.
+
+```{r}
+#| label: tbl-fit-1
+#| tbl-cap: Parameter estimates from Model 1 (Bayes).
+
+fit_1 |>
   parameters(effects = "all", centrality = "mean", dispersion = TRUE) |>
@@
     Mean = number(Mean, .01),
-    SD = number(SD, .001)
-  ) |>
-  select(Parameter, Mean, SD) |>
-  pivot_longer(c(Mean, SD), names_to = "Sample") |>
-  pivot_wider(names_from = Parameter, values_from = value)
-
-set.seed(191)
-post1 |>
-  slice_sample(n = 6) |>
-  arrange(as.integer(Sample)) |>
-  mutate(
-    across(-Sample, ~ number(., .01))
-  ) |>
-  add_row(post1_sum) |>
-  setNames(c("Sample", m1_parnames)) |>
-  apa_table(
-    span_text_columns = FALSE,
-    font_size = "small",
-    midrules = 6,
-    caption = "Six random samples from Model 1's parameters' posterior distributions, and their summaries.",
-    escape = FALSE,
-    placement = "h"
-  )
-```
-
-The goal of bayesian inference is the multivariate posterior probability distribution of the model's parameters.
-However, because closed-form solutions are not available for posterior distributions of many important types of statistical models, in practice bayesian methods rely on drawing many random samples from the posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016].
-Therefore, "estimate" means a large number of random samples from the posterior distribution.
-These samples can then be used to calculate, summarize, and visualize the posterior distribution of any desired quantity.
-
-In practice, one obtains (for example) 4,000 samples from the posterior distribution using computational algorithms [e.g.
-@standevelopmentteamStanModelingLanguage2023] through accessible software [e.g., @burknerBrmsPackageBayesian2017], and then summarizes them using familiar data processing techniques [e.g.
-@rcoreteamLanguageEnvironmentStatistical2023].
-Here, we used the R package brms [@burknerBrmsPackageBayesian2017] to specify the model and then draw random samples from its posterior distribution.
-The MCMC estimation algorithm completed in about 5 seconds on a modern laptop.
-We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
-
-Table \@ref(tab:tbl-samples-1) shows six random samples from the posterior distribution of Model 1's population-level parameters ("fixed" parameters in the frequentist nomenclature).
-The two bottom rows show their means and standard deviations (which correspond to frequentist standard errors).
-Note that because we used brms's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).
+    SD = number(SD, .01),
+    CI = str_glue("[{number(CI_low, .01)}, {number(CI_high, .01)}]")
+  ) |>
+  select(Parameter, Mean, SD, `95% CI` = CI) |> 
+  mutate(Parameter = m1_parnames) |> 
+  kable()
+```
 
@@
 
-Armed with the bayesian estimates (Table \@ref(tab:tbl-samples-1)), we now return to the distribution of valence effects in the population.
+Armed with the Bayesian draws (@tbl-fit-1), we can now return to the question of the distribution of valence effects in the population.
 We now have 4,000 samples from this heterogeneity distribution's posterior distribution.
-We first redraw the expected heterogeneity distribution from Figure \@ref(fig:fig-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:fig-2)A (thick dark blue curve).
-Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 random posterior samples of $\beta_1$ and $\tau_1$.
-From these curves we glean that the distribution of valence effects might well be narrower (i.e.
-less heterogeneity) or wider (more heterogeneity) than is suggested by the point estimates.
+We first redraw the expected heterogeneity distribution from @fig-1 using the posterior mean values of $\beta_1$ and $\tau_1$ in @fig-2 A (thick dark blue curve).
+Superimposed on that normal density curve are heterogeneity distributions calculated from 100 random posterior draws of $\beta_1$ and $\tau_1$.
+From these curves we can see that the true distribution of valence effects might well be less or more heterogeneous than is suggested by the expectation (point estimates).
 
@@
   geom_point()
-summary(fit_1)
 ```
@@
 #| label: fig-2
-#| fig.height: 3
-#| fig.width: 5.4
-#| fig.env: "figure*"
-#| fig.cap: "Bayesian estimates of the heterogeneity distribution of valence effects.
-A.
-Probability density function (PDF) curves.
-The thick line illustrates the expected (point estimate; posterior mean) PDF of the valence effect heterogeneity distribution.
-Thin lines show 100 random samples of the PDF's posterior distribution, and facilitate visual communication of uncertainty in the PDF and, therefore, the valence effects' heterogeneity distribution.
-Dark line segments are inside the [-0.1, 0.1] ROPE, light segments are outside.
-The y-axis is in arbitrary probability density units at each point of the x-axis.
-B.
-The heterogeneity distribution's posterior mean PDF curve (line), and 90\\%CI (ribbon).
+#| fig-width: 4
+#| fig-height: 1.8
+#| fig-cap: Bayesian estimates of the heterogeneity distribution of valence effects.
+#| apa-note: "A: Probability density function (PDF) curves.
+The thick line is the same expected PDF of valence effects from Figure 1.
+Thin lines show 100 PDFs calculated from random draws of $\\beta_1$ and $\\tau_1$ that collectively illustrate the uncertainty in the distribution's location and spread.
+B: Cumulative density function (CDF) curves, annotated as A.
 C.
-Cumulative density function (CDF) curves, annotated as A.
-D.
-Histogram of the posterior distribution of $p^-$ (CDF segments below zero): This value describes the proportion of individuals with negative valence effects on logRT.
-Point and interval representes the posterior mean and 95\\%CI.
-E.
-As C, but represents uncertainty with a 95\\% confidence ribbon.
-F.
-Approximate posterior density of $p^-$."
+Histogram of (draws from) the posterior distribution of $p^-$, which correspond to CDF segments below zero: This value describes the proportion of individuals with negative valence effects on logRT.
+Point and interval represents the posterior mean and 95\\%CI.
+Solid and dotted vertical lines in A and B highlight x-axis values of 0, and -0.1 and 0.1, respectively."
 
@@
   aes(x, cdf) +
-  ylab("Cumulative density") +
-  scale_color_manual(
-    values = c("dodgerblue1", "dodgerblue4")
-  ) +
+  ylab("Cumulative probability") +
   scale_x_continuous(
@@
     xintercept = 0, 
+    linewidth = .25
+  ) +
+  geom_vline(
+    xintercept = c(-0.1, 0.1), 
     lty = "dotted", 
-    linewidth = .25
+    linewidth = .2
   ) +
   geom_line(
-    linewidth = .6,
+    data = post1_density_samples,
+    aes(group = .draw),
+    alpha = .33,
+    linewidth = .15,
+    col = "dodgerblue1"
+  ) +
+  geom_line(
+    linewidth = .8,
     col = "dodgerblue4"
@@
   aes(y = pdf) +
-  geom_line(
-    data = post1_density_samples,
-    aes(group = Sample),
-    col = "dodgerblue1",
-    alpha = .33,
-    linewidth = .15
-  ) +
   scale_y_continuous(
@@
   ) +
-  ylab("Probability density") +
-  theme(
-    axis.title.x = element_blank()
-  )
-
-p_cdf_samples <- p_cdf_samples +
-  geom_line(
-    data = post1_density_samples,
-    aes(group = Sample, color = after_stat(between(x, -0.1, 0.1))),
-    alpha = .33,
-    linewidth = .15
-  )
-
-# Posterior mean and 95%CI pdf and cdf curves
-p_cdf_ribbon <- post1_density_ribbon |>
-  ggplot(aes(x = x)) +
-  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper) +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0),
-    breaks = extended_breaks()
-  ) +
-  scale_y_continuous(
-    breaks = extended_breaks(7),
-    expand = expansion(c(0.005, 0.005))
-  ) +
-  geom_vline(
-    xintercept = 0, 
-    lty = "dotted", 
-    linewidth = .25
-  ) +
-  geom_lineribbon(
-    linewidth = .5,
-    color = "dodgerblue4",
-    fill = alpha("dodgerblue1", .33)
-  ) +
-  theme(
-    legend.position = "none",
-    axis.title.y = element_blank()
-  )
-
-p_pdf_ribbon <- p_cdf_ribbon +
-  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
-  theme(
-    axis.title.x = element_blank(), 
-  )
+  ylab("Probability density")
 
 # Add summaries of CDF < 0 to CDF curves as inset plots
-p_p0 <- post1 |>
+p_p0_samples <- post1 |>
   mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |>
@@
   theme_void(base_size = 9) +
-  theme(
-    plot.tag.position = c(.55, .75), 
-    plot.tag = element_text(size = rel(1))
-  )
-p_p0_samples <- p_p0 + 
   stat_histinterval(
@@
   )
-p_p0_density <- p_p0 + 
-  stat_slabinterval(
-    aes(y = cdf(dist, 0)),
-    fill = alpha("dodgerblue1", 0.5),
-    color = "dodgerblue4",
-    side = "right",
-    .width = .95,
-    size = 1,
-    scale = 0.5,
-    justification = 0.025
-  )
+
 p_cdf_samples <- p_cdf_samples +
@@
     p_p0_samples,
-    0, 0, .2, 1, 
+    0, 0, .3, 1, 
     align_to = "panel", 
     on_top = TRUE
-  )
-p_cdf_ribbon <- p_cdf_ribbon +
-  inset_element(
-    p_p0_density,
-    0, 0, .2, 1, 
-    align_to = "panel", 
-    on_top = TRUE
-  )
-
-p_cdf <- p_cdf_samples | p_cdf_ribbon
-
-p_pdf <- p_pdf_samples | p_pdf_ribbon
-
-(p_pdf / p_cdf) +
+  ) +
+  theme(plot.tag.position = c(0.25, 0.75))
+
+(p_pdf_samples | p_cdf_samples) +
   plot_annotation(tag_levels = "A")
@@
 
-Moreover, the location of these curves differ: Some curves are further to the left (valence effect for the average person is more negative), and some further to the right (effect for the average person is more positive).
-The distribution of these curves represents our current knowledge about the heterogeneity distribution of valence effects in the population---given these data and this model---and can be more completely summarized graphically as a ribbon: Figure \@ref(fig:fig-2)B shows the 95%CI of the valence effect's probability density for each value of the effect on the x-axis.
-A sufficient description of heterogeneity must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
-
-However, depicting the heterogeneity distribution visually as a probability density function (PDF) curve has its drawbacks.
-First, it appears to us visually more challenging to read the degree of uncertainty from a PDF, even when the underlying parameters' uncertainty is summarized as a ribbon around the posterior mean (Figure \@ref(fig:fig-2)B).
-Second, for many applications, the y-axis is not informative: We do not care that the probability density of the curve is (for example) 3.0 at some specific value of the valence effect.
-
-Therefore, in Figure \@ref(fig:fig-2)C we depict the heterogeneity distribution as a *cumulative density* function (CDF).
-We again superimposed 100 random samples from the curve's posterior distribution on the posterior mean, and summarize the entire posterior distribution with a 95% credibility ribbon in Figure \@ref(fig:fig-2)D.
-We believe the CDF is particularly useful as a descriptor of heterogeneity, because the y-axis describes a directly interpretable quantity: Namely, the proportion of population with valence effects below some specific value.
+Some curves in @fig-2 A are further to the left (valence effect for the average person is more negative), and some further to the right (effect for the average person is more positive).
+Moreover, some curves are flatter and wider (effect varies more around the average in the population), and some are narrower and more peaked (effect varies less between individuals).
+The distribution of these curves represents our current knowledge about the heterogeneity distribution of valence effects in the population---given these data and this model.
+A sufficient description of heterogeneity therefore must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.
+
+Depicting the heterogeneity distribution as a probability density function (PDF) curve has its drawbacks.
+First, it appears to us visually more challenging to read the degree of uncertainty from a PDF.
+Second, for many applications, the y-axis is not informative: We typically do not care that the probability density of the curve is (for example) 3.0 at some specific value of the valence effect.
+
+Therefore, in @fig-2 B we depict the heterogeneity distribution as *cumulative distribution* function (CDF) curves based on 100 random posterior draws, together with the mean CDF in a darker color.
+We believe the CDF is a useful visualization tool because the y-axis describes a directly interpretable quantity: The proportion of the population with valence effects below some specific value.
 
@@
 However, a single interval cannot accommodate the uncertainty with which the underlying parameters are estimated.
-To carry uncertainty forward from model parameters to their transformations, such as the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
-Consequently, we get 4,000 samples from the $HI_{90}$s posterior distribution (Figure \@ref(fig:intervals)).
+To carry uncertainty forward from model parameters to the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
+Consequently, we get 4,000 draws from the $HI_{90}$s posterior distribution (@fig-intervals).
 
 Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
-The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (Figure \@ref(fig:intervals)B).
+The 95% most plausible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (@fig-intervals B).
 Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.
 
-Figure \@ref(fig:intervals)A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
-For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2023] the scalar descriptors discussed below.
+@fig-intervals A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
+For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2024] the scalar descriptors discussed below.
 
@@
 For example, we can calculate proportions with negative and positive effects by using zero as the critical value.
-In this example, we ask "What proportion of individuals in the population have negative effects of valence?"
+In this example, we ask "What proportion of individuals in the population endorse positive words faster than negative words?"
 
@@
 
-To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$.
-We show 100 posterior samples of the CDF in Figure \@ref(fig:fig-2)C with a vertical line superimposed at zero.
+To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
+We show 100 posterior draws of the CDF in @fig-2 B with a vertical line superimposed at zero.
 The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$).
-We also show a histogram of all 4,000 posterior samples of that proportion in the top left margin of Figure \@ref(fig:fig-2)C, with the associated 95%CI.
+We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-2 C, with its associated 95%CI.
 The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
-Alternatively, the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
+Stated differently, the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.
 
 Moreover, if theory allows defining a range of parameter values that are practically equivalent to zero (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically negligible effects.
-We colored the curves in Figure \@ref(fig:fig-2)C such that curve segments within [-0.1, 0.1] are highlighted with a darker color.
-Those darker line segments represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$).
+We added dotted vertical lines in @fig-2 to highlight the \[-0.1, 0.1\] interval, which serves as an example ROPE.
+Line segments within that interval represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$).
 To quantify uncertainty in $p^{ROPE}$ we then aggregate the segments' to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
-We note that the ROPE of [-0.1, 0.1] here was arbitrary and picked just to illustrate the example.
+We note that the ROPE of \[-0.1, 0.1\] here was arbitrary and picked just to illustrate the example.
 
@@
 
-Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$.
-Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (Figure \@ref(fig:hdist-5)B).
+Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$ (see @tbl-samples-1).
+@fig-ratio shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (panel B).
 The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
@@
 ```{r}
-#| label: hdist-5
-#| fig.height: 2.2
-#| fig.width: 4
-#| fig.cap: Panel A.
+#| label: fig-ratio
+#| fig-height: 2.2
+#| fig-width: 4
+#| fig-cap: Bivariate posterior of $\beta_1$, $\tau_1$, and their ratio.
+#| apa-note: Panel A.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
@@
 
-We have seen that---with these example data and this model---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects.
+We have seen that---with these example data and this model---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding how valence effects vary between people in the population.
 We will next see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing its magnitude across multiple populations.
 
-# Comparing heterogeneity
-
-Next, we move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across potentially multiple populations of different kinds of study units.
-To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2023 addressing differences in between person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment.
-In @mahVariabilitySubjectsFree2023's Experiment 3, 260 individuals studied a list of twenty target words.
+\clearpage
+
+# Comparing heterogeneity between populations
+
+We now move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of study units.
+To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2024 addressing differences in between-person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment.
+In @mahVariabilitySubjectsFree2024's Experiment 3, 260 individuals studied a list of twenty target words.
 After a short break, they then either freely recalled as many of the target words as they could (Free recall group, N = 123) or recalled as many target words as they could when prompted with related cue words (Cued recall group, N = 137).
 Thus, the Free and Cued recall tasks had different groups of participants but the same target words.
-We show a sample of these data in Table \@ref(tab:tab-dat2).
-
-```{r data-mah-lindsay-3, include = FALSE}
+We show a sample of these data in @tbl-dat2.
+
+```{r}
+#| label: data-mah-lindsay-3
+#| include: false
+
 # Experiment 3: task type is between-person
@@
 
-With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2023 found that participants who completed the cued recall task showed greater heterogeneity in memory performance than did the free recall group: The Cued:Free recall variance ratio was 1.33 (with a [1.14, 1.54] 95% bootstrap interval).
-@mahVariabilitySubjectsFree2023 also confirmed this result by comparing multilevel models that did and didn't allow for different between-person variabilities in memory ability for the two groups, and found similar results across three different experiments.
-
-```{r tab-dat2}
+With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2024 found that participants who completed the cued recall task were more heterogeneous in their memory performance than those in the free recall group: The Cued:Free recall between-person memory performance variance ratio was 1.33 (with a \[1.14, 1.54\] 95% bootstrap interval).
+@mahVariabilitySubjectsFree2024, across three experiments, confirmed this result by comparing models that did and did not allow for distinct between-person variabilities in each group.
+
+```{r}
+#| label: tbl-dat2
+#| tbl-cap: "Six rows of example dataset 2 (Mah \\& Lindsay, 2023; Exp 3)."
+
 dat |>
   slice(4:6, .by = task) |> 
-  apa_table(
-    span_text_columns = FALSE,
-    font_size = "small",
-    escape = FALSE,
-    digits = 0,
-    caption = "Six rows of example dataset 2 (Mah \\& Lindsay, 2023; Exp 3)."
-  )
-```
-
-We extend those analyses to illustrate how our descriptions of heterogeneity generalize to differences in heterogeneity in ways that inform audiences beyond variance ratios and model comparisons.
-We ask three questions about between-task differences in heterogeneity: (1) To what extent is memory performance more variable between people in the cued recall task compared to the free recall task?
-(2) To what extent is memory performance variability between target words different across cued recall and free recall tasks?
+  rename_with(~str_to_sentence(.x)) |> 
+  kable()
+```
+
+Let us now see how our earlier descriptions of heterogeneity can be usefully extended to group differences.
+We ask three questions about differences in heterogeneity: (1) To what extent is memory performance more variable *between people* in the cued recall task compared to the free recall task?
+(2) To what extent is memory performance more variable *between target words* in cued-versus-free recall tasks?
 And (3) How consistent is target word heterogeneity across the two tasks: Are target words associated with good memory performance in cued recall experiments the same words that are associated with good memory performance in free recall experiments?
 
-To answer these questions, we model the $i$th total recall accuracy in 1 to `r nrow(dat)`, of person $j$ in 1 to `r length(unique(dat$person))`, word $k$ in 1 to `r length(unique(dat$target))`, and task $m$ in {F (free recall), C (cued recall)} as Bernoulli distributed, where the probability of an accurate answer is determined by the rate parameter $\pi_{ijkm}$.
+To answer these questions, we model the $i$th total recall accuracy in 1 to `r nrow(dat)`, of person $j$ in 1 to `r length(unique(dat$person))`, word $k$ in 1 to `r length(unique(dat$target))`, and task $m$ in {F (free recall), C (cued recall)} as Bernoulli distributed, where the probability of an accurate answer is determined by the rate parameter $\pi$.
 As is common with generalized linear models, we model the rate parameter through a nonlinear link function.
@@
 
+```{=tex}
 \begin{align*}
 \tag{2}
-\text{Accuracy}_{ijkm} &\sim \operatorname{Bernoulli}\left(\pi_{ijkm}\right) \\
-\pi_{ijkm} &= \Phi\left(\eta_{ijkm}\right) \\
-\eta_{ijkm} &= \beta_{m} + \gamma_{jm} + \delta_{km} \\
+\text{Accuracy}_{ijkm} &\sim \operatorname{Bernoulli}\left(\pi_{jkm}\right) \\
+\pi_{jkm} &= \Phi\left(\eta_{jkm}\right) \\
+\eta_{jkm} &= \beta_{m} + \gamma_{jm} + \delta_{km} \\
 \gamma_{[m:F]} &\sim 
@@
 \end{align*}
-
+```
 This Model 2 of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$).
@@
 This reparameterization is useful, because it allows us to quantify heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference.
-Notice also that we model $\gamma$ using to independent normal distributions, and $\delta$ with a multivariate normal distribution.
+
+Note also that we model $\gamma$ using two independent normal distributions, and $\delta$ with a multivariate normal distribution.
 Because different persons participated in the two tasks, we cannot assess whether participant-specific abilities are correlated across the tasks.
@@
 
-```{r fit-brms-ml3}
+```{r}
+#| label: fit-brms-ml3
 #| message: false
@@
 
-We estimated Model 2 exactly as Model 1, by drawing 4,000 random samples from its posterior distribution [@burknerBrmsPackageBayesian2017].
+We estimated Model 2 exactly as Model 1, by taking 4,000 random draws from its posterior distribution [@burknerBrmsPackageBayesian2017].
 We then confirmed graphically and numerically that the estimation algorithm had converged, and that the model performed adequately using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
-We summarise the model's posterior distribution in Table \@ref(tab:tbl-samples-2).
-
-```{r}
-#| label: tbl-samples-2
-#| cache: true
+We summarise the model's posterior distribution in @tbl-fit-2.
+
+```{r}
+#| label: tbl-fit-2
+#| tbl-cap: Parameter estimates from Model 2.
 
 m2_parnames <- c(
-  "$\\beta_{m:F}$", "$\\beta_{m:C}$",
-  "$\\tau_{\\gamma_{m:F}}$", "$\\tau_{\\gamma_{m:C}}$",
-  "$\\tau_{\\delta_{m:F}}$", "$\\tau_{\\delta_{m:C}}$",
+  "$\\beta_{m:F}$", 
+  "$\\beta_{m:C}$",
+  "$\\tau_{\\gamma_{m:F}}$", 
+  "$\\tau_{\\gamma_{m:C}}$",
+  "$\\tau_{\\delta_{m:F}}$", 
+  "$\\tau_{\\delta_{m:C}}$",
   "$\\rho_\\delta$"
@@
   regex = TRUE
-) |> 
-  select(-.iteration, -.chain) |> 
-  relocate(.draw, .before = b_taskFree)
-
-set.seed(191)
-post2 |>
-  slice_sample(n = 6) |>
-  mutate(.draw = as.character(.draw)) |> 
-  add_row(
-    post2 |> 
-      summarise_draws(
-        Mean = ~mean(.),
-        SD = ~sd(.)
-      ) |> 
-      pivot_longer(c(Mean, SD), names_to = ".draw") |>
-      pivot_wider(names_from = variable, values_from = value)
-  ) |> 
-  setNames(c("Sample", m2_parnames)) |> 
-  apa_table(
-    span_text_columns = FALSE,
-    font_size = "scriptsize",
-    midrules = 6,
-    digits = 2,
-    caption = "Six random samples from Model 2's parameters' posterior distributions, and their summaries.",
-    escape = FALSE,
-    placement = "h"
-  )
+)
+
+fit2 |> 
+  parameters(effects = "all", centrality = "mean", dispersion = TRUE) |>
+  tibble() |>
+  mutate(
+    Mean = number(Mean, .01),
+    SD = number(SD, .001),
+    CI = str_glue("[{number(CI_low, .01)}, {number(CI_high, .01)}]")
+  ) |>
+  select(Parameter, Mean, SD, `95% CI` = CI) |> 
+  mutate(Parameter = m2_parnames) |> 
+  kable()
 ```
@@
 
-Descriptively, we reproduced @mahVariabilitySubjectsFree2023's finding that participants' memory performance was more heterogeneous in the cued recall than in the free recall experiment (columns 4 and 5 in Table \@ref(tab:tbl-samples-2)).
-We show the relevant estimated quantities, and the implied heterogeneity distributions in Figure \@ref(fig:fig-2-person).
-
-The top panel of Figure \@ref(fig:fig-2-person)A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall).
+Descriptively, we reproduced @mahVariabilitySubjectsFree2024's finding that participants' memory performance was more heterogeneous in the cued recall than in the free recall experiment (rows 3 and 4 in @tbl-fit-2).
+We show the relevant estimated quantities, and the implied heterogeneity distributions in @fig-2-person.
+
+The top panel of @fig-2-person A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall).
 Notice that the model's parameters refer to probits ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` probits in the free and cued recall conditions, respectively, the corresponding probabilities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued_p"]`.
-However, this conversion does not work in general and we therefore restrict descriptions to probits in what follows.
 
@@
 #| label: fig-2-person
-#| fig.height: 3.2
-#| fig.env: "figure*"
-#| fig.cap: "Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
-A.
+#| fig-height: 3.2
+#| fig-cap: Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
+#| apa-twocolumn: true
+#| apa-note: "A.
 Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
 Differences calculated as Cued - Free recall.
+$p^+$ indicates the estimated fraction of the population with memory performance above 0.5.
+Heterogeneity ratio indicates standard deviations divided with their respective means.
 B.
-Probability density (top) and cumulative density functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
-The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance."
+Probability density (top) and cumulative distribution functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
+The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above 0.5."
 
@@
       levels = c("b", "sd", "pd", "ratio"),
-      labels = c("Mean", "SD", "Pr(Above chance)", "Ratio")
+      labels = c("Mean", "SD", "p+", "Heterogeneity ratio")
     ),
@@
 
-Second, and more importantly, the second row in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+More importantly, the second row in @fig-2-person A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
 The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
 
-The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks ($p^+$), and their difference.
-The model estimates the proportion of individuals who perform above chance to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
-Notice that the this quantity refers to population proportions and is not a z-score.
-
-Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of Figure \@ref(fig:fig-2-person)A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
-(We truncated the Ratio axis at [-10, 10] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
+The third row of @fig-2-person A shows the estimated proportions of individuals whose memory performance exceeded 50% ($p^+$), and their difference.
+The model estimates the proportion of individuals who recall over 50% of items to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
+Note that this quantity refers to population proportions and is not a z-score.
+
+Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of @fig-2-person A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
+(We truncated the Ratio axis at \[-10, 10\] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
 The difference in ratios is very uncertain partly for the same reason.)
 
-We also depict the heterogeneity distributions as PDFs and CDFs in Figure \@ref(fig:fig-2-person)B.
-These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks: First, we see that the majority of the free recall CDF (green) is to the left of zero (chance), indicating that the majority of this population is predicted to perform worse than chance.
-This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-chance performance only for a proportion of `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
-Second, we see that the slope of the cued recall CDF (red) is less steep than that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.
+We also depict the heterogeneity distributions as PDFs and CDFs in @fig-2-person B.
+Unlike in @fig-2 that drew random draws of the functions' posteriors, to reduce overplotting @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light).
+These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks.
+
+First, we see that the majority of the free recall group's CDF (green) is to the left of zero (50% recall), indicating that the majority of this population is predicted to recall less than half of items.
+This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-50% performance only for a proportion of `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
+Second, we see that the slope of the cued recall CDF (red) is less steep and to the right to that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task, and `r post2_sum$x[post2_sum$variable=="pd_person_cued"]` individuals in that group are predicted to perform above 50%.
 
@@
 While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $HI_90$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
-Studying Figure \@ref(fig:fig-2-person)B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
+Studying @fig-2-person B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
 
@@
 However, examining heterogeneity in other sampled units can be both theoretically and methodologically important.
-Recognizing this difference, and for the sake of our example analyses we next turn to our second and third questions regarding potential differences and consistencies in between-target word heterogeneity.
+We next turn to our second and third questions regarding potential differences and consistencies in between-target word heterogeneity.
 
@@
 #| label: fig-2-target
-#| fig.height: 4
-#| fig.width: 7.4
-#| fig.env: "figure*"
-#| fig.cap: "Estimated between-target word heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
-A.
+#| fig-height: 4
+#| fig-width: 7.4
+#| apa-twocolumn: true
+#| fig-cap: Heterogeneity between target words in memory performance in Free recall and Cued recall tasks from Model 2.
+#| apa-note: "A.
 Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
@@
 B.
-Probability density (top) and cumulative density functions (bottom) of the two tasks' heterogeneity distributions (green: free recall, red: cued recall).
+Probability density (top) and cumulative distribution functions (bottom) of the two tasks' heterogeneity distributions (green: free recall, red: cued recall).
 The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance.
@@
       levels = c("sd", "pd", "ratio", "cor"),
-      labels = c("SD", "Pr(Above chance)", "Ratio", "Cor(Free, Cued)")
+      labels = c("SD", "p+", "Heterogeneity ratio", "Cor(Free, Cued)")
     ),
@@
 The results regarding differences in memory performance heterogeneity across target words were similar to those as observed regarding heterogeneity in people's memory performances.
-Figure \@ref(fig:fig-2-target)A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
-The interpretation of this difference is quantitatively similar to that observed about people above: Both people and target words exhibit greater memory task performance variability in the cued recall than in the free recall task.
-However, because @mahVariabilitySubjectsFree2023 used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
-
-Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than chance performance in the cued recall than in the free recall task.
+@fig-2-target A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
+The interpretation of this difference is quantitatively similar to that observed about people above: Both people and target words exhibit greater memory performance variability in the cued recall than in the free recall task.
+However, because @mahVariabilitySubjectsFree2024 used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
+
+Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than 50% accurate recall in the cued recall than in the free recall task.
 Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared nearly identical across the two tasks.
@@
 This analysis of target word heterogeneity afforded an additional piece of information, because the same target words were used across the two tasks, and thus allowed answering our third question.
-There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of Figure \@ref(fig:fig-2-target)A, and Figure \@ref(fig:fig-2-target)C).
+There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of @fig-2-target A, and @fig-2-target C).
 The posterior mean and 95%CI of this correlation was `r post2_sum$x[post2_sum$variable=="cor_target__taskFree__taskCued"]`.
+
 What this correlation means substantively is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task.
-@bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.
+(@bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.)
 
 Although outside our current scope, we note that our study of Model 2's results might indicate exciting new avenues for this line of inquiry.
-One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2023].
-Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (Figure \@ref(fig:fig-2-target)A).
-Moreover, we observed across two kinds of study units (people, target words) that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
+One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2024].
+Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (@fig-2-target A).
+Second, we observed across both people and target words that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
+Finally, given that we operationalized the stability of item difficulties as a correlation across tasks, it might be theoretically important to look for sets of stimuli where this positive correlation did not occur.
 
@@
 
-In the current work, we illustrated the use of practical descriptors of heterogeneity in psychological phenomena with examples drawn from social and cognitive psychology.
+In the current work, we illustrated the use of practical descriptors of heterogeneity with examples drawn from social and cognitive psychology.
 Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects in psychology---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
-
-Although prior work on developing metrics of heterogeneity, and placing experimental effect sizes in context of person-specific effects exists, it mostly has ignored estimation uncertainty and thus remained purely descriptive.
+Our currently proposed methods incorporate uncertainty into both modelling of and inferences about heterogeneity.
+
+Although prior work on developing metrics of heterogeneity, and placing experimental effect sizes in context of person-specific effects exists, it has ignored estimation uncertainty and thus remained purely descriptive.
 For example, @gricePersonsEffectSizes2020 describe a method whereby analysts simply count the number of individuals whose point estimate of an effect is concordant with a hypothesis.
@@
 
-We emphasized throughout that a probabilistic ("bayesian") approach is uniquely positioned to answer the needs of researchers interested in heterogeneity.
+Second [@schuetze2024, p.3] suggest that "past efforts to identify heterogeneous effects have yielded a disproportionate number of disappointing, uninterpretable, and non-replicable findings", and suggest low power as one potential antecedent.
+While perhaps an overstatement, one reason for why previous investigations of heterogeneity may have been suboptimal indeed relates to statistical power: By not duly incorporating and reporting on the uncertainty with which heterogeneity is estimated, investigations are more suspect for reporting substantial heterogeneity where it may not truly exist.
+
+Finally, @bolgerCausalProcessesPsychology2019 provided an extensive discussion of how heterogeneity can be estimated for causal effects in psychology.
+We directly built on that work to illustrate the benefits and how such descriptions can and should include representations of uncertainty.
+
+We emphasized throughout that a probabilistic (Bayesian) approach is uniquely positioned to answer the needs of researchers interested in heterogeneity.
 Bayesian methods allow carrying uncertainty forward from model parameters to descriptors of heterogeneity and beyond.
@@
 In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
-We further elucidate these points in our Online Supplementary Analysis [todo].
-
-We believe that psychology, broadly speaking, is methodologically and theoretically at a ripe position to taking a meta-theoretic step forward and try to incorporate an understanding of heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019].
-In order to do so, descriptions of heterogeneity must include measures of uncertainty, and we hope the techniques illustrated here help researchers do so.
+
+We believe that psychology, broadly speaking, is methodologically and theoretically ripe for incorporating effect heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019].
+To do so, descriptions of heterogeneity must include measures of uncertainty, and we hope the techniques illustrated here help researchers do so.
 
@@
 In our example analyses, we have brushed many important modelling decisions under the rug in order to focus on the main topic of heterogeneity.
-First, in example 1 we analyzed reaction times by simply log-transforming reaction times.
+First, in the first example, we analyzed reaction times by simply log-transforming reaction times.
 More informative analyses of RTs would make use of models that make more realistic assumptions about the data generating process underlying reaction time responses, but here we necessarily excluded this complication for reasons of brevity.
 
-Our exposition and interpretation of heterogeneity relied on a critical assumption in line with standard practices in multilevel and generalized linear mixed modelling; that of (multivariate) normality of the group-level parameters.
+Our exposition and interpretation of heterogeneity relies on a critical assumption in line with standard practices in multilevel and generalized linear mixed modelling; that of (multivariate) normality of the unit-level (person, item, etc) parameters.
 Assuming that random effects are normally distributed is a computationally and conceptually useful fiction, and we recognize that it is unlikely to hold exactly in real psychological phenomena.
@@
 
-All the code supporting this manuscript is available at <https://github.com/mvuorre/heterogeneity-uncertainty> and archived on Zenodo at [todo].
-We reused openly available data from @bolgerCausalProcessesPsychology2019 and @mahVariabilitySubjectsFree2023.
+All the code supporting this manuscript is available at <https://github.com/mvuorre/heterogeneity-uncertainty>.
+We reused openly available datasets [@bolgerCausalProcessesPsychology2019; @mahVariabilitySubjectsFree2024].
 
@@
 
-Conceptualization: MV, NB\
+\noindent Conceptualization: MV, NB\
 Formal Analysis: MV\
@@
 
-# References {-}
+# References {.unnumbered}
 
@@
 <!-- Prepare appendix section -->
+
+```{=tex}
 \clearpage
@@
 \onecolumn
-
+```
 # Appendix
@@
 ```{r}
-#| label: intervals
-#| fig.height: 2.4
-#| fig.cap: Bayesian estimates of the 90% heterogeneity interval of valence effects.
-A.
+#| label: fig-intervals
+#| fig-height: 2.4
+#| fig-cap: Bayesian estimates of the 90% heterogeneity interval of valence effects. 
+#| apa-note: A.
 Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$.
@@
   ) |>
-  select(Sample, hi_low, hi_high) |>
-  pivot_longer(-Sample) |>
+  select(hi_low, hi_high) |>
+  pivot_longer(everything()) |>
   ggplot(aes(value, group = name)) +
@@
 
-p3 <- distinct(post1_density_samples, Sample, b_valence1, sd_person__valence1) |>
+p3 <- post1_density_samples |> 
+  distinct(
+    .draw, 
+    b_valence1, 
+    sd_person__valence1
+) |>
   arrange(b_valence1) |>

### 2024-06-25 index.qmd
Revision and restructuring

--- 2024-06-07 ms.qmd
@@
 title: Communicating causal effect heterogeneity
-shorttitle: Communicating heterogeneity
+running-head: Communicating heterogeneity
 author: 
   - name: Matti Vuorre
-    corresponding: true
     orcid: 0000-0001-5052-066X
     email: mjvuorre@uvt.nl
+    corresponding: true
     affiliation: 
@@
   - name: Matthew Kay
+    orcid: 0000-0001-9446-0419
     affiliation:
@@
   - name: Niall Bolger
+    orcid: 0000-0001-8698-8117
     affiliation:
@@
     department: Department of Psychology
-abstract: "Advances in experimental, data collection, and analysis methods have brought population variability in psychological phenomena to the fore.
-Yet, current practices for interpreting such heterogeneity underestimate the uncertainty inevitable in any statistical summary.
-Causal effect heterogeneity is best thought of as distribution of effects with a mean (average person’s effect) and variance (between-person differences).
-This average or expected heterogeneity distribution can be further summarized as a specific heterogeneity interval [@bolgerCausalProcessesPsychology2019].
-However, because the underlying mean and variance parameters will be estimated with uncertainty in empirical studies, these expected distributions and intervals will underestimate the true range of effect values in the population.
-Using Bayesian hierarchical models, and with the aid of empirical datasets, we provide a walk-through of tools for effective heterogeneity reporting and display.
-We cover interval, proportion, and ratio measures of heterogeneity and their interpretation.
-These tools can be a spur to theory building, allowing researchers to widen their focus from population averages to population heterogeneity in psychological phenomena."
+abstract: |
+  Advances in experimental, data collection, and analysis methods have brought population variability in psychological phenomena to the fore.
+Yet, current practices for interpreting such heterogeneity do not appropriately treat the uncertainty inevitable in any statistical summary.
+Heterogeneity is best thought of as a distribution of features with a mean (average person’s effect) and variance (between-person differences).
+This expected heterogeneity distribution can be further summarized e.g.
+as a heterogeneity interval [@bolgerCausalProcessesPsychology2019].
+However, because empirical studies estimate the underlying mean and variance parameters with uncertainty, the expected distribution and interval will underestimate the actual range of plausible effects in the population.
+Using Bayesian hierarchical models, and with the aid of empirical datasets from social and cognitive psychology, we provide a walk-through of effective heterogeneity reporting and display tools that appropriately convey measures of uncertainty.
+We cover interval, proportion, and ratio measures of heterogeneity and their estimation and interpretation.
+These tools can be a spur to theory building, allowing researchers to widen their focus from population averages to population heterogeneity in psychological phenomena.
 keywords: [heterogeneity, uncertainty, variation, hierarchical model, statistics]
+authornote: |
+  This manuscript is not yet peer-reviewed.
 bibliography: bibliography.bib
-mainfont: Libertinus Serif
-numbered-lines: true
-floatsintext: true
-format:
-  apaquarto-pdf: 
-    documentmode: doc
-    keep-tex: true
-    link-citations: true
-  apaquarto-docx: default
+citeproc: true
+csl: apa.csl
 ---
@@
 library(ellipse)
+library(tinytable)
 library(cmdstanr)
@@
 opts_chunk$set(
-  fig.width = 6,
-  fig.align = "center",
   eval = TRUE,
@@
   error = FALSE,
-  message = FALSE
+  message = FALSE,
+  echo = FALSE
 )
 options(digits = 2)
+options(tinytable_tt_digits = 3)
 
@@
 
-When building and testing theories, psychologists have long focused on asking whether an effect exists or not, and what its magnitude might be.
+When building and testing theories, psychologists have long focused on asking whether an effect exists and what its magnitude might be.
 Yet, establishing that an independent variable affects a dependent variable, possibly to some specific extent, may not be a sufficient description of the phenomenon if the effect varies appreciably from one treatment unit (e.g.
@@
 One reason for the scarcity of reporting and sufficiently interpreting heterogeneity is that psychologists still commonly analyze data with models that obscure its assessment, such as traditional ANOVA [@bolgerCausalProcessesPsychology2019].
-However, more informative modeling is not the only challenge: Although more informative hierarchical (or multilevel, mixed-effects) models are becoming widespread, most users do not yet have the conceptual and practical tools to benefit from the greater explanatory power such models afford.
-
-When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022; @gricePersonsEffectSizes2020], graphical displays that don't yield numerical estimates of hypothetical data-generating mechanisms[@beck2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023].
-These, as we will show, provide an incomplete picture of variation: If (e.g.) a treatment effect is found for 60% or participants in a sample but the uncertainty inherent in that percentage is not communicated, we cannot make inferential conclusions about the effect's prevalence in the population.
-Therefore, to communicate heterogeneity effectively we need not only meaningful measures of it, but also effective methods for describing uncertainty in them.
-Our goal in this paper is to address this challenge by illustrating measures of heterogeneity and how to communicate it, both numerically and graphically, in ways that take uncertainty into account.
-
-<!-- Fix quarto rendering problem in documentmode: jou. -->
-
-<!-- \clearpage -->
-
-<!-- Fix -->
+However, more informative modeling is not the only challenge: Although more informative hierarchical (or multilevel, mixed-effects [@gelmanDataAnalysisUsing2007]) models are becoming widespread, many users do not yet have the conceptual and practical tools to benefit from the greater explanatory power such models afford.
+
+When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022; @gricePersonsEffectSizes2020], graphical displays that don't yield numerical estimates of hypothetical data-generating mechanisms [@beck2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023].
+These, as we will show, provide an incomplete picture of variation that is sometimes difficult to interpret: If (e.g.) a treatment effect is found for 60% or participants in a sample but the uncertainty inherent in that percentage is not communicated, we cannot make inferential conclusions about the effect's prevalence in the population.
+Therefore, to communicate heterogeneity effectively we need not only meaningful measures of it, but also effective methods for describing the associated uncertainties.
+Our goal in this paper is to address this challenge by illustrating measures of heterogeneity and how to communicate them, both numerically and graphically, in ways that take uncertainty into account.
 
 Our plan is as follows.
-First, we review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology and describe additional ways in which model parameters can be transformed to describe distributions of causal effects in the population.
-We review the concepts and computations underlying three heterogeneity metrics: Individual effect’s mean and standard deviation; their heterogeneity interval; and the prevalence proportion.
-We then move beyond summarizing expected degrees of heterogeneity to describing distributions of plausible degrees of heterogeneity.
+First, we review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology.
+We then describe additional ways in which model parameters can be transformed to describe distributions of causal effects.
+We review the concepts and computations underlying three heterogeneity metrics: The effect's mean and standard deviation in the population; the heterogeneity interval; and the prevalence proportion.
+Second, we move beyond summarizing expected degrees of heterogeneity that lack information about uncertainty to describing distributions of plausible degrees of heterogeneity.
+Such uncertainty distributions of population feature distributions are natural components of Bayesian hierarchical models and afford efficient tools for describing distributional uncertainty.
 Finally, we extend these methods to compare heterogeneity across different populations using an example dataset from cognitive psychology.
@@
 We first wrangled the data as in @bolgerCausalProcessesPsychology2019, which led to a sample of `r number(nrow(dat), big.mark = ",")` trials from `r length(unique(dat$person))` participants that were endorsed as self-descriptive.
-(We present our code in the online analysis supplement.) We show a sample of these data in @tbl-dat1.
+Our online analysis supplement (<https://osf.io/yp2gq>) includes the complete code to reproduce this manuscript and computations therein.
+We show a sample of these data in @tbl-dat1.
 
@@
 #| label: tbl-dat1
-#| tbl-cap: First six rows of example dataset 1.
-#| apa-note: Reaction time is log-transformed.
+#| tbl-cap: First six rows of example dataset 1 [@bolgerCausalProcessesPsychology2019].
 
@@
   head() |>
-  rename(`Reaction time` = rt) |> 
+  rename(`log(Reaction time)` = rt) |> 
   rename_with(~str_to_sentence(.x)) |> 
@@
 
-Then, we estimated the same statistical model (Model 1; equations 1.1-1.3).
-We modeled the log-transformed reaction time (logRT) of person $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials as indicated by the lack of subscripts:
-
-```{=tex}
-\begin{equation}
-\tag{1.1}
+Then, we estimated the same statistical model (@eq-m1-1 - [-@eq-m1-3]).
+We modeled the log-transformed reaction time of person $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials as indicated by the lack of subscripts:
+
+$$
 \text{logRT}_{ij} \sim \operatorname{Normal}\left(\eta_{ij}, \sigma^2\right).
-\end{equation}
-```
+$$ {#eq-m1-1}
+
 (We recognize that there are better alternatives to modeling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specified a model of the mean of the logRT distribution ($\eta_{ij}$) such that the regression coefficients captured our substantive questions:
 
-```{=tex}
-\begin{equation}
-\tag{1.2}
+$$
 \eta_{ij} = \beta_0 + \gamma_{0j} + \left(\beta_1 + \gamma_{1j}\right)\text{V}_{ij}.
-\end{equation}
-```
+$$ {#eq-m1-2}
+
 This equation includes two sets of parameters: The first set contains $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V).
@@
 
-```{=tex}
-\begin{equation}
-\tag{1.3}
+$$
 \begin{bmatrix} 
@@
 \right).
-\end{equation}
-```
+$$ {#eq-m1-3}
+
 In this equation, we assume that the person-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
@@
 The lme4 software package does not report a standard error or CI for (co)variance parameters by default, and we therefore calculated it by bootstrapping.
-The resulting 95% bootstrap CI of the valence effect's standard deviation was \[`r sd1[2]`, `r sd1[3]`\].
+The resulting 95% bootstrap CI of the valence effect's standard deviation was [`r sd1[2]`, `r sd1[3]`].
 
@@
 #| label: fig-1
-#| fig.height: 1.8
-#| apa-twocolumn: true
+#| fig-height: 2
+#| fig-width: 7
 #| fig-cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
@@
 However, this distribution is an incomplete description of heterogeneity for two reasons.
-First, it does not incorporate uncertainty in the two determinants of heterogeneity, that is $\beta_1$ and $tau_1$: If they are precisely estimated, i.e.
+First, it does not incorporate uncertainty in the two determinants of heterogeneity, that is $\beta_1$ and $\tau_1$: If they are precisely estimated, i.e.
 when uncertainty regarding them is negligible, the distribution and any quantities calculated from it characterize the population well.
@@
 Second, the distribution or its parameters do not, for many purposes, communicate heterogeneity in clear and actionable terms.
-Below, we introduce several metrics that e.g.
-directly describe where a given proportion of the slopes are expected to fall.
+Below, we introduce several metrics that directly describe e.g.
+where a given proportion of the slopes are expected to fall.
 
@@
 For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution.
-Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = \[`r hi90`\].
-In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval.
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
 We illustrate this interval in @fig-1 B.
@@
 
-The above HI summarizes where a given proportion of individuals' effects lie.
+The above HI summarizes where a given proportion of individuals' effects in the population are.
 In contrast, some applications might find it more informative to summarize proportions of effects above or below some critical value, or within some critical range.
@@
 In common applications, ROPE is used to statistically infer whether an estimated parameter, such as the effect of valence on logRT for the average person, is practically significant.
-But we can equally well use a theory-informed region of parameter values to describe and make inferences about the heterogeneity distribution of this effect in the population.
-
-For example, let us imagine that a theory states that valence effects in the interval \[-0.1, 0.1\] are practically equivalent to zero.
+But we can equally well use a theory-informed region of effect sizes to describe and make inferences about the heterogeneity distribution of this effect in the population.
+
+For example, let us imagine that a theory states that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
 To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
 In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero.
-Note that this statement's validity critically depends on the chosen interval.
+Note that this statement's validity critically depends on the chosen interval's theoretical validity.
 We visualize this probability in @fig-1 D.
@@
 
-[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that heterogeneity can be deemed noteworthy when the ratio of the standard deviation to the average effect is 0.25 or greater: A ratio greater than 1/4 indicates a $HI_{95}$ whose limits extend to effects one-half and one-and-a-half times that of the average effect.
+[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that heterogeneity can be deemed noteworthy when the ratio of the standard deviation to the average effect is 0.25 or greater: A ratio greater than 1/4 implies a $HI_{95}$ whose limits extend to effects one-half and one-and-a-half times that of the average effect.
 With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r ratio`, suggesting that the degree of heterogeneity in valence effects is noteworthy.
@@
 
-Assessments of heterogeneity involve combining information about fixed and random effects; to fully incorporate inferential uncertainty then requires accounting for their uncertainties in combination.
+Assessments of heterogeneity involve combining information about fixed and random effects; to fully incorporate inferential uncertainty then requires accounting for their joint uncertainties.
 Probabilistic, that is, Bayesian methods are uniquely able to address this challenge.
@@
 However, closed-form solutions are not available for multivariate posterior distributions of many important types of statistical models.
-Therefore in practice modern Bayesian methods rely on algorithms that yield many random draws from the multivariate posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016].
+Therefore, in practice modern Bayesian methods rely on algorithms that yield many random draws from the multivariate posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016].
 These draws can then be used to calculate, summarize, and visualize any desired quantity of the multivariate posterior such as means, variances, correlations, proportions above or below zero, and so on.
@@
 We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
+(These, and other details, are presented in our supplementary online analyses.)
 
@@
 We now have 4,000 samples from this heterogeneity distribution's posterior distribution.
+Effectively, then, we have among other quantities 4,000 samples from the heterogeneity distribution's posterior distribution.
 We first redraw the expected heterogeneity distribution from @fig-1 using the posterior mean values of $\beta_1$ and $\tau_1$ in @fig-2 A (thick dark blue curve).
@@
 #| label: fig-2
-#| fig-width: 4
-#| fig-height: 1.8
-#| fig-cap: Bayesian estimates of the heterogeneity distribution of valence effects.
-#| apa-note: "A: Probability density function (PDF) curves.
+#| fig-width: 5.4
+#| fig-height: 2.2
+#| fig-cap: "Bayesian estimates of the heterogeneity distribution of valence effects.
+A: Probability density function (PDF) curves.
 The thick line is the same expected PDF of valence effects from Figure 1.
@@
 
-@fig-intervals A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions.
+```{r}
+#| label: fig-intervals
+#| fig-height: 2.4
+#| fig-cap: Bayesian estimates of the 90% heterogeneity interval of valence effects.
+A.
+Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$.
+B.
+Histograms of 4,000 samples of the HI90 lower (left) and upper (right) limits with their posterior means and 95%CIs as points and intervals.
+C.
+100 random samples from the $HI_{90}$ posterior distribution, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
+
+p1 <- post1 |>
+  mutate(
+    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+  ) |>
+  ggplot(aes(hi_low, hi_high)) +
+  labs(x = "HI90 lower limit", y = "HI90 upper limit") +
+  geom_point(
+    alpha = .25,
+    shape = 1,
+    size = 0.75,
+    col = "dodgerblue1"
+  ) +
+  theme(aspect.ratio = 1)
+
+p2 <- post1 |>
+  mutate(
+    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+  ) |>
+  select(hi_low, hi_high) |>
+  pivot_longer(everything()) |>
+  ggplot(aes(value, group = name)) +
+  scale_y_continuous(
+    expand = expansion(c(0, 0.1))
+  ) +
+  scale_x_continuous(
+    "Valence effect"
+  ) +
+  stat_histinterval(
+    breaks = 100,
+    fill = alpha("dodgerblue1", .33),
+    color = "dodgerblue4",
+    .width = c(.95),
+    justification = .01
+  ) +
+theme(
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank()
+  )
+
+p3 <- post1_density_samples |> 
+  distinct(
+    .draw, 
+    b_valence1, 
+    sd_person__valence1
+) |>
+  arrange(b_valence1) |>
+  mutate(i = 1:n()) |>
+  mutate(
+    dist = dist_normal(b_valence1, sd_person__valence1)
+  ) |>
+  ggplot() +
+  scale_x_continuous(
+    "Valence effect",
+    expand = expansion(0.025),
+    breaks = extended_breaks()
+  ) +
+  scale_y_continuous(
+    expand = expansion(0.025)
+  ) +
+  labs(y = "Posterior draw") +
+  stat_pointinterval(
+    aes(xdist = dist, y = i),
+    .width = .9,
+    interval_size_range = c(.15, .3),
+    alpha = .5,
+    color = "dodgerblue1"
+  ) +
+  geom_pointinterval(
+    data = post1 |>
+      summarise(
+        across(c(b_valence1, sd_person__valence1), mean)
+      ) |>
+      mutate(
+        hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
+        hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
+      ),
+    aes(y = 50, x = b_valence1, xmin = hi_low, xmax = hi_high),
+    color = "dodgerblue4"
+  )
+
+(p1 | p2 | p3) +
+  plot_annotation(tag_levels = "A")
+```
+
+@fig-intervals further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome but also ignores potential correlations between the HI endpoints' posterior distributions (panel A).
 For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2024] the scalar descriptors discussed below.
@@
 Moreover, if theory allows defining a range of parameter values that are practically equivalent to zero (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically negligible effects.
-We added dotted vertical lines in @fig-2 to highlight the \[-0.1, 0.1\] interval, which serves as an example ROPE.
+We added dotted vertical lines in @fig-2 to highlight the [-0.1, 0.1] interval, which serves as an example ROPE.
 Line segments within that interval represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$).
 To quantify uncertainty in $p^{ROPE}$ we then aggregate the segments' to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
-We note that the ROPE of \[-0.1, 0.1\] here was arbitrary and picked just to illustrate the example.
+We note that the ROPE of [-0.1, 0.1] here was arbitrary and picked just to illustrate the example.
 
@@
 The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
-If we used the 1/4 rule of thumb suggested in @bolgerCausalProcessesPsychology2019, with these results we could say *with confidence* that heterogeneity in valence effects is notable.
+If we used the 1/4 rule of thumb suggested in @bolgerCausalProcessesPsychology2019, with these results we could say *with confidence* that heterogeneity in valence effects is notable (the entire 95%CI exceeds 0.25).
 
@@
 #| label: fig-ratio
-#| fig-height: 2.2
-#| fig-width: 4
+#| fig-height: 2.8
+#| fig-width: 4.4
 #| fig-cap: Bivariate posterior of $\beta_1$, $\tau_1$, and their ratio.
-#| apa-note: Panel A.
+Panel A.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
@@
 We now move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of study units.
-To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2024 addressing differences in between-person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment.
+To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2024 addressing differences in between-person variability (heterogeneity) in memory performance between a free recall memory task and a cued recall memory task.
 In @mahVariabilitySubjectsFree2024's Experiment 3, 260 individuals studied a list of twenty target words.
@@
 Thus, the Free and Cued recall tasks had different groups of participants but the same target words.
+The metric of memory performance in this study was the proportion correct.
 We show a sample of these data in @tbl-dat2.
@@
 
-With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2024 found that participants who completed the cued recall task were more heterogeneous in their memory performance than those in the free recall group: The Cued:Free recall between-person memory performance variance ratio was 1.33 (with a \[1.14, 1.54\] 95% bootstrap interval).
+With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2024 found that participants who completed the cued recall task were more heterogeneous in their memory performance---the proportion of target words correctly recalled---than those in the free recall group: The Cued:Free recall between-person memory performance variance ratio was 1.33 (with a [1.14, 1.54] 95% bootstrap interval).
 @mahVariabilitySubjectsFree2024, across three experiments, confirmed this result by comparing models that did and did not allow for distinct between-person variabilities in each group.
@@
 Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.
-
-```{=tex}
+We write this model as
+
+$$
 \begin{align*}
-\tag{2}
 \text{Accuracy}_{ijkm} &\sim \operatorname{Bernoulli}\left(\pi_{jkm}\right) \\
@@
 \end{align*}
-```
-This Model 2 of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$).
+$$ {#eq-m2-1}
+
+This Model 2 (@eq-m2-1) of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$).
 In addition, instead of coding the task type (free recall vs.
 cued recall) using predictor coding schemes such as contrast or dummy coding, we have index-coded task using subscripts $_{m:F}$ to stand for Free recall parameters, and $_{m:C}$ for parameters pertaining to the Cued recall task.
-This reparameterization is useful, because it allows us to quantify heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference.
+This reparameterization allows quantifying heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference.
 
@@
 
-Descriptively, we reproduced @mahVariabilitySubjectsFree2024's finding that participants' memory performance was more heterogeneous in the cued recall than in the free recall experiment (rows 3 and 4 in @tbl-fit-2).
+Descriptively, we reproduced @mahVariabilitySubjectsFree2024's finding that participants' memory performance was more heterogeneous in the cued recall task than in the free recall task (rows 3 and 4 in @tbl-fit-2).
 We show the relevant estimated quantities, and the implied heterogeneity distributions in @fig-2-person.
 
-The top panel of @fig-2-person A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall).
-Notice that the model's parameters refer to probits ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` probits in the free and cued recall conditions, respectively, the corresponding probabilities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued_p"]`.
+The top panel of @fig-2-person A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` probits in the free and cued recall conditions, respectively, the corresponding probabilities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued_p"]`.
+Notice that the model's parameters refer to probits (standard normal deviates, or "z-scores") because of the link function we used.
+Therefore, for example zero translates to 50% accuracy.
 
@@
 #| fig-height: 3.2
-#| fig-cap: Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
-#| apa-twocolumn: true
-#| apa-note: "A.
-Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
-Differences calculated as Cued - Free recall.
-$p^+$ indicates the estimated fraction of the population with memory performance above 0.5.
-Heterogeneity ratio indicates standard deviations divided with their respective means.
+#| fig-cap: "Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
+A.
+Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95%CIs.
+Differences are calculated as Cued - Free recall.
+$p^+$ indicates the proportion of the population whose proportion correct is predicted to be above 50%.
+Heterogeneity ratio indicates standard deviations divided with their respective means (we truncated this axis at [-7, 7] for clarity.
 B.
 Probability density (top) and cumulative distribution functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
-The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above 0.5."
+The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95%CIs, of the proportions of the populations with memory performance above 0.5."
 
@@
     ) |> fct_rev(),
-    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
+    value = if_else(type == "Heterogeneity ratio" & abs(value) > 7, NaN, value)
   ) |> 
@@
     "Parameter value",
-    expand = expansion(0.025)
-  ) +
-  geom_vline(
-    xintercept = 0, linewidth = .2, alpha = .25
+    expand = expansion(0.025),
+    breaks = extended_breaks(6)
   ) +
@@
 
-Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of @fig-2-person A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
-(We truncated the Ratio axis at \[-10, 10\] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations.
-The difference in ratios is very uncertain partly for the same reason.)
-
-We also depict the heterogeneity distributions as PDFs and CDFs in @fig-2-person B.
-Unlike in @fig-2 that drew random draws of the functions' posteriors, to reduce overplotting @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light).
-These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks.
+Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom row of @fig-2-person A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
+This heterogeneity ratio's mathematical equivalent is commonly known as the coefficient of variation (CV), which is used frequently in many areas of psychological research, such as psychophysics.
+In those areas, a common finding is that while there might be experimental effects on an individual's response distribution's mean or dispersion, the CV frequently remains stable across conditions (disperion tends to grow larger with increased stimulus strength, for example).
+Our use of the heterogeneity ratio in this example calls to mind those applications and findings regarding the coefficient of variation.
+
+Moreover, we truncated the Heterogeneity ratio panel's x-axis at [-7, 7] because ratios of two normal distributions with zero means are Cauchy distributed.
+Sampling from a Cauchy distribution frequently returns extreme draws because of the distribution's thick tails.
+Consequently, posterior draws of $\frac{\tau}{\beta}$ can approximate a Cauchy-distribution and therefore exhibit frequent extreme values.
+These extreme values would obscure the bulk of the distribution if the axis was not truncated.
+More colloquially, near-zero means will necessarily lead to infinite ratios, and consequently this coefficient can be very sensitive to small changes in the mean value.
+The difference in ratios is very uncertain for the same reason.
+
+We also depict the heterogeneity distribution's posterior distribution as a PDF and a CDF in @fig-2-person B.
+Unlike in @fig-2 where we drew random draws of the functions' posteriors as thin lines, to reduce overplotting @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light areas).
+These figures allow for concise and complementary descriptions of (differences in) heterogeneity in the two tasks.
+In other words, they allow visually comparing the population distributions of memory performance across the cued and free-recall tasks.
 
@@
 Between-person heterogeneity is typically the more theoretically important phenomenon for psychologists than differences in model parameters between other randomly sampled study units, such as stimuli.
-However, examining heterogeneity in other sampled units can be both theoretically and methodologically important.
+However, examining heterogeneity in other sampled units can be both theoretically and methodologically important [@juddTreatingStimuliRandom2012; @juddExperimentsMoreOne2017].
 We next turn to our second and third questions regarding potential differences and consistencies in between-target word heterogeneity.
@@
 #| fig-width: 7.4
-#| apa-twocolumn: true
-#| fig-cap: Heterogeneity between target words in memory performance in Free recall and Cued recall tasks from Model 2.
-#| apa-note: "A.
+#| fig-cap: "Heterogeneity between target words in memory performance in Free recall and Cued recall tasks from Model 2.
+A.
 Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
@@
     ) |> fct_rev(),
-    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
+    value = if_else(type == "Heterogeneity ratio" & abs(value) > 5, NaN, value)
   ) |> 
@@
     expand = expansion(0.1)
-  ) +
-  geom_vline(
-    xintercept = 0, linewidth = .2, alpha = .25
   ) +
@@
   aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
-  geom_vline(
-    xintercept = 0, linewidth = .2, alpha = .25
-  ) +
   scale_color_brewer(
@@
   ) +
-  # geom_segment(
-  #   data = dat2_target_coefs,
-  #   aes(x = Free, xend = .Free, y = Cued, yend = .Cued)
-  # ) +
-  # geom_point(
-  #   data = dat2_target_coefs,
-  #   aes(x = Free, y = Cued),
-  #   shape = 21, fill = "white"
-  # ) +
-  # geom_point(
-  #   data = dat2_target_coefs,
-  #   aes(x = .Free, y = .Cued),
-  #   color = "dodgerblue4"
-  # ) +
   theme(
@@
 
-The results regarding differences in memory performance heterogeneity across target words were similar to those as observed regarding heterogeneity in people's memory performances.
+The results regarding differences in heterogeneity across target words' proportions correct were similar to those as observed regarding heterogeneity in people's memory performances.
 @fig-2-target A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
@@
 Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than 50% accurate recall in the cued recall than in the free recall task.
-Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared nearly identical across the two tasks.
-
-This analysis of target word heterogeneity afforded an additional piece of information, because the same target words were used across the two tasks, and thus allowed answering our third question.
-There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of @fig-2-target A, and @fig-2-target C).
+Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared very similar across the two tasks.
+
+The design of @mahVariabilitySubjectsFree2024's study and our analysis of the dataset afforded an additional piece of information: Because the same target words were used across the two tasks, we could examine the consistency of target words' heterogeneity across the two tasks (question (3)).
+There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of @fig-2-target A and C).
 The posterior mean and 95%CI of this correlation was `r post2_sum$x[post2_sum$variable=="cor_target__taskFree__taskCued"]`.
 
-What this correlation means substantively is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task.
-(@bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.)
-
-Although outside our current scope, we note that our study of Model 2's results might indicate exciting new avenues for this line of inquiry.
-One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2024].
-Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (@fig-2-target A).
+This correlation's substantive interpretation is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task.
+(@bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.) For example, the tools presented here would facilitate seeking for theoretically interesting conditions where this consistency is violated.
+
+Our study of Model 2's results might indicate exciting new avenues for this line of inquiry.
+One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different recall strategies in the two tasks [@mahVariabilitySubjectsFree2024].
+Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that accuracy is more heterogeneous across target words in cued than in the free recall task (@fig-2-target A).
 Second, we observed across both people and target words that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
@@
 In the current work, we illustrated the use of practical descriptors of heterogeneity with examples drawn from social and cognitive psychology.
-Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects in psychology---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
+Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity.
 Our currently proposed methods incorporate uncertainty into both modelling of and inferences about heterogeneity.
 
-Although prior work on developing metrics of heterogeneity, and placing experimental effect sizes in context of person-specific effects exists, it has ignored estimation uncertainty and thus remained purely descriptive.
-For example, @gricePersonsEffectSizes2020 describe a method whereby analysts simply count the number of individuals whose point estimate of an effect is concordant with a hypothesis.
+Although prior work on developing metrics of heterogeneity, and placing experimental effect sizes in context of person-specific effects exists, it has largely ignored estimation uncertainty and thus remained purely descriptive.
+For example, @gricePersonsEffectSizes2020 describe a method whereby analysts count the number of individuals whose point estimate of an effect is concordant with a hypothesis.
 But such counting ignores estimation uncertainty in both the person-specific effects and variability among them.
 By accounting for these uncertainties, the methods described here go beyond description and allow inference to be drawn regarding populations and individuals with confidence.
+Moreover, counting individuals' parameters provides a description of individuals in the sample, rather than of the population, which was our focus.
 
@@
 
-We emphasized throughout that a probabilistic (Bayesian) approach is uniquely positioned to answer the needs of researchers interested in heterogeneity.
+We emphasized throughout that a probabilistic (Bayesian) approach is well-positioned to answer the needs of researchers interested in heterogeneity.
 Bayesian methods allow carrying uncertainty forward from model parameters to descriptors of heterogeneity and beyond.
@@
 In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
+However, some methods described here could be implemented with e.g.
+joint bootstrap methods, but in our view those require additional practical steps---bootstrapping, for one---and might therefore be less practical.
 
@@
 
-# Data and code availability
-
-All the code supporting this manuscript is available at <https://github.com/mvuorre/heterogeneity-uncertainty>.
-We reused openly available datasets [@bolgerCausalProcessesPsychology2019; @mahVariabilitySubjectsFree2024].
-
-# Author contributions
+# Disclosures
+
+## Data and code availability
+
+The online analysis supplement is readable at {{<meta site-url>}}.
+Our materials are available at GitHub ({{<meta repo-url>}}) and the OSF ({{<meta osf-url>}}).
+We reused openly available datasets from @bolgerCausalProcessesPsychology2019 and @mahVariabilitySubjectsFree2024.
+
+## Author contributions
 
@@
 
-\noindent Conceptualization: MV, NB\
+Conceptualization: MV, NB\
 Formal Analysis: MV\
@@
 
-# Competing interests
+## Competing interests
 
@@
 
-# References {.unnumbered}
-
-::: {#refs custom-style="Bibliography"}
+<!-- Format references better in non-html formats -->
+::: {.content-hidden when-format="html"}
+# References
+
+:::{.refs}
 :::
-
-<!-- Prepare appendix section -->
-
-```{=tex}
-\clearpage
-\setcounter{page}{1}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\onecolumn
-```
-# Appendix
-
-```{r}
-#| label: fig-intervals
-#| fig-height: 2.4
-#| fig-cap: Bayesian estimates of the 90% heterogeneity interval of valence effects. 
-#| apa-note: A.
-Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$.
-B.
-Histograms of 4,000 samples of the HI90 lower (left) and upper (right) limits with their posterior means and 95%CIs as points and intervals.
-C.
-100 random samples from the $HI_{90}$ posterior distribution, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
-
-p1 <- post1 |>
-  mutate(
-    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
-    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
-  ) |>
-  ggplot(aes(hi_low, hi_high)) +
-  labs(x = "HI90 lower limit", y = "HI90 upper limit") +
-  geom_point(
-    alpha = .25,
-    shape = 1,
-    size = 0.75,
-    col = "dodgerblue1"
-  ) +
-  theme(aspect.ratio = 1)
-
-p2 <- post1 |>
-  mutate(
-    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
-    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
-  ) |>
-  select(hi_low, hi_high) |>
-  pivot_longer(everything()) |>
-  ggplot(aes(value, group = name)) +
-  scale_y_continuous(
-    expand = expansion(c(0, 0.1))
-  ) +
-  scale_x_continuous(
-    "Valence effects"
-  ) +
-  stat_histinterval(
-    breaks = 100,
-    fill = alpha("dodgerblue1", .33),
-    color = "dodgerblue4",
-    .width = c(.95),
-    justification = .01
-  ) +
-theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-p3 <- post1_density_samples |> 
-  distinct(
-    .draw, 
-    b_valence1, 
-    sd_person__valence1
-) |>
-  arrange(b_valence1) |>
-  mutate(i = 1:n()) |>
-  mutate(
-    dist = dist_normal(b_valence1, sd_person__valence1)
-  ) |>
-  ggplot() +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0.025),
-    breaks = extended_breaks()
-  ) +
-  scale_y_continuous(
-    expand = expansion(0.025)
-  ) +
-  labs(y = "Posterior sample") +
-  stat_pointinterval(
-    aes(xdist = dist, y = i),
-    .width = .9,
-    interval_size_range = c(.15, .3),
-    alpha = .5,
-    color = "dodgerblue1"
-  ) +
-  geom_pointinterval(
-    data = post1 |>
-      summarise(
-        across(c(b_valence1, sd_person__valence1), mean)
-      ) |>
-      mutate(
-        hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
-        hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
-      ),
-    aes(y = 50, x = b_valence1, xmin = hi_low, xmax = hi_high),
-    color = "dodgerblue4"
-  )
-
-(p1 | p2 | p3) +
-  # coord_cartesian(xlim = c(-0.6, 0.3)) &
-  # scale_x_continuous(
-  #   "Valence effect",
-  #   expand = expansion(0),
-  #   breaks = extended_breaks(7)
-  # )) +
-  # plot_layout(widths = c(6, 4)) +
-  plot_annotation(tag_levels = "A")
-```
+:::
+
+<!-- Material after this only appears in html output -->
+::: {.content-visible when-format="html"}
+
+:::

### 2024-07-04 index.qmd
draft figures describing descriptor construction

@@
 csl: apa.csl
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 )
+
+dist_ml <- dist_normal(mu1n, sd1n)
 ```
@@
 B.
-The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area.
+The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the blue shaded area.
 C.
-Proportion of negative valence effects in the population (dark).
+Proportion of negative valence effects in the population (blue).
 D.
-Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).
-
-fig_0 <- ggplot() +
+Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; blue).
+
+# annotation definitions
+valence_dist_color = "gray25"
+valence_dist_samples_color = "gray65"
+desc_color = "dodgerblue1"
+desc_color_dark = "dodgerblue4"
+zero_color = alpha("black", 0.75)
+
+arrowhead <- \(...) arrow(length = unit(4, "pt"), type = "closed", ...)
+
+fig_1 <- ggplot() +
   aes(xdist = dist_normal(mu1n, sd1n)) +
@@
     axis.title.x = element_text(lineheight = rel(0.5))
-  ) +
-  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white")
+  )
+
 d1 <- dist_normal(mu1n, sd1n) |>
@@
 
-fig_1 <- fig_0 +
+fig_1_params <- fig_1 +
   stat_slab(
@@
     linewidth = .35,
-    arrow = arrow(length = unit(4, "pt"), type = "closed")
+    arrow = arrowhead()
   ) +
@@
     linewidth = .35,
-    arrow = arrow(length = unit(4, "pt"), type = "closed")
-  ) +
+    arrow = arrowhead()
+  ) +
+  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white") +
   theme(axis.title.y = element_text(angle = 90))
 
-fig_2 <- fig_0 +
+fig_1_descriptors <- fig_1 +
+  scale_fill_manual(values = c("grey90", alpha(desc_color, 0.45)), na.value = "white")
+
+fig_1_hi <- fig_1_descriptors +
   stat_slab(
@@
     linewidth = .35,
-    arrow = arrow(length = unit(4, "pt"), type = "closed", ends = "both")
-  )
-
-fig_3 <- fig_0 +
+    arrow = arrowhead(ends = "both")
+  )
+
+fig_1_p_neg <- fig_1_descriptors +
   stat_slab(
@@
 
-fig_4 <- fig_0 +
+fig_1_p_rope <- fig_1_descriptors +
   stat_slab(
@@
     x = 0.0, y = 0.5, hjust = 0,
-    size = 3,
+    size = 3.5,
     parse = TRUE
@@
 
-(fig_1 | fig_2 | fig_3 | fig_4) &
+(fig_1_params | fig_1_hi | fig_1_p_neg | fig_1_p_rope) &
   scale_y_continuous(
@@
 
+```{r}
+#| label: fig-hi-point
+#| fig-width: 2.7
+#| fig-height: 2.2
+#| fig-cap: "Construction of point estimates of the limits of the heterogeneity interval.
+To construct a $\\pi%$ heterogeneity interval, $HI_\\pi$, we use the inverse cumulative distribution function (CDF) of the heterogeneity distribution, $\\Phi^{-1}$.
+When $\\pi = 0.90$, we have $HI_90 = \\Phi^{-1}\\left((1 \\pm \\pi)/2 | \\beta_1, \\tau_1 \\right) = \\Phi^{-1}\\left([0.05, 0.95] | \\beta_1, \\tau_1 \\right)$."
+
+# Shared Y axis
+scale_y_01 <- scale_y_continuous(
+  breaks = breaks_extended(7),
+  expand = expansion(c(0.005, 0.005)),
+  limits = c(0, 1)
+)
+
+# shared x axis
+xmin <- -0.6
+xmax <- 0.3
+scale_x_valence <- \(..., breaks = extended_breaks()) scale_x_continuous(
+  ...,
+  limits = c(xmin, xmax),
+  expand = expansion(0),
+  breaks = breaks
+)
+
+# ML estimate of CDF
+p_ml_cdf <-
+  ggplot() +
+  aes(xdist = dist_ml) +
+  stat_slab(
+    aes(thickness = after_stat(thickness(cdf))),
+    scale = 1,
+    linewidth = .8,
+    color = valence_dist_color,
+    fill = NA
+  ) +
+  geom_vline(
+    xintercept = 0, 
+    linewidth = .25,
+    color = zero_color,
+    linetype = "dashed"
+  ) +
+  theme(legend.position = "none") +
+  scale_y_01 +
+  xlab("Valence effect") +
+  ylab(quote(Pr("Valence effect" <= italic(x))))
+
+# construction of HI_90 on the CDF
+p_ml_cdf_hi <- p_ml_cdf +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_ml, 0.05), y = 0.05, yend = 0.05),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.05), y = 0.05, yend = 0),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_ml, 0.95), y = 0.95, yend = 0.95),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.95), xend = quantile(dist_ml, 0.95), y = 0.95, yend = 0),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.95), y = 0.025, yend = 0.025),
+    color = desc_color,
+    arrow = arrowhead(ends = "both")
+  ) +
+  geom_text(
+    aes(x = mu1n, y = .04),
+    label = expression(bold(HI)[bold("90")]),
+    data = data.frame()[1,],
+    size = 3,
+    parse = TRUE,
+    color = desc_color,
+    vjust = 0
+  ) +
+  annotate("text",
+    x = xmin,
+    y = c(0.05, 0.95),
+    vjust = c(0.3, 0.7),
+    label = c(0.05, 0.95),
+    color = desc_color,
+    hjust = 1.1,
+    fontface = "bold",
+    size = 2.5
+  ) +
+  scale_x_valence() +
+  coord_cartesian(clip = "off", ylim = c(0,1))
+
+p_ml_cdf_hi
+```
+
 To calculate a heterogeneity interval, we first specify the desired probability limits.
 For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution.
-Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95] | \beta_1, \tau_1) = \Phi^{-1}([.05, .95] | $ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
 In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
-We illustrate this interval in @fig-1 B.
+We illustrate this interval in @fig-1 B and its construction in @fig-hi-point.
 
@@
 
-To answer, we pass zero (the critical value) and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
+```{r}
+#| label: fig-p-neg-point
+#| fig-width: 2.7
+#| fig-height: 2.2
+#| fig-cap: "Construction of the point estimate of $p^{-1}$, the proportion of individuals in the population with negative effects.
+To construct $p^{-1}$, we pass 0 to the cumulative distribution function (CDF) of the heterogeneity distribution."
+
+# construction of p^- on the CDF
+p_ml_cdf_p_neg <- p_ml_cdf +
+  geom_segment(
+    aes(x = 0, xend = 0, y = 0, yend = cdf(dist_ml, 0)),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = 0, xend = I(0), y = cdf(dist_ml, 0)),
+    color = desc_color,
+    arrow = arrowhead()
+  ) + 
+  annotate("text", 
+    x = 0, y = I(0),
+    label = "0", 
+    color = desc_color, 
+    vjust = 1.3, 
+    fontface = "bold", 
+    size = 3.5
+  ) +
+  geom_text( 
+    aes(x = I(0), y = cdf(dist_ml, 0)),
+    label = expression(bolditalic(p ^ "−")), 
+    color = desc_color, 
+    hjust = 1.2, 
+    fontface = "bold", 
+    size = 3.5
+  ) +
+  scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
+  coord_cartesian(clip = "off", ylim = c(0,1))
+
+p_ml_cdf_p_neg
+```
+
+To answer, we pass zero (the critical value) and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = Pr(\textrm{Valence effect} \le 0) = \Phi(0 | \beta_1, \tau_1) = \Phi(0 | $ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
 This number is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
-We illustrate this probability in @fig-1 C.
+We illustrate this probability in @fig-1 C and its construction using the CDF in @fig-p-neg-point.
 
@@
 For example, let us imagine that a theory states that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
-To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
+To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = Pr(-0.1 \le \textrm{Valence effect} \le 0.1) = \Phi(0.1 | \beta_1, \tau_1) - \Phi(-0.1 | \beta_1, \tau_1)$ = `r p_rope`.
 In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero.
@@
   regex = TRUE
-)
+) |>
+  mutate(dist_valence = dist_normal(b_valence1, sd_person__valence1))
 ```
@@
 set.seed(99)
-x <- seq(-0.6, 0.3, length = 301)
+x <- seq(xmin, xmax, length = 301)
 
@@
 
-post1_density_mean <- post1 |>
+post1_mean <- post1 |>
   mean_qi(b_valence1, sd_person__valence1) |>
+  mutate(dist_valence = dist_normal(b_valence1, sd_person__valence1))
+
+post1_density_mean <- post1_mean |>
   calculate_density(b_valence1, sd_person__valence1)
 
-post1_density_samples <- post1 |>
-  slice_sample(n = 100) |>
+post1_samples <- post1 |>
+  slice_sample(n = 100)
+
+post1_density_samples <- post1_samples |>
   calculate_density(b_valence1, sd_person__valence1)
@@
 Thin lines show 100 PDFs calculated from random draws of $\\beta_1$ and $\\tau_1$ that collectively illustrate the uncertainty in the distribution's location and spread.
-B: Cumulative density function (CDF) curves, annotated as A.
-C.
-Histogram of (draws from) the posterior distribution of $p^-$, which correspond to CDF segments below zero: This value describes the proportion of individuals with negative valence effects on logRT.
-Point and interval represents the posterior mean and 95\\%CI.
-Solid and dotted vertical lines in A and B highlight x-axis values of 0, and -0.1 and 0.1, respectively."
-
-# Shared Y axis
-scale_y_01 <- scale_y_continuous(
-  breaks = extended_breaks(7),
-  expand = expansion(c(0.005, 0.005)),
-  limits = c(0, 1)
-)
+B: Cumulative distribution function (CDF) curves, annotated as in A.
+The blue lines show the calculation of the point estimate of $p^- = Pr(\\textrm{Valence effect} \\le 0)$, the proportion of individuals with negative valence effects on logRT, using the mean CDF of valence effects."
 
 # Posterior samples of pdf and cdf curves
-p_cdf_samples <- post1_density_mean |>
+p_samples <- post1_density_mean |>
   ggplot() +
-  aes(x, cdf) +
-  ylab("Cumulative probability") +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0),
-    breaks = extended_breaks()
-  ) +
-  scale_y_01 +
-  geom_vline(
-    xintercept = 0, 
-    linewidth = .25
-  ) +
-  geom_vline(
-    xintercept = c(-0.1, 0.1), 
-    lty = "dotted", 
-    linewidth = .2
-  ) +
+  aes(x) +
   geom_line(
@@
     aes(group = .draw),
-    alpha = .33,
+    alpha = .3,
     linewidth = .15,
-    col = "dodgerblue1"
+    color = valence_dist_samples_color
   ) +
@@
     linewidth = .8,
-    col = "dodgerblue4"
-  ) +
-  theme(legend.position = "none")
-
-p_pdf_samples <- p_cdf_samples +
+    col = valence_dist_color
+  ) +
+  geom_vline(
+    xintercept = 0, 
+    linewidth = .25,
+    color = zero_color,
+    linetype = "dashed"
+  ) +
+  theme(legend.position = "none") +
+  xlab("Valence effect")
+
+p_cdf_samples <- p_samples +
+  aes(y = cdf) +
+  scale_x_valence() +
+  scale_y_01 +
+  ylab(quote(Pr("Valence effect" <= italic(x))))
+
+p_pdf_samples <- p_samples +
   aes(y = pdf) +
+  scale_x_valence() +
   scale_y_continuous(
@@
 
-# Add summaries of CDF < 0 to CDF curves as inset plots
-p_p0_samples <- post1 |>
-  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |>
-  ggplot() +
-  scale_x_continuous(
-    expand = expansion(c(0.0, 0.1))
-  ) +
-  scale_y_01 +
-  theme_void(base_size = 9) +
-  stat_histinterval(
-    aes(y = cdf(dist, 0)),
-    breaks = 50,
-    fill = alpha("dodgerblue1", 0.5),
-    color = "dodgerblue4",
-    side = "right",
-    .width = .95,
-    size = 1,
-    scale = 0.5,
-    justification = 0.025
-  )
-
-p_cdf_samples <- p_cdf_samples +
-  inset_element(
-    p_p0_samples,
-    0, 0, .3, 1, 
-    align_to = "panel", 
-    on_top = TRUE
-  ) +
-  theme(plot.tag.position = c(0.25, 0.75))
-
-(p_pdf_samples | p_cdf_samples) +
+p_pdf_samples + p_cdf_samples +
   plot_annotation(tag_levels = "A")
@@
 
-Therefore, in @fig-2 B we depict the heterogeneity distribution as *cumulative distribution* function (CDF) curves based on 100 random posterior draws, together with the mean CDF in a darker color.
+Therefore, in @fig-2 B we depict the heterogeneity distribution as *cumulative distribution function* (CDF) curves based on 100 random posterior draws, together with the mean CDF in a darker color.
 We believe the CDF is a useful visualization tool because the y-axis describes a directly interpretable quantity: The proportion of the population with valence effects below some specific value.
@@
 
+```{r}
+#| label: fig-hi-samples
+#| fig-width: 2.7
+#| fig-height: 2.2
+#| fig-cap: "Bayesian derivation of the posterior distribution for the lower and upper limits of the heterogeneity interval.
+Thin gray lines show 100 CDFs of the distribution of valence effects calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
+Blue lines show the calculation of the posterior distribution of $HI_{90} = \\left[\\Phi^{-1}(0.05 | \\beta_1, \\tau_1), \\Phi^{-1}(0.95 | \\beta_1, \\tau_1)\\right]$, which is also depicted as marginal histograms.
+Compare to the point estimate of the same interval endpoints in @fig-hi-point."
+
+# Add summaries of CDF < 0 to CDF curves as inset plots
+p_cdf_intervals <- p_cdf_samples +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_valence, 0.05), y = 0.05, yend = 0.05),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_valence, 0.05), xend = quantile(dist_valence, 0.05), y = 0.05, yend = 0),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0.95),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_valence, 0.95), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  stat_histinterval(
+    aes(x = quantile(dist_valence, 0.05)),
+    y = 0,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 0.1
+  ) +
+  stat_histinterval(
+    aes(x = quantile(dist_valence, 0.95)),
+    y = 0,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 0.1
+  ) +
+  annotate("text",
+    x = min(x),
+    y = c(0.05, 0.95),
+    vjust = c(0.3, 0.7),
+    label = c(0.05, 0.95),
+    color = desc_color,
+    hjust = 1.1,
+    fontface = "bold",
+    size = 2.5
+  ) +
+  scale_x_valence() +
+  scale_y_continuous(
+    breaks = breaks_extended(7),
+    expand = expansion(0)
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0,1))
+
+p_cdf_intervals
+```
+
+```{r}
+#| label: fig-rope-densities
+#| fig-width: 5.4
+#| fig-height: 2.8
+#| fig-cap: "Bayesian derivation of the posterior distribution for $p^{ROPE}$."
+
+post1_samples_rope <- post1_samples |> 
+  slice_head(n = 25)
+
+p_rope_densities <- post1_samples_rope |> 
+  ggplot() + 
+  stat_slab(
+    aes(y = factor(.draw), xdist = dist_valence, fill = after_stat(-0.1 < x & x < 0.1)),
+    color = valence_dist_samples_color,
+    alpha = 0.5,
+    scale = 2.5,
+    linewidth = 0.5
+  ) +
+  geom_vline(xintercept = c(-0.1, 0.1), zero_color, alpha = 0.1, linetype = "22") +
+  scale_fill_manual(values = c("TRUE" = desc_color, "FALSE" = alpha(valence_dist_samples_color, 0.3))) +
+  guides(fill = "none") +
+  coord_cartesian(clip = "off") +
+  scale_x_valence() +
+  scale_y_discrete(labels = NULL) +
+  theme(axis.ticks.y = element_blank()) +
+  ylab("Posterior draw") +
+  xlab("Valence effect")
+
+p_rope_props <- post1_samples_rope |> 
+  slice_head(n = 30) |> 
+  ggplot() + 
+  aes(y = factor(.draw), x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)) + 
+  geom_segment(aes(xend = 0, yend = factor(.draw)), color = desc_color) + 
+  geom_point(color = desc_color, size = 0.75) + 
+  scale_x_continuous(limits = c(0,1), expand = expansion(0)) +
+  scale_y_discrete(labels = NULL) +
+  theme(axis.ticks.y = element_blank()) +
+  ylab(NULL) +
+  xlab(quote(italic(p)^italic(ROPE))) +
+  stat_histinterval(
+    aes(x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)),
+    y = 0.1,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 3
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0.7, nrow(post1_samples_rope)))
+
+p_rope_densities + p_rope_props + plot_annotation(tag_levels = "A")
+```
+
 ### Proportion descriptors
@@
 In this example, we ask "What proportion of individuals in the population endorse positive words faster than negative words?"
+
+```{r}
+#| label: fig-p-neg
+#| fig-width: 2.7
+#| fig-height: 2.2
+#| fig-cap: "Bayesian derivation of the posterior distribution of $p^- = Pr(\\textrm{Valence effect} \\le 0)$, the proportion of individuals with negative valence effects on logRT.
+Thin gray lines show 100 CDFs of the distribution of valence effects calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
+Blue lines show the calculation of the posterior distribution of $p^-$, which is also depicted as a marginal histogram at the top left.
+The point estimate and interval represents the posterior mean and 95\\%CI.
+Compare to the point estimate of $p^-$ in @fig-2 B."
+
+# Add summaries of CDF < 0 to CDF curves as inset plots
+p_cdf_samples_with_p_neg <- p_cdf_samples +
+  geom_segment(
+    aes(x = 0, xend = 0, y = 0, yend = cdf(dist_valence, 0)),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = 0, xend = I(0), y = cdf(dist_valence, 0), yend = cdf(dist_valence, 0)),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  stat_histinterval(
+      aes(y = cdf(dist_valence, 0)),
+      x = I(0),
+      data = post1,
+      breaks = 50,
+      fill = alpha(desc_color, 0.35),
+      color = desc_color_dark,
+      side = "left",
+      .width = .95,
+      scale = 1,
+      width = 0.1
+  ) + 
+  annotate(
+    "text", 
+    x = 0, y = I(0), 
+    label = "0", 
+    color = desc_color, 
+    vjust = 1.3, 
+    fontface = "bold", 
+    size = 3.5
+  ) +
+  geom_text( 
+    aes(x = I(0), y = cdf(dist_valence, 0)),
+    data = post1_mean,
+    label = expression(bolditalic(p ^ "−")), 
+    color = desc_color, 
+    hjust = 1.4, 
+    fontface = "bold", 
+    size = 3.5
+  ) +
+  scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
+  coord_cartesian(clip = "off")
+
+p_cdf_samples_with_p_neg
+```
 
@@
 
-To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
-We show 100 posterior draws of the CDF in @fig-2 B with a vertical line superimposed at zero.
+To answer, we calculate $p^- = Pr(\textrm{Valence effect} \le 0) = \Phi(0 | \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
+We show 100 posterior draws of the CDF in @fig-p-neg with a vertical line superimposed at zero.
 The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$).
-We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-2 C, with its associated 95%CI.
+We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-p-neg, with its associated 95%CI.
 The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
@@
     "https://files.ca-1.osf.io/v1/resources/pfhu9/providers/osfstorage/?zip=",
+    mode = "wb",
     destfile = path
@@
 As is common with generalized linear models, we model the rate parameter through a nonlinear link function.
-In this example, we use the cumulative normal density function ($\Phi$, or "probit"), but other link functions are also available, such as the common logit.
+In this example, we use the cumulative normal distribution function ($\Phi$, or "probit"), but other link functions are also available, such as the common logit.
 Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.

### 2024-07-05 index.qmd
complete figures for descriptor construction

--- 2024-06-25 index.qmd
@@
 csl: apa.csl
+editor_options: 
+  chunk_output_type: console
 ---
@@
 #| fig-cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
-A.
-The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
+**A.** The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
 Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean.
-B.
-The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area.
-C.
-Proportion of negative valence effects in the population (dark).
-D.
-Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).
-
-fig_0 <- ggplot() +
+**B.** The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the blue shaded area.
+**C.** Proportion of negative valence effects in the population (blue).
+**D.** Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; blue).
+
+# annotation definitions
+valence_dist_color = "gray25"
+valence_dist_samples_color = "gray65"
+desc_color = "dodgerblue1"
+desc_color_dark = "dodgerblue4"
+zero_color = alpha("black", 0.75)
+
+arrowhead <- \(...) arrow(length = unit(4, "pt"), type = "closed", ...)
+
+fig_1 <- ggplot() +
   aes(xdist = dist_normal(mu1n, sd1n)) +
@@
     axis.title.y = element_blank(),
-    axis.title.x = element_text(lineheight = rel(0.5))
-  ) +
-  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white")
+    axis.title.x = element_text(lineheight = rel(0.5)),
+    plot.title.position = "plot"
+  ) +
+  scale_y_continuous(
+    expand = expansion(c(0.0, 0.05))
+  ) +
+  labs(
+    y = "Density",
+    x = "Valence effect"
+  ) +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "none"
+  )
+
 d1 <- dist_normal(mu1n, sd1n) |>
@@
     linewidth = .35,
-    arrow = arrow(length = unit(4, "pt"), type = "closed")
-  ) +
+    arrow = arrowhead()
+  ) +
+  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white") +
+  labs(subtitle = "A. Heterogeneity distribution") +
   theme(axis.title.y = element_text(angle = 90))
 
-fig_2 <- fig_0 +
+fig_1_descriptors <- fig_1 +
+  scale_fill_manual(values = c("grey90", alpha(desc_color, 0.45)), na.value = "white")
+
+fig_1_interval <- fig_1_descriptors +
   stat_slab(
@@
     size = 3,
-    parse = TRUE
+    parse = TRUE,
+    color = desc_color_dark
   ) +
@@
     linewidth = .35,
-    arrow = arrow(length = unit(4, "pt"), type = "closed", ends = "both")
-  )
-
-fig_3 <- fig_0 +
+    arrow = arrowhead(ends = "both"),
+    color = desc_color_dark
+  ) +
+  labs(subtitle = "B. Heterogeneity interval")
+
+fig_1_pneg <- fig_1_descriptors +
   stat_slab(
@@
     size = 3,
-    parse = TRUE
-  )
-
-fig_4 <- fig_0 +
+    parse = TRUE,
+    color = desc_color_dark
+  ) +
+  labs(subtitle = "C. Proportion negative")
+
+fig_1_prope <- fig_1_descriptors +
   stat_slab(
@@
     x = 0.0, y = 0.5, hjust = 0,
-    size = 3,
-    parse = TRUE
-  )
-
-(fig_1 | fig_2 | fig_3 | fig_4) &
-  scale_y_continuous(
-    expand = expansion(c(0.0, 0.05))
-  ) &
-  labs(
-    y = "Density",
-    x = "Valence effect"
-  ) &
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "none"
-  ) &
-  plot_annotation(tag_levels = "A")
+    size = 3.5,
+    parse = TRUE,
+    color = desc_color_dark
+  ) +
+  labs(subtitle = "D. Proportion in ROPE")
+
+fig_1_params | fig_1_interval | fig_1_pneg | fig_1_prope
 ```
@@
 
-To calculate a heterogeneity interval, we first specify the desired probability limits.
-For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution.
-Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
-In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
-We illustrate this interval in @fig-1 B.
+```{r}
+#| label: fig-ml-interval
+#| fig-width: 4.5
+#| fig-height: 2.2
+#| fig-cap: "Construction of point estimates of the limits of the heterogeneity interval.
+**A.** While $HI_{90}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF), $\\Phi(x; \\beta_1, \\tau_1)$.
+To construct a 90% heterogeneity interval, we pass 0.05 and 0.95 to the inverse of the CDF: $HI_{90} = \\Phi^{-1}\\left([0.05, 0.95]; \\beta_1, \\tau_1 \\right)$."
+
+# Shared Y axis
+scale_y_01 <- scale_y_continuous(
+  breaks = c(0, 0.5, 1),
+  expand = expansion(c(0.005, 0.005)),
+  limits = c(0, 1)
+)
+
+# shared x axis
+xmin <- -0.6
+xmax <- 0.3
+scale_x_valence <- \(..., breaks = extended_breaks()) scale_x_continuous(
+  ...,
+  limits = c(xmin, xmax),
+  expand = expansion(0),
+  breaks = breaks
+)
+
+# ML estimate of CDF
+fig_ml_cdf <-
+  ggplot() +
+  aes(xdist = dist_ml) +
+  stat_slab(
+    aes(thickness = after_stat(thickness(cdf))),
+    scale = 1,
+    linewidth = .8,
+    color = valence_dist_color,
+    fill = NA
+  ) +
+  geom_text(
+    aes(x = I(1), y = 0.97),
+    label = expression(bold(Phi)),
+    vjust = 1,
+    hjust = 1,
+    family = "serif"
+  ) +
+  theme(legend.position = "none") +
+  scale_y_01 +
+  xlab("Valence effect") +
+  ylab(quote(Pr("Valence effect" <= italic(x)))) +
+  theme(plot.title.position = "plot")
+
+# construction of HI_90 on the CDF
+fig_ml_cdf_interval <- fig_ml_cdf +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_ml, 0.05), y = 0.05, yend = 0.05),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.05), y = 0.05, yend = 0),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_ml, 0.95), y = 0.95, yend = 0.95),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.95), xend = quantile(dist_ml, 0.95), y = 0.95, yend = 0),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.95), y = 0.025, yend = 0.025),
+    color = desc_color,
+    arrow = arrowhead(ends = "both")
+  ) +
+  geom_text(
+    aes(x = mu1n, y = .04),
+    label = expression(bold(HI)[bold("90")]),
+    data = data.frame()[1,],
+    size = 3,
+    parse = TRUE,
+    color = desc_color,
+    vjust = 0
+  ) +
+  annotate("text",
+    x = xmin,
+    y = c(0.05, 0.95),
+    vjust = c(0.3, 0.7),
+    label = c(0.05, 0.95),
+    color = desc_color,
+    hjust = 1.1,
+    fontface = "bold",
+    size = 2.5
+  ) +
+  scale_x_valence() +
+  coord_cartesian(clip = "off", ylim = c(0,1)) +
+  labs(subtitle = TeX("B. $HI_{90}$ on the CDF"))
+
+(
+  (fig_1_interval + labs(subtitle = TeX("A. $HI_{90}$ on the PDF"))) +
+  fig_ml_cdf_interval
+) &
+  theme(axis.title.y = element_text(angle = 90))
+```
+
+To calculate a heterogeneity interval, we first specify the desired probability limits: 
+for a $\pi\%$ interval, we use the limits $(1 \pm \pi)/2$.
+Thus, for a 90% interval, we use 0.05 and 0.95, which together define the central 90% of the distribution.
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95]; \beta_1, \tau_1) = \Phi^{-1}([.05, .95];$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the point estimates of the mean ($\beta_1$) and standard deviation ($\tau_1$): We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
+We illustrate this interval in @fig-ml-interval.
 
@@
 
-To answer, we pass zero (the critical value) and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
+```{r}
+#| label: fig-ml-pneg
+#| fig-width: 4.5
+#| fig-height: 2.2
+#| fig-cap: "Construction of the point estimate of the proportion of negative effects.
+**A.** While $p^{-}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF): $p^{-} = \\Phi(0; \\beta_1, \\tau_1)$."
+
+zero_line = geom_vline(
+  xintercept = 0,
+  linetype = "22",
+  color = zero_color,
+  alpha = 0.25
+)
+
+# construction of p^- on the CDF
+fig_ml_cdf_pneg <- fig_ml_cdf +
+  zero_line +
+  geom_segment(
+    aes(x = 0, xend = 0, y = 0, yend = cdf(dist_ml, 0)),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = 0, xend = I(0), y = cdf(dist_ml, 0)),
+    color = desc_color,
+    arrow = arrowhead()
+  ) + 
+  annotate("text", 
+    x = 0, y = I(0),
+    label = "0", 
+    color = desc_color, 
+    vjust = 1.35, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  geom_text( 
+    aes(x = I(0), y = cdf(dist_ml, 0)),
+    label = expression(bold(p ^ "−")), 
+    color = desc_color, 
+    hjust = 1.2,
+    vjust = 0.6,
+    fontface = "bold", 
+    size = 3
+  ) +
+  scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
+  coord_cartesian(clip = "off", ylim = c(0,1)) +
+  labs(subtitle = TeX("B. $p^-$ on the CDF"))
+
+(
+  (fig_1_pneg + zero_line + labs(subtitle = TeX("A. $p^-$ on the PDF"))) +
+  fig_ml_cdf_pneg
+) &
+  theme(axis.title.y = element_text(angle = 90))
+```
+
+To answer, we pass zero (the critical value) and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): 
+$p^{-} = Pr(\textrm{Valence effect} \le 0) = \Phi(0; \beta_1, \tau_1) = \Phi(0;$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
 This number is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
-We illustrate this probability in @fig-1 C.
+We illustrate this probability in @fig-1 C and its construction using the CDF in @fig-ml-pneg.
 
@@
 
+```{r}
+#| label: fig-ml-prope
+#| fig-width: 4.5
+#| fig-height: 2.2
+#| fig-cap: "Construction of the point estimate of the proportion of effects in the ROPE.
+**A.** While $p^{ROPE}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF): $p^{ROPE} = \\Phi(\\epsilon; \\beta_1, \\tau_1) - \\Phi(- \\epsilon; \\beta_1, \\tau_1)$."
+
+rope_line = geom_vline(
+  xintercept = c(-0.1, 0.1),
+  linetype = "22",
+  color = zero_color,
+  alpha = 0.25
+)
+
+# construction of p^- on the CDF
+fig_ml_cdf_prope <- fig_ml_cdf +
+  rope_line +
+  geom_segment(
+    aes(x = -0.1, xend = -0.1, y = 0, yend = cdf(dist_ml, -0.1)),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = -0.1, xend = I(0), y = cdf(dist_ml, -0.1)),
+    color = desc_color,
+    arrow = arrowhead()
+  ) + 
+  geom_segment(
+    aes(x = 0.1, xend = 0.1, y = 0, yend = cdf(dist_ml, 0.1)),
+    color = desc_color
+  ) +
+  geom_segment(
+    aes(x = 0.1, xend = I(0), y = cdf(dist_ml, 0.1)),
+    color = desc_color,
+    arrow = arrowhead()
+  ) + 
+  geom_segment(
+    aes(x = I(0.09), xend = I(0.09), y = cdf(dist_ml, -0.1), yend = cdf(dist_ml, 0.1)),
+    color = desc_color,
+    arrow = arrowhead(ends = "both")
+  ) +
+  annotate("text", 
+    x = -0.1, y = I(0),
+    label = "−ε",
+    color = desc_color, 
+    vjust = 1.35, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  annotate("text", 
+    x = 0.1, y = I(0),
+    label = "+ε", 
+    color = desc_color, 
+    vjust = 1.35, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  geom_text( 
+    aes(x = I(0.12), y = mean(cdf(dist_ml, c(-0.1, 0.1))[[1]])),
+    label = expression(bold(p ^ "ROPE")), 
+    color = desc_color, 
+    hjust = 0,
+    vjust = 0.6,
+    fontface = "bold", 
+    size = 3
+  ) +
+  scale_x_valence() +
+  coord_cartesian(clip = "off", ylim = c(0,1)) +
+  labs(subtitle = TeX("B. $p^{ROPE}$ on the CDF"))
+
+(
+  (fig_1_prope + rope_line + labs(subtitle = TeX("A. $p^{ROPE}$ on the PDF"))) +
+  fig_ml_cdf_prope
+) &
+  theme(axis.title.y = element_text(angle = 90))
+```
+
 For example, let us imagine that a theory states that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero.
-To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`.
+To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = Pr(-0.1 \le \textrm{Valence effect} \le 0.1) = \Phi(0.1; \beta_1, \tau_1) - \Phi(-0.1; \beta_1, \tau_1)$ = `r p_rope`.
 In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero.
 Note that this statement's validity critically depends on the chosen interval's theoretical validity.
-We visualize this probability in @fig-1 D.
+We visualize this probability in @fig-ml-prope.
 
@@
 
-Armed with the Bayesian draws (@tbl-fit-1), we can now return to the question of the distribution of valence effects in the population.
-We now have 4,000 samples from this heterogeneity distribution's posterior distribution.
-Effectively, then, we have among other quantities 4,000 samples from the heterogeneity distribution's posterior distribution.
-We first redraw the expected heterogeneity distribution from @fig-1 using the posterior mean values of $\beta_1$ and $\tau_1$ in @fig-2 A (thick dark blue curve).
+Armed with the joint Bayesian posterior distribution of all model parameters (@tbl-fit-1), we can now return to the question of the distribution of valence effects in the population.
+We now have 4,000 draws from the posterior distribution of the heterogeneity distribution.
+Within each draw from the posterior distribution, we can perform any calculation we did previously on only the point estimate of the heterogeneity distribution.
+
+We first recompute the expected heterogeneity distribution from @fig-1 using the posterior mean values of $\beta_1$ and $\tau_1$ in @fig-2 A (thick black curve).
 Superimposed on that normal density curve are heterogeneity distributions calculated from 100 random posterior draws of $\beta_1$ and $\tau_1$.
 From these curves we can see that the true distribution of valence effects might well be less or more heterogeneous than is suggested by the expectation (point estimates).
+We can perform the same exercise on the CDF as well (@fig-2 B).
 
@@
 #| label: fig-2
-#| fig-width: 5.4
+#| fig-width: 4.5
 #| fig-height: 2.2
-#| fig-cap: "Bayesian estimates of the heterogeneity distribution of valence effects.
-A: Probability density function (PDF) curves.
-The thick line is the same expected PDF of valence effects from Figure 1.
+#| fig-cap: "Uncertainty in Bayesian estimates of the heterogeneity distribution of valence effects.
+**A.** Probability density function (PDF) curves.
+The thick line is the same expected PDF of valence effects from @fig-1.
 Thin lines show 100 PDFs calculated from random draws of $\\beta_1$ and $\\tau_1$ that collectively illustrate the uncertainty in the distribution's location and spread.
-B: Cumulative density function (CDF) curves, annotated as A.
-C.
-Histogram of (draws from) the posterior distribution of $p^-$, which correspond to CDF segments below zero: This value describes the proportion of individuals with negative valence effects on logRT.
-Point and interval represents the posterior mean and 95\\%CI.
-Solid and dotted vertical lines in A and B highlight x-axis values of 0, and -0.1 and 0.1, respectively."
-
-# Shared Y axis
-scale_y_01 <- scale_y_continuous(
-  breaks = extended_breaks(7),
-  expand = expansion(c(0.005, 0.005)),
-  limits = c(0, 1)
-)
+**B.** Cumulative distribution function (CDF) curves, annotated as in A."
 
 # Posterior samples of pdf and cdf curves
-p_cdf_samples <- post1_density_mean |>
+p_samples <- post1_density_mean |>
   ggplot() +
-  aes(x, cdf) +
-  ylab("Cumulative probability") +
-  scale_x_continuous(
-    "Valence effect",
-    expand = expansion(0),
-    breaks = extended_breaks()
-  ) +
-  scale_y_01 +
-  geom_vline(
-    xintercept = 0, 
-    linewidth = .25
-  ) +
-  geom_vline(
-    xintercept = c(-0.1, 0.1), 
-    lty = "dotted", 
-    linewidth = .2
-  ) +
+  aes(x) +
   geom_line(
@@
     linewidth = .8,
-    col = "dodgerblue4"
-  ) +
-  theme(legend.position = "none")
-
-p_pdf_samples <- p_cdf_samples +
+    col = valence_dist_color
+  ) +
+  xlab("Valence effect") +
+  theme(
+    legend.position = "none",
+    plot.title.position = "plot"
+  )
+
+p_pdf_samples <- p_samples +
   aes(y = pdf) +
+  scale_x_valence() +
   scale_y_continuous(
@@
   ) +
-  ylab("Probability density")
-
-# Add summaries of CDF < 0 to CDF curves as inset plots
-p_p0_samples <- post1 |>
-  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |>
-  ggplot() +
-  scale_x_continuous(
-    expand = expansion(c(0.0, 0.1))
-  ) +
+  ylab("Probability density") +
+  labs(subtitle = "A. Uncertainty in the PDF")
+
+p_cdf_samples <- p_samples +
+  aes(y = cdf) +
+  scale_x_valence() +
   scale_y_01 +
-  theme_void(base_size = 9) +
-  stat_histinterval(
-    aes(y = cdf(dist, 0)),
-    breaks = 50,
-    fill = alpha("dodgerblue1", 0.5),
-    color = "dodgerblue4",
-    side = "right",
-    .width = .95,
-    size = 1,
-    scale = 0.5,
-    justification = 0.025
-  )
-
-p_cdf_samples <- p_cdf_samples +
-  inset_element(
-    p_p0_samples,
-    0, 0, .3, 1, 
-    align_to = "panel", 
-    on_top = TRUE
-  ) +
-  theme(plot.tag.position = c(0.25, 0.75))
-
-(p_pdf_samples | p_cdf_samples) +
-  plot_annotation(tag_levels = "A")
+  ylab(quote(Pr("Valence effect" <= italic(x)))) +
+  labs(subtitle = "B. Uncertainty in the CDF")
+
+p_pdf_samples + p_cdf_samples
 ```
@@
 To carry uncertainty forward from model parameters to the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
-Consequently, we get 4,000 draws from the $HI_{90}$s posterior distribution (@fig-intervals).
+Consequently, we get 4,000 draws from the posterior distribution of $HI_{90}$ (@fig-cdfs-intervals).
+
+```{r}
+#| label: fig-cdfs-intervals
+#| fig-width: 2.5
+#| fig-height: 2.2
+#| fig-cap: "Uncertainty in the limits of the lower and upper limits of the heterogeneity interval (compare to the point estimates in @fig-ml-interval B).
+Thin gray lines show 100 CDFs of the heterogeneity distribution calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
+Blue lines show the calculation of the posterior distribution of $HI_{90} = \\Phi^{-1}([0.05, 0.95]; \\beta_1, \\tau_1)$, which is also depicted as marginal histograms.
+Point estimates and intervals represent the posterior mean and 95\\%CI."
+
+# Add summaries of CDF < 0 to CDF curves as inset plots
+p_cdf_intervals <- p_cdf_samples +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_valence, 0.05), y = 0.05, yend = 0.05),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_valence, 0.05), xend = quantile(dist_valence, 0.05), y = 0.05, yend = 0),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  geom_segment(
+    aes(x = I(0), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0.95),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = quantile(dist_valence, 0.95), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  stat_histinterval(
+    aes(x = quantile(dist_valence, 0.05)),
+    y = 0,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 0.1
+  ) +
+  stat_histinterval(
+    aes(x = quantile(dist_valence, 0.95)),
+    y = 0,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 0.1
+  ) +
+  annotate("text",
+    x = min(x),
+    y = c(0.05, 0.95),
+    vjust = c(0.3, 0.7),
+    label = c(0.05, 0.95),
+    color = desc_color,
+    hjust = 1.1,
+    fontface = "bold",
+    size = 2.5
+  ) +
+  scale_x_valence() +
+  scale_y_continuous(
+    breaks = c(0, 0.5, 1),
+    expand = expansion(0)
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0,1)) +
+  theme(plot.title.position = "panel") +
+  labs(subtitle = TeX("Uncertainty in the limits of $HI_{90}$"))
+
+p_cdf_intervals
+```
 
 Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
-The 95% most plausible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (@fig-intervals B).
+The 95% most plausible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (marginal histograms in @fig-cdfs-intervals, or @fig-intervals B).
 Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.
@@
 #| label: fig-intervals
-#| fig-height: 2.4
-#| fig-cap: Bayesian estimates of the 90% heterogeneity interval of valence effects.
-A.
-Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$.
-B.
-Histograms of 4,000 samples of the HI90 lower (left) and upper (right) limits with their posterior means and 95%CIs as points and intervals.
-C.
-100 random samples from the $HI_{90}$ posterior distribution, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
+#| fig-width: 5.4
+#| fig-height: 2.7
+#| fig-cap: Correlation in Bayesian estimates of the 90% heterogeneity interval of valence effects.
+**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $HI_{90}$.
+**B.** Histograms of 4,000 draws of the $HI_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
+**C.** 100 random draws from the posterior distribution of $HI_{90}$, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
 
@@
   ggplot(aes(hi_low, hi_high)) +
-  labs(x = "HI90 lower limit", y = "HI90 upper limit") +
+  labs(x = TeX("$HI_{90}$ lower limit"), y = TeX("$HI_{90}$ upper limit")) +
   geom_point(
@@
     size = 0.75,
-    col = "dodgerblue1"
-  ) +
-  theme(aspect.ratio = 1)
+    col = desc_color
+  ) +
+  theme(
+    aspect.ratio = 1,
+    plot.title.position = "plot"
+  ) +
+  labs(subtitle = TeX("A. Correlation in the limits of $HI_{90}$"))
 
@@
   ) +
-  scale_x_continuous(
-    "Valence effect"
-  ) +
+  scale_x_valence() +
   stat_histinterval(
     breaks = 100,
-    fill = alpha("dodgerblue1", .33),
-    color = "dodgerblue4",
+    fill = alpha(desc_color, .33),
+    color = desc_color_dark,
     .width = c(.95),
@@
     axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
+    axis.ticks.y = element_blank(),
+    plot.title.position = "plot"
+  ) +
+  coord_cartesian(clip = "off") +
+  xlab(NULL) +
+  labs(subtitle = TeX("B. Uncertainty in the limits of $HI_{90}$"))
 
@@
   ggplot() +
-  scale_x_continuous(
-    "Valence effect",
+  scale_x_valence("Valence effect") +
+  scale_y_continuous(
     expand = expansion(0.025),
-    breaks = extended_breaks()
-  ) +
-  scale_y_continuous(
-    expand = expansion(0.025)
+    labels = NULL
   ) +
@@
     alpha = .5,
-    color = "dodgerblue1"
+    color = desc_color
   ) +
@@
     aes(y = 50, x = b_valence1, xmin = hi_low, xmax = hi_high),
-    color = "dodgerblue4"
-  )
-
-(p1 | p2 | p3) +
-  plot_annotation(tag_levels = "A")
-```
-
-@fig-intervals further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome but also ignores potential correlations between the HI endpoints' posterior distributions (panel A).
+    color = desc_color_dark
+  ) +
+  theme(plot.title.position = "plot", axis.ticks.y = element_blank()) +
+  labs(subtitle = TeX("C. 100 draws from the distribution of $HI_{90}$"))
+
+p1 + 
+((p2 + p3) + plot_layout(heights = c(0.3, 0.7)))
+  # plot_annotation(tag_levels = "A")
+```
+
+@fig-intervals further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome but also ignores potential correlations between the posterior distributions of the HI endpoints (panel A).
 For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2024] the scalar descriptors discussed below.
@@
 In this example, we ask "What proportion of individuals in the population endorse positive words faster than negative words?"
+
+```{r}
+#| label: fig-cdfs-pneg
+#| fig-width: 2.5
+#| fig-height: 2.2
+#| fig-cap: "Uncertainty in the proportion of negative effects (compare to the point estimate in @fig-ml-pneg B).
+Thin gray lines show 100 CDFs of the heterogeneity distribution calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
+Blue lines show the calculation of the posterior distribution of $p^-$, which is also depicted as a marginal histogram at the top left.
+The point estimate and interval represents the posterior mean and 95\\%CI."
+
+# Add summaries of CDF < 0 to CDF curves as inset plots
+p_cdf_samples_with_p_neg <- p_cdf_samples +
+  geom_segment(
+    aes(x = 0, xend = 0, y = 0, yend = cdf(dist_valence, 0)),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.05
+  ) +
+  geom_segment(
+    aes(x = 0, xend = I(0), y = cdf(dist_valence, 0), yend = cdf(dist_valence, 0)),
+    data = post1_samples,
+    color = desc_color,
+    alpha = 0.1
+  ) +
+  stat_histinterval(
+      aes(y = cdf(dist_valence, 0)),
+      x = I(0),
+      data = post1,
+      breaks = 50,
+      fill = alpha(desc_color, 0.35),
+      color = desc_color_dark,
+      side = "left",
+      .width = .95,
+      scale = 1,
+      width = 0.1
+  ) + 
+  annotate(
+    "text", 
+    x = 0, y = I(0), 
+    label = "0", 
+    color = desc_color, 
+    vjust = 1.35, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  geom_text( 
+    aes(x = I(0), y = cdf(dist_valence, 0)),
+    data = post1_mean,
+    label = expression(bold(p ^ "−")), 
+    color = desc_color, 
+    hjust = 1.5, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
+  coord_cartesian(clip = "off") +
+  theme(plot.title.position = "panel") +
+  labs(subtitle = TeX("Uncertainty in $p^{-}"))
+
+p_cdf_samples_with_p_neg
+```
 
@@
 
-To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
-We show 100 posterior draws of the CDF in @fig-2 B with a vertical line superimposed at zero.
+To answer, we calculate $p^- = Pr(\textrm{Valence effect} \le 0) = \Phi(0; \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
+We show 100 posterior draws of the CDF in @fig-cdfs-pneg with a vertical line superimposed at zero.
 The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$).
-We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-2 C, with its associated 95%CI.
+We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-cdfs-pneg, with its associated 95%CI.
 The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
@@
 
+```{r}
+#| label: fig-pdfs-rope
+#| fig-width: 4.6
+#| fig-height: 2.4
+#| fig-cap: "Uncertainty in the proportion of effects in the ROPE (compare to the point estimate in @fig-ml-prope).
+**A.** 25 PDFs of heterogeneity distributions drawn from the joint distribution of $\\beta_1$ and $\\tau_1$.
+In each PDF, the area under the curve within the ROPE of $[-0.1, 0.1]$ is highlighted in blue.
+**B.** Line segments and points show the size of the area in the ROPE for each corresponding PDF in A.
+Marginal histogram (bottom) shows the posterior distribution of $p^{ROPE}$ calculated from all 4,000 draws from the joint distribution of $\\beta_1$ and $\\tau_1$.
+The point estimate and interval represents the posterior mean and 95\\%CI."
+
+rope_n <- 25
+
+post1_samples_rope <- post1_samples |> 
+  slice_head(n = rope_n)
+
+p_rope_densities <- post1_samples_rope |> 
+  ggplot() + 
+  stat_slab(
+    aes(y = factor(.draw), xdist = dist_valence, fill = after_stat(-0.1 < x & x < 0.1)),
+    color = valence_dist_samples_color,
+    alpha = 0.5,
+    scale = 2.5,
+    linewidth = 0.5
+  ) +
+  annotate("text", 
+    x = -0.1, y = I(0),
+    label = "−ε",
+    color = desc_color, 
+    vjust = 1.15, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  annotate("text", 
+    x = 0.1, y = I(0),
+    label = "+ε", 
+    color = desc_color, 
+    vjust = 1.15, 
+    fontface = "bold", 
+    size = 3
+  ) +
+  geom_vline(xintercept = c(-0.1, 0.1), zero_color, alpha = 0.1, linetype = "22") +
+  scale_fill_manual(values = c("TRUE" = desc_color, "FALSE" = alpha(valence_dist_samples_color, 0.3))) +
+  guides(fill = "none") +
+  coord_cartesian(clip = "off") +
+  scale_x_valence() +
+  scale_y_discrete(labels = NULL) +
+  theme(
+    axis.ticks.y = element_blank(),
+    plot.title.position = "plot"
+  ) +
+  ylab("Posterior draw") +
+  xlab("Valence effect") +
+  labs(subtitle = "A. PDFs of heterogeneity distributions")
+
+p_rope_props <- post1_samples_rope |> 
+  slice_head(n = 30) |> 
+  ggplot() + 
+  aes(y = factor(.draw), x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)) + 
+  geom_segment(aes(xend = 0, yend = factor(.draw)), color = desc_color) + 
+  geom_point(color = desc_color, size = 0.75) + 
+  scale_x_continuous(limits = c(0,1), expand = expansion(0)) +
+  scale_y_discrete(labels = NULL) +
+  theme(
+    axis.ticks.y = element_blank(),
+    plot.title.position = "plot"
+  ) +
+  ylab(NULL) +
+  xlab(TeX("$p^{ROPE}$")) +
+  labs(subtitle = TeX("B. Uncertainty in $p^{ROPE}$")) +
+  stat_histinterval(
+    aes(x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)),
+    y = 0.1,
+    data = post1,
+    breaks = 50,
+    fill = alpha(desc_color, 0.35),
+    color = desc_color_dark,
+    side = "bottom",
+    .width = .95,
+    scale = 1,
+    height = 3
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0.7, nrow(post1_samples_rope)))
+
+p_rope_densities + p_rope_props
+```
+
 Moreover, if theory allows defining a range of parameter values that are practically equivalent to zero (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically negligible effects.
-We added dotted vertical lines in @fig-2 to highlight the [-0.1, 0.1] interval, which serves as an example ROPE.
-Line segments within that interval represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$).
-To quantify uncertainty in $p^{ROPE}$ we then aggregate the segments' to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
+Dotted vertical lines in @fig-pdfs-rope A highlight the [-0.1, 0.1] interval, which serves as an example ROPE.
+The area under the PDF within that interval represent the proportion of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$), and each line in @fig-pdfs-rope B depicts the corresponding area under the curve from @fig-pdfs-rope A.
+To quantify uncertainty in $p^{ROPE}$ we then aggregate these values to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero.
 We note that the ROPE of [-0.1, 0.1] here was arbitrary and picked just to illustrate the example.

### 2024-07-05 index.qmd
improved p_rope ridgelines

@@
     aes(y = factor(.draw), xdist = dist_valence, fill = after_stat(-0.1 < x & x < 0.1)),
-    color = valence_dist_samples_color,
-    alpha = 0.5,
+    color = "gray90",
+    # alpha = 0.5,
     scale = 2.5,
@@
   geom_vline(xintercept = c(-0.1, 0.1), zero_color, alpha = 0.1, linetype = "22") +
-  scale_fill_manual(values = c("TRUE" = desc_color, "FALSE" = alpha(valence_dist_samples_color, 0.3))) +
+  scale_fill_manual(values = c("TRUE" = colorspace::lighten(desc_color, 0.35, space = "HLS"), "FALSE" = "gray75")) +
   guides(fill = "none") +

### 2024-07-05 index.qmd
fix typesetting of HI_90

@@
 First, we can use the point estimates in @tbl-lmer to construct an expected *heterogeneity interval* that describes the range within which a certain percentage of the population's slopes are expected to fall [@bolgerCausalProcessesPsychology2019].
-To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$).
+To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($\textit{HI}_{95}$).
 However, because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties.
@@
 #| fig-cap: "Construction of point estimates of the limits of the heterogeneity interval.
-**A.** While $HI_{90}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF), $\\Phi(x; \\beta_1, \\tau_1)$.
-To construct a 90% heterogeneity interval, we pass 0.05 and 0.95 to the inverse of the CDF: $HI_{90} = \\Phi^{-1}\\left([0.05, 0.95]; \\beta_1, \\tau_1 \\right)$."
+**A.** While $\\textit{HI}_{90}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF), $\\Phi(x; \\beta_1, \\tau_1)$.
+To construct a 90% heterogeneity interval, we pass 0.05 and 0.95 to the inverse of the CDF: $\\textit{HI}_{90} = \\Phi^{-1}\\left([0.05, 0.95]; \\beta_1, \\tau_1 \\right)$."
 
@@
 Thus, for a 90% interval, we use 0.05 and 0.95, which together define the central 90% of the distribution.
-Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $HI_{90} = \Phi^{-1}([.05, .95]; \beta_1, \tau_1) = \Phi^{-1}([.05, .95];$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $\textit{HI}_{90} = \Phi^{-1}([.05, .95]; \beta_1, \tau_1) = \Phi^{-1}([.05, .95];$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
 In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the point estimates of the mean ($\beta_1$) and standard deviation ($\tau_1$): We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
@@
 
-[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that heterogeneity can be deemed noteworthy when the ratio of the standard deviation to the average effect is 0.25 or greater: A ratio greater than 1/4 implies a $HI_{95}$ whose limits extend to effects one-half and one-and-a-half times that of the average effect.
+[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that heterogeneity can be deemed noteworthy when the ratio of the standard deviation to the average effect is 0.25 or greater: A ratio greater than 1/4 implies a $\textit{HI}_{95}$ whose limits extend to effects one-half and one-and-a-half times that of the average effect.
 With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r ratio`, suggesting that the degree of heterogeneity in valence effects is noteworthy.
@@
 The expected normal distribution of valence effects and its transformations (@fig-1) ignore uncertainty inherent in the estimated parameters.
-That is, we calculated $HI_{90}$, $p^-$, and the other heterogeneity measures from the point estimates $\beta_1$ = `r mu1[1]` and $\tau_1$ = `r sd1[1]`.
+That is, we calculated $\textit{HI}_{90}$, $p^-$, and the other heterogeneity measures from the point estimates $\beta_1$ = `r mu1[1]` and $\tau_1$ = `r sd1[1]`.
 We did not use any information about the precision, or uncertainty, with which these parameters were estimated.
@@
 Above, we described the heterogeneity interval as a range of values where a specific percentage of the population's slopes are expected to fall (e.g.
-$HI_{90}$ for a 90% heterogeneity interval).
+$\textit{HI}_{90}$ for a 90% heterogeneity interval).
 However, a single interval cannot accommodate the uncertainty with which the underlying parameters are estimated.
-To carry uncertainty forward from model parameters to the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
-Consequently, we get 4,000 draws from the posterior distribution of $HI_{90}$ (@fig-cdfs-intervals).
+To carry uncertainty forward from model parameters to the $\textit{HI}_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
+Consequently, we get 4,000 draws from the posterior distribution of $\textit{HI}_{90}$ (@fig-cdfs-intervals).
 
@@
 Thin gray lines show 100 CDFs of the heterogeneity distribution calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
-Blue lines show the calculation of the posterior distribution of $HI_{90} = \\Phi^{-1}([0.05, 0.95]; \\beta_1, \\tau_1)$, which is also depicted as marginal histograms.
+Blue lines show the calculation of the posterior distribution of $\\textit{HI}_{90} = \\Phi^{-1}([0.05, 0.95]; \\beta_1, \\tau_1)$, which is also depicted as marginal histograms.
 Point estimates and intervals represent the posterior mean and 95\\%CI."
@@
 Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
-The 95% most plausible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (marginal histograms in @fig-cdfs-intervals, or @fig-intervals B).
+The 95% most plausible lower bounds of $\textit{HI}_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (marginal histograms in @fig-cdfs-intervals, or @fig-intervals B).
 Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.
@@
 #| fig-cap: Correlation in Bayesian estimates of the 90% heterogeneity interval of valence effects.
-**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $HI_{90}$.
-**B.** Histograms of 4,000 draws of the $HI_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
-**C.** 100 random draws from the posterior distribution of $HI_{90}$, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
+**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $\textit{HI}_{90}$.
+**B.** Histograms of 4,000 draws of the $\textit{HI}_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
+**C.** 100 random draws from the posterior distribution of $\textit{HI}_{90}$, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
 
@@
     scale = 1,
-    height = 3
+    height = 4
   ) +
@@
 Finally, we turn to the heterogeneity interval (HI).
-The $HI_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task.
-While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $HI_90$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
+The $\textit{HI}_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task.
+While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $\textit{HI}_{90}$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
 Studying @fig-2-person B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.

### 2024-08-19 index.qmd
Draft revisions

@@
 csl: apa.csl
-editor_options: 
-  chunk_output_type: console
 ---
@@
 # Output options
+is_html <- knitr::is_html_output()
 opts_chunk$set(
@@
 The estimated standard deviation of valence effects in the population is `r sd1[1]` log seconds.
-The lme4 software package does not report a standard error or CI for (co)variance parameters by default, and we therefore calculated it by bootstrapping.
+The lme4 software package does not report a standard error or CI for (co)variance parameters by default, and we therefore calculated it by bootstrapping, using lme4's `confint(..., method = "boot")` method.
 The resulting 95% bootstrap CI of the valence effect's standard deviation was [`r sd1[2]`, `r sd1[3]`].
@@
 To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($\textit{HI}_{95}$).
-However, because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties.
-The appropriate percentage to describe with a heterogeneity interval is determined by the substantive and communicative aims at hand; for our illustration 90% seemed reasonable.
+However, because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on 90% heterogeneity interval, and reserve 95% to descriptions of uncertainty, such as confidence or credibility intervals.
+(The appropriate percentage to describe with a heterogeneity interval is determined by the substantive and communicative aims at hand; for our illustration 90% seemed reasonable.)
 
@@
 Assessments of heterogeneity involve combining information about fixed and random effects; to fully incorporate inferential uncertainty then requires accounting for their joint uncertainties.
-Probabilistic, that is, Bayesian methods are uniquely able to address this challenge.
+Bayesian computational methods are uniquely able to address this challenge.
 Modern Bayesian methods, by obtaining draws from the joint posterior distribution of all model parameters presumed to underlie the observed data, allow incorporating posterior uncertainty in combinations of parameters such as the ones highlighted above [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020].
@@
 
-```{r}
+```{r echo = is_html, include = is_html}
 #| label: ppcheck-1
-#| eval: false
+#| fig-cap: "Graphical check of Model 1 performance and posterior predictive density."
 
 # Quick model checking
-pairs(fit_1$fit, pars = variables(fit_1)[1:6])
-mcmc_plot(fit_1, type = "rhat_hist")
-pp_check(fit_1)
+pb <- mcmc_plot(fit_1, type = "rhat_hist") + 
+  scale_x_continuous(labels = ~number(., .001))
+pc <- pp_check(fit_1)
+pb | pc
 ```
@@
 #| label: tbl-samples-1
-#| tbl-cap: Random draws from $\beta_1$, $\tau_1$, and their ratio's posterior.
+#| tbl-cap: Random draws from the posterior distributions of $\beta_1$, $\tau_1$, and their ratio.
 
@@
 We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
-(These, and other details, are presented in our supplementary online analyses.)
+(We present these and other details in our supplementary online analyses.)
 
@@
 ```{r}
-#| label: person-specific-valence-effects
-#| eval: false
-
-# Just sketching a figure of person-specific valence slopes
-coef(fit_lmer)$person |> 
-  as.data.frame() |> 
-  rownames_to_column("person") |> 
-  left_join(
-    coef(fit_1)$person |> 
-      as.data.frame() |> 
-      rownames_to_column("person"),
-    by = join_by(person)
-) |> 
-  select(person, MLE = valence1, Bayes = Estimate.valence1) |> 
-  pivot_longer(-person) |> 
-  ggplot(aes(value, name)) +
-  geom_line(aes(group = person)) +
-  geom_point()
-```
-
-```{r}
 #| label: fig-2
@@
 Thin lines show 100 PDFs calculated from random draws of $\\beta_1$ and $\\tau_1$ that collectively illustrate the uncertainty in the distribution's location and spread.
+Vertical lines on x-axis are estimated slopes for individuals in the sample (posterior means of $\\gamma_1$).
 **B.** Cumulative distribution function (CDF) curves, annotated as in A."
@@
     expand = expansion(c(0.005, 0.005)),
+  ) +
+  geom_rug(
+    data = as_tibble(coef(fit_1)$person[,,"valence1"]) |>
+      rownames_to_column("person"),
+    aes(x = Estimate, y = 0), length = unit(4, "pt"), linewidth = 0.25
   ) +
@@
 
-\clearpage
-
 # Comparing heterogeneity between populations
 
-We now move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of study units.
+We now move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across multiple populations of two different study units: Persons and stimuli.
 To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2024 addressing differences in between-person variability (heterogeneity) in memory performance between a free recall memory task and a cued recall memory task.
@@
 
-Let us now see how our earlier descriptions of heterogeneity can be usefully extended to group differences.
+Let us now see how our earlier descriptions of heterogeneity can be usefully extended to potential differences between populations.
+In addition, we extend the inquiry to incorporate heterogeneity across another source of variance: The target words used in the study [@juddTreatingStimuliRandom2012; @juddExperimentsMoreOne2017].
 We ask three questions about differences in heterogeneity: (1) To what extent is memory performance more variable *between people* in the cued recall task compared to the free recall task?
@@
 To answer these questions, we model the $i$th total recall accuracy in 1 to `r nrow(dat)`, of person $j$ in 1 to `r length(unique(dat$person))`, word $k$ in 1 to `r length(unique(dat$target))`, and task $m$ in {F (free recall), C (cued recall)} as Bernoulli distributed, where the probability of an accurate answer is determined by the rate parameter $\pi$.
-As is common with generalized linear models, we model the rate parameter through a nonlinear link function.
-In this example, we use the cumulative normal distribution function ($\Phi$, or "probit"), but other link functions are also available, such as the common logit.
-Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.
+As is common with generalized linear models, we model $\pi$ using a nonlinear link function.
+In this example, we use the cumulative normal density function ($\Phi$, or probit link), but other link functions could also have been used, such as the logit.
+We then specify the "linear predictor" $\eta$ of this function as a linear combination of the fixed and random effects.
 We write this model as
@@
 
-This Model 2 (@eq-m2-1) of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$).
+This Model 2 (@eq-m2-1) of memory performance is similar to our Model 1 of valence effects above but contains two sources of heterogeneity: persons, whose parameters we represent with $\gamma$, and target words, whose parameters we represent with $\delta$.
 In addition, instead of coding the task type (free recall vs.
-cued recall) using predictor coding schemes such as contrast or dummy coding, we have index-coded task using subscripts $_{m:F}$ to stand for Free recall parameters, and $_{m:C}$ for parameters pertaining to the Cued recall task.
-This reparameterization allows quantifying heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference.
-
-Note also that we model $\gamma$ using two independent normal distributions, and $\delta$ with a multivariate normal distribution.
+cued recall) using predictor coding schemes such as contrast or dummy coding, we have index-coded task using subscripts $_{m:F}$ for Free recall parameters and $_{m:C}$ for Cued recall parameters.
+This parameterization allows us to quantify heterogeneity in memory performance separately for the two tasks.
+
+Note also that while we model $\gamma$ using independent normal distributions for each task, we model $\delta$ with correlated normal distributions.
 Because different persons participated in the two tasks, we cannot assess whether participant-specific abilities are correlated across the tasks.
@@
 
-```{r}
+```{r echo = is_html, include = is_html}
 #| label: ppcheck-2
-#| eval: false
+#| fig-cap: "Graphical check of Model 2 performance and posterior predictive density."
 
 # Quick model checking
-pairs(fit2$fit, pars = variables(fit2)[1:7])
-mcmc_plot(fit2, type = "rhat_hist")
-pp_check(fit2, type = "bars_grouped", group = "task", ndraws = 100)
+pa <- mcmc_plot(fit2, type = "rhat_hist") +
+  scale_x_continuous(labels = ~number(., .001))
+pb <- pp_check(fit2, type = "bars_grouped", group = "task", ndraws = 100)
+pa | pb
 ```
@@
 Descriptively, we reproduced @mahVariabilitySubjectsFree2024's finding that participants' memory performance was more heterogeneous in the cued recall task than in the free recall task (rows 3 and 4 in @tbl-fit-2).
-We show the relevant estimated quantities, and the implied heterogeneity distributions in @fig-2-person.
+We show the relevant estimated quantities and the implied heterogeneity distributions in @fig-2-person.
 
@@
   curve_interval(.width = 0.9)
-```
-
-```{r}
-#| label: model-2-ratio
-#| include: false
-
-# Some ratios are weird
-post2 |> 
-  select(
-    starts_with("b_"), 
-    starts_with("sd_person"), 
-    starts_with("ratio_person"), 
-    -ends_with("_difference")
-  ) |> 
-  select(
-    contains("free")
-  ) |> 
-  filter(ratio_person_free > 10)
+
+tmp_person_dist <- bind_rows(
+  "Free" = post2 |>
+    select(b = b_taskFree, sd = `sd_person__Intercept:taskFree`),
+  "Cued" = post2 |>
+    select(b = b_taskCued, sd = `sd_person__Intercept:taskCued`),
+  .id = "Task"
+) |> 
+  mutate(dist_accuracy = dist_normal(b, sd))
 ```
@@
 $p^+$ indicates the proportion of the population whose proportion correct is predicted to be above 50%.
-Heterogeneity ratio indicates standard deviations divided with their respective means (we truncated this axis at [-7, 7] for clarity.
+Heterogeneity ratio indicates standard deviations divided with their respective means (we truncated this axis at [-5, 5] for clarity).
 B.
@@
     ) |> fct_rev(),
-    value = if_else(type == "Heterogeneity ratio" & abs(value) > 7, NaN, value)
+    value = if_else(type == "Heterogeneity ratio" & abs(value) > 5, NaN, value)
   ) |> 
@@
   stat_histinterval(
-    slab_alpha = .25,
+    slab_alpha = 0.35,
     breaks = 50,
@@
   ggplot(aes(x = x)) +
-  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
+  aes(y = cdf, col = Task, fill = Task, ymin = cdf.lower, ymax = cdf.upper) +
   geom_vline(
@@
 
-p2_cdf_person_inset <- post2 |> 
-  select(pd_person_free, pd_person_cued) |> 
-  pivot_longer(everything()) |> 
-  ggplot() +
-  scale_y_01 +
-  scale_x_continuous(
-    expand = expansion(0)
-  ) +
-  scale_color_brewer(
-    "Recall task",
-    palette = "Dark2",
-    aesthetics = c("color", "fill"),
-    direction = -1
-  ) +
-  stat_slab(
-    aes(y = value, fill = name),
-    side = "right",
+p2_cdf_person <- p2_cdf_person + 
+    stat_histinterval(
+    aes(y = cdf(dist_accuracy, 0), ymin = NULL, ymax = NULL),
+    x = I(0),
+    data = tmp_person_dist,
+    breaks = 50,
+    slab_alpha = 0.35,
+    side = "left",
     .width = .95,
-    size = 1,
-    scale = 0.5,
-    alpha = .2,
-    justification = 0.025
-  ) +
-  stat_pointinterval(
-    aes(y = value, col = name),
-    side = "right",
-    .width = .95,
-    size = 1,
-    justification = 0.025
-  ) +
-  theme_void() +
-  theme(legend.position = "none")
-
-p2_cdf_person <- p2_cdf_person +
-  inset_element(
-    p2_cdf_person_inset,
-    0, 0, .2, 1, 
-    align_to = "panel", 
-    on_top = TRUE
-  )
+    scale = 0.8,
+    size = 0.5,
+    width = 0.1
+  ) +
+  coord_cartesian(clip = "off")
 
@@
 In those areas, a common finding is that while there might be experimental effects on an individual's response distribution's mean or dispersion, the CV frequently remains stable across conditions (disperion tends to grow larger with increased stimulus strength, for example).
-Our use of the heterogeneity ratio in this example calls to mind those applications and findings regarding the coefficient of variation.
 
@@
 We also depict the heterogeneity distribution's posterior distribution as a PDF and a CDF in @fig-2-person B.
-Unlike in @fig-2 where we drew random draws of the functions' posteriors as thin lines, to reduce overplotting @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light areas).
+Unlike in @fig-2 where we represented random draws of the functions' posteriors as thin lines to reduce overplotting, @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light areas).
 These figures allow for concise and complementary descriptions of (differences in) heterogeneity in the two tasks.
@@
 
-## Comparing between-target heterogeneity across tasks
+## Comparing target word heterogeneity across tasks
 
@@
   curve_interval(.width = 0.9)
+
+tmp_target_dist <- bind_rows(
+  "Free" = post2 |>
+    select(b = b_taskFree, sd = `sd_target__taskFree`),
+  "Cued" = post2 |>
+    select(b = b_taskCued, sd = `sd_target__taskCued`),
+  .id = "Task"
+) |> 
+  mutate(dist_accuracy = dist_normal(b, sd))
 ```
@@
 Posterior mean (dark), and 100 posterior draws (light) of the correlation between target words' proportions correct in the free (x-axis) and cued recall (y-axis) tasks.
-Ellipses indicate the 90th percentile of the bivariate normal distribution."
+Ellipses indicate the 90th percentile of the bivariate normal distribution.
+Small black circles are estimated parameters for target words in the sample (posterior means of $\\gamma$)."
 
@@
     size = 1,
-    slab_alpha = 0.2,
+    slab_alpha = 0.35,
     justification = 0.01
@@
 
-p2_cdf_target_inset <- post2 |> 
-  select(pd_target_free, pd_target_cued) |> 
-  pivot_longer(everything()) |> 
-  ggplot() +
-  scale_y_01 +
-  scale_x_continuous(
-    expand = expansion(0)
-  ) +
-  scale_color_brewer(
-    "Recall task",
-    palette = "Dark2",
-    aesthetics = c("color", "fill"),
-    direction = -1
-  ) +
-  stat_slab(
-    aes(y = value, fill = name),
-    side = "right",
+p2_cdf_target <- p2_cdf_target + 
+    stat_histinterval(
+    aes(y = cdf(dist_accuracy, 0), ymin = NULL, ymax = NULL),
+    x = I(0),
+    data = tmp_target_dist,
+    breaks = 50,
+    slab_alpha = 0.35,
+    side = "left",
     .width = .95,
-    size = 1,
-    scale = 0.5,
-    alpha = .2,
-    justification = 0.025
-  ) +
-  stat_pointinterval(
-    aes(y = value, col = name),
-    side = "right",
-    .width = .95,
-    size = 1,
-    justification = 0.025
-  ) +
-  theme_void() +
-  theme(legend.position = "none")
-
-p2_cdf_target <- p2_cdf_target +
-  inset_element(
-    p2_cdf_target_inset,
-    0, 0, .2, 1, 
-    align_to = "panel", 
-    on_top = TRUE
-  )
+    scale = 0.8,
+    size = 0.5,
+    width = 0.1
+  ) +
+  coord_cartesian(clip = "off")
 
@@
   ) +
+  # Estimates for each target word
+  geom_point(
+    data = dat2_target_coefs,
+    aes(x = .Free, y = .Cued),
+    shape = 1
+  ) +
   geom_path(
@@
 
-The results regarding differences in heterogeneity across target words' proportions correct were similar to those as observed regarding heterogeneity in people's memory performances.
+Differences in between target-word heterogeneity were similar to those observed for between-person heterogeneity.
 @fig-2-target A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
-The interpretation of this difference is quantitatively similar to that observed about people above: Both people and target words exhibit greater memory performance variability in the cued recall than in the free recall task.
-However, because @mahVariabilitySubjectsFree2024 used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
-
-Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than 50% accurate recall in the cued recall than in the free recall task.
-Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared very similar across the two tasks.
+Thus, both people and target words exhibit greater memory performance variability in the cued recall than in the free recall task.
+Moreover, this difference holds even when the exact same units---target words, in this example---are used in the two different tasks.
+
+As was the case for between-person heterogeneity, the model predicts the proportion of words that elicit greater than 50% accurate recall to be greater in the cued recall than in the free recall task.
+The ratio of the heterogeneity distribution's standard deviation to its mean again appeared very similar across the two tasks.
 
@@
 One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different recall strategies in the two tasks [@mahVariabilitySubjectsFree2024].
-Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that accuracy is more heterogeneous across target words in cued than in the free recall task (@fig-2-target A).
-Second, we observed across both people and target words that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
+Because we find target words, too, to be more heterogeneous in the Cued task (@fig-2-target A), recall strategies and differences therein might further depend on target words.
+We also observed across both people and target words that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.
 Finally, given that we operationalized the stability of item difficulties as a correlation across tasks, it might be theoretically important to look for sets of stimuli where this positive correlation did not occur.
@@
 
-We emphasized throughout that a probabilistic (Bayesian) approach is well-positioned to answer the needs of researchers interested in heterogeneity.
+We emphasized throughout that the Bayesian approach is well-positioned to answer the needs of researchers interested in heterogeneity.
 Bayesian methods allow carrying uncertainty forward from model parameters to descriptors of heterogeneity and beyond.
 The resulting metrics are useful because they not only convey analysts' expectations regarding heterogeneity, but more fully convey their states of knowledge regarding heterogeneity, including degrees of certainty.
-In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
+In addition, Bayesian modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
 However, some methods described here could be implemented with e.g.

### 2024-08-19 index.qmd
Improve plot consistency
I was unable to align and fit all tags and subtitles so took latter off.

@@
       line = element_line(linewidth = .25),
-      plot.tag = element_text(size = rel(1))
+      plot.tag = element_text(size = rel(1)),
+      plot.tag.location = "margin"
     )
@@
     axis.title.y = element_blank(),
-    axis.title.x = element_text(lineheight = rel(0.5)),
-    plot.title.position = "plot"
+    axis.title.x = element_text(lineheight = rel(0.5))
   ) +
@@
   scale_fill_manual(values = c("grey90", "grey80"), na.value = "white") +
-  labs(subtitle = "A. Heterogeneity distribution") +
-  theme(axis.title.y = element_text(angle = 90))
+  theme(
+    axis.title.y = element_text(angle = 90)
+  )
 
@@
     color = desc_color_dark
-  ) +
-  labs(subtitle = "B. Heterogeneity interval")
+  )
 
@@
     color = desc_color_dark
-  ) +
-  labs(subtitle = "C. Proportion negative")
+  )
 
@@
     color = desc_color_dark
-  ) +
-  labs(subtitle = "D. Proportion in ROPE")
-
-fig_1_params | fig_1_interval | fig_1_pneg | fig_1_prope
+  )
+
+(fig_1_params | fig_1_interval | fig_1_pneg | fig_1_prope) +
+  plot_annotation(tag_levels = "A")
 ```
@@
   xlab("Valence effect") +
-  ylab(quote(Pr("Valence effect" <= italic(x)))) +
-  theme(plot.title.position = "plot")
+  ylab(quote(Pr("Valence effect" <= italic(x))))
 
@@
   scale_x_valence() +
-  coord_cartesian(clip = "off", ylim = c(0,1)) +
-  labs(subtitle = TeX("B. $HI_{90}$ on the CDF"))
-
-(
-  (fig_1_interval + labs(subtitle = TeX("A. $HI_{90}$ on the PDF"))) +
-  fig_ml_cdf_interval
-) &
-  theme(axis.title.y = element_text(angle = 90))
+  coord_cartesian(clip = "off", ylim = c(0,1))
+
+(fig_1_interval | fig_ml_cdf_interval) &
+  theme(axis.title.y = element_text(angle = 90)) &
+  plot_annotation(tag_levels = "A")
 ```
@@
   scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
-  coord_cartesian(clip = "off", ylim = c(0,1)) +
-  labs(subtitle = TeX("B. $p^-$ on the CDF"))
-
-(
-  (fig_1_pneg + zero_line + labs(subtitle = TeX("A. $p^-$ on the PDF"))) +
-  fig_ml_cdf_pneg
-) &
-  theme(axis.title.y = element_text(angle = 90))
+  coord_cartesian(clip = "off", ylim = c(0,1))
+
+((fig_1_pneg + zero_line) + fig_ml_cdf_pneg) &
+  theme(axis.title.y = element_text(angle = 90)) &
+  plot_annotation(tag_levels = "A")
 ```
@@
   scale_x_valence() +
-  coord_cartesian(clip = "off", ylim = c(0,1)) +
-  labs(subtitle = TeX("B. $p^{ROPE}$ on the CDF"))
+  coord_cartesian(clip = "off", ylim = c(0,1))
 
 (
-  (fig_1_prope + rope_line + labs(subtitle = TeX("A. $p^{ROPE}$ on the PDF"))) +
+  (fig_1_prope + rope_line) +
   fig_ml_cdf_prope
 ) &
-  theme(axis.title.y = element_text(angle = 90))
+  theme(axis.title.y = element_text(angle = 90)) &
+  plot_annotation(tag_levels = "A")
 ```
@@
   ) +
-  ylab("Probability density") +
-  labs(subtitle = "A. Uncertainty in the PDF")
+  ylab("Probability density")
 
@@
   scale_y_01 +
-  ylab(quote(Pr("Valence effect" <= italic(x)))) +
-  labs(subtitle = "B. Uncertainty in the CDF")
-
-p_pdf_samples + p_cdf_samples
+  ylab(quote(Pr("Valence effect" <= italic(x))))
+
+(p_pdf_samples + p_cdf_samples) +
+  plot_annotation(tag_levels = "A")
 ```
@@
   ) +
-  coord_cartesian(clip = "off", ylim = c(0,1)) +
-  theme(plot.title.position = "panel") +
-  labs(subtitle = TeX("Uncertainty in the limits of $HI_{90}$"))
+  coord_cartesian(clip = "off", ylim = c(0,1))
 
@@
 #| fig-cap: Correlation in Bayesian estimates of the 90% heterogeneity interval of valence effects.
-**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $\textit{HI}_{90}$.
+**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $\textit{HI}_{90}$ showing their correlation.
 **B.** Histograms of 4,000 draws of the $\textit{HI}_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
@@
   theme(
-    aspect.ratio = 1,
-    plot.title.position = "plot"
-  ) +
-  labs(subtitle = TeX("A. Correlation in the limits of $HI_{90}$"))
+    aspect.ratio = 1
+  )
 
@@
   coord_cartesian(clip = "off") +
-  xlab(NULL) +
-  labs(subtitle = TeX("B. Uncertainty in the limits of $HI_{90}$"))
+  xlab(NULL)
 
@@
   ) +
-  theme(plot.title.position = "plot", axis.ticks.y = element_blank()) +
-  labs(subtitle = TeX("C. 100 draws from the distribution of $HI_{90}$"))
+  theme(plot.title.position = "plot", axis.ticks.y = element_blank())
 
 p1 + 
-((p2 + p3) + plot_layout(heights = c(0.3, 0.7)))
-  # plot_annotation(tag_levels = "A")
+((p2 + p3) + plot_layout(heights = c(0.3, 0.7))) +
+  plot_annotation(tag_levels = "A")
 ```
@@
   coord_cartesian(clip = "off") +
-  theme(plot.title.position = "panel") +
-  labs(subtitle = TeX("Uncertainty in $p^{-}"))
+  theme(plot.title.position = "panel")
 
@@
   ylab("Posterior draw") +
-  xlab("Valence effect") +
-  labs(subtitle = "A. PDFs of heterogeneity distributions")
+  xlab("Valence effect")
 
@@
   xlab(TeX("$p^{ROPE}$")) +
-  labs(subtitle = TeX("B. Uncertainty in $p^{ROPE}$")) +
   stat_histinterval(
@@
 
-p_rope_densities + p_rope_props
+(p_rope_densities + p_rope_props) +
+  plot_annotation(tag_levels = "A")
 ```
@@
 #| fig-cap: Bivariate posterior of $\beta_1$, $\tau_1$, and their ratio.
-Panel A.
+**A**.
 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
-B.
+**B**.
 Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\%CI.
@@
   )
+
 (p1 | p2) &
@@
 #| fig-cap: "Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2.
-A.
+**A**.
 Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95%CIs.
@@
 Heterogeneity ratio indicates standard deviations divided with their respective means (we truncated this axis at [-5, 5] for clarity).
-B.
+**B**.
 Probability density (top) and cumulative distribution functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
@@
 
-p2_pcdf_person <- p2_pdf_person / p2_cdf_person
+p2_pcdf_person <- (p2_pdf_person / p2_cdf_person)
 
@@
 #| fig-cap: "Heterogeneity between target words in memory performance in Free recall and Cued recall tasks from Model 2.
-A.
+**A**.
 Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs.
 Differences calculated as Cued - Free recall.
-B.
+**B**.
 Probability density (top) and cumulative distribution functions (bottom) of the two tasks' heterogeneity distributions (green: free recall, red: cued recall).
 The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance.
-C.
+**C**.
 Posterior mean (dark), and 100 posterior draws (light) of the correlation between target words' proportions correct in the free (x-axis) and cued recall (y-axis) tasks.

### 2024-08-20 index.qmd
Code cleaning

@@
 These tools can be a spur to theory building, allowing researchers to widen their focus from population averages to population heterogeneity in psychological phenomena.
-keywords: [heterogeneity, uncertainty, variation, hierarchical model, statistics]
+keywords: [heterogeneity, uncertainty, variation, multilevel model, statistics, visualization]
 authornote: |
@@
   message = FALSE,
-  echo = FALSE
+  echo = is_html
 )
@@
 # Plotting options
+plot_font <- ifelse(is_html, "Fira Sans", "Libertinus Sans")
+update_geom_defaults("text", list(family = plot_font))
 theme_set(
-  theme_classic(base_size = 9) +
+  theme_classic(base_size = 10, base_family = plot_font) +
     theme(
@@
 Finally, we extend these methods to compare heterogeneity across different populations using an example dataset from cognitive psychology.
+We present the computational notebook supporting this manuscript as an online supplement at {{<meta site-url>}}.
 
@@
 #| label: data-bolgeretal
-#| include: false
 
@@
 contrasts(dat$valence) <- c(-0.5, 0.5)
-contrasts(dat$valence)
 ```
@@
 #| label: tbl-dat1
-#| tbl-cap: First six rows of example dataset 1 [@bolgerCausalProcessesPsychology2019].
+#| tbl-cap: "First six rows of example dataset 1 [@bolgerCausalProcessesPsychology2019]."
 
@@
 The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$, have the subscript $j$ to indicate that they are person-specific deviations from the average intercept and slope, respectively.
-That is, $\beta_0 + \gamma_{01}$ is the intercept (average reaction time) for person $j=1$.
+That is, $\beta_0 + \gamma_{01}$ is the intercept (average log reaction time) for person $j=1$.
 In frequentist texts, these are typically called "random" parameters because they are modeled as varying randomly according to a specified distribution.
@@
 
-In this equation, we assume that the person-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
+In this equation, we assume that the person-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in #eq-m1-2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*).
 Perhaps confusingly, $\tau$s and $\rho$ are also sometimes called random effects because they describe random (co)variations of the person-specific effects.
-To be clear, despite this naming convention they are features of the population, not of any one group or individual.
+To be clear, despite this naming convention they are features of the population, not of any one individual.
 
@@
 #| label: tbl-lmer
-#| tbl-cap: Parameter estimates from Model 1 (ML).
+#| tbl-cap: "Parameter estimates from Model 1 (ML)."
 
@@
 #| label: descriptors
-#| include: false
 
@@
 #| fig-width: 7
-#| fig-cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
-**A.** The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
+#| fig-cap: "Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1.
+**A.** The normal density curve defined by the point estimates of the valence effect distribution's mean ($\\beta_1$) and standard deviation ($\\tau_1$).
 Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean.
-**B.** The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the blue shaded area.
+**B.** The 90\\% Heterogeneity Interval as represented by a line segment with arrows, and the blue shaded area.
 **C.** Proportion of negative valence effects in the population (blue).
-**D.** Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; blue).
+**D.** Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; blue)."
 
 # annotation definitions
-valence_dist_color = "gray25"
-valence_dist_samples_color = "gray65"
-desc_color = "dodgerblue1"
-desc_color_dark = "dodgerblue4"
-zero_color = alpha("black", 0.75)
+valence_dist_color <- "gray25"
+valence_dist_samples_color <- "gray65"
+desc_color <- "dodgerblue1"
+desc_color_dark <- "dodgerblue4"
+zero_color <- alpha("black", 0.75)
 
@@
 fig_1_descriptors <- fig_1 +
-  scale_fill_manual(values = c("grey90", alpha(desc_color, 0.45)), na.value = "white")
+  scale_fill_manual(
+    values = c("grey90", alpha(desc_color, 0.45)), 
+    na.value = "white"
+  )
 
@@
 
-### Interval descriptors
+## Interval descriptors
 
@@
 #| fig-cap: "Construction of point estimates of the limits of the heterogeneity interval.
-**A.** While $\\textit{HI}_{90}$ can be depicted on the probability density function (PDF), **B.** its construction is easier to depict on the cumulative distribution function (CDF), $\\Phi(x; \\beta_1, \\tau_1)$.
+While $\\textit{HI}_{90}$ can be depicted on the probability density function (PDF; **A.**), its construction is easier to depict on the cumulative distribution function (CDF; **B.**), $\\Phi(x; \\beta_1, \\tau_1)$.
 To construct a 90% heterogeneity interval, we pass 0.05 and 0.95 to the inverse of the CDF: $\\textit{HI}_{90} = \\Phi^{-1}\\left([0.05, 0.95]; \\beta_1, \\tau_1 \\right)$."
@@
   geom_segment(
-    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.05), y = 0.05, yend = 0),
+    aes(
+      x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.05), 
+      y = 0.05, yend = 0
+    ),
     color = desc_color
@@
   geom_segment(
-    aes(x = quantile(dist_ml, 0.95), xend = quantile(dist_ml, 0.95), y = 0.95, yend = 0),
+    aes(
+      x = quantile(dist_ml, 0.95), xend = quantile(dist_ml, 0.95), 
+      y = 0.95, yend = 0
+    ),
     color = desc_color
@@
   geom_segment(
-    aes(x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.95), y = 0.025, yend = 0.025),
+    aes(
+      x = quantile(dist_ml, 0.05), xend = quantile(dist_ml, 0.95), 
+      y = 0.025, yend = 0.025
+    ),
     color = desc_color,
@@
     label = expression(bold(HI)[bold("90")]),
-    data = data.frame()[1,],
+    data = data.frame()[1, ],
     size = 3,
@@
   scale_x_valence() +
-  coord_cartesian(clip = "off", ylim = c(0,1))
+  coord_cartesian(clip = "off", ylim = c(0, 1))
 
@@
 Thus, for a 90% interval, we use 0.05 and 0.95, which together define the central 90% of the distribution.
-Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $\textit{HI}_{90} = \Phi^{-1}([.05, .95]; \beta_1, \tau_1) = \Phi^{-1}([.05, .95];$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
+Then, we pass those limits and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to get the interval: $\textit{HI}_{90} = \Phi^{-1}((1 \pm \pi)/2; \beta_1, \tau_1) = \Phi^{-1}([.05, .95];$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`].
 In words, this function calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the point estimates of the mean ($\beta_1$) and standard deviation ($\tau_1$): We expect 90% of valence effects in the population to fall in the [`r hi90`] interval.
@@
 
-### Proportion descriptors
+## Proportion descriptors
 
@@
 In contrast, some applications might find it more informative to summarize proportions of effects above or below some critical value, or within some critical range.
-For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" In other words, we ask a question of *prevalence*: What proportion of the heterogeneity distribution is below zero?
+For example, we might ask "What proportion of individuals in the population endorse faster to positively valenced words?" In other words, we ask a question of *prevalence*: What proportion of the heterogeneity distribution is below zero?
 We label this quantity $p^-$ for proportion of population with negative effects.
@@
 
-zero_line = geom_vline(
+zero_line <- geom_vline(
   xintercept = 0,
@@
   ) + 
-  annotate("text", 
+  annotate(
+    "text", 
     x = 0, y = I(0),
@@
   scale_x_valence(breaks = c(-0.4, -0.2, 0.2)) +
-  coord_cartesian(clip = "off", ylim = c(0,1))
+  coord_cartesian(clip = "off", ylim = c(0, 1))
 
@@
 $p^{-} = Pr(\textrm{Valence effect} \le 0) = \Phi(0; \beta_1, \tau_1) = \Phi(0;$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`.
-This number is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects.
-We illustrate this probability in @fig-1 C and its construction using the CDF in @fig-ml-pneg.
+This number is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population who are expected to endorse positive words faster than negative words.
+We illustrate this probability and its construction using the CDF in @fig-ml-pneg.
 
@@
 
-rope_line = geom_vline(
+rope_line <- geom_vline(
   xintercept = c(-0.1, 0.1),
@@
   geom_segment(
-    aes(x = I(0.09), xend = I(0.09), y = cdf(dist_ml, -0.1), yend = cdf(dist_ml, 0.1)),
+    aes(
+      x = I(0.09), xend = I(0.09), 
+      y = cdf(dist_ml, -0.1), yend = cdf(dist_ml, 0.1)
+    ),
     color = desc_color,
@@
   scale_x_valence() +
-  coord_cartesian(clip = "off", ylim = c(0,1))
-
-(
-  (fig_1_prope + rope_line) +
-  fig_ml_cdf_prope
-) &
+  coord_cartesian(clip = "off", ylim = c(0, 1))
+
+((fig_1_prope + rope_line) + fig_ml_cdf_prope) &
   theme(axis.title.y = element_text(angle = 90)) &
@@
 Note that this statement's validity critically depends on the chosen interval's theoretical validity.
-We visualize this probability in @fig-ml-prope.
-
-### Ratio descriptors
+We visualize this probability, and how it is constructed using the CDF, in @fig-ml-prope.
+
+## Ratio descriptors
 
@@
 
-The expected normal distribution of valence effects and its transformations (@fig-1) ignore uncertainty inherent in the estimated parameters.
+The expected normal distribution of valence effects and its transformations ignore uncertainty inherent in the estimated parameters.
 That is, we calculated $\textit{HI}_{90}$, $p^-$, and the other heterogeneity measures from the point estimates $\beta_1$ = `r mu1[1]` and $\tau_1$ = `r sd1[1]`.
@@
 
-```{r echo = is_html, include = is_html}
+```{r include = is_html}
 #| label: ppcheck-1
+#| fig-height: 2.4
 #| fig-cap: "Graphical check of Model 1 performance and posterior predictive density."
@@
 #| label: tbl-samples-1
-#| tbl-cap: Random draws from the posterior distributions of $\beta_1$, $\tau_1$, and their ratio.
+#| tbl-cap: "Random draws from the posterior distributions of $\\beta_1$, $\\tau_1$, and their ratio."
 
@@
 @dplyr2023; @rcoreteamLanguageEnvironmentStatistical2023].
-Here, we used the R package brms [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] to specify the model and then draw random samples from its posterior distribution.
+Here, we used the R package brms [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] to specify the model and then sample random draws from its posterior distribution.
 The MCMC estimation algorithm completed in about 5 seconds on a modern laptop.
 We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].
-(We present these and other details in our supplementary online analyses.)
+(We present these and other details in our [online supplement](https://mvuorre.github.io/heterogeneity-uncertainty/).)
 
@@
 #| label: tbl-fit-1
-#| tbl-cap: Parameter estimates from Model 1 (Bayes).
+#| tbl-cap: "Parameter estimates from Model 1 (Bayes)."
 
@@
 
-Armed with the joint Bayesian posterior distribution of all model parameters (@tbl-fit-1), we can now return to the question of the distribution of valence effects in the population.
-We now have 4,000 draws from the posterior distribution of the heterogeneity distribution.
+Armed with the joint Bayesian posterior distribution of all model parameters, we can now returned to the question of the distribution of valence effects in the population.
+We have 4,000 draws from the posterior distribution of the heterogeneity distribution.
 Within each draw from the posterior distribution, we can perform any calculation we did previously on only the point estimate of the heterogeneity distribution.
@@
 From these curves we can see that the true distribution of valence effects might well be less or more heterogeneous than is suggested by the expectation (point estimates).
-We can perform the same exercise on the CDF as well (@fig-2 B).
+We then perform the same exercise on the CDF as well (@fig-2 B).
 
@@
   geom_rug(
-    data = as_tibble(coef(fit_1)$person[,,"valence1"]) |>
+    data = as_tibble(coef(fit_1)$person[, , "valence1"]) |>
       rownames_to_column("person"),
@@
 However, a single interval cannot accommodate the uncertainty with which the underlying parameters are estimated.
-To carry uncertainty forward from model parameters to the $\textit{HI}_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
+To carry uncertainty forward from model parameters to the $\textit{HI}_{90}$, we repeat the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we redo the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$.
 Consequently, we get 4,000 draws from the posterior distribution of $\textit{HI}_{90}$ (@fig-cdfs-intervals).
@@
 Thin gray lines show 100 CDFs of the heterogeneity distribution calculated from random draws of $\\beta_1$ and $\\tau_1$, as in @fig-2 B.
-Blue lines show the calculation of the posterior distribution of $\\textit{HI}_{90} = \\Phi^{-1}([0.05, 0.95]; \\beta_1, \\tau_1)$, which is also depicted as marginal histograms.
+Blue lines show the calculation of the posterior distribution of $\\textit{HI}_{90} = \\Phi^{-1}([0.05, 0.95]; \\beta_1, \\tau_1)$, whose limits are also depicted as marginal histograms below the x-axis.
 Point estimates and intervals represent the posterior mean and 95\\%CI."
@@
   geom_segment(
-    aes(x = quantile(dist_valence, 0.05), xend = quantile(dist_valence, 0.05), y = 0.05, yend = 0),
+    aes(
+      x = quantile(dist_valence, 0.05), 
+      xend = quantile(dist_valence, 0.05), 
+      y = 0.05, yend = 0
+    ),
     data = post1_samples,
@@
   geom_segment(
-    aes(x = I(0), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0.95),
+    aes(
+      x = I(0), xend = quantile(dist_valence, 0.95), 
+      y = 0.95, yend = 0.95
+    ),
     data = post1_samples,
@@
   geom_segment(
-    aes(x = quantile(dist_valence, 0.95), xend = quantile(dist_valence, 0.95), y = 0.95, yend = 0),
+    aes(
+      x = quantile(dist_valence, 0.95), 
+      xend = quantile(dist_valence, 0.95), 
+      y = 0.95, yend = 0
+    ),
     data = post1_samples,
@@
   ) +
-  coord_cartesian(clip = "off", ylim = c(0,1))
+  coord_cartesian(clip = "off", ylim = c(0, 1))
 
@@
 Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds.
-The 95% most plausible lower bounds of $\textit{HI}_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (marginal histograms in @fig-cdfs-intervals, or @fig-intervals B).
+The 95% most plausible lower bounds of $\textit{HI}_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (marginal histograms in @fig-cdfs-intervals and @fig-intervals B).
 Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.
@@
 #| fig-height: 2.7
-#| fig-cap: Correlation in Bayesian estimates of the 90% heterogeneity interval of valence effects.
-**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $\textit{HI}_{90}$ showing their correlation.
-**B.** Histograms of 4,000 draws of the $\textit{HI}_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
-**C.** 100 random draws from the posterior distribution of $\textit{HI}_{90}$, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.
+#| fig-cap: "Correlation in Bayesian estimates of the 90% heterogeneity interval of valence effects.
+**A.** Scatterplot of 4,000 posterior draws of the lower (x-axis) and upper (y-axis) limits of $\\textit{HI}_{90}$ showing their correlation.
+**B.** Histograms of 4,000 draws of the $\\textit{HI}_{90}$ lower (left) and upper (right) limits with their posterior means and 95% CIs as points and intervals.
+**C.** 100 random draws from the posterior distribution of $\\textit{HI}_{90}$, with the posterior mean heterogeneity interval superimposed in a darker shade of blue."
 
@@
   arrange(b_valence1) |>
-  mutate(i = 1:n()) |>
+  mutate(i = seq_len(n())) |>
   mutate(
@@
 For example, we can calculate proportions with negative and positive effects by using zero as the critical value.
-In this example, we ask "What proportion of individuals in the population endorse positive words faster than negative words?"
+In this example, we asked "What proportion of individuals in the population endorse positive words faster than negative words?"
 
@@
   geom_segment(
-    aes(x = 0, xend = 0, y = 0, yend = cdf(dist_valence, 0)),
+    aes(
+      x = 0, xend = 0, 
+      y = 0, yend = cdf(dist_valence, 0)
+    ),
     data = post1_samples,
@@
   geom_segment(
-    aes(x = 0, xend = I(0), y = cdf(dist_valence, 0), yend = cdf(dist_valence, 0)),
+    aes(
+      x = 0, xend = I(0), 
+      y = cdf(dist_valence, 0), yend = cdf(dist_valence, 0)
+    ),
     data = post1_samples,
@@
   stat_histinterval(
-      aes(y = cdf(dist_valence, 0)),
+    aes(
+      y = cdf(dist_valence, 0)),
       x = I(0),
@@
     pd_r = str_glue(
-      "{percent(pd, .1)} [{percent(pd.lower, .1)}, {percent(pd.upper, .1)}]"
+      "{percent(pd, .1)} ",
+      "[{percent(pd.lower, .1)}, {percent(pd.upper, .1)}]"
     ),
     pd_r_n = str_glue(
-      "{percent(1-pd, .1)} [{percent(1-pd.upper, .1)}, {percent(1-pd.lower, .1)}]"
+      "{percent(1-pd, .1)} ",
+      "[{percent(1-pd.upper, .1)}, {percent(1-pd.lower, .1)}]"
     ),
     rope_r = str_glue(
-      "{percent(rope, .1)} [{percent(rope.lower, .1)}, {percent(rope.upper, .1)}]"
+      "{percent(rope, .1)} ",
+      "[{percent(rope.lower, .1)}, {percent(rope.upper, .1)}]"
     ),
     rope_r_n = str_glue(
-      "{percent(1-rope, .1)} [{percent(1-rope.upper, .1)}, {percent(1-rope.lower, .1)}]"
+      "{percent(1-rope, .1)} ",
+      "[{percent(1-rope.upper, .1)}, {percent(1-rope.lower, .1)}]"
     ),
     ratio_r = str_glue(
-      "{number(ratio, .01)} [{number(ratio.lower, .01)}, {number(ratio.upper, .01)}]"
+      "{number(ratio, .01)} ",
+      "[{number(ratio.lower, .01)}, {number(ratio.upper, .01)}]"
     )
@@
 To answer, we calculate $p^- = Pr(\textrm{Valence effect} \le 0) = \Phi(0; \beta_1, \tau_1)$ for each posterior draw of $\beta_1$ and $\tau_1$.
-We show 100 posterior draws of the CDF in @fig-cdfs-pneg with a vertical line superimposed at zero.
+@fig-cdfs-pneg shows 100 posterior draws of the CDF with a vertical line superimposed at zero.
 The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$).
-We also show a histogram of all 4,000 posterior draws of that proportion in the top left margin of @fig-cdfs-pneg, with its associated 95%CI.
+We also show a histogram of all 4,000 posterior draws of that proportion on the y-axis of @fig-cdfs-pneg, with its associated 95%CI.
 The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`.
@@
   stat_slab(
-    aes(y = factor(.draw), xdist = dist_valence, fill = after_stat(-0.1 < x & x < 0.1)),
+    aes(
+      y = factor(.draw), 
+      xdist = dist_valence, 
+      fill = after_stat(-0.1 < x & x < 0.1)
+    ),
     color = "gray90",
-    # alpha = 0.5,
     scale = 2.5,
-    linewidth = 0.5
+    linewidth = 0.25
   ) +
@@
   ) +
-  geom_vline(xintercept = c(-0.1, 0.1), zero_color, alpha = 0.1, linetype = "22") +
-  scale_fill_manual(values = c("TRUE" = colorspace::lighten(desc_color, 0.35, space = "HLS"), "FALSE" = "gray75")) +
+  geom_vline(
+    xintercept = c(-0.1, 0.1), 
+    color = zero_color, 
+    alpha = 0.1, 
+    linetype = "22"
+  ) +
+  scale_fill_manual(
+    values = c(
+      "TRUE" = colorspace::lighten(desc_color, 0.35, space = "HLS"), 
+      "FALSE" = "gray75"
+    )
+  ) +
   guides(fill = "none") +
@@
   ggplot() + 
-  aes(y = factor(.draw), x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)) + 
+  aes(
+    y = factor(.draw), 
+    x = cdf(dist_valence, 0.1) - cdf(dist_valence, -0.1)
+  ) + 
   geom_segment(aes(xend = 0, yend = factor(.draw)), color = desc_color) + 
   geom_point(color = desc_color, size = 0.75) + 
-  scale_x_continuous(limits = c(0,1), expand = expansion(0)) +
+  scale_x_continuous(limits = c(0, 1), expand = expansion(0)) +
   scale_y_discrete(labels = NULL) +
@@
 
-### Ratio descriptors
+## Ratio descriptors
 
 Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$ (see @tbl-samples-1).
-@fig-ratio shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (panel B).
+@fig-ratio A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (panel B).
 The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence.
@@
 #| fig-width: 4.4
-#| fig-cap: Bivariate posterior of $\beta_1$, $\tau_1$, and their ratio.
+#| fig-cap: "Bivariate posterior of $\\beta_1$, $\\tau_1$, and their ratio.
 **A**.
-4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$).
+4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\\beta_1$) and standard deviation ($\\tau_1$).
 **B**.
-Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\%CI.
+Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its location $\\frac{\\tau_1}{\\beta_1}$, and its posterior mean and 95\\%CI."
 
@@
 #| label: data-mah-lindsay-3
-#| include: false
 
@@
     person = factor(idnumber),
-    task = factor(test_type, levels = c("FR", "CR"), labels = c("Free", "Cued")),
+    task = factor(
+      test_type, 
+      levels = c("FR", "CR"), 
+      labels = c("Free", "Cued")
+    ),
     accuracy
@@
 As is common with generalized linear models, we model $\pi$ using a nonlinear link function.
-In this example, we use the cumulative normal density function ($\Phi$, or probit link), but other link functions could also have been used, such as the logit.
+In this example, we use the cumulative normal distribution function ($\Phi$, or probit link), but other link functions could also have been used, such as the logit.
 We then specify the "linear predictor" $\eta$ of this function as a linear combination of the fixed and random effects.
@@
 
-```{r echo = is_html, include = is_html}
+```{r include = is_html}
 #| label: ppcheck-2
+#| fig-height: 2.8
 #| fig-cap: "Graphical check of Model 2 performance and posterior predictive density."
@@
 #| label: tbl-fit-2
-#| tbl-cap: Parameter estimates from Model 2.
+#| tbl-cap: "Parameter estimates from Model 2."
 
@@
       `sd_target__taskCued` - `sd_target__taskFree`,
+    sd_target_ratio = 
+      `sd_target__taskCued` / `sd_target__taskFree`,
     pd_target_difference = pd_target_cued - pd_target_free,
@@
   mutate(
-    x = str_glue("{number(mean, .01)} [{number(q2.5, .01)}, {number(q97.5, .01)}]")
+    x = str_glue(
+      "{number(mean, .01)} ",
+      "[{number(q2.5, .01)}, {number(q97.5, .01)}]")
   )
@@
 Therefore, for example zero translates to 50% accuracy.
+
+More importantly, the second row in @fig-2-person A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
+The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
+Note that our estimate of the heterogeneity difference is greater, and associated with greater uncertainty, than what was originally reported by @mahVariabilitySubjectsFree2024, potentially because our model also includes heterogeneity across target words.
+
+The third row of @fig-2-person A shows the estimated proportions of individuals whose memory performance exceeded 50% ($p^+$), and their difference between tasks.
+The model estimates the proportion of individuals who recall over 50% of items to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
+Note that this quantity refers to population proportions and is not a z-score.
 
@@
 Probability density (top) and cumulative distribution functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall).
-The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95%CIs, of the proportions of the populations with memory performance above 0.5."
+The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95%CIs, of the proportions of the populations with memory performance above 0.5.
+Densities, points, and intervals on the x-axis of the bottom panel indicate approximate posterior densities, with means and 95%CIs, of the 90\\% heterogeneity interval's lower (left) and upper (right) bounds."
 
@@
     legend.position = "none",
+    panel.spacing = unit(0, "pt"),
     axis.title.y = element_blank()
@@
     legend.position = "none",
+    panel.spacing = unit(0, "pt"),
     panel.border = element_rect(fill = NA)
@@
   stat_histinterval(
-    aes(y = cdf(dist_accuracy, 0), ymin = NULL, ymax = NULL),
+    aes(y = 1 - cdf(dist_accuracy, 0), ymin = NULL, ymax = NULL),
     x = I(0),
@@
   ) +
-  coord_cartesian(clip = "off")
+  stat_histinterval(
+    aes(x = quantile(dist_accuracy, 0.05), ymin = NULL, ymax = NULL),
+    y = I(0),
+    data = tmp_person_dist,
+    slab_alpha = 0.35,
+    breaks = 50,
+    side = "bottom",
+    .width = .95,
+    size = 0.5,
+    scale = 0.8,
+    height = 0.35
+  ) +
+    stat_histinterval(
+    aes(x = quantile(dist_accuracy, 0.95), ymin = NULL, ymax = NULL),
+    y = I(0),
+    data = tmp_person_dist,
+    slab_alpha = 0.35,
+    breaks = 50,
+    side = "bottom",
+    .width = .95,
+    size = 0.5,
+    scale = 0.8,
+    height = 0.35
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0, 1))
 
@@
 
-More importantly, the second row in @fig-2-person A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall).
-The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).
-
-The third row of @fig-2-person A shows the estimated proportions of individuals whose memory performance exceeded 50% ($p^+$), and their difference.
-The model estimates the proportion of individuals who recall over 50% of items to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task.
-Note that this quantity refers to population proportions and is not a z-score.
-
 Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom row of @fig-2-person A shows that the degree of relative heterogeneity is virtually identical across the two tasks.
 This heterogeneity ratio's mathematical equivalent is commonly known as the coefficient of variation (CV), which is used frequently in many areas of psychological research, such as psychophysics.
-In those areas, a common finding is that while there might be experimental effects on an individual's response distribution's mean or dispersion, the CV frequently remains stable across conditions (disperion tends to grow larger with increased stimulus strength, for example).
-
-Moreover, we truncated the Heterogeneity ratio panel's x-axis at [-7, 7] because ratios of two normal distributions with zero means are Cauchy distributed.
+In those areas, a common finding is that while there might be experimental effects on an individual's response distribution's mean or dispersion, the CV frequently remains stable across conditions (dispersion tends to grow larger with increased stimulus strength, for example).
+Our results show that this coefficient of variation in individual's memory abilities remains stable across conditions that lead to different average memory performances.
+
+We truncated the Heterogeneity ratio panel's x-axis at [-5, 5] because ratios of two normal distributions with zero means are Cauchy distributed.
 Sampling from a Cauchy distribution frequently returns extreme draws because of the distribution's thick tails.
@@
 We also depict the heterogeneity distribution's posterior distribution as a PDF and a CDF in @fig-2-person B.
-Unlike in @fig-2 where we represented random draws of the functions' posteriors as thin lines to reduce overplotting, @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light areas).
+Unlike in @fig-2 where we represented random draws of the functions' posteriors as thin lines, @fig-2-person instead aggregates the posteriors to means (dark line) and 95% credibility ribbons (light areas) to reduce overplotting.
 These figures allow for concise and complementary descriptions of (differences in) heterogeneity in the two tasks.
@@
 This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-50% performance only for a proportion of `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population.
-Second, we see that the slope of the cued recall CDF (red) is less steep and to the right to that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task, and `r post2_sum$x[post2_sum$variable=="pd_person_cued"]` individuals in that group are predicted to perform above 50%.
+Second, we see that the slope of the cued recall CDF (red) is less steep and to the right to that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.
+The proportion of individuals in the cued recall task who are predicted to perform above 50% was `r post2_sum$x[post2_sum$variable=="pd_person_cued"]`.
 
 Finally, we turn to the heterogeneity interval (HI).
-The $\textit{HI}_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task.
-While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $\textit{HI}_{90}$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
-Studying @fig-2-person B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.
+The $\textit{HI}_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task (leftmost green and red densities on bottom x-axis of @fig-2-person, respectively).
+While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles (rightmost green and red densities on bottom x-axis of @fig-2-person, respectively) differed at the 95% credibility level (the Cued recall upper $\textit{HI}_{90}$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater).
+Studying @fig-2-person B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performance in the cued recall condition (although this difference was not credibly different from zero).
 
@@
 The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance.
+Densities, points, and intervals on the x-axis of the bottom panel indicate approximate posterior densities, with means and 95%CIs, of the 90\\% heterogeneity interval's lower (left) and upper (right) bounds.
 **C**.
@@
   ) +
-  coord_cartesian(clip = "off")
+  stat_histinterval(
+    aes(x = quantile(dist_accuracy, 0.05), ymin = NULL, ymax = NULL),
+    y = I(0),
+    data = tmp_target_dist,
+    slab_alpha = 0.35,
+    breaks = 50,
+    side = "bottom",
+    .width = .95,
+    size = 0.5,
+    scale = 0.8,
+    height = 0.3
+  ) +
+    stat_histinterval(
+    aes(x = quantile(dist_accuracy, 0.95), ymin = NULL, ymax = NULL),
+    y = I(0),
+    data = tmp_target_dist,
+    slab_alpha = 0.35,
+    breaks = 50,
+    side = "bottom",
+    .width = .95,
+    size = 0.5,
+    scale = 0.8,
+    height = 0.3
+  ) +
+  coord_cartesian(clip = "off", ylim = c(0, 1))
 
@@
 
-(wrap_elements(full = p2_pars_target) | wrap_elements(full = p2_pcdf_target) | p2_ellipse) +
+(
+  wrap_elements(full = p2_pars_target) | 
+    wrap_elements(full = p2_pcdf_target) | 
+    p2_ellipse
+) +
   plot_annotation(tag_levels = "A")
@@
 Differences in between target-word heterogeneity were similar to those observed for between-person heterogeneity.
-@fig-2-target A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task.
+@fig-2-target A shows that heterogeneity in memory performance was greater when words appeared in the cued recall task (the standard deviation was `r post2_sum$x[post2_sum$variable=="sd_target_difference"]` probits greater in the cued recall task [ratio: `r post2_sum$x[post2_sum$variable=="sd_target_ratio"]`]).
 Thus, both people and target words exhibit greater memory performance variability in the cued recall than in the free recall task.
@@
 
-As was the case for between-person heterogeneity, the model predicts the proportion of words that elicit greater than 50% accurate recall to be greater in the cued recall than in the free recall task.
-The ratio of the heterogeneity distribution's standard deviation to its mean again appeared very similar across the two tasks.
+As was the case for between-person heterogeneity, the model predicts the proportion of words that elicit greater than 50% accurate recall to be greater in the cued recall (`r post2_sum$x[post2_sum$variable=="pd_target_cued"]`) than in the free recall task (`r post2_sum$x[post2_sum$variable=="pd_target_free"]`; difference: `r post2_sum$x[post2_sum$variable=="pd_target_difference"]`).
+The ratio of the heterogeneity distribution's standard deviation to its mean was again very similar across the two tasks (`r post2_sum$x[post2_sum$variable=="ratio_target_difference"]`).
 
@@
 In addition, Bayesian modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups.
+Our online supplement illustrates these techniques in detail.
 However, some methods described here could be implemented with e.g.
@@
 
-The online analysis supplement is readable at {{<meta site-url>}}.
-Our materials are available at GitHub ({{<meta repo-url>}}) and the OSF ({{<meta osf-url>}}).
+The online supplement is at {{<meta site-url>}}.
+Our materials are available on GitHub ({{<meta repo-url>}}) and the OSF ({{<meta osf-url>}}).
 We reused openly available datasets from @bolgerCausalProcessesPsychology2019 and @mahVariabilitySubjectsFree2024.
@@
 
-<!-- Format references better in non-html formats -->
+<!-- HTML format automatically adds "References" in an appendix. -->
 ::: {.content-hidden when-format="html"}
@@
 :::
-
-<!-- Material after this only appears in html output -->
-::: {.content-visible when-format="html"}
-
-:::

### 2024-08-26 index.qmd
Fix figure

@@
 p2_pars_target <- post2 |> 
-  select(contains("target")) |> 
+  select(contains("target"), -sd_target_ratio) |> 
   rename_with(
@@
     ) |> fct_rev(),
-    value = if_else(type == "Heterogeneity ratio" & abs(value) > 5, NaN, value)
+    value = if_else(
+      type == "Heterogeneity ratio" & abs(value) > 5, 
+      NaN, 
+      value
+    )
   ) |> 
@@
   ggplot(aes(x = x)) +
-  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
+  aes(
+    y = cdf, ymin = cdf.lower, ymax = cdf.upper, 
+    col = Task, fill = Task
+  ) +
   scale_color_brewer(

### 2024-09-03 index.qmd
Re-render site

@@
 library(ellipse)
-library(tinytable)
 library(cmdstanr)
@@
   brms.backend = Sys.getenv("BRMS_BACKEND", "rstan"), 
-  brms.threads = as.numeric(Sys.getenv("MAX_CORES"), 1),
+  brms.threads = as.numeric(Sys.getenv("BRMS_THREADS"), 1),
   mc.cores = as.numeric(Sys.getenv("MAX_CORES"), 4)
@@
   file = "models/brm-fit-1",
-  control = list(adapt_delta = .95)
 )
@@
 pb | pc
+```
+
+```{r include = is_html}
+post1 |> 
+  head() |> 
+  as_tibble() |> 
+  select(-contains(".")) |> 
+  select(2, 4, 7) |> 
+  mutate(
+    hi90 = str_glue(
+      "[{number(quantile(dist_valence, .05), .01)}, ",
+      "{number(quantile(dist_valence, .95), .01)}]"
+      ),
+    pp = cdf(dist_valence, 0),
+    prope = cdf(dist_valence, .1) - cdf(dist_valence, -.1),
+    ratio = sd_person__valence1 / b_valence1
+  ) |> 
+  mutate(dist_valence = as.character(dist_valence)) |> 
+  # setNames(c(m1_parnames, "$p(\\beta_1)$")) |> 
+  setNames(
+    c(
+      m1_parnames[c(2, 4)], 
+      "$p(\\beta_1)$", "$HI_{90}$", "$p^+$", "$p^{ROPE}$", "$\\frac{\\tau_1}{\\beta_1}$"
+    )
+  ) |> 
+  kable(digits = 2)
 ```
@@
 #| label: fig-2-target
-#| fig-height: 4
+#| fig-height: 3.6
 #| fig-width: 7.4
@@
 ::: {.content-hidden when-format="html"}
-# References
+# References {.unnumbered}
 
